### 2019-05-29 old_outline.tex
finished about 2/3 of the outline for the prelim;

\documentclass{article}
\title{Prelim Outline}
\author{Shivesh Pathak}
\usepackage[margin=1.0in]{geometry}
\usepackage{graphicx}
\usepackage{float}
\begin{document}

\section{Introduction and methods}
\textit{The goal of the introduction is to show the reader that the project is valuable. Should be x pages, other 5-x page will be methods.}

\paragraph{Overall goal}
We want to understand the low energy behavior of the cuprate superconductors.
\\
\textit{This paragraph will be outlining the various phenomenon attributed with the cuprates, e.g. the many phases.}

\paragraph{Barrier to achieving goal}
The strong coupling between magnetic, charge, and lattice degrees of freedom in the cuprates makes writing down an accurate low energy effective theory very challenging.
\\
\textit{This paragraph will outline the many experiments that show the strong coupling between these many degrees of freedom.}

\paragraph{Current state of the art}
Many theoretical approaches to generating a low energy model for the cuprates have been attempted with varying degrees of success.
\\
\textit{This paragraph will outline the various approaches people have tried to generate effective theories. Hole doped, electron doped, pen and paper, computational, energetics, one- and two-body properties.}

\paragraph{Advancement to the state of the art}
We propose that a density matrix downfolding approach using \textit{ab initio} quantum Monte Carlo calculations can generate an effective model which can accurately describe low-energy properties and energetics for the infinite layer electron-doped cuprates at zero temperature and dopings of x=0, 0.125, 0.25.
\\
\textit{This paragraph will describe in detail each of the conditions in the thesis sentence.}

\paragraph{Density matrix down folding} 
The density matrix downfolding method allows us to downfold a first-principles Hamiltonian into a low-energy effective theory by mapping the downfolding problem to a linear regression problem which uses \textit{ab-initio} energies and reduced density matrices of wave functions sampled from a low-energy subspace of the full Hilbert space.
\\
\textit{This paragraph should outline the details of the DMD method, with parameter derivatives. No reference to QMC yet.}

\paragraph{Fixed node diffusion Monte Carlo}
We will use fixed node diffusion Monte Carlo (FN-DMC) to generate wave functions in our chosen low-energy subspace and to accurately calculate \textit{ab-initio} energies and reduced density matrices.
\\
\textit{This paragraph should outline how we will use FN-DMC (with DFT) to generate the low energy states, their energies, and their RDMs}

%\begin{figure}[H]
%\centering
%\includegraphics[width=0.6\textwidth]{Figures/I1-phase_diagram.pdf}
%\caption{\label{fig1} Phase diagram of Sr$_{1-x}$La$_x$CuO$_2$ for %various electron doping \textit{x} and temperature \textit{T}. Figure %taken from [ref].}
%\end{figure}

%\begin{figure}[H]
%\centering
%\includegraphics[width=0.9\textwidth]{Figures/I2-DMD_diagram.pdf}
%\caption{\label{fig2} Schematic drawing of the the density matrix %downfolding method, describing how the wave function samples, total %energies, and descriptor values are mapped onto a linear regression %problem.}
%\end{figure}

\pagebreak

\section{Preliminary results and discussion}
\textit{The goal of the preliminary results is to indicate that I have the tools to actually complete the proposed research.}

\paragraph{NO determinants: Introduction} In order to become acquainted with our code  and QMC algorithms in general I worked on implementing and testing non-orthogonal orbital optimization in QWalk. 
\\
\textit{This paragraph should introduce the idea of non orthogonal orbital optimization.}

\paragraph{NO determinants: Work} The main work involved comparing the FN-DMC energies and single particle densities when using trial wave functions with and without non-orthogonal orbital optimization.
\\
\textit{This paragraph should include details about how we actually went about doing these calculations.}

\paragraph{NO determinants: Results and Discussion} We found that in addition to the improvements from orthogonal optimization, non-orthogonal orbital optimization yields improvements to the ground state FN-DMC energies and single particle properties.
\\
\textit{This paragraph should discuss the actual results from the calculations outlined above.}

\paragraph{SCO SCF: Introduction} To get a grasp on the relevant energy scales, model terms, and values for model parameters before trying the computationally expensive DMD with FN-DMC we attempted an \textit{ad hoc} fitting procedure using only DFT eigenvalues and total energies. 
\\
\textit{This paragraph should talk about the fitting procedure that we are planning on using.}

\paragraph{SCO SCF: Work} The task at hand was to generate low-energy states for the three different dopings x=0, 0.125, and 0.25 on a 2$\sqrt(2)$x2$\sqrt(2)$x1 unit cell and use them in our \textit{ad hoc} fitting procedure to generate an approximate effective model.
\\
\textit{This paragraph should talk about the procedural work of generating these low energy states and the mean field approximation for fitting. REMEMBER TO TALK ABOUT DROPPED STATES}

\paragraph{SCO SCF: Results and discussion} We find that a four parameter model with NN hopping, NNN hopping, AF Heisenberg exchange, and a FM NN Kondo term using only Cu 3d$_{x^2-y^2}$ orbitals can accurately describe the DFT band structures and total energies of low-energy states at T=0 and x=0, 0.125 and 0.25.
\\
\textit{This paragraph should talk about the results of the work from above.}

\paragraph{SCO DMD: Basis elements} Our first step into DMD with FN-DMC was the generation of appropriate basis elements for our effective model Hamiltonian which we handled, for all three dopings, using intrinsic atomic orbitals.
\\
\textit{This paragraph should talk about the IAO analysis that we have.}

%\begin{figure}[H]
%\centering
%\includegraphics[width=0.6\textwidth]{Figures/R1-mean_field.pdf}
%\caption{\label{fig3} Schematic diagram showing the method for mean-%field approximation we used in our DFT model fitting calculations.}
%\end{figure}

%\begin{figure}[H]
%\centering
%\includegraphics[width=0.9\textwidth]{Figures/R2-pareto.pdf}
%\caption{\label{fig4} Results of the \textit{ad hoc} fitting procedure just using the DFT eigenvalues and total energies for three different dopings.}
%\end{figure}

%\begin{figure}[H]
%\centering
%\includegraphics[width=0.9\textwidth]{Figures/R3-iao_basis.pdf}
%\caption{\label{fig5} Relaxation of the IAOs for the spin majority and minority channel under doping for the eighth doped flip state. The majority spin IAO can be used as a basis for the fixed spin moments and the minority spin IAO for the itinerant electrons.}
%\end{figure}
\pagebreak

\section{Proposed work}
\paragraph{SCO DMD: Low energy states} Low energy states for the x=0 system can be constructed by using DFT states with different spin textures multiplied by a 3-body Jastrow factor as trial wave functions in FN-DMC.
\\
\textit{This paragraph should talk about low energy states for x=0 and the DMD procedure (no parameter derivatives).}

\paragraph{SCO DMD: Low energy states} For the doped systems, low energy states will be constructed using multi-determinant expansions composed of low energy DFT states and singly excited determinants multiplied by a 3-body Jastrow factor as trial wave functions in FN-DMC.
\\
\textit{This paragraph should talk about the linear combination states and parameter derivatives.}

\paragraph{SCO: Band structure} In addition to using them in the DMD procedure, we will use these low energy multi-determinant expansions to generate effective band structures for the doped states.
\\
\textit{This paragraph should talk about the independent project of using parameter derivatives for band structure on large systems.}

\paragraph{SCO DMD: Descriptors} After calculating the RDMs on our IAO basis using FN-DMC, the relevant descriptors in our model can be initially restricted through symmetry considerations.
\\
\textit{This paragraph should talk about symmetry restriction of descriptors.}

\paragraph{SCO DMD: Descriptors} Further restriction of relevant descriptors will be accomplished by using a greedy selection algorithm.
\\
\textit{Talk about the greedy algorithm.}

\paragraph{SCO DMD:  FSE and K-points} QMC calculations on solids are accompanied by finite size effects which we will deal with, if necessary, using either a supercell extrapolation, k-point (twist) averaging, or both.
\\
\textit{Talk about how we would be able to diagnose whether FSE is important, then talk about how each of these methods work.}

 \paragraph{SCO DMD: Reach goal} A longer term goal is to use the DMD method to generate a single low-energy model that can accurately describe the low-energy properties of both superconducting SCO and an isostructural non-super conducting nickelate.
 \\
 \textit{This should talk about the goal of showing that DMD is actually accurate in its predictions, and is an interesting scientific problem not just a calculation we did.}

%\begin{figure}[H]
%\centering
%\includegraphics[width=0.9\textwidth]{Figures/P1-fese_greed.pdf}
%\caption{\label{fig6} Results for greedy algorithm procedure for choosing %descriptors showing the decrease in $R^2$ as more descriptors are added for the FeSe molecule [ref].}
%\end{figure}

%\begin{figure}[H]
%\centering
%\includegraphics[width=0.9\textwidth]{Figures/P2-Ni_Cu.pdf}
%\caption{\label{fig7} The primitive unit cells of the isostructural materials LNCO [ref] and LSCO [ref], the prior which is not super conducting all the way to temperatures of 30mK [ref].}
%\end{figure}
\pagebreak

\section{Conclusion and Summary}
\paragraph{SCO DMD} We propose that a DMD method using \textit{ab initio} QMC calculations can generate an effective model which can describe the low-energy properties of the electron doped cuprate SCO with a level of accuracy previously unattainable.

\paragraph{SCO Reach goal} We hope to extend our model fitting calculation to gain insight into why materials like nickelates, which are isostructural to the cuprates, do not superconduct.

\end{document}

### 2019-05-29 outline.tex
finished about 2/3 of the outline for the prelim;

\documentclass{article}
\title{Prelim Outline}
\author{Shivesh Pathak}
\usepackage[margin=1.0in]{geometry}
\usepackage{graphicx}
\usepackage{float}
\begin{document}

\section{Introduction and methods}
\textit{The goal of the introduction is to show the reader that the project is valuable. Should be x pages, other 5-x page will be methods.}

\paragraph{Overall goal} We want to be able to systematically develop low-energy effective models for strongly correlated electron systems.

\paragraph{Barrier to achieving goal} Non-perturbative interactions between various degrees of freedom - for example, lattice, spin and charge - makes writing down low-energy effective theories for strongly correlated materials very challenging.

\paragraph{Current state of the art} ??

\paragraph{Advancement to the state of the art} We propose using a density matrix downfolding (DMD) procedure to develop low-energy effective theories of the 1-band nearest-neighbor Hubbard model on a square lattice as a stepping stone towards usage on \textit{ab-initio} Hamiltonians of strongly correlated materials.

\paragraph{Density matrix down folding} 
The DMD procedure allows us to develop low-energy effective theories for physical systems in a systematic manner beginning with a higher energy Hamiltonian, like the \textit{ab-initio} Hamiltonian $H_{ab}$.

\paragraph{Fixed node diffusion Monte Carlo}
We will use fixed node diffusion Monte Carlo (FN-DMC) to generate wave functions in our chosen low-energy subspace and to accurately calculate \textit{ab-initio} energies and reduced density matrices.

\pagebreak

\section{Preliminary results and discussion}
\textit{The goal of the preliminary results is to indicate that I have the tools to actually complete the proposed research.}

\paragraph{CuO: Introduction} As a first step towards developing accurate models for extended systems like solids using DMD, we constructed a many-body effective model for the CuO molecule which accurately describes the eigenstates and spectra seen in experiment up to 2eV above the ground state.

\paragraph{CuO: Methods} In order to fit a model using DMD we had to construct a low energy space (LES), choose a sampling scheme for states within the LES resulting in our sample space (SS), and develop a basis which can accurately describe variations within the LES.

\paragraph{CuO: Results and Discussion} The resultant model from our analysis accurately describes the various eigenstates as measured in experiment up to 2eV, and is quantitatively and qualitatively more accurate than competing methods like density functional theory.

\paragraph{Hubbard: Preliminary results} Our preliminary calculations on the 2x2 1-band Hubbard model at half-filling indicate the emergence of a gapped low-energy subspace for U/t $\ge$ 10 described well by a nearest neighbor Heisenberg model, and a set of models with 0 $<$ U/t $\leq$ 4 which are poorly described by either a non-interacting fermion or Heisenberg model.

\section{Proposed work}
\paragraph{Hubbard stuff?}

\section{Conclusion and Summary}

\end{document}

### 2019-05-31 outline.tex
really rough outline, send to lucas for comments

@@
 \usepackage{float}
+\usepackage{xcolor}
 \begin{document}
@@
 \section{Introduction and methods}
-\textit{The goal of the introduction is to show the reader that the project is valuable. Should be x pages, other 5-x page will be methods.}
-
 \paragraph{Overall goal} We want to be able to systematically develop low-energy effective models for strongly correlated electron systems.
@@
 
-\paragraph{Current state of the art} ??
+\paragraph{Current state of the art} \textcolor{red}{Talk about methods people use for model fitting now. Will fill out as I do literature review.}
 
-\paragraph{Advancement to the state of the art} We propose using a density matrix downfolding (DMD) procedure to develop low-energy effective theories of the 1-band nearest-neighbor Hubbard model on a square lattice as a stepping stone towards usage on \textit{ab-initio} Hamiltonians of strongly correlated materials.
+\paragraph{Advancement to the state of the art} We propose using a density matrix downfolding (DMD) procedure to develop low-energy effective theories of the 1-band nearest-neighbor Hubbard model on a square lattice as a stepping stone towards effective theories starting from \textit{ab-initio} Hamiltonians of strongly correlated materials.
 
@@
 
-\pagebreak
-
 \section{Preliminary results and discussion}
@@
 
-\paragraph{CuO: Methods} In order to fit a model using DMD we had to construct a low energy space (LES), choose a sampling scheme for states within the LES resulting in our sample space (SS), and develop a basis which can accurately describe variations within the LES.
+\paragraph{CuO: Methods} In order to fit a model using DMD we had to construct a low energy space (LES), choose a sampling scheme for states within the LES resulting in our sample space (SS), develop a basis which can accurately describe variations within the LES, identify and select descriptors built out of these basis elements, and appropriately fit the data from our SS onto our selected descriptors.
 
@@
 
-\paragraph{Hubbard: Preliminary results} Our preliminary calculations on the 2x2 1-band Hubbard model at half-filling indicate the emergence of a gapped low-energy subspace for U/t $\ge$ 10 described well by a nearest neighbor Heisenberg model, and a set of models with 0 $<$ U/t $\leq$ 4 which are poorly described by either a non-interacting fermion or Heisenberg model.
+\paragraph{Hubbard: Preliminary results} Our preliminary calculations on the 2x2 half-filled 1-band Hubbard model map out the limiting behavior in the U-T model space by investigating the effectiveness of the single descriptor models.
 
 \section{Proposed work}
-\paragraph{Hubbard stuff?}
+\paragraph{Hubbard: Effective models at half-filling} We want to develop low-energy effective theories for the half-filled Hubbard model over the space U: 0 - 12, 1/$k_B$T: 0 - 3, particularly on the medium coupling region of U: 2 - 6 and 1/$k_B$T $\sim$ 1 where perturbative effective theories seem to break down.
 
-\section{Conclusion and Summary}
+\paragraph{Hubbard: Effective models away from half-filling} A farther reaching goal would be to develop effective theories in the U-T-$\delta$ space, where more complex behavior similar to the cuprates have been found.
+
+\paragraph{Hubbard: Finite size effects} Our initial calculations can be done using exact diagonalization, but memory limitations will require us to use approximate quantum Monte Carlo methods to calculate energies and density matrix elements for states in our LES.
+
+\section{Proposed timeline}
+\textit{Do this after proposed work nailed down}.
 

### 2019-05-31 outline.tex
.

@@
 
-\paragraph{Advancement to the state of the art} We propose using a density matrix downfolding (DMD) procedure to develop low-energy effective theories of the 1-band nearest-neighbor Hubbard model on a square lattice as a stepping stone towards effective theories starting from \textit{ab-initio} Hamiltonians of strongly correlated materials.
+\paragraph{Advancement to the state of the art} We propose using a density matrix downfolding (DMD) procedure to develop low-energy effective theories of the 1-band nearest-neighbor Hubbard model on a square lattice as a stepping stone towards building effective theories for strongly correlated materials.
 
@@
 \section{Preliminary results and discussion}
-\textit{The goal of the preliminary results is to indicate that I have the tools to actually complete the proposed research.}
+\paragraph{C2 Non-orthogonal: Brief summary} In order to become acquainted with our code and QMC algorithms in general I worked on implementing non-orthogonal orbital optimization for QMC trial wave functions and testing the effectiveness of these trial functions in calculating the total ground state energy of a C$_2$ molecule.
 
@@
 \paragraph{Hubbard: Preliminary results} Our preliminary calculations on the 2x2 half-filled 1-band Hubbard model map out the limiting behavior in the U-T model space by investigating the effectiveness of the single descriptor models.
+\textcolor{red}{Will be changed based on how far along we get with preliminary results.}
 
 \section{Proposed work}
-\paragraph{Hubbard: Effective models at half-filling} We want to develop low-energy effective theories for the half-filled Hubbard model over the space U: 0 - 12, 1/$k_B$T: 0 - 3, particularly on the medium coupling region of U: 2 - 6 and 1/$k_B$T $\sim$ 1 where perturbative effective theories seem to break down.
+\paragraph{Hubbard: Effective models at half-filling} We want to develop low-energy effective theories for the half-filled Hubbard model over the space U: 0 - 12, 1/$k_B$T: 0 - 3 using the DMD method, particularly on the medium coupling region of U: 2 - 6 and 1/$k_B$T $\sim$ 1 where perturbative effective theories break down and a metal-insulator transition exists.
 
-\paragraph{Hubbard: Effective models away from half-filling} A farther reaching goal would be to develop effective theories in the U-T-$\delta$ space, where more complex behavior similar to the cuprates have been found.
+\paragraph{Hubbard: Effective models away from half-filling} A farther reaching goal would be to develop effective theories in the U-T-$\delta$ space using DMD, where more complex behavior similar to the cuprates have been found.
 
-\paragraph{Hubbard: Finite size effects} Our initial calculations can be done using exact diagonalization, but memory limitations will require us to use approximate quantum Monte Carlo methods to calculate energies and density matrix elements for states in our LES.
+\paragraph{Hubbard: Finite size effects} Our initial calculations necessary for DMD can be done using exact diagonalization, but memory limitations will require us to use approximate methods like lattice quantum Monte Carlo to calculate energies and density matrix elements for states in our LES.
 

### 2019-06-05 draft.tex
added introduction for draft!

\documentclass{article}
\usepackage{graphicx}
\usepackage{xcolor}
\usepackage{subcaption}
\usepackage[margin=1.0in]{geometry}
\usepackage{float}
\usepackage{ulem}
\usepackage{amsmath}
\usepackage{mathtools}

\begin{document}
\section{Introduction}
A core problem of condensed matter physics is understanding how to systematically develop low-energy effective models for strongly correlated electron systems. In the forefront of modern research are high temperature superconductors, heavy fermion materials, materials with strong spin-orbit coupling effects, and more for which simple effective theories have not yet been found. For example, cuprate family of high-Tc superconductors have an intricate phase diagram in the doping-temperature space \textbf{ref} with a robust anti-ferromagnetic (AFM) charge-transfer insulator state near zero doping, a prominent superconducting dome in the 0.1 - 0.2 hole doped region at low temperatures with psuedo-gap and non-Fermi liquid metallic phases at higher temperatures and Fermi liquid like behavior in the overdoped area, of which only the AFM insulating and Fermi liquid phases have well understood mechanism and models \textbf{ref}. The iron-pnictides also have superconducting domes, non-Fermi liquid like metallic phases, a spin-density-wave antiferommagnetic phase, all under various dopings and temperatures \textbf{ref}. As in the cuprate case, models for these phases have not been fully fleshed out \textbf{ref}. Heavy-fermion materials have a purported quantum critical point between an AFM and Fermi-liquid phase which houses a superconducting dome, leading to another phase diagram for which building model Hamiltonians is very challenging \textbf{ref}. Materials with strong spin-orbit coupling effects exhibit a host of properties like \textbf{LIST THESE HERE!}.

The fundamental barrier in writing down low-energy theories for strong correlated materials is the existence of non-perturbative interactions between various degrees of freedom - for example, lattice, spin and charge. Consider just the cuprate case for now, the two phases which are well understood contain a singular important degree of freedom in the low-energy space. Near zero doping the cuprates are charge-transfer insulators with an optical gap around 2 eV \textbf{ref}, indicating that the lowest energy excitations are primarily spin-like, allowing for a simple nearest-neighbor spin-only model like an AFM Heisenberg model to describe the low-energy physics quite accurately. In the heavily overdoped side, the lowest energy excitations look like well defined Fermi-liquid quasiparticles \textbf{ref}, indicating our lowest energy excitations are electron-like, and a simple non-interacting electron model (with renormalized masses, etc) will suffice. In the intermediate doping region where the low-energy excitations cannot be easily sectioned into spin-like or electron-like only, we see the emergence of exotic phases and where the most difficulties arise in writing down low-energy effective models. On top of that, this entire discussion did not include any lattice effects like structural phase transitions which many strongly-correlated materials have as well \textbf{ref}. This discussion can be extended to the many other strongly correlated electron materials presented above as well.

\textbf{Current state of the art.}

We propose using a density matrix downfolding (DMD) procedure\textbf{ref} to develop low-energy effective theories of the 1-band nearest-neighbor Hubbard model on a square lattice as a stepping stone towards building effective theories for strongly correlated materials. The Hubbard model is a simple model with nearest-neighbor hopping and a local on-site Coulomb repulsion, parameterized by a single ratio U/t: 
\begin{equation}
H_\text{hub} = -t \sum_{\langle i,j \rangle,\sigma}( c_{i^\dagger,\sigma} c_{j,\sigma} + h.c.) + U \sum_i n_{i\uparrow} n_{i\downarrow}
\label{hubbard}.
\end{equation}
The solutions in 1-d are well understood via the Bethe ansatz \textbf{ref}, however the 2-d square lattice Hubbard model has no known exact solution.  In the very strong coupling limit $U/t \rightarrow \infty$ an emergent low-energy spin-space exists with an effective nearest-neighbor AFM Heisenberg Hamiltonian \textbf{ref}. In the region $U = 0$ a trivial non-interacting theory holds. Between these two limits, in particular the intermediate coupling region $1 < U < 7$, no well established effective theory exists. In this region there is a metal-insulator transition at $U_c/t \sim 4$ \textbf{ref}, and the pseudogap opens up at finite temperature around $U/t \sim 7$, leading to the development of a Mott gap for large $U/t > 11$ \textbf{ref}. In addition to the half-filling phase diagram, this model exhibits a plethora of interesting ground state properties when doped including stripe phases and d-wave pairing \textbf{ref}. The relative simplicity of the Hubbard model and the lack of a well-established low-energy theory in a broad intermediate coupling region lends the model as a useful playground for our new downfolding procedure in preparation for tackling realistic strongly correlated electron systems.

The DMD procedure allows us to develop low-energy effective theories
for physical systems in a systematic manner beginning with a higher energy Hamiltonian H on a Hilbert space $\mathcal{H}$. When trying to develop effective theories for realistic systems H is usually the \textit{ab-initio} Hamiltonian
\begin{equation}
H_\text{ab} = -\frac{1}{2} \sum_{i} \nabla_i^2 - \sum_{i,I}\frac{Z}{|r_i - R_I|} + \sum_{i<j}\frac{1}{|r_i - r_j|} 
\label{ab}
\end{equation} 
where the sum $i$ is over electrons and $I$ over nuclei, and the unit for energy is Hartree (Ha). Typically we work under the Born-Oppenheimer approximation and use pseudo-potentials as an efficient replacement for core electrons. This is not always the case, for example when downfolding a second quantized model like shown in equation \ref{hubbard}. The core idea behind the DMD procedure is functional fitting, in particular approximating the total energy functional on a particular subspace of the full Hilbert space. Mathematically:

\begin{equation}
(E[\Psi] \equiv \langle \Psi | H |\Psi \rangle, \mathcal{H}) \rightarrow (E_{eff}[\Psi], \mathcal{LE})\text{ s.t. } E_{eff}[\Psi] = E[\Psi] + \epsilon[\Psi] \ \forall \ |\Psi\rangle \in \mathcal{LE}.
\label{DMD}
\end{equation}
where $\epsilon$ represents an error within our effective theory. The fitting procedure is conducted similarly to fitting any function: 1) Choose a $\mathcal{LE}$ which is spanned by the N lowest energy eigenvectors within $\mathcal{H}$, 2) sample the space to get a sample set $\mathcal{SS} \in \mathcal{LE}$, 3) construct a set of potential models with some variable parameters $\mathcal{MS}$, and 4) regress these models using your sampled data to fit the variable parameters. The density-matrix part of DMD comes from the particular choice of functional parameterization: 
\begin{equation}
E_{eff}[\Psi] = \sum_i c_i d_i[\Psi],\ d_i[\Psi] = \langle \Psi |\hat{d}_i|\Psi \rangle
\end{equation}
where $\hat{d}_i$ is a Hermitian operator in our Hilbert space and many times chosen to be a (sum of) 1- or 2-body density matrices. Common choices of $\hat{d}_i$ are $n_k,\ c_j^\dagger c_k + h.c.,\ n_{k\uparrow}n_{k\downarrow},\ \vec{S}_k \cdot \vec{S}_l,\ $ etc. With this parameterization one can even extract an effective Hamiltonian from the effective energy functional: $H_{eff} = \sum_i c_i \hat{d}_i$ which can be solved and studied. Mapping a downfolding problem onto a regression problem allows us to develop methodologies and systems for building effective theories. In addition we can quantiatively validate our proposed effective theories using tools from data science and statistics. Not only is this method pleasing to statisticians, it is in principle exact. Suppose we want to build a model on $\mathcal{LE}$ composed of the N lowest-energy eigenstates. Let us consider a potential model $H_{eff} = \sum_{i=1}^{ N} c_i |\Psi_i\rangle \langle \Psi_i|$ where the sum is over the N lowest eigenstates with variable coefficients $\{c\}$. If we select \textit{any} linearly independent set of N samples from $\mathcal{LE}$, eigenstates are not required, an ordinary linear regression will lead to the optimal coefficients $c_i = E_i$, the eigenvalues of the eigenvectors $|\Psi_i\rangle$, and our effective model will be exact. In practice we do not know all the eigenstates of H so identifying and sampling $\mathcal{LE}$ is of primary concern, especially for strongly correlated electron systems where the low-energy degrees of freedom may be very complicated collective excitations. I will discuss one approach we had to sampling the low-energy space when talking about our model for the CuO molecule later on. If we do not know the eigenstates we also need to choose descriptors $d_i$ which can compactly represent the variation in properties within $\mathcal{LE}$. Typically one will collect all reasonable $d_i$ like occupation, hopping, exchange coupling, on-site Coulomb repulsion, Hund's coupling, etc ad consider only those $d_i$ which vary within $\mathcal{SS}$. One can then use tools from data science like principal component analysis, step-wise AIC/BIC, regularized regression (e.g. LASSO) for model selection and fitting.

When downfolding \textit{ab-initio} Hamiltonians we will use fixed node diffusion Monte Carlo (FN-DMC) to generate wave functions in our chosen low-energy subspace and to accurately calculate energies and reduced density matrices. Diffusion Monte Carlo (DMC) is a quantum Monte Carlo method which projects out the ground state of a Hamiltonian given some initial trial wave function. Consider a trial wave function $|\psi\rangle$ and a Hamiltonian $H$ with ground state $|\phi_0\rangle$. We apply the projector $e^{-\tau (H-E_0)}$ as $\tau \rightarrow \infty$ to $|\psi \rangle$
\begin{equation}
\lim_{\tau \rightarrow \infty} e^{-\tau (H-E_0)} |\psi\rangle \propto \langle \phi_0|\psi\rangle |\phi_0\rangle,
\end{equation}
projecting out the ground state $|\phi_0\rangle$ of $H$ as long as the trial wave function we choose is not orthogonal to the ground state. The stochastic implementation involves moving samples from the trial function $\Psi_T(R)$ using the Green function $G(R, R^\prime, \tau) = \langle R | e^{-\tau(H - E_T)} | R^\prime \rangle$. Since the Hamiltonian has two parts, $H = T + V$ a Trotter expansion is used to approximate the Green function $G(R, R^\prime, \tau) = \langle R | e^{-\tau(H - E_T)} | R^\prime \rangle \sim e^{-\tau(V(R) + V(R^\prime))/2} \langle R| e^{-\tau(T - E_T)}|R^\prime \rangle + O(\tau^2) $ which leads to a time step error in the calculation if the time step between moves $\tau$ is too large. This implementation when used on a Hamiltonian with fermions also leads to a fermion sign problem. We deal with this sign problem through a fixed-node approximation, where the nodal surface of the projected wave function is forced to match that of the initial trial wave function. This approximation makes FN-DMC variational, and will only return the exact ground state of $H$ if the nodal surfaces of $|\psi\rangle$ and $|\phi_0\rangle$ are identical. Therefore the FN-DMC calculation has two major systematic errors that have to be dealt with, the fixed-node error and the time-step error. The time-step error is typically handled through a time-step extrapolation, where multiple FN-DMC calculations are conducted at sequentially smaller $\tau$ and the relevant expectation values are extrapolated to $\tau \rightarrow 0$ \cite{Needs2010}. Fixed-node error can be reduced by using optimized many-body trial wave functions of the (multi-)Slater-Jastrow form seen in \cite{PhysRevLett.98.110201}. We can use the final FN-DMC state to calculate the expectation values of operators like the Hamiltonian or reduced density matrices. For an operator $\hat{O}$ we will use a mixed estimator to calculate the expectation value $\langle \Psi_{DMC} |\hat{O} | \Psi_T \rangle/\langle \Psi_{DMC} | \Psi_T \rangle$ where $\Psi_{DMC}$ is the projected FN-DMC state starting with the trial wave function $\Psi_T$, as this estimator allows us to use the importance sampled distribution $f = \Psi_{DMC}\Psi_T$ for our Monte Carlo estimates. For operators which do not commute with the Hamiltonian H the mixed estimator has an error which can be corrected to first order by subtracting twice the VMC estimate of the expectation value \cite{ceperley_kalos_1979}. Wagner provides the details for calculating reduced density matrix elements in QMC \cite{doi:10.1063/1.4793531}.
\end{document}

### 2019-06-12 draft.tex
finished preliminary work, 5.5 pages

@@
 \usepackage{mathtools}
+\usepackage{wrapfig}
 
@@
 
-The DMD procedure allows us to develop low-energy effective theories
-for physical systems in a systematic manner beginning with a higher energy Hamiltonian H on a Hilbert space $\mathcal{H}$.
-When trying to develop effective theories for realistic systems H is usually the \textit{ab-initio} Hamiltonian
+The DMD procedure allows us to develop low-energy effective theories for physical systems in a systematic manner beginning with the \textit{ab-initio} Hamiltonian:
 \begin{equation}
-H_\text{ab} = -\frac{1}{2} \sum_{i} \nabla_i^2 - \sum_{i,I}\frac{Z}{|r_i - R_I|} + \sum_{i<j}\frac{1}{|r_i - r_j|} 
-\label{ab}
+\hat{H}_\text{ab} = -\frac{1}{2} \sum_{i} \nabla_i^2 - \sum_{i,I}\frac{Z}{|r_i - R_I|} + \sum_{i<j}\frac{1}{|r_i - r_j|}
+\label{eq:Hab}
 \end{equation}
-where the sum $i$ is over electrons and $I$ over nuclei, and the unit for energy is Hartree (Ha).
-Typically we work under the Born-Oppenheimer approximation and use pseudo-potentials as an efficient replacement for core electrons.
-This is not always the case, for example when downfolding a second quantized model like shown in equation \ref{hubbard}.
-The core idea behind the DMD procedure is functional fitting, in particular approximating the total energy functional on a particular subspace of the full Hilbert space.
-Mathematically:
+where $i$ indicates electrons and $I$ nuclei, and the unit of energy is Hartree (Ha).
+In this paper we will be working under the Born-Oppenheimer approximation and core electrons will be treated using pseudo-potentials.
+The procedure defines a method for downfolding the \textit{ab-initio} energy functional and Hilbert space onto an effective energy functional defined on a low-energy subspace of the Hilbert space 
+\begin{equation}
+(E_\text{ab}, \mathcal{H}) \xrightarrow{\text{DMD}} (E_\text{eff}, \mathcal{LE} \subset \mathcal{H}), \text{ s.t. }
+\forall |\Psi\rangle \in \mathcal{LE}, \ E_\text{eff}[\Psi] - E_\text{ab}[\Psi] = \epsilon[\Psi] + E_0.
+\label{eq:DMD}
+\end{equation} 
+The energy functionals can be written as expectations over Hamiltonian operators $E_\text{ab} = \langle \Psi | \hat{H}_\text{ab} |\Psi \rangle$, $E_\text{eff} = \langle \Psi | \hat{H}_\text{eff} |\Psi \rangle$, $\epsilon$ represents the error in our effective theory, and $E_0$ is a constant energy shift.
+Broadly the DMD method consists of four steps which will be discussed in detail later: 
+A) Defining the space of low-energy excitations which $E_\text{eff}$ should be able to describe accurately.
+B) Sampling, at minimum, a set of states in $\mathcal{H}$ which probe the chosen low-energy excitations in a statistically independent manner. The selected states need not be eigenstates. 
+C) Constructing a set of candidate models which are linear in reduced density matrix (RDM) descriptors with variable coefficients $\{c\}$
+\begin{equation}
+E_\text{eff} = \sum_k c_k \langle \Psi | \hat{d}_k |\Psi \rangle + E_0.
+\label{eq:Eeff}
+\end{equation}
+Here $\hat{d}_k$ are Hermitian operators typically composed of 1- or 2-RDM elements. 
+Examples of possible $\hat{d}_k$ are number operators $\hat{n}_i$ and exchange operators $\vec{S}_i \cdot \vec{S}_j$.
+Additionally, a single particle basis must be built to express RDM elements on and should be chosen such that variations among states in $\mathcal{LE}$ are mostly described by variations in the projection of these states onto the basis.
+D) Using the samples to fit the coefficients $\{c\}$ by linear regression such that $E_\text{eff}$ satisfies the conditions in \eqref{eq:DMD} with the smallest error $\epsilon$ possible. 
+The fitting procedure is typically an ordinary linear regression, however, one may use a custom cost function in the regression as well. 
+In order to conduct the linear regression for a functional like \eqref{eq:Eeff}, one needs to calculate for each state sampled from $\mathcal{LE}$ the expectation values of $\hat{H}_\text{ab}$ and every descriptor $\{\hat{d}_k\}$ used in the regression.
+In principle, if conducted with no approximations, the DMD method is rigorously guaranteed to obtain exact results, namely $\epsilon[\Psi] = 0$.
 
+When downfolding $\hat{H}_\text{ab}$ we will use FN-DMC to access states in $\mathcal{LE}$ and accurately calculate their \textit{ab-initio} energies and RDM elements.
+Diffusion Monte Carlo (DMC) is a quantum Monte Carlo method which projects out the ground state of a Hamiltonian given some initial trial wave function.
+Consider a trial wave function $|\Psi_T\rangle$ and the Hamiltonian $\hat{H}_\text{ab}$ with ground state $|\Phi_0\rangle$.
+We apply the projector $e^{-\tau \hat{H}_\text{ab}}$ as $\tau \rightarrow \infty$ to $|\Psi_T \rangle$
 \begin{equation}
-(E[\Psi] \equiv \langle \Psi | H |\Psi \rangle, \mathcal{H}) \rightarrow (E_{eff}[\Psi], \mathcal{LE})\text{ s.t. } E_{eff}[\Psi] = E[\Psi] + \epsilon[\Psi] \ \forall \ |\Psi\rangle \in \mathcal{LE}.
-\label{DMD}
+\lim_{\tau \rightarrow \infty} e^{-\tau \hat{H}_\text{ab}} |\Psi_T\rangle 
+\equiv \lim_{\tau \rightarrow \infty} |\Psi_\text{DMC}(\tau)\rangle \propto \langle \Phi_0|\Psi_T\rangle |\Phi_0\rangle,
 \end{equation}
-where $\epsilon$ represents an error within our effective theory.
-The fitting procedure is conducted similarly to fitting any function: 1) Choose a $\mathcal{LE}$ which is spanned by the N lowest energy eigenvectors within $\mathcal{H}$, 2) sample the space to get a sample set $\mathcal{SS} \in \mathcal{LE}$, 3) construct a set of potential models with some variable parameters $\mathcal{MS}$, and 4) regress these models using your sampled data to fit the variable parameters.
-The density-matrix part of DMD comes from the particular choice of functional parameterization:
+projecting out the ground state as long as the trial wave function we choose is not orthogonal to the ground state. 
+The stochastic implementation involves moving samples from the trial function $\Psi_T(R)$ using the Green function $G(R, R^\prime, \tau) = \langle R | e^{-\tau(\hat{H}_\text{ab} - E_T)} | R^\prime \rangle$.
+Since $H_\text{ab} = T + V$, kinetic and potential energy terms, the Green function is approximated by a Trotter expansion $G(R, R^\prime, \tau) = \langle R | e^{-\tau(\hat{H} - E_T)} | R^\prime \rangle \sim \Big[e^{-d\tau(V(R) + V(R^\prime))/2} \langle R| e^{-d\tau(\hat{T} - E_T)}|R^\prime \rangle + O(d\tau^2) \Big]^N $ where $d\tau = \tau/N$.
+This expansion can be interpreted as an interative procedure where samples are moved N times with a small timestep $d\tau$ until convergence.
+The constant $E_T$ is a trial energy used to control the normalization of $\Psi_\text{DMC}(\tau, R)$ and is updated at each move.
+This implementation, however, suffers from a fermion sign problem. 
+We deal with this sign problem through a fixed-node approximation, where the nodal surface of $\Psi_\text{DMC}(\tau, R)$ is forced to match that of the initial trial wave function for all $\tau$.
+This approximation makes FN-DMC variational, and will only return the exact ground state of $\hat{H}_\text{ab}$ if the nodal surfaces of $|\Psi_T\rangle$ and $|\Phi_0\rangle$ are identical.
+We make use of this variational principle to access low-energy states within $\mathcal{H}$ by choosing trial wave functions with nodes that differ from the nodes of $|\Phi_0 \rangle$.
+To calculate the expectation value of an operator $\hat{A}$ in FN-DMC we will use a mixed estimator $\langle \Psi_{DMC}(\tau) |\hat{A} | \Psi_T \rangle/\langle \Psi_\text{DMC}(\tau) | \Psi_T \rangle$ which allows us to use the importance sampled distribution $f = \Psi_\text{DMC}(\tau)\Psi_T$ for our Monte Carlo estimates.
+Details for calculating reduced density matrix elements in FN-DMC can be found in Wagner.
+
+\pagebreak
+\section{Preliminary Work}
+\subsection{Non-orthogonal determinants in FN-DMC trial wave functions}
+In order to become acquainted with our code QWalk \cite{WAGNER20093390} and QMC algorithms in general I worked on implementing and testing multi-Slater-Jastrow trial functions with optimized non-orthogonal determinants (MSJ+NO) in FN-DMC \cite{Pathak2018}.
+We assessed the efficiency and compactness of this new trial function by calculating the ground state energy and single particle densities of a C$_2$ molecule using FN-DMC and comparing to the results when using multi-Slater-Jastrow trial functions with optimized orthogonal determinant trial functions (MSJ+O).
+The workflow involved constructing the un-optimized trial wave functions, optimizing the parameters using an energy optimization method \cite{Toulouse2007}, and finally using the optimized trial functions in an FN-DMC calculation.
+We found that the FN-DMC energy calculated using an MSJ+NO trial function with only 24 determinants was lower than the FN-DMC energy using an MSJ+O trial function with 55 determinants. 
+Further, the FN-DMC charge density calculated using MSJ+NO trial functions had stronger bonding character than when using MSJ+O trial functions, a reasonable result as introducing correlations into trial functions allows for electrons to avoid each other while still occupying the same bonding region.
+Our results indicated that using non-orthogonal determinants may lead to more compact multi-Slater-Jastrow trial wave functions for small molecules.
+
+\subsection{Effective theory for CuO molecule using DMD}
+As a first step towards developing accurate models for extended systems like solids using DMD, we constructed a many-body effective model for the CuO molecule which accurately describes the eigenstates and spectra seen in experiment up to 2eV above the ground state.
+The neutral CuO molecule has been a subject of intense theoretical and experimental study due to the complex structure of its low-energy space.
+While the low-energy excitations in the doublet sector of the CuO molecule are well understood, singlet selection rules for spectroscopic measurements and the doublet ground state leave the quartet sector of excitations unexplored.
+We construct our $\mathcal{LE}$ based on recent anion photoelectron spectroscopy (APES) measurements of the low-lying excited states of the nuetral CuO molecule. 
+The lowest energy excitations measured, which lie below 2eV, involve primarily the O 2p$_z$, 2p$_\pi$ and Cu 4s orbitals, which are fully-filled, half-filled and empty in the ground state respectively.
+Beginning at 2.2 eV are excitations which remove a single electron from the fully filled Cu 3d shell and add them to either the O 2p$_\pi$ or Cu 4s orbitals.
+The neutral copper atom also carries a Hund's coupling between the d and s orbitals of about $\sim$ 0.5 eV, the effects of which would be measurable among the states in $\mathcal{LE}$ as long as states of both doublet and quartet states are included.
+In order to probe all of these low-energy excitations we select a low-energy space
 \begin{equation}
-E_{eff}[\Psi] = \sum_i c_i d_i[\Psi],\ d_i[\Psi] = \langle \Psi |\hat{d}_i|\Psi \rangle
+\mathcal{LE} = \text{Span(}\{ |\Psi \rangle | \langle \Psi | \hat{n}_{3d} | \Psi \rangle \ge 9,\ \hat{H}_\text{ab}|\Psi\rangle = E |\Psi\rangle \}\text{)}.
+\label{eq:LE}
 \end{equation}
-where $\hat{d}_i$ is a Hermitian operator in our Hilbert space and many times chosen to be a (sum of) 1- or 2-body density matrices.
-Common choices of $\hat{d}_i$ are $n_k,\ c_j^\dagger c_k + h.c.,\ n_{k\uparrow}n_{k\downarrow},\ \vec{S}_k \cdot \vec{S}_l,\ $ etc.
-With this parameterization one can even extract an effective Hamiltonian from the effective energy functional: $H_{eff} = \sum_i c_i \hat{d}_i$ which can be solved and studied.
-Mapping a downfolding problem onto a regression problem allows us to develop methodologies and systems for building effective theories.
-In addition we can quantiatively validate our proposed effective theories using tools from data science and statistics.
-Not only is this method pleasing to statisticians, it is in principle exact.
-Suppose we want to build a model on $\mathcal{LE}$ composed of the N lowest-energy eigenstates.
-Let us consider a potential model $H_{eff} = \sum_{i=1}^{ N} c_i |\Psi_i\rangle \langle \Psi_i|$ where the sum is over the N lowest eigenstates with variable coefficients $\{c\}$.
-If we select \textit{any} linearly independent set of N samples from $\mathcal{LE}$, eigenstates are not required, an ordinary linear regression will lead to the optimal coefficients $c_i = E_i$, the eigenvalues of the eigenvectors $|\Psi_i\rangle$, and our effective model will be exact.
-In practice we do not know all the eigenstates of H so identifying and sampling $\mathcal{LE}$ is of primary concern, especially for strongly correlated electron systems where the low-energy degrees of freedom may be very complicated collective excitations.
-I will discuss one approach we had to sampling the low-energy space when talking about our model for the CuO molecule later on.
-If we do not know the eigenstates we also need to choose descriptors $d_i$ which can compactly represent the variation in properties within $\mathcal{LE}$.
-Typically one will collect all reasonable $d_i$ like occupation, hopping, exchange coupling, on-site Coulomb repulsion, Hund's coupling, etc ad consider only those $d_i$ which vary within $\mathcal{SS}$.
-One can then use tools from data science like principal component analysis, step-wise AIC/BIC, regularized regression (e.g.
-LASSO) for model selection and fitting.
 
-When downfolding \textit{ab-initio} Hamiltonians we will use fixed node diffusion Monte Carlo (FN-DMC) to generate wave functions in our chosen low-energy subspace and to accurately calculate energies and reduced density matrices.
-Diffusion Monte Carlo (DMC) is a quantum Monte Carlo method which projects out the ground state of a Hamiltonian given some initial trial wave function.
-Consider a trial wave function $|\psi\rangle$ and a Hamiltonian $H$ with ground state $|\phi_0\rangle$.
-We apply the projector $e^{-\tau (H-E_0)}$ as $\tau \rightarrow \infty$ to $|\psi \rangle$
+Our sample space was generated by using multi-Slater-Jastrow trial wavefunctions in FN-DMC. 
+In principle we could probe every low-energy excitation in $\mathcal{LE}$ in a statistically independent manner by just sampling the eigenstates within $\mathcal{LE}$ and including various non-eigenstates within $\mathcal{LE}$ as well to ensure robust statistics.
+In practice we do not have access to these eigenstates and approximate methods are required to sample the low-energy space. 
+Consider a trial wavefunction of the form $e^{J_i} |\text{D}_i\rangle$, where $e^{J_i}$ is a three-body Jastrow factor and $|\text{D}_i\rangle$ is a determinant of some single particle orbitals.
+If the nodes of $|\text{D}_i\rangle$ are similar to the nodes of an eigenstate $|\Phi_i\rangle \in \mathcal{LE}$, then the properties and energy calculated from FN-DMC using this trial wavefunction will approximate those of the eigenstate, with the degree of approximation depending on the quality of the trial function nodal structure.
+If we were able to find for each $|\Phi_i \rangle \in \mathcal{LE}$ such a $|\text{D}_i\rangle$ then we can approximate any state in $\mathcal{LE}$ as
 \begin{equation}
-\lim_{\tau \rightarrow \infty} e^{-\tau (H-E_0)} |\psi\rangle \propto \langle \phi_0|\psi\rangle |\phi_0\rangle,
+|\Psi\rangle = \sum_i c_i |\Phi_i\rangle \sim \sum_i c_i \lim_{\tau \rightarrow \infty} e^{-\tau \hat{H}_\text{ab}} (e^{J_i} |D_i\rangle) \xrightarrow{J_i = J} \lim_{\tau \rightarrow \infty} e^{-\tau \hat{H}_\text{ab}} (e^{J}\sum_{i} c_i|\text{D}_i\rangle).
+\label{eq:sampling}
 \end{equation}
-projecting out the ground state $|\phi_0\rangle$ of $H$ as long as the trial wave function we choose is not orthogonal to the ground state.
-The stochastic implementation involves moving samples from the trial function $\Psi_T(R)$ using the Green function $G(R, R^\prime, \tau) = \langle R | e^{-\tau(H - E_T)} | R^\prime \rangle$.
-Since the Hamiltonian has two parts, $H = T + V$ a Trotter expansion is used to approximate the Green function $G(R, R^\prime, \tau) = \langle R | e^{-\tau(H - E_T)} | R^\prime \rangle \sim e^{-\tau(V(R) + V(R^\prime))/2} \langle R| e^{-\tau(T - E_T)}|R^\prime \rangle + O(\tau^2) $ which leads to a time step error in the calculation if the time step between moves $\tau$ is too large.
-This implementation when used on a Hamiltonian with fermions also leads to a fermion sign problem.
-We deal with this sign problem through a fixed-node approximation, where the nodal surface of the projected wave function is forced to match that of the initial trial wave function.
-This approximation makes FN-DMC variational, and will only return the exact ground state of $H$ if the nodal surfaces of $|\psi\rangle$ and $|\phi_0\rangle$ are identical.
-Therefore the FN-DMC calculation has two major systematic errors that have to be dealt with, the fixed-node error and the time-step error.
-The time-step error is typically handled through a time-step extrapolation, where multiple FN-DMC calculations are conducted at sequentially smaller $\tau$ and the relevant expectation values are extrapolated to $\tau \rightarrow 0$ \cite{Needs2010}.
-Fixed-node error can be reduced by using optimized many-body trial wave functions of the (multi-)Slater-Jastrow form seen in \cite{PhysRevLett.98.110201}.
-We can use the final FN-DMC state to calculate the expectation values of operators like the Hamiltonian or reduced density matrices.
-For an operator $\hat{O}$ we will use a mixed estimator to calculate the expectation value $\langle \Psi_{DMC} |\hat{O} | \Psi_T \rangle/\langle \Psi_{DMC} | \Psi_T \rangle$ where $\Psi_{DMC}$ is the projected FN-DMC state starting with the trial wave function $\Psi_T$, as this estimator allows us to use the importance sampled distribution $f = \Psi_{DMC}\Psi_T$ for our Monte Carlo estimates.
-For operators which do not commute with the Hamiltonian H the mixed estimator has an error which can be corrected to first order by subtracting twice the VMC estimate of the expectation value \cite{ceperley_kalos_1979}.
-Wagner provides the details for calculating reduced density matrix elements in QMC \cite{doi:10.1063/1.4793531}.
+The last equality is valid if $\forall i\  J_i = J$, allowing us to map $|\Psi \rangle \in \mathcal{LE}$ to the end result of an FN-DMC calculation using a multi-Slater-Jastrow trial wavefunction. 
+We use symmetry-targeted unrestricted Kohn-Sham (UKS) to generate our determinants $|\text{D}_i\rangle$ using a B3LYP functional, which has been shown to accurately reproduce ground state nodal properties of transition-metal oxide molecules, a Trail-Needs pseudopotential, and the VTZ Trail-Needs basis using the package PySCF.
+This method allows us to access almost every excitation in $\mathcal{LE}$, however, there are some cases of two or more low-energy states within $\mathcal{LE}$ which have identical S$_z$ and number of electrons per irrep.
+An example of this dilemma would be the ground state and the state with an excitation from Cu 3d$_{xz} \rightarrow $  O 2p$_x$.
+In these few cases the higher energy excited $|\text{D}_i\rangle$ is inaccessible and the corresponding $|\Phi_i\rangle$ is excluded from the sum in \eqref{eq:sampling}, restricting our sample space to a subspace of $\mathcal{LE}$.
+We will discuss the consequences of this restriction later on.
+A single three-body Jastrow factor $J$ was used for every $|\text{D}_i\rangle$ and was optimized on the lowest energy UKS state using a linear energy optimization method.
+The FN-DMC calculations were conducted with T-moves, to make the calculation variational while using pseudopotentials, with a timestep of $\tau = 0.01$.
+The coefficients $\{c_j\}$ for our sample states were chosen via a shell-sampling method. 
+We begin by fixing the coefficient $c_0 = \sqrt{w}$ where $w \in \{1.0, 0.8, ..., 0.2\}$. 
+For each choice of $w$ we sample n = 5 states by randomly selecting the unassigned coefficients from a uniform distribution such that $\sum_i c_i^2 = 1$. 
+This procedure generates shells of samples at decreasing distance from the eigenstate $|\Phi_0\rangle$. 
+We can then loop over all $i$ to generate shells near each eigenstate.
+
+Parameterizing our effective model requires constructing a basis on which to calculate our \textit{ab-initio} 1-/2-RDM elements and selecting a set of descriptors from which we would like to build our model.
+We constructed two such bases: a localized intrinsic atomic orbital (IAO) basis and a molecular orbital (MO) basis.
+The IAO basis was built on all Cu 3d, 4s and O 2p molecular orbitals from the $|\text{D}_i\rangle$ used to construct our sample states.
+The MO basis consists of the Cu 3d, 4s and O 2p molecular orbitals from the lowest energy restricted open-shell Kohn-Sham (ROKS) solution with $x \leftrightarrow y$ spatial symmetry, a O 2p$_z \rightarrow$ O 2p$_\pi$ excitation.
+As these two bases are used to describe the same space, they are related by an orthogonal transformation which will allow us to map between them if necessary.
+In order to describe the low-energy degrees of freedom in the CuO molecule our model will require occupation energies and hybridizations between the Cu 3d, 4s and O 2p atomic orbitals which are most compactly represented in an MO basis.
+The lack of doubly occupied Cu 4s states in the low-energy spectrum of CuO indicate the necessity of a Hubbard U term on the Cu 4s.
+This is corroborated by the fact that the FN-DMC energy of the lowest-lying sampled state with a doubly occupied Cu 4s orbital is 7 eV above the lowest energy sampled state.
+Further, we include a Hund's coupling between the Cu 4s and Cu 3d orbitals in our model based on the existence of a similar Hund's coupling 
+for the copper atom.
+These two interaction terms are best represented in the IAO basis as they are intuitively interactions between localized orbitals.
+In addition to these important parameters, we will consider models which contain any number of hybridizations between the MOs to account for orbital relaxation between excited states in the CuO molecule.
+In total our space of canditate models is then 
+\begin{equation}
+\begin{split}
+\{\bar{n}_{4s}, \bar{n}_{p_\pi}, \bar{n}_{p_z}, \hat{n}_{4s\uparrow} \hat{n}_{4s\downarrow},\sum_{i \in \{xy, xz, ...\}}\vec{S}_{4s}\cdot \vec{S}_{d_i}\} \ + \\
+\text{P}(\bar{c}_{d_\pi}^\dagger \bar{c}_{p_\pi} + h.c.,\ \bar{c}_{d_z^2}^\dagger \bar{c}_{p_z} + h.c.,\ \bar{c}_{4s}^\dagger \bar{c}_{p_z} + h.c.,\ \bar{c}_{d_z^2}^\dagger \bar{c}_{4s} + h.c.)
+\end{split}
+\label{eq:models}
+\end{equation}
+where P denotes the power set and operators are defined on the IAO basis unless denoted by a bar, in which case the operator is built on the MO basis.
+The coefficients for each term above will be denoted $\bar{\epsilon}_{4s},\ \bar{\epsilon}_{p_\pi},\ \bar{\epsilon}_{p_z},\ U_s,\ J_{sd},\ \bar{t}_\pi,\ \bar{t}_{dz},\ \bar{t}_{sz},\ \bar{t}_{ds}$.
+The operator $\bar{n}_{3d}$ is not included in any models as it is linearly dependent on all other occupation energies through the relationship $\bar{n}_{3d} + \bar{n}_{p_z} + \bar{n}_{p_\pi} + \bar{n}_{4s} = N_\text{elec} = \text{15}$.
+The symbol $\pi$ denotes a contraction over $x, y$ for p orbitals and $xz,\ yz$ for d orbitals.
+The symbol $\delta$, introduced later, denotes a contraction over $xy,\ x^2-y^2$ for d orbitals.
+
+After fitting each of the potential models in \eqref{eq:models} using ordinary linear regression (OLS) and solving the resultant models using exact diagonalization, we find a set of models which describe the energy functional on our sample set accurately but whose eigenstates and spectra differ as a consequence of intruder states below an energy of 2eV.
+The sixteen different models all regress with R$^2 >$ 0.95 on our sample data, however some models are more robust than others. 
+Models which have coefficients that are zero within 95\% confidence intervals, computed through bootstrap estimates, are removed from consideration.
+Any models which have an average errorbar in the first 20 exactly diagonalized eigenvalues, also estimated by bootstrapping, greater than 0.20 eV are ignored.
+Each of the remaining three models describe our sample set accurately, have robust non-zero coefficients, and have small error bars in eigenvalues of states which we care about the most.
+However, the eigenstates below 2eV differ greatly between these models as illustrated in Figure \ref{fig:InitED}.
+
+\begin{figure}[H]
+\begin{center}
+\includegraphics[width=0.7\linewidth]{../cuo/qwalk/old/ub3lyp_s1_/analysis/figs/init_ed.pdf}
+\end{center}
+\caption{Results of exact diagonalization showing the energies and properties of the first twenty eigenstates for three candidate models fit using ordinary linear regression.
+The ground state and first excited state are colored and have enlarged markers for clarity.
+Error bars are 95\% CI calculated by a bootstrap estimate.}
+\label{fig:InitED}
+\end{figure}
+
+All three models have a first excited state around 1eV, but only in the minimal model does this state fall near the sample set used for the fitting, and is the only model in which this state corresponds to the correct excitation O 2$p_z \rightarrow $ O 2p$_\pi$ as seen in experiment.
+In contrast, the first excited state for the other two models are visibly distant from our sample set and do not correspond to the experimental first excited state.
+The large distance of these eigenstates from our sample set indicates that their predicted eigenvalues may suffer from large extrapolation errors, and as such we will call them \textit{intruder} states.
+In Figure \ref{fig:Intruder} we present all eigenstates among the three potential models which we classify as intruders using a k-nearest neighbor approach with k = 5.
+
+Given that our intruder states correspond to states within $\mathcal{LE}$ but outside our sample set, they must live within the span of the eigenstates which were excluded in our sampling scheme as described before.
+This is evidenced by the fact that two of the intruder states correspond explicitly to excitations we could not access: Cu 3d$_{z^2} \rightarrow $ O 2p$_\pi$ and  Cu 3d$_\pi \rightarrow $ O 2p$_\pi$.
+In fact, these are the only two singly excited states which we could not access with our sampling scheme.
+While inaccessible, we can approximate these states by single rigid MO excitations above the UKS ground state, which results in states with a FN-DMC energy above 4eV.
+Including a generous 2eV reduction in FN-DMC energy due to orbital relaxation, we estimate that these two states should lie at least 2eV in energy. 
+Since all the other inaccessible eigenstates are double or higher order excitations, our very approximate exploration of the inaccessible parts of $\mathcal{LE}$ leaves us with a reasonable prior: intruder states should lie above 2eV in energy in our effective theories.
+We can enforce this prior by adding a term to our ordinary linear regression cost function
+\begin{equation}
+\text{Cost} = \sum_{i} (E_\text{eff}[\Psi_i] - E_\text{ab}[\Psi_i])^2 + \lambda \sum_{p}\text{QHL}(2 - E_\text{eff}[\Psi_p]),\ \text{QHL}(x) = \Theta(x)x^2
+\label{eq:cost}
+\end{equation}
+where $\lambda>0$ is a parameter which can be varied, QHL is a quadratic hinge loss, $\Theta$ is the Heaviside step function, the index $i$ is over our sampled states and $p$ over the selected intruder states.
+Increasing the value of $\lambda$ leads to a stronger penalty for models which do not satisfy our prior.
+
+\begin{figure}[H]
+\begin{center}
+\includegraphics[width=0.7\linewidth]{../cuo/qwalk/old/ub3lyp_s1_/analysis/figs/intruder.pdf}
+\end{center}
+\caption{Collection of unique intruder states from our three potential models selected using a k-nearest neighbor approach with k = 5.}
+\label{fig:Intruder}
+\end{figure}
+
+Using our new cost function we find a set of models which accurately describe our sample data, have very similar spectra and eigenvectors, and do not contain intruder states below the 2eV prior. 
+Shown in Figure \ref{fig:Prior} are the R$^2$ scores and QHL for the three candidate models fit at $\lambda \in [0,20]$. 
+The R$^2$ score stays above 0.95 for all models and $\lambda$, and the QHL converges quickly to nearly zero at $\lambda = 20$ for all three models. 
+We conclude that all three models at $\lambda = 20$ should be considered good models as they satisfy our prior and also explain our sample set accurately. 
+Also shown are $E_\text{eff}, E_\text{ab}$ for our samples using the minimal model at $\lambda = 20$ to illustrate the high fit quality even with a large $\lambda$.
+Figure \ref{fig:FinalED} presents the properties and eigenvalues of the lowest 30 eigenstates for the three fit models at $\lambda = 20$. 
+In comparison one can view Figure \ref{fig:InitED} as the same figure for fits at $\lambda = 0$. 
+At $\lambda = 20$ there are no outliers below 2eV in any of the three fit models. 
+The additional constraint from asserting a prior has seemingly caused a single effective theory to emerge from our set of candidate models.
+This conclusion is verified by looking at the values of the fit parameters for our three models, presented in Table 1.
+The parameter values for the three different models fit at $\lambda = 20$ are nearly identical within error bars and indicate that asserting our prior ensures that all candidate models regress to a single low-energy effective theory.
+All occupation energies are relative to $\epsilon_{d_\delta}$.
+
+Our final regressed model accurately reproduces properties and energies of the lowest lying eigenstates as seen in modern APES experiments on the CuO molecule.
+From Figure \ref{fig:FinalED} we see that the ground state of our effective theory has fully filled Cu 3d shell, nearly two electrons in O 2p$_z$ orbital, three in the O 2p$_\pi$ and around a quarter of an electron in the Cu 4s orbital.
+An approximate electronic configuration for the ground state is therefore
+be assigned 1$\sigma^2$1$\pi^4$1$\delta^4$2$\sigma^2$2$\pi^3$3$\sigma^0$ where the orbitals associated with the assignment are Cu 3d$_{z^2}$, Cu 3d$_\pi$, Cu 3d$_\delta$, O 2p$_z$, O 2p$_\pi$, and Cu 4s.
+This state also has a total spin projection S$_z = \frac{1}{2}$.
+This matches the well studied ground state electron configuration for the molecule $^2\Pi$.
+There are two excited states around 1eV in energy, the first of which is a state with the following single excitation O 2p$_z \rightarrow$ O 2p$_\pi$, has the configuration 1$\sigma^2$1$\pi^4$1$\delta^4$2$\sigma^2$2$\pi^4$4$\sigma^0$, and S$_z = \frac{1}{2}$.
+According to experiment there is also such a state at 1eV and is usually denoted as $^2X$.
+A second excited state at 1.2 eV exists which corresponds to an excitation 
+O 2p$_\pi \rightarrow$ Cu 4s, has configuration 1$\sigma^2$1$\pi^4$1$\delta^4$2$\sigma^2$2$\pi^2$4$\sigma^1$, and can either be S$_z = \frac{1}{2}, \frac{3}{2}$. 
+This is also the lowest energy eigenstate which can be a spin quartet.
+\textbf{ref} agree that this excitation should be the lowest spin quartet excitation, but argue that this state should be at an energy of 1.9 eV. 
+Our model has a different spin quartet state with a full Cu d shell at 1.9 eV, corresponding to an excitation O 2p$_z \rightarrow$ Cu 4s, with configuration 1$\sigma^2$1$\pi^4$1$\delta^4$2$\sigma^1$2$\pi^4$4$\sigma^1$.
+There are another host of eigenstates for our models at 2 eV corresponding to singles excitations of the form Cu 3d $\rightarrow$ O 2p$_\pi$. 
+This matches measurements which indicate that the lowest energy eigenstates with a single hole in the Cu 3d shell lie at 2.2 eV above the ground state.
+
+\begin{figure}[H]
+\begin{center}
+\includegraphics[width=0.7\linewidth]{../cuo/qwalk/old/ub3lyp_s1_/analysis/figs/prior_and_regr.pdf}
+\end{center}
+\caption{On the left, scores of the three candidate models at various $\lambda$ when fitting our effective theories using the cost function \eqref{eq:cost}.
+On the right we show predicted versus calculated energies for our low-energy sampled states for the minimal model fit at $\lambda = 20$.}
+\label{fig:Prior}
+\end{figure}
+
+\begin{table}[H]
+\begin{center}
+\begin{tabular}{l|llllll}
+&Min & Min & Min \\
+& & $ + \bar{c}_{d_\pi}^\dagger \bar{c}_{p_\pi}$ & $ + \bar{c}_{d_{z^2}}^\dagger \bar{c}_{p_z}$ \\ $\ $ \\ \hline 
+$\epsilon_{d_{z^2}}$& 0.32(2)& 0.32(1)& 0.29(2)\\
+$\epsilon_{d_\pi}$& 0.19(1)& 0.10(2)& 0.19(1)\\
+$\epsilon_{d_\delta}$& 0.0 & 0.0 & 0.0\\
+$\epsilon_{4s}$& 2.2(4)& 2.23(3)& 2.19(5)\\
+$\epsilon_{p_\pi}$& 1.7(3)& 1.8(2)& 1.7(3)\\
+$\epsilon_{p_z}$& 0.96(4)& 0.97(3)& 0.99(5)\\
+$t_\pi$& -0.57(1)& -0.45(2)& -0.57(1)\\
+$t_{dz}$& 0.55(3)& 0.56(2)& 0.54(2)\\
+$t_{sz}$& 0.87(2)& 0.89(2)& 0.85(2)\\
+$t_{ds}$& 0.44(1)& 0.45(1)& 0.45(2)\\
+$J_{sd}$& -0.6(1)& -0.8(1)& -0.6(1)\\
+$U_s$& 3.8(2)& 3.7(2)& 3.9(2))\\
+\end{tabular}
+\\ $\ $
+\\
+\end{center}
+Table 1: Parameters in eV for our three potential models when fit using \eqref{eq:cost} at $\lambda = 20$. Error bars are 95\% CI calculated using a bootstrap estimate.
+\end{table}
+
+\begin{figure}[H]
+\centering
+\includegraphics[width=0.7\linewidth]{../cuo/qwalk/old/ub3lyp_s1_/analysis/figs/final_ed.pdf}
+\caption{Results of exact diagonalization showing the energies and properties of the first thirty eigenstates for three candidate models using linear regression with cost function \ref{eq:cost} and $\lambda = 20$.
+The ground state and first excited state are colored and have enlarged markers for clarity.
+Error bars are 95\% CI calculated by a bootstrap estimate.}
+\label{fig:FinalED}
+\end{figure}
+
 \end{document}

### 2019-06-12 draft.tex
finished a pragraph of proposed, a bit to go!

@@
 In the heavily overdoped side, the lowest energy excitations look like well defined Fermi-liquid quasiparticles \textbf{ref}, indicating our lowest energy excitations are electron-like, and a simple non-interacting electron model (with renormalized masses, etc) will suffice.
-In the intermediate doping region where the low-energy excitations cannot be easily sectioned into spin-like or electron-like only, we see the emergence of exotic phases and where the most difficulties arise in writing down low-energy effective models.
+In the intermediate doping region where the low-energy excitations cannot be easily sectioned into spin-like or electron-like only, we see the emergence of exotic phases and where most difficulties arise in writing down low-energy effective models.
 On top of that, this entire discussion did not include any lattice effects like structural phase transitions which many strongly-correlated materials have as well \textbf{ref}.
-This discussion can be extended to the many other strongly correlated electron materials presented above as well.
 
@@
 
-We propose using a density matrix downfolding (DMD) procedure\textbf{ref} to develop low-energy effective theories of the 1-band nearest-neighbor Hubbard model on a square lattice as a stepping stone towards building effective theories for strongly correlated materials.
+We propose using a density matrix downfolding (DMD) procedure\textbf{ref} to systematically develop low-energy effective theories of the 1-band nearest-neighbor Hubbard model on a square lattice as a stepping stone towards building effective theories for strongly correlated materials.
 The Hubbard model is a simple model with nearest-neighbor hopping and a local on-site Coulomb repulsion, parameterized by a single ratio U/t:
@@
 The MO basis consists of the Cu 3d, 4s and O 2p molecular orbitals from the lowest energy restricted open-shell Kohn-Sham (ROKS) solution with $x \leftrightarrow y$ spatial symmetry, a O 2p$_z \rightarrow$ O 2p$_\pi$ excitation.
+We determine whether each basis is valid by looking at how much the trace of 1-RDM in FN-DMC on that basis differs from the expected value of 15 electrons.
+We find that within 95\% confidence the trace is 14.7-14.8 (MO) and 14.6 - 14.7 (IAO), indicating that the major variation between states in $\mathcal{LE}$ can be described by either single particle basis.
+As these two bases are used to describe the same space, they are related by a  transformation which will allow us to map between them if necessary.
 As these two bases are used to describe the same space, they are related by an orthogonal transformation which will allow us to map between them if necessary.
@@
 
+\section{Proposed work}
+We want to develop low-energy effective theories for the
+half-filled Hubbard model over the space U: 0 - 12, 1/k$_B$T: 0 - 3 using the DMD method, particularly focusing on the medium coupling region of U: 2 - 6 and 1/k$_B$T $\sim$ 1 where perturbative effective theories break down and a metal-insulator transition exists.
+Similar to the case of the \textit{ab-initio} downfolding, we will need to construct low-energy spaces, sample the constructed spaces, build effective theories and respective bases, and fit our models.
+Unlike \textit{ab-initio} downfolding, the Hubbard model downfolding has significantly fewer degrees of freedom, and therefore the DMD procedure can be done in a much more controlled fashion.
+We will begin from a starting point of highest control, namely for system sizes small enough where we can exactly diagonalize the Hubbard model.
+In the situation where we have full access to the eigenstates and spectra of a system, selecting $\mathcal{LE}$ can be done exactly as the span of some chosen $N$ lowest energy eigenstates. 
+We can further guarantee that our sample set fully encapsulates all low-energy degrees of freedom by simply sampling states parameterized as $|\Psi\rangle = \sum_{j=1}^N c_j |\Phi_j\rangle$ where $|\Phi_j\rangle$ are eigenstates of the Hubbard Hamiltonian.
+This can be seen in contrast to the CuO fitting where $\mathcal{LE}$ was defined implicitly through a number occupation descriptor and the sampling of $\mathcal{LE}$ was incomplete leading to intruder states in our initial effective theories.
+Basis construction for the Hubbard Hamiltonian is also explicitly clear as any possible single particle basis we can use is just an orthogonal transformation of the basis the Hubbard model is defined on, and is guaranteed to be a complete single particle basis.
+Again, this can be compared to the case of an \textit{ab-initio} system like the CuO molecule where bases are only approximately complete due to the huge number of degrees of freedom in the system.
+In order to include temperature into the mix we can work with the full model Hilbert space such that $\mathcal{LE} = \mathcal{H}$, but generate samples within our space using a Monte Carlo sampling method according to a Boltzmann probability distribution with variable parameter $\beta = 1/k_BT$.
+This way, our sample space will still be able to probe every excitation within a certain temperature range, the samples will be biased towards lower energy states according to the Boltzmann weight, and hence the fit model will also be biased towards describing more accurately the energy functional on the states with highest weight.
+The Monte Carlo sampling may need to be altered further in order to deal with entropic effects when the density of states is very low in an energy range of interest and relatively high elsewhere.
+
+Show in \ref{fig:Hubbard} is ... \textbf{SHOW OUR PRELIMINARY RESULTS}
+
+In order to make any statements about effective theories for the Hubbard model we will have to work with larger system sizes and approximate methods like lattice quantum Monte Carlo to generate low-energy states for the DMD procedure.
+\textbf{TALK ABOUT FINITE SIZE SCALING AND APPROXIMATE METHODS (liken to FN-DMC for ab initio)}
+
+A farther reaching goal would be to develop effective theories in the U-T-$\delta$ (doping) space for the Hubbard model, where more complex behavior similar to the cuprates superconductors have been found.
+
 \end{document}

### 2019-06-12 draft.tex
need to just finish state of the art and finite size/approximate methods. done!

@@
 In the region $U = 0$ a trivial non-interacting theory holds.
-Between these two limits, in particular the intermediate coupling region $1 < U < 7$, no well established effective theory exists.
+Between these two limits, in particular the intermediate coupling region $2 < U < 8$, no well established effective theory exists.
 In this region there is a metal-insulator transition at $U_c/t \sim 4$ \textbf{ref}, and the pseudogap opens up at finite temperature around $U/t \sim 7$, leading to the development of a Mott gap for large $U/t > 11$ \textbf{ref}.
@@
 \section{Proposed work}
-We want to develop low-energy effective theories for the
-half-filled Hubbard model over the space U: 0 - 12, 1/k$_B$T: 0 - 3 using the DMD method, particularly focusing on the medium coupling region of U: 2 - 6 and 1/k$_B$T $\sim$ 1 where perturbative effective theories break down and a metal-insulator transition exists.
+Our goal is to develop low-energy effective theories for the
+Hubbard model over the space U-T-$\delta$ using the DMD method, particularly focusing on the medium coupling regions and moderate dopings where perturbative effective theories break down.
 Similar to the case of the \textit{ab-initio} downfolding, we will need to construct low-energy spaces, sample the constructed spaces, build effective theories and respective bases, and fit our models.
@@
 
-Show in \ref{fig:Hubbard} is ... \textbf{SHOW OUR PRELIMINARY RESULTS}
+In order to make any statements about effective theories for the Hubbard model we will have to work with larger system sizes and approximate methods like lattice quantum Monte Carlo to generate low-energy states for the DMD procedure.
 
-In order to make any statements about effective theories for the Hubbard model we will have to work with larger system sizes and approximate methods like lattice quantum Monte Carlo to generate low-energy states for the DMD procedure.
-\textbf{TALK ABOUT FINITE SIZE SCALING AND APPROXIMATE METHODS (liken to FN-DMC for ab initio)}
+We conclude by showing some very preliminary results on model fitting for the half-filled, 2x2 square Hubbard model for U: 0 - 12 which put our approach to model fitting with exact diagonalization at T$\geq 0$ into practice.
+We start off by exactly diagonalizing each model, define $\mathcal{LE} = \mathcal{H}$, and treat as our full sample space as just the eigenstates of the model.
+We then fit three single parameter models to the eigenstates: A) 
+A "t" only model $t \sum_{\langle i, j \rangle, \sigma} c_{i\sigma}^\dagger c_{j\sigma} + h.c.$, B) A "U" only model $U \sum_{i} n_{i\uparrow} n_{i\downarrow}$, and C) A "J" only model $J \sum_{\langle i,j \rangle} \vec{S}_i \cdot \vec{S}_j$ using a weighted linear regression with weights $e^{-\beta E}$.
+The basis for all of these operators are just the bare Hubbard Hamiltonian basis.
+We have restricted ourselves to eigenstates only as we have not developed a sufficient low-energy sampling scheme yet which deals with the entropy issue discussed earlier.
+The weighted linear regression approximates the effect of generating samples using a Markov Chain Monte Carlo method with a Boltzmann distribution.
+For small U the t-only model is an accurate effective theory as the temperature increases, which is to be expected in this perturbative limit.
+If one were to look at large U and high temperature a U-only model is quite good since the dominant and relevant energy scale at higher temperatures is related to double occupation.
+At large U and low temperature we see that the expected Heisenberg model accurately describes the low-energy excitations. 
+In this region a t-only model also works quite accurately since for large U the hopping descriptor $\sum_{\langle i, j \rangle, \sigma} c_{i\sigma}^\dagger c_{j\sigma} + h.c.$ is correlated to the exchange operator $J \sum_{\langle i,j \rangle} \vec{S}_i \cdot \vec{S}_j$ for the lowest energy excitations in the system.
+It should be noted, however, that this t-only model would have to be explicitly written to only act on states with no double occupancy, an assumption inherent in the Heisenberg model. 
+Finally, in the region beginning at 4 $<$ U $<$ 8 and moderate $\beta$: 0.1 - 0.4 we see that none of the candidate single parameter models can accurately describe the low-energy excitations of the system.
+This poorly explained region indicates the breakdown of perturbative expansions of the Hubbard model about the strong- or weak-coupling limits.
+In order to make more conclusive statements we would have to actually sample the Hilbert space to construct a more robust estimate of the best models for different $U, \beta$.
+We would be particularly interested in whether an effective theory for this region exists and how this region and corresponding effective theory change as we increase the system size.
 
-A farther reaching goal would be to develop effective theories in the U-T-$\delta$ (doping) space for the Hubbard model, where more complex behavior similar to the cuprates superconductors have been found.
+\begin{figure}[H]
+\centering
+\includegraphics[width=0.7\linewidth]{../hubbard/eff_eig_prelim.pdf}
+\caption{Results of weighted single parameter model fits to the eigenstates of a Hubbard model for various U and temperature. Darker hues indicate a higher R$^2$ score.}
+\label{fig:Hubbard}
+\end{figure}
 

### 2019-06-13 draft.tex
made 1.5 linespace
made 12pt font. will need to trim preliminary work a lot
finished approximate methods + finite size scaling, timeline. need to trim a lot!

@@
-\documentclass{article}
+\documentclass[12pt]{article}
 \usepackage{graphicx}
@@
 \usepackage{wrapfig}
+\renewcommand{\baselinestretch}{1.5}
 
@@
 A core problem of condensed matter physics is understanding how to systematically develop low-energy effective models for strongly correlated electron systems.
-In the forefront of modern research are high temperature superconductors, heavy fermion materials, materials with strong spin-orbit coupling effects, and more for which simple effective theories have not yet been found.
-For example, cuprate family of high-Tc superconductors have an intricate phase diagram in the doping-temperature space \textbf{ref} with a robust anti-ferromagnetic (AFM) charge-transfer insulator state near zero doping, a prominent superconducting dome in the 0.1 - 0.2 hole doped region at low temperatures with psuedo-gap and non-Fermi liquid metallic phases at higher temperatures and Fermi liquid like behavior in the overdoped area, of which only the AFM insulating and Fermi liquid phases have well understood mechanism and models \textbf{ref}.
-The iron-pnictides also have superconducting domes, non-Fermi liquid like metallic phases, a spin-density-wave antiferommagnetic phase, all under various dopings and temperatures \textbf{ref}.
-As in the cuprate case, models for these phases have not been fully fleshed out \textbf{ref}.
-Heavy-fermion materials have a purported quantum critical point between an AFM and Fermi-liquid phase which houses a superconducting dome, leading to another phase diagram for which building model Hamiltonians is very challenging \textbf{ref}.
-Materials with strong spin-orbit coupling effects exhibit a host of properties like \textbf{LIST THESE HERE!}.
+In the forefront of modern research are materials like high temperature superconductors and heavy fermion systems for which simple effective theories have not yet been found.
+For example, the cuprate family of high-Tc superconductors have an intricate phase diagram in the doping-temperature space  with a robust anti-ferromagnetic (AFM) charge-transfer insulator state near zero doping, a prominent superconducting dome in the 0.1 - 0.2 hole doped region at low temperatures with psuedo-gap and non-Fermi liquid metallic phases at higher temperatures and Fermi liquid like behavior in the overdoped area, of which only the AFM insulating and Fermi liquid phases have well understood mechanism and models.
+The iron-pnictides also have superconducting domes, non-Fermi liquid like metallic phases, a spin-density-wave antiferommagnetic phase, all under various dopings and temperatures.
+As in the cuprate case, models for these phases have not been fully fleshed out.
+Heavy-fermion materials have a purported quantum critical point between an AFM and Fermi-liquid phase which houses a superconducting dome, leading to another phase diagram for which building model Hamiltonians is very challenging.
 
@@
 Consider just the cuprate case for now, the two phases which are well understood contain a singular important degree of freedom in the low-energy space.
-Near zero doping the cuprates are charge-transfer insulators with an optical gap around 2 eV \textbf{ref}, indicating that the lowest energy excitations are primarily spin-like, allowing for a simple nearest-neighbor spin-only model like an AFM Heisenberg model to describe the low-energy physics quite accurately.
-In the heavily overdoped side, the lowest energy excitations look like well defined Fermi-liquid quasiparticles \textbf{ref}, indicating our lowest energy excitations are electron-like, and a simple non-interacting electron model (with renormalized masses, etc) will suffice.
+Near zero doping the cuprates are charge-transfer insulators with an optical gap around 2 eV , indicating that the lowest energy excitations are primarily spin-like, allowing for a simple nearest-neighbor spin-only model like an AFM Heisenberg model to describe the low-energy physics quite accurately.
+In the heavily overdoped side, the lowest energy excitations look like well defined Fermi-liquid quasiparticles , indicating our lowest energy excitations are electron-like, and a simple non-interacting electron model (with renormalized masses, etc) will suffice.
 In the intermediate doping region where the low-energy excitations cannot be easily sectioned into spin-like or electron-like only, we see the emergence of exotic phases and where most difficulties arise in writing down low-energy effective models.
-On top of that, this entire discussion did not include any lattice effects like structural phase transitions which many strongly-correlated materials have as well \textbf{ref}.
+On top of that, this entire discussion did not include any lattice effects like structural phase transitions which many strongly-correlated materials have as well.
 
-\textbf{Current state of the art.}
+Common approaches to building suitable model Hamiltonians for strongly correlated materials involve using single particle theories like density functional theory (DFT).
+One-body terms are typically obtained by projecting the solutions from the DFT band structure onto a localized single-particle basis, giving an estimate of hoppings and occupations on a lattice model.
+Two-body terms in the model are developed by assuming a screened Coulomb interaction based on constrained DFT or RPA.
+Aside from this method, Lowdin methods coupled to a stochastic method and canonical transformations have also been used to develop effective theories.
+These methods have the drawback that the constructed effective theory typically does not come with any estimate of the quality of the model, therefore making model validation a phenomenological task which can lead to unclear biases in the model construction.
 
-We propose using a density matrix downfolding (DMD) procedure\textbf{ref} to systematically develop low-energy effective theories of the 1-band nearest-neighbor Hubbard model on a square lattice as a stepping stone towards building effective theories for strongly correlated materials.
+We propose using a density matrix downfolding (DMD) procedure to systematically develop low-energy effective theories of the 1-band nearest-neighbor Hubbard model on a square lattice as a stepping stone towards building effective theories for strongly correlated materials.
 The Hubbard model is a simple model with nearest-neighbor hopping and a local on-site Coulomb repulsion, parameterized by a single ratio U/t:
@@
 \end{equation}
-The solutions in 1-d are well understood via the Bethe ansatz \textbf{ref}, however the 2-d square lattice Hubbard model has no known exact solution.
-In the very strong coupling limit $U/t \rightarrow \infty$ an emergent low-energy spin-space exists with an effective nearest-neighbor AFM Heisenberg Hamiltonian \textbf{ref}.
+The solutions in 1-d are well understood via the Bethe ansatz , however the 2-d square lattice Hubbard model has no known exact solution.
+In the very strong coupling limit $U/t \rightarrow \infty$ an emergent low-energy spin-space exists with an effective nearest-neighbor AFM Heisenberg Hamiltonian.
 In the region $U = 0$ a trivial non-interacting theory holds.
 Between these two limits, in particular the intermediate coupling region $2 < U < 8$, no well established effective theory exists.
-In this region there is a metal-insulator transition at $U_c/t \sim 4$ \textbf{ref}, and the pseudogap opens up at finite temperature around $U/t \sim 7$, leading to the development of a Mott gap for large $U/t > 11$ \textbf{ref}.
-In addition to the half-filling phase diagram, this model exhibits a plethora of interesting ground state properties when doped including stripe phases and d-wave pairing \textbf{ref}.
+In this region there is a metal-insulator transition at $U_c/t \sim 4$ , and the pseudogap opens up at finite temperature around $U/t \sim 7$, leading to the development of a Mott gap for large $U/t > 11$.
+In addition to the half-filling phase diagram, this model exhibits a plethora of interesting ground state properties when doped including stripe phases and d-wave pairing.
 The relative simplicity of the Hubbard model and the lack of a well-established low-energy theory in a broad intermediate coupling region lends the model as a useful playground for our new downfolding procedure in preparation for tackling realistic strongly correlated electron systems.
@@
 
-\pagebreak
 \section{Preliminary Work}
@@
 The coefficients $\{c_j\}$ for our sample states were chosen via a shell-sampling method. 
-We begin by fixing the coefficient $c_0 = \sqrt{w}$ where $w \in \{1.0, 0.8, ..., 0.2\}$. 
+We begin by fixing the coefficient $c_0 = \sqrt{w}$ where $w \in \{1.0, 0.8,..., 0.2\}$. 
 For each choice of $w$ we sample n = 5 states by randomly selecting the unassigned coefficients from a uniform distribution such that $\sum_i c_i^2 = 1$. 
@@
 \begin{split}
-\{\bar{n}_{4s}, \bar{n}_{p_\pi}, \bar{n}_{p_z}, \hat{n}_{4s\uparrow} \hat{n}_{4s\downarrow},\sum_{i \in \{xy, xz, ...\}}\vec{S}_{4s}\cdot \vec{S}_{d_i}\} \ + \\
+\{\bar{n}_{4s}, \bar{n}_{p_\pi}, \bar{n}_{p_z}, \hat{n}_{4s\uparrow} \hat{n}_{4s\downarrow},\sum_{i \in \{xy, xz,...\}}\vec{S}_{4s}\cdot \vec{S}_{d_i}\} \ + \\
 \text{P}(\bar{c}_{d_\pi}^\dagger \bar{c}_{p_\pi} + h.c.,\ \bar{c}_{d_z^2}^\dagger \bar{c}_{p_z} + h.c.,\ \bar{c}_{4s}^\dagger \bar{c}_{p_z} + h.c.,\ \bar{c}_{d_z^2}^\dagger \bar{c}_{4s} + h.c.)
@@
 This is also the lowest energy eigenstate which can be a spin quartet.
-\textbf{ref} agree that this excitation should be the lowest spin quartet excitation, but argue that this state should be at an energy of 1.9 eV. 
+ agree that this excitation should be the lowest spin quartet excitation, but argue that this state should be at an energy of 1.9 eV. 
 Our model has a different spin quartet state with a full Cu d shell at 1.9 eV, corresponding to an excitation O 2p$_z \rightarrow$ Cu 4s, with configuration 1$\sigma^2$1$\pi^4$1$\delta^4$2$\sigma^1$2$\pi^4$4$\sigma^1$.
@@
 
-In order to make any statements about effective theories for the Hubbard model we will have to work with larger system sizes and approximate methods like lattice quantum Monte Carlo to generate low-energy states for the DMD procedure.
+In order to make any statements about effective theories for the Hubbard model we will have to work with larger system sizes and approximate methods like auxiliary field quantum Monte Carlo (AFQMC) to generate low-energy states for the DMD procedure.
+The auxiliary field quantum Monte Carlo (AFQMC) method has been used extensively in accurate calculations on both \textit{ab-initio} and lattice systems.
+AFQMC begins with a trial wave function which is projected using a Trotter expansion of the operator $e^{-\tau H}$ using a stochastic method.
+The final projected state from AFQMC can be used to calculate expectation values of operators using a similar mixed-estimator method as FN-DMC.
+The AFQMC method is especially useful for the Hubbard model as the auxiliary fields are real with potential values $\pm$ 1, only N such fields need to be considered, and at least at half-filling there is no sign problem and the method is numerically exact.
+Away from half-filling the effects of the fermion sign problem are especially potent and the nodes of the projected wave function are fixed to those of the trial wave function making the method variational in the same way FN-DMC is.
+In the weak and strong coupling regimes typical choices for trial wave functions are non-interacting and Gutzwiller trial functions respectively, but other post-Gutzwiller and generalized Hartree Fock (GHF) trial functions have been shown to give accurate results in the moderate coupling regimes.
+A rigorous prescription for finite size extrapolation specifically for the 1-band square lattice Hubbard model will be followed to extend our conclusions to the thermodynamic limit.
 
@@
 
+\subsection{Proposed timeline: 2 years}
+\begin{enumerate}
+\item Half-filled model ED
+	\begin{enumerate}
+	\item Constructing a sampling scheme for $\mathcal{LE}$: 2 months
+	\item DMD for 2x2, 2$\sqrt{2}$x2$\sqrt{2}$, 3x3, 4x4
+	system sizes: 4 months
+	\end{enumerate}
+\item Half-filled model AFQMC
+	\begin{enumerate}
+	\item Learning and benchmarking calculations: 1 month
+	\item Extending sampling scheme for $\mathcal{LE}$: 1 month
+	\item DMD calculations up to 16x16 cells: 4 months
+	\item Finite size scaling of results: 2 months
+	\end{enumerate}
+\item Doped model ED
+	\begin{enumerate}
+	\item DMD for  2x2, 2$\sqrt{2}$x2$\sqrt{2}$, 3x3, 4x4
+	system sizes, various dopings: 4 months
+	\end{enumerate}
+\item Doped model AFQMC
+	\begin{enumerate}
+	\item Constructing a sampling scheme - trial functions for constrained
+	path: 2 months
+	\item DMD calculations up to 16x16 cells: 4 months
+	\end{enumerate}
+\end{enumerate}
+
 \end{document}

### 2019-06-13 draft.tex
reduced to 13 pages! references and maybe bigger/fewer figures?

@@
 \usepackage{xcolor}
-\usepackage{subcaption}
+\usepackage{subfigure}
 \usepackage[margin=1.0in]{geometry}
@@
 The iron-pnictides also have superconducting domes, non-Fermi liquid like metallic phases, a spin-density-wave antiferommagnetic phase, all under various dopings and temperatures.
-As in the cuprate case, models for these phases have not been fully fleshed out.
 Heavy-fermion materials have a purported quantum critical point between an AFM and Fermi-liquid phase which houses a superconducting dome, leading to another phase diagram for which building model Hamiltonians is very challenging.
@@
 Consider just the cuprate case for now, the two phases which are well understood contain a singular important degree of freedom in the low-energy space.
-Near zero doping the cuprates are charge-transfer insulators with an optical gap around 2 eV , indicating that the lowest energy excitations are primarily spin-like, allowing for a simple nearest-neighbor spin-only model like an AFM Heisenberg model to describe the low-energy physics quite accurately.
-In the heavily overdoped side, the lowest energy excitations look like well defined Fermi-liquid quasiparticles , indicating our lowest energy excitations are electron-like, and a simple non-interacting electron model (with renormalized masses, etc) will suffice.
+Near zero doping the cuprates are charge-transfer insulators with localized spin-1/2 low-energy degrees of freedom.
+In the heavily overdoped side, the lowest energy excitations look like well defined Fermi-liquid quasiparticles.
 In the intermediate doping region where the low-energy excitations cannot be easily sectioned into spin-like or electron-like only, we see the emergence of exotic phases and where most difficulties arise in writing down low-energy effective models.
@@
 Between these two limits, in particular the intermediate coupling region $2 < U < 8$, no well established effective theory exists.
-In this region there is a metal-insulator transition at $U_c/t \sim 4$ , and the pseudogap opens up at finite temperature around $U/t \sim 7$, leading to the development of a Mott gap for large $U/t > 11$.
+In this region there is a metal-insulator transition at $U_c/t \sim 4$, and the pseudogap opens up at finite temperature around $U/t \sim 7$, leading to the development of a Mott gap for large $U/t > 11$.
 In addition to the half-filling phase diagram, this model exhibits a plethora of interesting ground state properties when doped including stripe phases and d-wave pairing.
@@
 
-The DMD procedure allows us to develop low-energy effective theories for physical systems in a systematic manner beginning with the \textit{ab-initio} Hamiltonian:
-\begin{equation}
-\hat{H}_\text{ab} = -\frac{1}{2} \sum_{i} \nabla_i^2 - \sum_{i,I}\frac{Z}{|r_i - R_I|} + \sum_{i<j}\frac{1}{|r_i - r_j|}
-\label{eq:Hab}
-\end{equation}
-where $i$ indicates electrons and $I$ nuclei, and the unit of energy is Hartree (Ha).
-In this paper we will be working under the Born-Oppenheimer approximation and core electrons will be treated using pseudo-potentials.
+The DMD procedure allows us to systematically develop low-energy effective theories for any Hamiltonian and Hilbert space $(H, \mathcal{H})$.
 The procedure defines a method for downfolding the \textit{ab-initio} energy functional and Hilbert space onto an effective energy functional defined on a low-energy subspace of the Hilbert space 
 \begin{equation}
-(E_\text{ab}, \mathcal{H}) \xrightarrow{\text{DMD}} (E_\text{eff}, \mathcal{LE} \subset \mathcal{H}), \text{ s.t. }
-\forall |\Psi\rangle \in \mathcal{LE}, \ E_\text{eff}[\Psi] - E_\text{ab}[\Psi] = \epsilon[\Psi] + E_0.
+(E, \mathcal{H}) \xrightarrow{\text{DMD}} (E_\text{eff}, \mathcal{LE} \subset \mathcal{H}), \text{ s.t. }
+\forall |\Psi\rangle \in \mathcal{LE}, \ E_\text{eff}[\Psi] - E[\Psi] = \epsilon[\Psi] + E_0.
 \label{eq:DMD}
 \end{equation} 
-The energy functionals can be written as expectations over Hamiltonian operators $E_\text{ab} = \langle \Psi | \hat{H}_\text{ab} |\Psi \rangle$, $E_\text{eff} = \langle \Psi | \hat{H}_\text{eff} |\Psi \rangle$, $\epsilon$ represents the error in our effective theory, and $E_0$ is a constant energy shift.
+The energy functionals can be written as expectations over Hamiltonian operators $E = \langle \Psi | \hat{H} |\Psi \rangle$, $E_\text{eff} = \langle \Psi | \hat{H}_\text{eff} |\Psi \rangle$, $\epsilon$ represents the error in our effective theory, and $E_0$ is a constant energy shift.
 Broadly the DMD method consists of four steps which will be discussed in detail later: 
@@
 The fitting procedure is typically an ordinary linear regression, however, one may use a custom cost function in the regression as well. 
-In order to conduct the linear regression for a functional like \eqref{eq:Eeff}, one needs to calculate for each state sampled from $\mathcal{LE}$ the expectation values of $\hat{H}_\text{ab}$ and every descriptor $\{\hat{d}_k\}$ used in the regression.
+In order to conduct the linear regression for a functional like \eqref{eq:Eeff}, one needs to calculate for each state sampled from $\mathcal{LE}$ the expectation values of $\hat{H}$ and every descriptor $\{\hat{d}_k\}$ used in the regression.
 In principle, if conducted with no approximations, the DMD method is rigorously guaranteed to obtain exact results, namely $\epsilon[\Psi] = 0$.
+For real world systems we will work with $H = H_\text{ab}$, the \textit{ab-initio} Hamiltonian under the Born-Oppenheimer approximation with psuedopotentials:
+\begin{equation}
+\hat{H}_\text{ab} = -\frac{1}{2} \sum_{i} \nabla_i^2 - \sum_{i,I}\frac{Z}{|r_i - R_I|} + \sum_{i<j}\frac{1}{|r_i - r_j|}
+\label{eq:Hab}
+\end{equation}
+where $i$ indicates electrons and $I$ nuclei, and the unit of energy is Hartree (Ha).
 
@@
 The constant $E_T$ is a trial energy used to control the normalization of $\Psi_\text{DMC}(\tau, R)$ and is updated at each move.
-This implementation, however, suffers from a fermion sign problem. 
-We deal with this sign problem through a fixed-node approximation, where the nodal surface of $\Psi_\text{DMC}(\tau, R)$ is forced to match that of the initial trial wave function for all $\tau$.
+This implementation suffers from a fermion sign problem which is dealt with through a fixed-node approximation where the nodal surface of $\Psi_\text{DMC}(\tau, R)$ is forced to match that of the initial trial wave function for all $\tau$.
 This approximation makes FN-DMC variational, and will only return the exact ground state of $\hat{H}_\text{ab}$ if the nodal surfaces of $|\Psi_T\rangle$ and $|\Phi_0\rangle$ are identical.
@@
 We construct our $\mathcal{LE}$ based on recent anion photoelectron spectroscopy (APES) measurements of the low-lying excited states of the nuetral CuO molecule. 
-The lowest energy excitations measured, which lie below 2eV, involve primarily the O 2p$_z$, 2p$_\pi$ and Cu 4s orbitals, which are fully-filled, half-filled and empty in the ground state respectively.
-Beginning at 2.2 eV are excitations which remove a single electron from the fully filled Cu 3d shell and add them to either the O 2p$_\pi$ or Cu 4s orbitals.
-The neutral copper atom also carries a Hund's coupling between the d and s orbitals of about $\sim$ 0.5 eV, the effects of which would be measurable among the states in $\mathcal{LE}$ as long as states of both doublet and quartet states are included.
-In order to probe all of these low-energy excitations we select a low-energy space
+The lowest lying excitations involve electrons in the Cu 3d, Cu 4s and O 2p orbitals.
+Given that the lowest energy excitations contain a fully-filled Cu 3d shell or a single hole in this shell we define our low-energy space as:
 \begin{equation}
@@
 Our sample space was generated by using multi-Slater-Jastrow trial wavefunctions in FN-DMC.
-In principle we could probe every low-energy excitation in $\mathcal{LE}$ in a statistically independent manner by just sampling the eigenstates within $\mathcal{LE}$ and including various non-eigenstates within $\mathcal{LE}$ as well to ensure robust statistics.
-In practice we do not have access to these eigenstates and approximate methods are required to sample the low-energy space. 
-Consider a trial wavefunction of the form $e^{J_i} |\text{D}_i\rangle$, where $e^{J_i}$ is a three-body Jastrow factor and $|\text{D}_i\rangle$ is a determinant of some single particle orbitals.
-If the nodes of $|\text{D}_i\rangle$ are similar to the nodes of an eigenstate $|\Phi_i\rangle \in \mathcal{LE}$, then the properties and energy calculated from FN-DMC using this trial wavefunction will approximate those of the eigenstate, with the degree of approximation depending on the quality of the trial function nodal structure.
-If we were able to find for each $|\Phi_i \rangle \in \mathcal{LE}$ such a $|\text{D}_i\rangle$ then we can approximate any state in $\mathcal{LE}$ as
+The parameterization for our low-energy trial wave functions was
 \begin{equation}
-|\Psi\rangle = \sum_i c_i |\Phi_i\rangle \sim \sum_i c_i \lim_{\tau \rightarrow \infty} e^{-\tau \hat{H}_\text{ab}} (e^{J_i} |D_i\rangle) \xrightarrow{J_i = J} \lim_{\tau \rightarrow \infty} e^{-\tau \hat{H}_\text{ab}} (e^{J}\sum_{i} c_i|\text{D}_i\rangle).
+e^{J}\sum_{i} c_i|\text{D}_i\rangle = \sum_{i} c_i (e^J |\text{D}_i \rangle) \xrightarrow{FN-DMC} \ \sim \sum_i c_i |\Phi_i \rangle \in \mathcal{LE}
 \label{eq:sampling}
 \end{equation}
-The last equality is valid if $\forall i\  J_i = J$, allowing us to map $|\Psi \rangle \in \mathcal{LE}$ to the end result of an FN-DMC calculation using a multi-Slater-Jastrow trial wavefunction. 
-We use symmetry-targeted unrestricted Kohn-Sham (UKS) to generate our determinants $|\text{D}_i\rangle$ using a B3LYP functional, which has been shown to accurately reproduce ground state nodal properties of transition-metal oxide molecules, a Trail-Needs pseudopotential, and the VTZ Trail-Needs basis using the package PySCF.
-This method allows us to access almost every excitation in $\mathcal{LE}$, however, there are some cases of two or more low-energy states within $\mathcal{LE}$ which have identical S$_z$ and number of electrons per irrep.
-An example of this dilemma would be the ground state and the state with an excitation from Cu 3d$_{xz} \rightarrow $  O 2p$_x$.
-In these few cases the higher energy excited $|\text{D}_i\rangle$ is inaccessible and the corresponding $|\Phi_i\rangle$ is excluded from the sum in \eqref{eq:sampling}, restricting our sample space to a subspace of $\mathcal{LE}$.
-We will discuss the consequences of this restriction later on.
-A single three-body Jastrow factor $J$ was used for every $|\text{D}_i\rangle$ and was optimized on the lowest energy UKS state using a linear energy optimization method.
-The FN-DMC calculations were conducted with T-moves, to make the calculation variational while using pseudopotentials, with a timestep of $\tau = 0.01$.
+where $e^J$ is a three-body Jastrow factor and $|\text{D}_i\rangle$ is a Slater determinant whose nodes accurately approximate those of an eigenstate $|\Phi_i \rangle \in \mathcal{LE}$.
+Under the FN-DMC projection these trial wave functions map to states which are very close to linear combinations of eigenstates in $\mathcal{LE}$.
+We used symmetry-targeted unrestricted Kohn-Sham (UKS) to generate our determinants $|\text{D}_i\rangle$ using a B3LYP functional, a Trail-Needs pseudopotential, and the VTZ Trail-Needs basis using the package PySCF.
+This method allows us to access almost every excitation in $\mathcal{LE}$, except the few cases when two states have identical S$_z$ and number of electrons per irrep.
+An example would be the ground state and the state with an excitation from Cu 3d$_{xz} \rightarrow $  O 2p$_x$, where the latter would be inaccessible.
+The FN-DMC calculations were conducted with T-moves to make the calculation variational while using pseudopotentials, and a timestep of $\tau = 0.01$.
 The coefficients $\{c_j\}$ for our sample states were chosen via a shell-sampling method. 
@@
 For each choice of $w$ we sample n = 5 states by randomly selecting the unassigned coefficients from a uniform distribution such that $\sum_i c_i^2 = 1$. 
-This procedure generates shells of samples at decreasing distance from the eigenstate $|\Phi_0\rangle$. 
-We can then loop over all $i$ to generate shells near each eigenstate.
+We can then loop over all $i$ to generate shells of decreasing distance near each eigenstate $|\Phi_i\rangle$.
 
-Parameterizing our effective model requires constructing a basis on which to calculate our \textit{ab-initio} 1-/2-RDM elements and selecting a set of descriptors from which we would like to build our model.
-We constructed two such bases: a localized intrinsic atomic orbital (IAO) basis and a molecular orbital (MO) basis.
-The IAO basis was built on all Cu 3d, 4s and O 2p molecular orbitals from the $|\text{D}_i\rangle$ used to construct our sample states.
-The MO basis consists of the Cu 3d, 4s and O 2p molecular orbitals from the lowest energy restricted open-shell Kohn-Sham (ROKS) solution with $x \leftrightarrow y$ spatial symmetry, a O 2p$_z \rightarrow$ O 2p$_\pi$ excitation.
-We determine whether each basis is valid by looking at how much the trace of 1-RDM in FN-DMC on that basis differs from the expected value of 15 electrons.
-We find that within 95\% confidence the trace is 14.7-14.8 (MO) and 14.6 - 14.7 (IAO), indicating that the major variation between states in $\mathcal{LE}$ can be described by either single particle basis.
-As these two bases are used to describe the same space, they are related by a  transformation which will allow us to map between them if necessary.
-As these two bases are used to describe the same space, they are related by an orthogonal transformation which will allow us to map between them if necessary.
-In order to describe the low-energy degrees of freedom in the CuO molecule our model will require occupation energies and hybridizations between the Cu 3d, 4s and O 2p atomic orbitals which are most compactly represented in an MO basis.
-The lack of doubly occupied Cu 4s states in the low-energy spectrum of CuO indicate the necessity of a Hubbard U term on the Cu 4s.
-This is corroborated by the fact that the FN-DMC energy of the lowest-lying sampled state with a doubly occupied Cu 4s orbital is 7 eV above the lowest energy sampled state.
-Further, we include a Hund's coupling between the Cu 4s and Cu 3d orbitals in our model based on the existence of a similar Hund's coupling 
-for the copper atom.
-These two interaction terms are best represented in the IAO basis as they are intuitively interactions between localized orbitals.
-In addition to these important parameters, we will consider models which contain any number of hybridizations between the MOs to account for orbital relaxation between excited states in the CuO molecule.
+Parameterizing our effective model requires selecting a set of descriptors from which we would like to build our model and a basis on which to build our descriptors.
+Occupation energies of and hybridization between Cu 3d, Cu 4s and O 2p are necessary to describe the low-energy excitations in CuO and are best represented by the occupation energies in a molecular orbital basis (MO) which would neatly package both effects.
+The lack of doubly occupied Cu 4s states in the low-energy spectrum of CuO and the existence of a large Hund's coupling between the Cu 4s and 3d orbitals in the Cu atom indicate the necessity for a Hubbard U$_s$ and Hund's J$_{sd}$ respectively.
+These two interaction terms are best represented in a localized intrinsic atomic orbital (IAO).
+To account for orbital relaxation between excited states in the CuO molecule we will consider including additional hybridization between MOs.
 In total our space of canditate models is then 
@@
 The coefficients for each term above will be denoted $\bar{\epsilon}_{4s},\ \bar{\epsilon}_{p_\pi},\ \bar{\epsilon}_{p_z},\ U_s,\ J_{sd},\ \bar{t}_\pi,\ \bar{t}_{dz},\ \bar{t}_{sz},\ \bar{t}_{ds}$. 
-The operator $\bar{n}_{3d}$ is not included in any models as it is linearly dependent on all other occupation energies through the relationship $\bar{n}_{3d} + \bar{n}_{p_z} + \bar{n}_{p_\pi} + \bar{n}_{4s} = N_\text{elec} = \text{15}$.
-The symbol $\pi$ denotes a contraction over $x, y$ for p orbitals and $xz,\ yz$ for d orbitals.
-The symbol $\delta$, introduced later, denotes a contraction over $xy,\ x^2-y^2$ for d orbitals.
+All energies will be relative to $\bar{\epsilon_{3d}} = 0$. 
+The symbol $\pi$ denotes a contraction over $x, y$ for p orbitals and $xz,\ yz$ for d orbitals. The symbol $\delta$, introduced later, denotes a contraction over $xy,\ x^2-y^2$ for d orbitals.
 
-After fitting each of the potential models in \eqref{eq:models} using ordinary linear regression (OLS) and solving the resultant models using exact diagonalization, we find a set of models which describe the energy functional on our sample set accurately but whose eigenstates and spectra differ as a consequence of intruder states below an energy of 2eV.
-The sixteen different models all regress with R$^2 >$ 0.95 on our sample data, however some models are more robust than others. 
-Models which have coefficients that are zero within 95\% confidence intervals, computed through bootstrap estimates, are removed from consideration.
-Any models which have an average errorbar in the first 20 exactly diagonalized eigenvalues, also estimated by bootstrapping, greater than 0.20 eV are ignored.
-Each of the remaining three models describe our sample set accurately, have robust non-zero coefficients, and have small error bars in eigenvalues of states which we care about the most.
-However, the eigenstates below 2eV differ greatly between these models as illustrated in Figure \ref{fig:InitED}.
+After fitting each of the potential models in \eqref{eq:models} using ordinary least squares (OLS) linear regression we find three models which describe the energy functional on our sample set accurately with low variance, but whose eigenstates and spectra differ as a consequence of intruder states below an energy of 2eV as illustrated in Figure \ref{fig:InitED}.
+All three models have a first excited state around 1eV, but only in the minimal model does this state fall near the sample set used for the fitting.
+The large distance of the eigenstates for the other two models from our sample set indicates that their predicted eigenvalues may suffer from large extrapolation errors, and as such we will call them \textit{intruder} states.
+In Figure \ref{fig:Intruder} we present all eigenstates among the three potential models which we classify as intruders using a k-nearest neighbor approach with k = 5.
 
 \begin{figure}[H]
-\begin{center}
-\includegraphics[width=0.7\linewidth]{../cuo/qwalk/old/ub3lyp_s1_/analysis/figs/init_ed.pdf}
-\end{center}
-\caption{Results of exact diagonalization showing the energies and properties of the first twenty eigenstates for three candidate models fit using ordinary linear regression.
+	\centering
+	\subfigure[Results of exact diagonalization showing the energies and properties of the first twenty eigenstates for three candidate models fit using ordinary linear regression.
 The ground state and first excited state are colored and have enlarged markers for clarity.
-Error bars are 95\% CI calculated by a bootstrap estimate.}
+Error bars are 95\% CI calculated by a bootstrap estimate.]{
 	 	\label{fig:InitED}
+		\includegraphics[width=0.45\textwidth]{../cuo/qwalk/old/ub3lyp_s1_/analysis/figs/init_ed.pdf}
+	}
+	\qquad
+	\subfigure[Collection of unique intruder states from our three potential models selected using a k-nearest neighbor approach with k = 5.]{	
+		\label{fig:Intruder}
+		\includegraphics[width=0.45\textwidth]{../cuo/qwalk/old/ub3lyp_s1_/analysis/figs/intruder.pdf}
+	}
+	\caption{Results of OLS fitting on sampled data and selected intruder states.}
 \end{figure}
-
-All three models have a first excited state around 1eV, but only in the minimal model does this state fall near the sample set used for the fitting, and is the only model in which this state corresponds to the correct excitation O 2$p_z \rightarrow $ O 2p$_\pi$ as seen in experiment.
-In contrast, the first excited state for the other two models are visibly distant from our sample set and do not correspond to the experimental first excited state.
-The large distance of these eigenstates from our sample set indicates that their predicted eigenvalues may suffer from large extrapolation errors, and as such we will call them \textit{intruder} states.
-In Figure \ref{fig:Intruder} we present all eigenstates among the three potential models which we classify as intruders using a k-nearest neighbor approach with k = 5.
 
@@
 This is evidenced by the fact that two of the intruder states correspond explicitly to excitations we could not access: Cu 3d$_{z^2} \rightarrow $ O 2p$_\pi$ and  Cu 3d$_\pi \rightarrow $ O 2p$_\pi$.
-In fact, these are the only two singly excited states which we could not access with our sampling scheme.
-While inaccessible, we can approximate these states by single rigid MO excitations above the UKS ground state, which results in states with a FN-DMC energy above 4eV.
+We can approximate these states by single rigid MO excitations above the UKS ground state, which results in states with a FN-DMC energy above 4eV.
 Including a generous 2eV reduction in FN-DMC energy due to orbital relaxation, we estimate that these two states should lie at least 2eV in energy. 
-Since all the other inaccessible eigenstates are double or higher order excitations, our very approximate exploration of the inaccessible parts of $\mathcal{LE}$ leaves us with a reasonable prior: intruder states should lie above 2eV in energy in our effective theories.
-We can enforce this prior by adding a term to our ordinary linear regression cost function
+Since all the other inaccessible eigenstates are double or higher order excitations, we assert a prior: intruder states should lie above 2eV in energy in our effective theories, enforced via an augmented cost function
 \begin{equation}
@@
 where $\lambda>0$ is a parameter which can be varied, QHL is a quadratic hinge loss, $\Theta$ is the Heaviside step function, the index $i$ is over our sampled states and $p$ over the selected intruder states.
-Increasing the value of $\lambda$ leads to a stronger penalty for models which do not satisfy our prior.
 
 \begin{figure}[H]
-\begin{center}
-\includegraphics[width=0.7\linewidth]{../cuo/qwalk/old/ub3lyp_s1_/analysis/figs/intruder.pdf}
-\end{center}
-\caption{Collection of unique intruder states from our three potential models selected using a k-nearest neighbor approach with k = 5.}
-\label{fig:Intruder}
+	\centering
+	\subfigure[On the left, scores of the three candidate models at various $\lambda$ when fitting our effective theories using the cost function \eqref{eq:cost}.
+On the right we show predicted versus calculated energies for our low-energy sampled states for the minimal model fit at $\lambda = 20$.]{
+	 	\label{fig:Prior}
+		\includegraphics[width=0.45\textwidth]{../cuo/qwalk/old/ub3lyp_s1_/analysis/figs/prior_and_regr.pdf}
+	}
+	\qquad
+	\subfigure[Results of exact diagonalization showing the energies and properties of the first thirty eigenstates for three candidate models using linear regression with cost function \ref{eq:cost} and $\lambda = 20$.
+The ground state and first excited state are colored and have enlarged markers for clarity.
+Error bars are 95\% CI calculated by a bootstrap estimate.]{
+		\label{fig:FinalED}
+		\includegraphics[width=0.45\textwidth]{../cuo/qwalk/old/ub3lyp_s1_/analysis/figs/final_ed.pdf}
+	}
+	\caption{Results of linear regression using augmented cost function \eqref{eq:cost}.}
 \end{figure}
 
-Using our new cost function we find a set of models which accurately describe our sample data, have very similar spectra and eigenvectors, and do not contain intruder states below the 2eV prior. 
+Using our new cost function we find a set of models which accurately describe our sample data, have very similar spectra and eigenvectors to eachother and experiment, and do not contain intruder states below the 2eV prior.
 Shown in Figure \ref{fig:Prior} are the R$^2$ scores and QHL for the three candidate models fit at $\lambda \in [0,20]$. 
-The R$^2$ score stays above 0.95 for all models and $\lambda$, and the QHL converges quickly to nearly zero at $\lambda = 20$ for all three models. 
-We conclude that all three models at $\lambda = 20$ should be considered good models as they satisfy our prior and also explain our sample set accurately. 
+We believe that all three models at $\lambda = 20$ should be considered good models as they satisfy our prior and also explain our sample set accurately. 
 Also shown are $E_\text{eff}, E_\text{ab}$ for our samples using the minimal model at $\lambda = 20$ to illustrate the high fit quality even with a large $\lambda$.
 Figure \ref{fig:FinalED} presents the properties and eigenvalues of the lowest 30 eigenstates for the three fit models at $\lambda = 20$. 
-In comparison one can view Figure \ref{fig:InitED} as the same figure for fits at $\lambda = 0$. 
-At $\lambda = 20$ there are no outliers below 2eV in any of the three fit models. 
 The additional constraint from asserting a prior has seemingly caused a single effective theory to emerge from our set of candidate models.
 This conclusion is verified by looking at the values of the fit parameters for our three models, presented in Table 1.
-The parameter values for the three different models fit at $\lambda = 20$ are nearly identical within error bars and indicate that asserting our prior ensures that all candidate models regress to a single low-energy effective theory.
 All occupation energies are relative to $\epsilon_{d_\delta}$.
-
-Our final regressed model accurately reproduces properties and energies of the lowest lying eigenstates as seen in modern APES experiments on the CuO molecule.
-From Figure \ref{fig:FinalED} we see that the ground state of our effective theory has fully filled Cu 3d shell, nearly two electrons in O 2p$_z$ orbital, three in the O 2p$_\pi$ and around a quarter of an electron in the Cu 4s orbital.
-An approximate electronic configuration for the ground state is therefore
-be assigned 1$\sigma^2$1$\pi^4$1$\delta^4$2$\sigma^2$2$\pi^3$3$\sigma^0$ where the orbitals associated with the assignment are Cu 3d$_{z^2}$, Cu 3d$_\pi$, Cu 3d$_\delta$, O 2p$_z$, O 2p$_\pi$, and Cu 4s.
-This state also has a total spin projection S$_z = \frac{1}{2}$.
-This matches the well studied ground state electron configuration for the molecule $^2\Pi$.
-There are two excited states around 1eV in energy, the first of which is a state with the following single excitation O 2p$_z \rightarrow$ O 2p$_\pi$, has the configuration 1$\sigma^2$1$\pi^4$1$\delta^4$2$\sigma^2$2$\pi^4$4$\sigma^0$, and S$_z = \frac{1}{2}$.
-According to experiment there is also such a state at 1eV and is usually denoted as $^2X$.
-A second excited state at 1.2 eV exists which corresponds to an excitation 
-O 2p$_\pi \rightarrow$ Cu 4s, has configuration 1$\sigma^2$1$\pi^4$1$\delta^4$2$\sigma^2$2$\pi^2$4$\sigma^1$, and can either be S$_z = \frac{1}{2}, \frac{3}{2}$. 
-This is also the lowest energy eigenstate which can be a spin quartet.
- agree that this excitation should be the lowest spin quartet excitation, but argue that this state should be at an energy of 1.9 eV. 
-Our model has a different spin quartet state with a full Cu d shell at 1.9 eV, corresponding to an excitation O 2p$_z \rightarrow$ Cu 4s, with configuration 1$\sigma^2$1$\pi^4$1$\delta^4$2$\sigma^1$2$\pi^4$4$\sigma^1$.
-There are another host of eigenstates for our models at 2 eV corresponding to singles excitations of the form Cu 3d $\rightarrow$ O 2p$_\pi$. 
-This matches measurements which indicate that the lowest energy eigenstates with a single hole in the Cu 3d shell lie at 2.2 eV above the ground state.
-
-\begin{figure}[H]
-\begin{center}
-\includegraphics[width=0.7\linewidth]{../cuo/qwalk/old/ub3lyp_s1_/analysis/figs/prior_and_regr.pdf}
-\end{center}
-\caption{On the left, scores of the three candidate models at various $\lambda$ when fitting our effective theories using the cost function \eqref{eq:cost}.
-On the right we show predicted versus calculated energies for our low-energy sampled states for the minimal model fit at $\lambda = 20$.}
-\label{fig:Prior}
-\end{figure}
+In the doublet sector the ground and first excited state of any of the three models match experiment, as well as the block of states at 2eV corresponding to excitations out of the Cu 3d shell.
+The lowest energy quartet states are at 1, 2eV respectively and are in line with the observation of a quartet state at 2eV in APES measurements.
 
@@
 \begin{center}
+\begin{tabular}{l|lllll}
+Model &$\epsilon_{d_{z^2}}$ & $\epsilon_{d_\pi}$ & $\epsilon_{4s}$ & $\epsilon_{p_\pi}$ & $\epsilon_{p_z}$ \\ \hline 
+Min & 0.32(2)& 0.19(1)& 2.2(4)& 1.7(3)& 0.96(4)\\
+Min $ +\ \bar{c}_{d_\pi}^\dagger \bar{c}_{p_\pi}$& 0.32(1)& 0.10(2)& 2.23(3)& 1.8(2)& 0.97(3)\\
+Min $ +\ \bar{c}_{d_{z^2}}^\dagger \bar{c}_{p_z}$& 0.29(2)& 0.19(1)& 2.19(5)& 1.7(3)& 0.99(5)\\
+\end{tabular} \\
+
 \begin{tabular}{l|llllll}
-&Min & Min & Min \\
-& & $ + \bar{c}_{d_\pi}^\dagger \bar{c}_{p_\pi}$ & $ + \bar{c}_{d_{z^2}}^\dagger \bar{c}_{p_z}$ \\ $\ $ \\ \hline 
-$\epsilon_{d_{z^2}}$& 0.32(2)& 0.32(1)& 0.29(2)\\
-$\epsilon_{d_\pi}$& 0.19(1)& 0.10(2)& 0.19(1)\\
-$\epsilon_{d_\delta}$& 0.0 & 0.0 & 0.0\\
-$\epsilon_{4s}$& 2.2(4)& 2.23(3)& 2.19(5)\\
-$\epsilon_{p_\pi}$& 1.7(3)& 1.8(2)& 1.7(3)\\
-$\epsilon_{p_z}$& 0.96(4)& 0.97(3)& 0.99(5)\\
-$t_\pi$& -0.57(1)& -0.45(2)& -0.57(1)\\
-$t_{dz}$& 0.55(3)& 0.56(2)& 0.54(2)\\
-$t_{sz}$& 0.87(2)& 0.89(2)& 0.85(2)\\
-$t_{ds}$& 0.44(1)& 0.45(1)& 0.45(2)\\
-$J_{sd}$& -0.6(1)& -0.8(1)& -0.6(1)\\
-$U_s$& 3.8(2)& 3.7(2)& 3.9(2))\\
-\end{tabular}
-\\ $\ $
-\\
+Model &$t_\pi$ & $t_{dz}$ & $t_{sz}$ & $t_{ds}$ & $J_{sd}$ & $U_s$ \\ \hline 
+Min &  -0.57(1)& 0.55(3)& 0.87(2)& 0.44(1)& -0.63(9)& 3.8(2)\\
+Min $ +\ \bar{c}_{d_\pi}^\dagger \bar{c}_{p_\pi}$& -0.45(2)& 0.56(2)& 0.89(2)& 0.45(1)& -0.78(7)& 3.7(2)\\
+Min $ +\ \bar{c}_{d_{z^2}}^\dagger \bar{c}_{p_z}$& -0.57(1)& 0.54(2)& 0.85(2)& 0.45(2)& -0.58(3)& 3.9(2)\\
+\end{tabular} \\
+
 \end{center}
-Table 1: Parameters in eV for our three potential models when fit using \eqref{eq:cost} at $\lambda = 20$. Error bars are 95\% CI calculated using a bootstrap estimate.
+Table 1: Parameters in eV for our three potential models when fit using \eqref{eq:cost} at $\lambda = 20$, energies relative to $\epsilon_{d_\delta} = 0$.
+Error bars are 95\% CI calculated using a bootstrap estimate.
 \end{table}
-
-\begin{figure}[H]
-\centering
-\includegraphics[width=0.7\linewidth]{../cuo/qwalk/old/ub3lyp_s1_/analysis/figs/final_ed.pdf}
-\caption{Results of exact diagonalization showing the energies and properties of the first thirty eigenstates for three candidate models using linear regression with cost function \ref{eq:cost} and $\lambda = 20$.
-The ground state and first excited state are colored and have enlarged markers for clarity.
-Error bars are 95\% CI calculated by a bootstrap estimate.}
-\label{fig:FinalED}
-\end{figure}
 
@@
 Similar to the case of the \textit{ab-initio} downfolding, we will need to construct low-energy spaces, sample the constructed spaces, build effective theories and respective bases, and fit our models.
-Unlike \textit{ab-initio} downfolding, the Hubbard model downfolding has significantly fewer degrees of freedom, and therefore the DMD procedure can be done in a much more controlled fashion.
 We will begin from a starting point of highest control, namely for system sizes small enough where we can exactly diagonalize the Hubbard model.
-In the situation where we have full access to the eigenstates and spectra of a system, selecting $\mathcal{LE}$ can be done exactly as the span of some chosen $N$ lowest energy eigenstates. 
-We can further guarantee that our sample set fully encapsulates all low-energy degrees of freedom by simply sampling states parameterized as $|\Psi\rangle = \sum_{j=1}^N c_j |\Phi_j\rangle$ where $|\Phi_j\rangle$ are eigenstates of the Hubbard Hamiltonian.
-This can be seen in contrast to the CuO fitting where $\mathcal{LE}$ was defined implicitly through a number occupation descriptor and the sampling of $\mathcal{LE}$ was incomplete leading to intruder states in our initial effective theories.
-Basis construction for the Hubbard Hamiltonian is also explicitly clear as any possible single particle basis we can use is just an orthogonal transformation of the basis the Hubbard model is defined on, and is guaranteed to be a complete single particle basis.
-Again, this can be compared to the case of an \textit{ab-initio} system like the CuO molecule where bases are only approximately complete due to the huge number of degrees of freedom in the system.
-In order to include temperature into the mix we can work with the full model Hilbert space such that $\mathcal{LE} = \mathcal{H}$, but generate samples within our space using a Monte Carlo sampling method according to a Boltzmann probability distribution with variable parameter $\beta = 1/k_BT$.
-This way, our sample space will still be able to probe every excitation within a certain temperature range, the samples will be biased towards lower energy states according to the Boltzmann weight, and hence the fit model will also be biased towards describing more accurately the energy functional on the states with highest weight.
-The Monte Carlo sampling may need to be altered further in order to deal with entropic effects when the density of states is very low in an energy range of interest and relatively high elsewhere.
+In this case selecting $\mathcal{LE}$ can be done exactly as the span of some chosen $N$ lowest energy eigenstates. 
+We can guarantee that our sample set fully encapsulates all low-energy degrees of freedom by sampling states parameterized as $|\Psi\rangle = \sum_{j=1}^N c_j |\Phi_j\rangle$ where $|\Phi_j\rangle$ are eigenstates of the Hubbard Hamiltonian.
+Basis construction for the Hubbard Hamiltonian is also explicitly clear as any possible single particle basis we can use is just an orthogonal transformation of the basis the Hubbard model is defined on.
+In order to include temperature into the mix we can work with the full model Hilbert space such that $\mathcal{LE} = \mathcal{H}$ but generate samples within our space using a Monte Carlo sampling method according to a Boltzmann probability distribution with variable parameter $\beta = 1/k_BT$, biasing our final model towards describing more accurately the energy functional on the states with highest weight.
 
@@
 
+\begin{figure}[H]
+\centering
+\includegraphics[width=0.6\linewidth]{../hubbard/eff_eig_prelim.pdf}
+\caption{Results of weighted single parameter model fits to the eigenstates of a Hubbard model for various U and temperature. Darker hues indicate a higher R$^2$ score.}
+\label{fig:Hubbard}
+\end{figure}
+
 We conclude by showing some very preliminary results on model fitting for the half-filled, 2x2 square Hubbard model for U: 0 - 12 which put our approach to model fitting with exact diagonalization at T$\geq 0$ into practice.
@@
 The basis for all of these operators are just the bare Hubbard Hamiltonian basis.
-We have restricted ourselves to eigenstates only as we have not developed a sufficient low-energy sampling scheme yet which deals with the entropy issue discussed earlier.
 The weighted linear regression approximates the effect of generating samples using a Markov Chain Monte Carlo method with a Boltzmann distribution.
 For small U the t-only model is an accurate effective theory as the temperature increases, which is to be expected in this perturbative limit.
-If one were to look at large U and high temperature a U-only model is quite good since the dominant and relevant energy scale at higher temperatures is related to double occupation.
+At large U and high temperature a U-only model is quite good since the dominant and relevant energy scale at higher temperatures is related to double occupation.
 At large U and low temperature we see that the expected Heisenberg model accurately describes the low-energy excitations. 
 In this region a t-only model also works quite accurately since for large U the hopping descriptor $\sum_{\langle i, j \rangle, \sigma} c_{i\sigma}^\dagger c_{j\sigma} + h.c.$ is correlated to the exchange operator $J \sum_{\langle i,j \rangle} \vec{S}_i \cdot \vec{S}_j$ for the lowest energy excitations in the system.
-It should be noted, however, that this t-only model would have to be explicitly written to only act on states with no double occupancy, an assumption inherent in the Heisenberg model. 
+It should be noted that this t-only model would have to be explicitly written to only act on states with no double occupancy, an assumption inherent in the Heisenberg model. 
 Finally, in the region beginning at 4 $<$ U $<$ 8 and moderate $\beta$: 0.1 - 0.4 we see that none of the candidate single parameter models can accurately describe the low-energy excitations of the system.
@@
 
-\begin{figure}[H]
-\centering
-\includegraphics[width=0.7\linewidth]{../hubbard/eff_eig_prelim.pdf}
-\caption{Results of weighted single parameter model fits to the eigenstates of a Hubbard model for various U and temperature. Darker hues indicate a higher R$^2$ score.}
-\label{fig:Hubbard}
-\end{figure}
-
-\subsection{Proposed timeline: 2 years}
-\begin{enumerate}
-\item Half-filled model ED
-	\begin{enumerate}
-	\item Constructing a sampling scheme for $\mathcal{LE}$: 2 months
-	\item DMD for 2x2, 2$\sqrt{2}$x2$\sqrt{2}$, 3x3, 4x4
-	system sizes: 4 months
-	\end{enumerate}
-\item Half-filled model AFQMC
-	\begin{enumerate}
-	\item Learning and benchmarking calculations: 1 month
-	\item Extending sampling scheme for $\mathcal{LE}$: 1 month
-	\item DMD calculations up to 16x16 cells: 4 months
-	\item Finite size scaling of results: 2 months
-	\end{enumerate}
-\item Doped model ED
-	\begin{enumerate}
-	\item DMD for  2x2, 2$\sqrt{2}$x2$\sqrt{2}$, 3x3, 4x4
-	system sizes, various dopings: 4 months
-	\end{enumerate}
-\item Doped model AFQMC
-	\begin{enumerate}
-	\item Constructing a sampling scheme - trial functions for constrained
-	path: 2 months
-	\item DMD calculations up to 16x16 cells: 4 months
-	\end{enumerate}
-\end{enumerate}
-
+\subsection{Proposed timeline}
+We believe that developing effective theories for the half-filled Hubbard model should take 1 year. 
+Six months should be spent on the exact diagonalization calculations. 
+The first two months would be spent developing an efficient sampling scheme for $\mathcal{LE}$.
+The following four months will be geared towards DMD for increasing system sizes from 2x2 to 4x4 lattices and will be heavily exploratory.
+The next six months will be focused on extending our exploration to the thermodynamic system.
+The first two months will be dedicated to learning the AFQMC method and extending our ED sampling scheme to AFQMC.
+The following four months will be focused on using AFQMC DMD to construct effective theories up to 16x16 system size and extrapolating our results for the half-filled model to the thermodynamic limit.
+This timeline can then be copied as we extend our analysis to the doped model. 
+The only change being the two months previously dedicated to learning AFQMC will be focused instead on developing efficient trial wave functions for constrained path calculations.
+In total our goal is to finish the project within two years.
 \end{document}

### 2019-06-14 draft.tex
finished title page

@@
 \renewcommand{\baselinestretch}{1.5}
+\author{Shivesh Pathak, Lucas Wagner}
+\title{Developing effective theories for the square lattice Hubbard model using density matrix downfolding}
 
 \begin{document}
+\maketitle
+\begin{abstract}
+A core problem of condensed matter physics is understanding how to systematically develop low-energy effective models for strongly correlated electron systems. 
+We believe that a density matrix downfolding (DMD) method can be used to systematically build effective theories for such systems beginning with a high energy theory like the \textit{ab-initio} Hamiltonian H$_\text{ab}$.
+We present as a preliminary result an effective theory developed using DMD which can accurately reproduce the spectra and properties of the lowest lying excitations of the CuO molecule.
+As a step towards bulk strongly correlated materials, we propose developing effective theories using DMD for the simplest toy system which exhibits strongly correlated behavior: the 1-band square lattice Hubbard model for various couplings U, temperatures T and dopings $\delta$.
+\end{abstract}
+\pagebreak
+
 \section{Introduction}
@@
 
-When downfolding $\hat{H}_\text{ab}$ we will use FN-DMC to access states in $\mathcal{LE}$ and accurately calculate their \textit{ab-initio} energies and RDM elements.
+When downfolding $\hat{H}_\text{ab}$ we will use fixed-node diffusion Monte Carlo (FN-DMC) to access states in $\mathcal{LE}$ and accurately calculate their \textit{ab-initio} energies and RDM elements.
 Diffusion Monte Carlo (DMC) is a quantum Monte Carlo method which projects out the ground state of a Hamiltonian given some initial trial wave function.

### 2019-06-24 draft.tex
moved figures around, compilable now

@@
 	 	\label{fig:InitED}
-		\includegraphics[width=0.45\textwidth]{../cuo/qwalk/old/ub3lyp_s1_/analysis/figs/init_ed.pdf}
+		\includegraphics[width=0.45\textwidth]{figs/init_ed.pdf}
 	}
@@
 		\label{fig:Intruder}
-		\includegraphics[width=0.45\textwidth]{../cuo/qwalk/old/ub3lyp_s1_/analysis/figs/intruder.pdf}
+		\includegraphics[width=0.45\textwidth]{figs/intruder.pdf}
 	}
@@
 	 	\label{fig:Prior}
-		\includegraphics[width=0.45\textwidth]{../cuo/qwalk/old/ub3lyp_s1_/analysis/figs/prior_and_regr.pdf}
+		\includegraphics[width=0.45\textwidth]{figs/prior_and_regr.pdf}
 	}
@@
 		\label{fig:FinalED}
-		\includegraphics[width=0.45\textwidth]{../cuo/qwalk/old/ub3lyp_s1_/analysis/figs/final_ed.pdf}
+		\includegraphics[width=0.45\textwidth]{figs/final_ed.pdf}
 	}
@@
 \centering
-\includegraphics[width=0.6\linewidth]{../hubbard/eff_eig_prelim.pdf}
+\includegraphics[width=0.6\linewidth]{figs/eff_eig_prelim.pdf}
 \caption{Results of weighted single parameter model fits to the eigenstates of a Hubbard model for various U and temperature. Darker hues indicate a higher R$^2$ score.}

### 2019-06-24 draft.tex
cleaned up proposed work and a bit of intro. talks about what new thing we can do, what i expect to see, and what success looks like to me

@@
 We present as a preliminary result an effective theory developed using DMD which can accurately reproduce the spectra and properties of the lowest lying excitations of the CuO molecule.
-As a step towards bulk strongly correlated materials, we propose developing effective theories using DMD for the simplest toy system which exhibits strongly correlated behavior: the 1-band square lattice Hubbard model for various couplings U, temperatures T and dopings $\delta$.
+As a step towards bulk strongly correlated materials, we propose developing effective theories using DMD for the simplest toy system which exhibits strongly correlated behavior: the 1-band square lattice Hubbard model at half-filling for various couplings U and temperatures T.
 \end{abstract}
@@
 
-We propose using a density matrix downfolding (DMD) procedure to systematically develop low-energy effective theories of the 1-band nearest-neighbor Hubbard model on a square lattice as a stepping stone towards building effective theories for strongly correlated materials.
+We propose using a density matrix downfolding (DMD) procedure to systematically develop low-energy effective theories of the 1-band nearest-neighbor Hubbard model on a square lattice for various couplings U and temperatures T at half-filling as a stepping stone towards building effective theories for strongly correlated materials.
 The Hubbard model is a simple model with nearest-neighbor hopping and a local on-site Coulomb repulsion, parameterized by a single ratio U/t: 
@@
 \end{equation}
-The solutions in 1-d are well understood via the Bethe ansatz , however the 2-d square lattice Hubbard model has no known exact solution.  
-In the very strong coupling limit $U/t \rightarrow \infty$ an emergent low-energy spin-space exists with an effective nearest-neighbor AFM Heisenberg Hamiltonian.
-In the region $U = 0$ a trivial non-interacting theory holds.
-Between these two limits, in particular the intermediate coupling region $2 < U < 8$, no well established effective theory exists. 
-In this region there is a metal-insulator transition at $U_c/t \sim 4$, and the pseudogap opens up at finite temperature around $U/t \sim 7$, leading to the development of a Mott gap for large $U/t > 11$.
-In addition to the half-filling phase diagram, this model exhibits a plethora of interesting ground state properties when doped including stripe phases and d-wave pairing. 
-The relative simplicity of the Hubbard model and the lack of a well-established low-energy theory in a broad intermediate coupling region lends the model as a useful playground for our new downfolding procedure in preparation for tackling realistic strongly correlated electron systems.
+The solutions in 1-d are well understood via the Bethe ansatz, however, the 2-d square lattice Hubbard model has no known exact solution.  
+A few well accepted facts do exist for this model, such as the existence of long-range anti-ferromagnetic (AFM) order for any U$>$0 at T = 0 and a Mott-Heisenberg limit as $U/t \rightarrow \infty$ at low temperatures.
+The T = 0 diagram of the model at half-filling has been a long-standing source of controversy between two groups:  those who believe that for any U $>$ 0 the model exhibits Mott-Heisenberg behavior, and those who believe that there is a transition at finite U$^*$ between Slater and Mott-Heisenberg behavior.
+While recent numerical and theoretical calculations tend to agree with the latter position, non-negligible evidence for the prior does exist. 
+Further, modern finite temperature calculations indicate thermal transitions in the weak to moderate coupling region between the Slater phase at T = 0, pseudogap phase below T$_{pg}$ and a Fermi liquid phase above.
+A finite temperature psuedogap to Mott insulator transition has also been found between the weak to strong coupling limits.
+The nature of these transitions as phase transitions or simply crossovers is poorly understood.
+The relative simplicity of the half-filled Hubbard model, potential complexity of its phase diagram in U-T space,  and the lack of well-established low-energy theories for moderate U/t lends the model as a useful playground for our new downfolding procedure in preparation for tackling realistic strongly correlated electron systems.
 
@@
 \section{Proposed work}
-Our goal is to develop low-energy effective theories for the
-Hubbard model over the space U-T-$\delta$ using the DMD method, particularly focusing on the medium coupling regions and moderate dopings where perturbative effective theories break down.
-Similar to the case of the \textit{ab-initio} downfolding, we will need to construct low-energy spaces, sample the constructed spaces, build effective theories and respective bases, and fit our models.
-We will begin from a starting point of highest control, namely for system sizes small enough where we can exactly diagonalize the Hubbard model.
+My goal is to develop low-energy effective theories for the square lattice 1-band Hubbard model in order to map out the phases and transitions in the U-T space.
+Recent attempts at solving this problem involve constructing single particle properties like densities of states using approximate numerical or effective theory techniques to classify different phases.
+While instructive, reliance on single particle properties to identify phases instead of studying changes in the many-body Hilbert space and low-energy effective theory can lead to ambiguities regarding the nature of the phases.
+I believe the DMD method can remove these ambiguities as the method explicitly involves studying changes in the low-energy Hilbert space and provides us with accurate many-body effective theories which can precisely differentiate different phases.
+
+My first step would be to show whether the model supports more than one phase at T/t = 0 with respect to U, and if so what the features of these phases and the transition between them are.
+I will investigate in particular the hypothesis that for $U/t<<1$ the model exhibits gapped Slater AFM behavior and for $U/t>>1$ Mott-Heisenberg behavior and whether the transition between them is a genuine phase transition or simply a crossover due to rearrangement of a few states.
+To do so I will initially develop effective theories using DMD at two points U/t = 2, 12 and T/t = 0, allowing me to argue the existence or non-existence of multiple phases and investigate the properties of each phase.
+I will follow by developing theories for various intermediate U in order to locate and study the transition between the two phases, if it exists, focusing on the hypothesis of a crossover U$^*/t\sim 4 - 5$ between the Slater and Mott-Heisenberg regimes.
+Repeating this exercise for finite temperature I will build four more models at U/t = 2, 12 and T/t = 0.05, 0.20 to investigate the finite temperature phases at strong- and weak-coupling, in particular the claim of pseudogap and Fermi liquid phases at T/t = 0.05, 0.20 at weak coupling.
+Filling in the U-T diagram will allow me to locate and study thermal transitions in the system. 
+Building the T$>$0 effective theories will require special attention as they must satisfy the Mermin-Wagner theorem.
+
+Similar to the case of the \textit{ab-initio} downfolding, I will need to construct low-energy spaces, sample the constructed spaces, build effective theories and respective bases, and fit candidate models.
+I will begin from a starting point of highest control, namely for system sizes small enough where we can exactly diagonalize the Hubbard model.
 In this case selecting $\mathcal{LE}$ can be done exactly as the span of some chosen $N$ lowest energy eigenstates. 
-We can guarantee that our sample set fully encapsulates all low-energy degrees of freedom by sampling states parameterized as $|\Psi\rangle = \sum_{j=1}^N c_j |\Phi_j\rangle$ where $|\Phi_j\rangle$ are eigenstates of the Hubbard Hamiltonian.
+One can guarantee that the sample set fully encapsulates all low-energy degrees of freedom by sampling states parameterized as $|\Psi\rangle = \sum_{j=1}^N c_j |\Phi_j\rangle$ where $|\Phi_j\rangle$ are eigenstates of the Hubbard Hamiltonian.
 Basis construction for the Hubbard Hamiltonian is also explicitly clear as any possible single particle basis we can use is just an orthogonal transformation of the basis the Hubbard model is defined on.
-In order to include temperature into the mix we can work with the full model Hilbert space such that $\mathcal{LE} = \mathcal{H}$ but generate samples within our space using a Monte Carlo sampling method according to a Boltzmann probability distribution with variable parameter $\beta = 1/k_BT$, biasing our final model towards describing more accurately the energy functional on the states with highest weight.
-
-In order to make any statements about effective theories for the Hubbard model we will have to work with larger system sizes and approximate methods like auxiliary field quantum Monte Carlo (AFQMC) to generate low-energy states for the DMD procedure.
-The auxiliary field quantum Monte Carlo (AFQMC) method has been used extensively in accurate calculations on both \textit{ab-initio} and lattice systems.
-AFQMC begins with a trial wave function which is projected using a Trotter expansion of the operator $e^{-\tau H}$ using a stochastic method.
-The final projected state from AFQMC can be used to calculate expectation values of operators using a similar mixed-estimator method as FN-DMC.
-The AFQMC method is especially useful for the Hubbard model as the auxiliary fields are real with potential values $\pm$ 1, only N such fields need to be considered, and at least at half-filling there is no sign problem and the method is numerically exact.
-Away from half-filling the effects of the fermion sign problem are especially potent and the nodes of the projected wave function are fixed to those of the trial wave function making the method variational in the same way FN-DMC is.
-In the weak and strong coupling regimes typical choices for trial wave functions are non-interacting and Gutzwiller trial functions respectively, but other post-Gutzwiller and generalized Hartree Fock (GHF) trial functions have been shown to give accurate results in the moderate coupling regimes.
+In order to include temperature into the mix one can work with the full model Hilbert space such that $\mathcal{LE} = \mathcal{H}$ but generate samples within the space using a Monte Carlo sampling method according to a Boltzmann probability distribution with variable parameter $\beta = 1/k_BT$, biasing the final model towards describing more accurately the energy functional on the states with highest weight.
+Finite size extrapolation to the thermodynamic limit requires working with lattice sizes for which exact diagonalization is no longer feasible, and approximate methods like lattice variational Monte Carlo (VMC) will be employed to sample low-energy states as well as calculate energies and RDM elements.
+In the weak and strong coupling regimes typical choices for varitional wave functions are non-interacting and Gutzwiller trial functions respectively, but other post-Gutzwiller and generalized Hartree Fock (GHF) trial functions have been shown to give accurate results in the moderate coupling regimes.
 A rigorous prescription for finite size extrapolation specifically for the 1-band square lattice Hubbard model will be followed to extend our conclusions to the thermodynamic limit.
@@
 
-We conclude by showing some very preliminary results on model fitting for the half-filled, 2x2 square Hubbard model for U: 0 - 12 which put our approach to model fitting with exact diagonalization at T$\geq 0$ into practice.
-We start off by exactly diagonalizing each model, define $\mathcal{LE} = \mathcal{H}$, and treat as our full sample space as just the eigenstates of the model.
-We then fit three single parameter models to the eigenstates: A) 
+I will conclude by showing some very preliminary results on model fitting for the half-filled, 2x2 square Hubbard model for U: 0 - 12 which put the outlined approach to model fitting with exact diagonalization at T$\geq 0$ into practice.
+I started off by exactly diagonalizing each model, defining $\mathcal{LE} = \mathcal{H}$, and treating the full sample space as just the eigenstates of the model.
+I then fit three single parameter models to the eigenstates: A) 
 A "t" only model $t \sum_{\langle i, j \rangle, \sigma} c_{i\sigma}^\dagger c_{j\sigma} + h.c.$, B) A "U" only model $U \sum_{i} n_{i\uparrow} n_{i\downarrow}$, and C) A "J" only model $J \sum_{\langle i,j \rangle} \vec{S}_i \cdot \vec{S}_j$ using a weighted linear regression with weights $e^{-\beta E}$.
@@
 The weighted linear regression approximates the effect of generating samples using a Markov Chain Monte Carlo method with a Boltzmann distribution.
-For small U the t-only model is an accurate effective theory as the temperature increases, which is to be expected in this perturbative limit.
-At large U and high temperature a U-only model is quite good since the dominant and relevant energy scale at higher temperatures is related to double occupation.
-At large U and low temperature we see that the expected Heisenberg model accurately describes the low-energy excitations. 
-In this region a t-only model also works quite accurately since for large U the hopping descriptor $\sum_{\langle i, j \rangle, \sigma} c_{i\sigma}^\dagger c_{j\sigma} + h.c.$ is correlated to the exchange operator $J \sum_{\langle i,j \rangle} \vec{S}_i \cdot \vec{S}_j$ for the lowest energy excitations in the system.
-It should be noted that this t-only model would have to be explicitly written to only act on states with no double occupancy, an assumption inherent in the Heisenberg model. 
-Finally, in the region beginning at 4 $<$ U $<$ 8 and moderate $\beta$: 0.1 - 0.4 we see that none of the candidate single parameter models can accurately describe the low-energy excitations of the system.
-This poorly explained region indicates the breakdown of perturbative expansions of the Hubbard model about the strong- or weak-coupling limits.
-In order to make more conclusive statements we would have to actually sample the Hilbert space to construct a more robust estimate of the best models for different $U, \beta$.
-We would be particularly interested in whether an effective theory for this region exists and how this region and corresponding effective theory change as we increase the system size.
+At low-temperatures there is a crossover between the weak coupling regime, best described by a non-interacting t-only model, and the strong coupling regime, best described by a J-only model, around U$^*/t \sim 5$ providing some evidence of a Slater to Mott-Heisenberg transition.
+In the crossover region T = 0 and U = 4 - 8 both models perform poorly and is a region of particular interest when developing our low-energy theories.
+At high temperatures there is a crossover between a non-interacting t-only and U-only model which can also be investigated.
 
 \subsection{Proposed timeline}
-We believe that developing effective theories for the half-filled Hubbard model should take 1 year. 
-Six months should be spent on the exact diagonalization calculations. 
-The first two months would be spent developing an efficient sampling scheme for $\mathcal{LE}$.
-The following four months will be geared towards DMD for increasing system sizes from 2x2 to 4x4 lattices and will be heavily exploratory.
-The next six months will be focused on extending our exploration to the thermodynamic system.
-The first two months will be dedicated to learning the AFQMC method and extending our ED sampling scheme to AFQMC.
-The following four months will be focused on using AFQMC DMD to construct effective theories up to 16x16 system size and extrapolating our results for the half-filled model to the thermodynamic limit.
-This timeline can then be copied as we extend our analysis to the doped model. 
-The only change being the two months previously dedicated to learning AFQMC will be focused instead on developing efficient trial wave functions for constrained path calculations.
-In total our goal is to finish the project within two years.
+I believe that developing effective theories for the half-filled Hubbard model should take 2 years. 
+16 months should be spent on T = 0 calculations.
+The first 8 months will be focused on exact diagonalization.
+In this period, the first two months would be spent developing an efficient sampling scheme for $\mathcal{LE}$.
+The following three months will be geared towards DMD for increasing system sizes from 2x2 to 4x4 lattices for U = 2, 12 in order to identify the existence of multiple zero temperature phases and will be primarily exploratory.
+The next three months will be focused on studying intermediate U and the transition between the two phases, if they exist.
+The following 8 months will be focused on finite size extrapolation.
+The first two months here will be dedicated to extending our ED sampling scheme to VMC.
+The following six months will involve extending our finite size results to the thermodynamic limit.
+At the end of the first 16 months I should have a phase diagram at T = 0 with effective theories at various U.
+The last 8 month chunk will be dedicated to the T $>$ 0 model fitting.
+The first four months will be dedicated to exact diagonalization calculations and should be a relatively simple extension of the T = 0 model fitting with samples drawn according to an additional importance criteria $e^{-\beta E}$.
+The last four months will be focused on extending our T = 0 VMC scheme to T $>$ 0 and conducting finite size extrapolation.
+At the end of the 2 years I should have a phase diagram in the U - T space with effective theories at various values of U, T and an understanding of the phases and transitions between them.
 \end{document}

### 2019-06-25 draft.tex
updated draft based on Lucas comments. need to updated CuO bit later

@@
 \renewcommand{\baselinestretch}{1.5}
-\author{Shivesh Pathak, Lucas Wagner}
+\usepackage[utf8]{inputenc}
+\author{Shivesh Pathak}
 \title{Developing effective theories for the square lattice Hubbard model using density matrix downfolding}
@@
 \end{abstract}
+Committee: Lucas Wagner, ...
 \pagebreak
@@
 The fundamental barrier in writing down low-energy theories for strong correlated materials is the existence of non-perturbative interactions between various degrees of freedom - for example, lattice, spin and charge.
-Consider just the cuprate case for now, the two phases which are well understood contain a singular important degree of freedom in the low-energy space. 
-Near zero doping the cuprates are charge-transfer insulators with localized spin-1/2 low-energy degrees of freedom.
-In the heavily overdoped side, the lowest energy excitations look like well defined Fermi-liquid quasiparticles.
+Consider just the cuprate case for now, the two phases which are relatively well understood contain a singular important degree of freedom in the low-energy space. 
+Near zero doping the cuprates are charge-transfer insulators with localized spin-1/2 low-energy degrees of freedom and resemble a Heisenberg antiferromagnet away from the magnetic Brillouin zone. 
+In the heavily overdoped side, the lowest energy single-particle excitations look like well defined quasiparticles resembling a Fermi liquid apart from non-standard plasmonic modes. 
 In the intermediate doping region where the low-energy excitations cannot be easily sectioned into spin-like or electron-like only, we see the emergence of exotic phases and where most difficulties arise in writing down low-energy effective models.
@@
 Two-body terms in the model are developed by assuming a screened Coulomb interaction based on constrained DFT or RPA.
-Aside from this method, Lowdin methods coupled to a stochastic method and canonical transformations have also been used to develop effective theories.
+Aside from this method, L\''{o}wdin methods coupled to a stochastic method and canonical transformations have also been used to develop effective theories.
 These methods have the drawback that the constructed effective theory typically does not come with any estimate of the quality of the model, therefore making model validation a phenomenological task which can lead to unclear biases in the model construction.
@@
 We assessed the efficiency and compactness of this new trial function by calculating the ground state energy and single particle densities of a C$_2$ molecule using FN-DMC and comparing to the results when using multi-Slater-Jastrow trial functions with optimized orthogonal determinant trial functions (MSJ+O).
-The workflow involved constructing the un-optimized trial wave functions, optimizing the parameters using an energy optimization method \cite{Toulouse2007}, and finally using the optimized trial functions in an FN-DMC calculation.
-We found that the FN-DMC energy calculated using an MSJ+NO trial function with only 24 determinants was lower than the FN-DMC energy using an MSJ+O trial function with 55 determinants. 
+The workflow involved constructing the un-optimized trial wave functions, optimizing the parameters using an energy optimization method \cite{Toulouse2007}, and finally using the optimized trial functions in both variational Monte Carlo (VMC) and FN-DMC calculations.
+We found the average improvement in variational energy when using optimized orthogonal determinants is $\langle E_\text{MSJ} - E_\text{MSJ+O} \rangle = $ 0.32 eV with a standard deviation of 0.08 eV, with an additional improvement from the non-orthogonal determinants of 0.031 eV with a standard deviation of 0.021 eV.
+The average improvement in the FN-DMC energy with the MSJ+O trial wave functions was 0.14 eV (standard deviation 0.03 eV) with an additional reduction of 0.032 eV (standard deviation 0.019 eV) when using the MSJ+NO trial wave function.
+While the benefit using the MSJ+NO trial wave function seems small, we saw that the FN-DMC energy calculated using an MSJ+NO trial function with only 24 determinants was lower than the FN-DMC energy using an MSJ+O trial function with 55 determinants.
 Further, the FN-DMC charge density calculated using MSJ+NO trial functions had stronger bonding character than when using MSJ+O trial functions, a reasonable result as introducing correlations into trial functions allows for electrons to avoid each other while still occupying the same bonding region.
@@
 \includegraphics[width=0.6\linewidth]{figs/eff_eig_prelim.pdf}
-\caption{Results of weighted single parameter model fits to the eigenstates of a Hubbard model for various U and temperature. Darker hues indicate a higher R$^2$ score.}
+\caption{Results of weighted single parameter model fits to the eigenstates of a Hubbard model for various U and temperature.
+Darker hues indicate a higher R$^2$ score.
+Vertical dashed line denotes the bulk Slater to Mott-Heisenberg transition U$^*$/t = 4.7 calculated in [ref].}
 \label{fig:Hubbard}
@@
 I then fit three single parameter models to the eigenstates: A) 
-A "t" only model $t \sum_{\langle i, j \rangle, \sigma} c_{i\sigma}^\dagger c_{j\sigma} + h.c.$, B) A "U" only model $U \sum_{i} n_{i\uparrow} n_{i\downarrow}$, and C) A "J" only model $J \sum_{\langle i,j \rangle} \vec{S}_i \cdot \vec{S}_j$ using a weighted linear regression with weights $e^{-\beta E}$.
+A "effective t" model $t_\text{eff} \sum_{\langle i, j \rangle, \sigma} c_{i\sigma}^\dagger c_{j\sigma} + h.c.$, B) An "effective U" model $U_\text{eff} \sum_{i} n_{i\uparrow} n_{i\downarrow}$, and C) An "effective J" model $J_\text{eff} \sum_{\langle i,j \rangle} \vec{S}_i \cdot \vec{S}_j$ using a weighted linear regression with weights $e^{-\beta E}$.
 The basis for all of these operators are just the bare Hubbard Hamiltonian basis.
 The weighted linear regression approximates the effect of generating samples using a Markov Chain Monte Carlo method with a Boltzmann distribution.
-At low-temperatures there is a crossover between the weak coupling regime, best described by a non-interacting t-only model, and the strong coupling regime, best described by a J-only model, around U$^*/t \sim 5$ providing some evidence of a Slater to Mott-Heisenberg transition.
+At low-temperatures there is a crossover between the weak coupling regime, best described by an effective t model, and the strong coupling regime, best described by an effective J model, around U$^*/t \sim 5$ providing some evidence of a Slater to Mott-Heisenberg transition.
 In the crossover region T = 0 and U = 4 - 8 both models perform poorly and is a region of particular interest when developing our low-energy theories.
-At high temperatures there is a crossover between a non-interacting t-only and U-only model which can also be investigated.
+At high temperatures there is a crossover between the effective t and effective U models which can also be investigated.
 

### 2019-07-25 draft.tex
updated methods section. need to reoutline/rewrite intro and proposed work. Need to plug in CuO updated doc into preliminary work. Intro + proposed: Graphene + Conventional SC

@@
 In the forefront of modern research are materials like high temperature superconductors and heavy fermion systems for which simple effective theories have not yet been found. 
-For example, the cuprate family of high-Tc superconductors have an intricate phase diagram in the doping-temperature space  with a robust anti-ferromagnetic (AFM) charge-transfer insulator state near zero doping, a prominent superconducting dome in the 0.1 - 0.2 hole doped region at low temperatures with psuedo-gap and non-Fermi liquid metallic phases at higher temperatures and Fermi liquid like behavior in the overdoped area, of which only the AFM insulating and Fermi liquid phases have well understood mechanism and models.
+For example, the cuprate family of high-Tc superconductors have an intricate phase diagram in the doping-temperature space with anti-ferromagnetic (AFM) insulating, superconducting, psuedo-gap, "bad metal", and Fermi-liquid like phases of which only the first and last are well understood.
 The iron-pnictides also have superconducting domes, non-Fermi liquid like metallic phases, a spin-density-wave antiferommagnetic phase, all under various dopings and temperatures. 
@@
 In the intermediate doping region where the low-energy excitations cannot be easily sectioned into spin-like or electron-like only, we see the emergence of exotic phases and where most difficulties arise in writing down low-energy effective models.
-On top of that, this entire discussion did not include any lattice effects like structural phase transitions which many strongly-correlated materials have as well.
 
@@
 
+\textcolor{red}{
 We propose using a density matrix downfolding (DMD) procedure to systematically develop low-energy effective theories of the 1-band nearest-neighbor Hubbard model on a square lattice for various couplings U and temperatures T at half-filling as a stepping stone towards building effective theories for strongly correlated materials.
@@
 The relative simplicity of the half-filled Hubbard model, potential complexity of its phase diagram in U-T space,  and the lack of well-established low-energy theories for moderate U/t lends the model as a useful playground for our new downfolding procedure in preparation for tackling realistic strongly correlated electron systems.
+}
 
-The DMD procedure allows us to systematically develop low-energy effective theories for any Hamiltonian and Hilbert space $(H, \mathcal{H})$.
-The procedure defines a method for downfolding the \textit{ab-initio} energy functional and Hilbert space onto an effective energy functional defined on a low-energy subspace of the Hilbert space 
+\section{Methods}
+\subsection{Density Matrix Downfolding (DMD)}
+\textbf{We begin by defining what we mean by a low-energy effective Hamiltonian and downfolding.}
+Consider a Hamiltonian $\hat{H}$ defined on a Hilbert space $\mathcal{H}$.
+A low-energy effective Hamiltonian is an operator $\hat{H}_\text{eff}$ which can accurately reproduce the spectra, up to a constant, and eigenstates of $\hat{H}$ on a subpsace $\mathcal{LE}$ of the Hilbert space $\mathcal{H}$.
+Here $\mathcal{LE}$ is taken to be the span of the N lowest energy eigenstates of $\hat{H}$, and $\hat{H}_\text{eff}$ is a linear combination of Hermitian operators $\hat{d}_k$ with a constant energy shift $E_0$
 \begin{equation}
-(E, \mathcal{H}) \xrightarrow{\text{DMD}} (E_\text{eff}, \mathcal{LE} \subset \mathcal{H}), \text{ s.t. }
-\forall |\Psi\rangle \in \mathcal{LE}, \ E_\text{eff}[\Psi] - E[\Psi] = \epsilon[\Psi] + E_0.
+\hat{H}_\text{eff} = \sum_{k} g_k \hat{d}_k  + E_0.
+\label{eq:Heff}
+\end{equation}
+The objective of Hamiltonian downfolding is to construct such an $\hat{H}_\text{eff}$ given $\hat{H}, \mathcal{H}$.
+
+\textbf{The key insight of DMD is that the Hamiltonian downfolding problem can be mapped onto a linear regression problem.}
+It has been shown that the above definition of $\hat{H}_\text{eff}$ is equivalent to the following: a low-energy effective Hamiltonian is an operator on a subspace $\mathcal{LE}$ of $\mathcal{H}$ such that
+\begin{equation}
+\begin{split}
+\forall |\Psi\rangle \in \mathcal{LE},\ (E[\Psi] \equiv \langle \Psi|\hat{H} | \Psi \rangle)  = \\ (E_\text{eff}[\Psi] \equiv \langle \Psi | \hat{H}_\text{eff} | \Psi \rangle) + \epsilon[\Psi]
+\end{split}
 \label{eq:DMD}
 \end{equation}
-The energy functionals can be written as expectations over Hamiltonian operators $E = \langle \Psi | \hat{H} |\Psi \rangle$, $E_\text{eff} = \langle \Psi | \hat{H}_\text{eff} |\Psi \rangle$, $\epsilon$ represents the error in our effective theory, and $E_0$ is a constant energy shift.
-Broadly the DMD method consists of four steps which will be discussed in detail later: 
-A) Defining the space of low-energy excitations which $E_\text{eff}$ should be able to describe accurately.
-B) Sampling, at minimum, a set of states in $\mathcal{H}$ which probe the chosen low-energy excitations in a statistically independent manner. The selected states need not be eigenstates. 
-C) Constructing a set of candidate models which are linear in reduced density matrix (RDM) descriptors with variable coefficients $\{c\}$
+where $\epsilon[\Psi]$ is the error in the effective theory.
+Given the general linear form of $\hat{H}_\text{eff}$ seen in \eqref{eq:Heff}, 
+the task of building an effective Hamiltonian then reduces to fitting a linear model that minimizes the error $\epsilon[\Psi]$ over $\mathcal{LE}$.
+The independent variables (descriptors) in the fitting are expectation values of the operators $d_k[\Psi] \equiv \langle \Psi |\hat{d}_k|\Psi \rangle$ and the target variable is $E[\Psi]$.
+
+\textbf{This linear regression problem can be tackled in three steps, beginning with sampling states from $\mathcal{LE}$.}
+Ideally the linear regression would be conducted over the entire space $\mathcal{LE}$. 
+In practice, the space is too large, and instead a representative sample set is drawn and used for fitting.
+For each sampled state $|\Psi_s\rangle$, the quantities $E[\Psi_s], \{d_k[\Psi_s]\}$ are calculated to be used in the linear regression.
+Optimal sampling schemes vary between systems and can significantly reduce the number of samples required to achieve robust statistics.
+
+\textbf{Generally, however, one cannot sample the true low-energy subspace and low-energy wave functions are sampled from an approximate subspace $\mathcal{LE}^\prime$.}
+A model fit to an arbitrary subspace of $\mathcal{H}$ will generally fail in describing the low-energy excitations of $\hat{H}$.
+Rather, the choice of approximate subspace must be made carefully to ensure the fit model is transferable to the true low-energy space.
+We describe in detail later how we selected $\mathcal{LE}^\prime$ for the CuO molecule and the scheme used to draw samples from it.
+
+\textbf{Next, a set of candidate descriptors which form the effective Hamiltonian are selected.}
+Descriptors will typically be selected if their values co-vary with the energy over the set of sampled states. 
+Knowledge of the low-energy degrees of the freedom from experiment and physical intuition can help restrict the set of candidate descriptors further.
+In principle there is no restriction on the form of the candidate operators other than Hermiticity, but second quantized operators are commonly used.
+
+\textbf{Typically, candidate operators are written in second quantized form, and an appropriate single particle basis to express these operators must be constructed.}
+An appropriate choice of single particle basis can simplify the representation of the effective theory and typically depends on the system of interest.
+Common choices include Bloch waves and molecular orbitals for delocalized basis elements or Wannier and intrinsic atomic orbitals for localized ones.
+The selected single particle basis can be validated by ensuring it accurately describes single particle properties among the sampled states like the total number of electrons in the system.
+
+\textbf{ Finally, the sampled data are used to fit the coefficients $\{g_k\}$ by linear regression.}
+A typical fitting workflow involves feature selection, model regression and model validation.
+Feature selection methods for choosing appropriate descriptors can range from wrapper methods like orthogonal matching pursuit and principal component analysis to embedded methods like LASSO, ridge and elastic net regularizations.
+Fitting the effective Hamiltonian usually involves an ordinary least squares linear regression, but the cost function can be altered as in the cases of the embedded methods above or weighted least squares linear regression.
+Model validation can be carried out by calculating cross validated single parameter measures like R$^2$ scores which carry information about goodness of fit and overfitting.
+
+\textbf{Importantly, the fit effective Hamiltonian carries with it a quantitative measure of its validity.}
+The ability to quantify the accuracy of an effective Hamiltonian sets DMD apart from contemporary downfolding approaches which typically report only a final model.
+Given the statistical nature of the fitting procedure, model quality assessment can be carried out and reported in exhaustive detail if one is interested.
+Typically, however, simple single parameter measures like cross validated R$^2$ scores and root mean squared errors are used to quantify the accuracy of the effective Hamiltonian.
+
+\subsection{Fixed-node diffusion Monte Carlo (FN-DMC)}
+\textbf{Diffusion Monte Carlo (DMC) is a quantum Monte Carlo method which projects out the ground state of a real-space Hamiltonian given some initial trial wave function.}
+Consider a trial wave function $|\Psi_T\rangle$ and the Hamiltonian $\hat{H}$ with ground state $|\Phi_0\rangle$. 
+Applying the projector $e^{-\tau \hat{H}}$ as $\tau \rightarrow \infty$ to $|\Psi_T \rangle$
 \begin{equation}
-E_\text{eff} = \sum_k c_k \langle \Psi | \hat{d}_k |\Psi \rangle + E_0.
-\label{eq:Eeff}
-\end{equation}
-Here $\hat{d}_k$ are Hermitian operators typically composed of 1- or 2-RDM elements. 
-Examples of possible $\hat{d}_k$ are number operators $\hat{n}_i$ and exchange operators $\vec{S}_i \cdot \vec{S}_j$.
-Additionally, a single particle basis must be built to express RDM elements on and should be chosen such that variations among states in $\mathcal{LE}$ are mostly described by variations in the projection of these states onto the basis.
-D) Using the samples to fit the coefficients $\{c\}$ by linear regression such that $E_\text{eff}$ satisfies the conditions in \eqref{eq:DMD} with the smallest error $\epsilon$ possible. 
-The fitting procedure is typically an ordinary linear regression, however, one may use a custom cost function in the regression as well. 
-In order to conduct the linear regression for a functional like \eqref{eq:Eeff}, one needs to calculate for each state sampled from $\mathcal{LE}$ the expectation values of $\hat{H}$ and every descriptor $\{\hat{d}_k\}$ used in the regression.
-In principle, if conducted with no approximations, the DMD method is rigorously guaranteed to obtain exact results, namely $\epsilon[\Psi] = 0$.
-For real world systems we will work with $H = H_\text{ab}$, the \textit{ab-initio} Hamiltonian under the Born-Oppenheimer approximation with psuedopotentials:
-\begin{equation}
-\hat{H}_\text{ab} = -\frac{1}{2} \sum_{i} \nabla_i^2 - \sum_{i,I}\frac{Z}{|r_i - R_I|} + \sum_{i<j}\frac{1}{|r_i - r_j|}
-\label{eq:Hab}
-\end{equation}
-where $i$ indicates electrons and $I$ nuclei, and the unit of energy is Hartree (Ha).
-
-When downfolding $\hat{H}_\text{ab}$ we will use fixed-node diffusion Monte Carlo (FN-DMC) to access states in $\mathcal{LE}$ and accurately calculate their \textit{ab-initio} energies and RDM elements.
-Diffusion Monte Carlo (DMC) is a quantum Monte Carlo method which projects out the ground state of a Hamiltonian given some initial trial wave function.
-Consider a trial wave function $|\Psi_T\rangle$ and the Hamiltonian $\hat{H}_\text{ab}$ with ground state $|\Phi_0\rangle$.
-We apply the projector $e^{-\tau \hat{H}_\text{ab}}$ as $\tau \rightarrow \infty$ to $|\Psi_T \rangle$
-\begin{equation}
-\lim_{\tau \rightarrow \infty} e^{-\tau \hat{H}_\text{ab}} |\Psi_T\rangle 
+\lim_{\tau \rightarrow \infty} e^{-\tau \hat{H}} |\Psi_T\rangle 
 \equiv \lim_{\tau \rightarrow \infty} |\Psi_\text{DMC}(\tau)\rangle \propto \langle \Phi_0|\Psi_T\rangle |\Phi_0\rangle,
 \end{equation}
-projecting out the ground state as long as the trial wave function we choose is not orthogonal to the ground state. 
-The stochastic implementation involves moving samples from the trial function $\Psi_T(R)$ using the Green function $G(R, R^\prime, \tau) = \langle R | e^{-\tau(\hat{H}_\text{ab} - E_T)} | R^\prime \rangle$.
-Since $H_\text{ab} = T + V$, kinetic and potential energy terms, the Green function is approximated by a Trotter expansion $G(R, R^\prime, \tau) = \langle R | e^{-\tau(\hat{H} - E_T)} | R^\prime \rangle \sim \Big[e^{-d\tau(V(R) + V(R^\prime))/2} \langle R| e^{-d\tau(\hat{T} - E_T)}|R^\prime \rangle + O(d\tau^2) \Big]^N $ where $d\tau = \tau/N$.
+projects out the ground state as long as the trial wave function is not orthogonal to the ground state. 
+DMC provides a stochastic implementation of this projector method and involves moving samples from the trial function $\Psi_T(R)$ using the Green function $G(R, R^\prime, \tau) = \langle R | e^{-\tau(\hat{H} - E_T)} | R^\prime \rangle$.
+Since $\hat{H} = \hat{T} + \hat{V}$, kinetic and potential energy terms, the Green function is approximated by a Trotter expansion $G(R, R^\prime, \tau) = \langle R | e^{-\tau(\hat{H} - E_T)} | R^\prime \rangle \sim \Big[e^{-d\tau(\frac{V(R) + V(R^\prime)}{2} - E_T)} \langle R| e^{-d\tau\hat{T}}|R^\prime \rangle + O(d\tau^2) \Big]^N $ where $d\tau = \tau/N$.
 This expansion can be interpreted as an interative procedure where samples are moved N times with a small timestep $d\tau$ until convergence.
 The constant $E_T$ is a trial energy used to control the normalization of $\Psi_\text{DMC}(\tau, R)$ and is updated at each move.
-This implementation suffers from a fermion sign problem which is dealt with through a fixed-node approximation where the nodal surface of $\Psi_\text{DMC}(\tau, R)$ is forced to match that of the initial trial wave function for all $\tau$.
-This approximation makes FN-DMC variational, and will only return the exact ground state of $\hat{H}_\text{ab}$ if the nodal surfaces of $|\Psi_T\rangle$ and $|\Phi_0\rangle$ are identical.
-We make use of this variational principle to access low-energy states within $\mathcal{H}$ by choosing trial wave functions with nodes that differ from the nodes of $|\Phi_0 \rangle$.
-To calculate the expectation value of an operator $\hat{A}$ in FN-DMC we will use a mixed estimator $\langle \Psi_{DMC}(\tau) |\hat{A} | \Psi_T \rangle/\langle \Psi_\text{DMC}(\tau) | \Psi_T \rangle$ which allows us to use the importance sampled distribution $f = \Psi_\text{DMC}(\tau)\Psi_T$ for our Monte Carlo estimates.
+
+\textbf{DMC, however, suffers from a fermion sign problem which is alleviated via a fixed-node approximation.}
+Under the fixed-node approximation the nodal surface of $\Psi_\text{DMC}(\tau, R)$ is forced to match that of the initial trial wave function for all $\tau$.
+This approximation makes FN-DMC variational, and will only return the exact ground state of $\hat{H}$ if the nodal surfaces of $|\Psi_T\rangle$ and $|\Phi_0\rangle$ are identical.
+
+\textbf{We take advantage of the variational nature of FN-DMC to sample the low-energy states necessary for DMD.}
+Consider a set of trial wave functions with varying nodal structures.
+The final projected $\Psi_\text{DMC}$ for these different trial functions will be the lowest energy states in $\mathcal{H}$ for the given nodal structures.
+If the initial trial wave functions are appropriately chosen, the final projected states will be low-energy states within $\Psi_\text{DMC}$.
+We calculate the expectation values $E[\Psi_\text{DMC}], \{d_k[\Psi_\text{DMC}]\}$ necessary for the model regression using a mixed estimator.
 Details for calculating reduced density matrix elements in FN-DMC can be found in Wagner.

### 2019-07-26 draft.tex
worked on propsed outline. ready for writing, has TBLG stuff now

@@
 \section{Introduction}
+Systematically developing model Hamiltonians which can accurately describe the energies and properties of the lowest lying eigenstates for strongly correlated electron systems is a pressing problem in modern condensed matter physics.
+
+The necessity for non-perturbative, many-body models in describing the low-lying excited states of strongly correlated electron systems poses a difficult challenge when developing effective theories for real world systems.
+
+Common approaches to building suitable model Hamiltonians for strongly correlated materials involve using single particle theories like density functional theory (DFT).
+
+We propose using a DMD procedure to develop accurate many-body effective theories for single (SLG) and bi-layer graphene (BLG) systems as a step towards building a many-body model for twisted bilayer graphene (TBLG).
+
 A core problem of condensed matter physics is understanding how to systematically develop low-energy effective models for strongly correlated electron systems. 
@@
 
-\textcolor{red}{
-We propose using a density matrix downfolding (DMD) procedure to systematically develop low-energy effective theories of the 1-band nearest-neighbor Hubbard model on a square lattice for various couplings U and temperatures T at half-filling as a stepping stone towards building effective theories for strongly correlated materials.
-The Hubbard model is a simple model with nearest-neighbor hopping and a local on-site Coulomb repulsion, parameterized by a single ratio U/t: 
-\begin{equation}
-H_\text{hub} = -t \sum_{\langle i,j \rangle,\sigma}( c_{i^\dagger,\sigma} c_{j,\sigma} + h.c.) + U \sum_i n_{i\uparrow} n_{i\downarrow}
-\label{hubbard}.
-\end{equation}
-The solutions in 1-d are well understood via the Bethe ansatz, however, the 2-d square lattice Hubbard model has no known exact solution.  
-A few well accepted facts do exist for this model, such as the existence of long-range anti-ferromagnetic (AFM) order for any U$>$0 at T = 0 and a Mott-Heisenberg limit as $U/t \rightarrow \infty$ at low temperatures.
-The T = 0 diagram of the model at half-filling has been a long-standing source of controversy between two groups:  those who believe that for any U $>$ 0 the model exhibits Mott-Heisenberg behavior, and those who believe that there is a transition at finite U$^*$ between Slater and Mott-Heisenberg behavior.
-While recent numerical and theoretical calculations tend to agree with the latter position, non-negligible evidence for the prior does exist. 
-Further, modern finite temperature calculations indicate thermal transitions in the weak to moderate coupling region between the Slater phase at T = 0, pseudogap phase below T$_{pg}$ and a Fermi liquid phase above.
-A finite temperature psuedogap to Mott insulator transition has also been found between the weak to strong coupling limits.
-The nature of these transitions as phase transitions or simply crossovers is poorly understood.
-The relative simplicity of the half-filled Hubbard model, potential complexity of its phase diagram in U-T space,  and the lack of well-established low-energy theories for moderate U/t lends the model as a useful playground for our new downfolding procedure in preparation for tackling realistic strongly correlated electron systems.
-}
+We propose using a density matrix downfolding (DMD) procedure to develop an accurate low-energy effective theory for twisted bilayer graphene (TBLG) beginning from the \textit{ab-initio} Hamiltonian.
+TBLG exhibits strong correlation effects like unconventional superconductivity and many-body insulating states for extremely small, gate-tunable, dopings near charge neutrality for certain magic twist angles.
+These strong correlation effects arise from the emergence of very flat bands near the Fermi level as one approaches the magic angles.
+While some DFT calculations have generated effective single particle theories, no accurate many-body model Hamiltonian for TBLG which can describe these strong electronic correlations exists.
+As discussed below, DMD provides us the unique ability to generate such accurate many-body models directly from the \textit{ab-initio} Hamiltonian.
 
@@
 \subsection{Density Matrix Downfolding (DMD)}
-\textbf{We begin by defining what we mean by a low-energy effective Hamiltonian and downfolding.}
+We begin by defining what we mean by a low-energy effective Hamiltonian and downfolding.
 Consider a Hamiltonian $\hat{H}$ defined on a Hilbert space $\mathcal{H}$.
@@
 
-\textbf{The key insight of DMD is that the Hamiltonian downfolding problem can be mapped onto a linear regression problem.}
+The key insight of DMD is that the Hamiltonian downfolding problem can be mapped onto a linear regression problem.
 It has been shown that the above definition of $\hat{H}_\text{eff}$ is equivalent to the following: a low-energy effective Hamiltonian is an operator on a subspace $\mathcal{LE}$ of $\mathcal{H}$ such that
@@
 
-\textbf{This linear regression problem can be tackled in three steps, beginning with sampling states from $\mathcal{LE}$.}
+This linear regression problem can be tackled in three steps, beginning with sampling states from $\mathcal{LE}$.
 Ideally the linear regression would be conducted over the entire space $\mathcal{LE}$. 
@@
 
-\textbf{Generally, however, one cannot sample the true low-energy subspace and low-energy wave functions are sampled from an approximate subspace $\mathcal{LE}^\prime$.}
+Generally, however, one cannot sample the true low-energy subspace and low-energy wave functions are sampled from an approximate subspace $\mathcal{LE}^\prime$.
 A model fit to an arbitrary subspace of $\mathcal{H}$ will generally fail in describing the low-energy excitations of $\hat{H}$.
@@
 
-\textbf{Next, a set of candidate descriptors which form the effective Hamiltonian are selected.}
+Next, a set of candidate descriptors which form the effective Hamiltonian are selected.
 Descriptors will typically be selected if their values co-vary with the energy over the set of sampled states. 
@@
 
-\textbf{Typically, candidate operators are written in second quantized form, and an appropriate single particle basis to express these operators must be constructed.}
+Typically, candidate operators are written in second quantized form, and an appropriate single particle basis to express these operators must be constructed.
 An appropriate choice of single particle basis can simplify the representation of the effective theory and typically depends on the system of interest.
@@
 
-\textbf{ Finally, the sampled data are used to fit the coefficients $\{g_k\}$ by linear regression.}
+Finally, the sampled data are used to fit the coefficients $\{g_k\}$ by linear regression.
 A typical fitting workflow involves feature selection, model regression and model validation.
@@
 
-\textbf{Importantly, the fit effective Hamiltonian carries with it a quantitative measure of its validity.}
+Importantly, the fit effective Hamiltonian carries with it a quantitative measure of its validity.
 The ability to quantify the accuracy of an effective Hamiltonian sets DMD apart from contemporary downfolding approaches which typically report only a final model.
@@
 \subsection{Fixed-node diffusion Monte Carlo (FN-DMC)}
-\textbf{Diffusion Monte Carlo (DMC) is a quantum Monte Carlo method which projects out the ground state of a real-space Hamiltonian given some initial trial wave function.}
+Diffusion Monte Carlo (DMC) is a quantum Monte Carlo method which projects out the ground state of a real-space Hamiltonian given some initial trial wave function.
 Consider a trial wave function $|\Psi_T\rangle$ and the Hamiltonian $\hat{H}$ with ground state $|\Phi_0\rangle$. 
@@
 
-\textbf{DMC, however, suffers from a fermion sign problem which is alleviated via a fixed-node approximation.}
+DMC, however, suffers from a fermion sign problem which is alleviated via a fixed-node approximation.
 Under the fixed-node approximation the nodal surface of $\Psi_\text{DMC}(\tau, R)$ is forced to match that of the initial trial wave function for all $\tau$.
@@
 
-\textbf{We take advantage of the variational nature of FN-DMC to sample the low-energy states necessary for DMD.}
+We take advantage of the variational nature of FN-DMC to sample the low-energy states necessary for DMD.
 Consider a set of trial wave functions with varying nodal structures.
@@
 \section{Proposed work}
-My goal is to develop low-energy effective theories for the square lattice 1-band Hubbard model in order to map out the phases and transitions in the U-T space.
-Recent attempts at solving this problem involve constructing single particle properties like densities of states using approximate numerical or effective theory techniques to classify different phases.
-While instructive, reliance on single particle properties to identify phases instead of studying changes in the many-body Hilbert space and low-energy effective theory can lead to ambiguities regarding the nature of the phases.
-I believe the DMD method can remove these ambiguities as the method explicitly involves studying changes in the low-energy Hilbert space and provides us with accurate many-body effective theories which can precisely differentiate different phases.
+My proposed plan involves building a sequence of model Hamiltonians for increasing complicated systems using DMD, working towards an accurate many-body model for TBLG.
 
-My first step would be to show whether the model supports more than one phase at T/t = 0 with respect to U, and if so what the features of these phases and the transition between them are.
-I will investigate in particular the hypothesis that for $U/t<<1$ the model exhibits gapped Slater AFM behavior and for $U/t>>1$ Mott-Heisenberg behavior and whether the transition between them is a genuine phase transition or simply a crossover due to rearrangement of a few states.
-To do so I will initially develop effective theories using DMD at two points U/t = 2, 12 and T/t = 0, allowing me to argue the existence or non-existence of multiple phases and investigate the properties of each phase.
-I will follow by developing theories for various intermediate U in order to locate and study the transition between the two phases, if it exists, focusing on the hypothesis of a crossover U$^*/t\sim 4 - 5$ between the Slater and Mott-Heisenberg regimes.
-Repeating this exercise for finite temperature I will build four more models at U/t = 2, 12 and T/t = 0.05, 0.20 to investigate the finite temperature phases at strong- and weak-coupling, in particular the claim of pseudogap and Fermi liquid phases at T/t = 0.05, 0.20 at weak coupling.
-Filling in the U-T diagram will allow me to locate and study thermal transitions in the system. 
-Building the T$>$0 effective theories will require special attention as they must satisfy the Mermin-Wagner theorem.
+The first system I will develop a model for is the benzene molecule.
 
-Similar to the case of the \textit{ab-initio} downfolding, I will need to construct low-energy spaces, sample the constructed spaces, build effective theories and respective bases, and fit candidate models.
-I will begin from a starting point of highest control, namely for system sizes small enough where we can exactly diagonalize the Hubbard model.
-In this case selecting $\mathcal{LE}$ can be done exactly as the span of some chosen $N$ lowest energy eigenstates. 
-One can guarantee that the sample set fully encapsulates all low-energy degrees of freedom by sampling states parameterized as $|\Psi\rangle = \sum_{j=1}^N c_j |\Phi_j\rangle$ where $|\Phi_j\rangle$ are eigenstates of the Hubbard Hamiltonian.
-Basis construction for the Hubbard Hamiltonian is also explicitly clear as any possible single particle basis we can use is just an orthogonal transformation of the basis the Hubbard model is defined on.
-In order to include temperature into the mix one can work with the full model Hilbert space such that $\mathcal{LE} = \mathcal{H}$ but generate samples within the space using a Monte Carlo sampling method according to a Boltzmann probability distribution with variable parameter $\beta = 1/k_BT$, biasing the final model towards describing more accurately the energy functional on the states with highest weight.
-Finite size extrapolation to the thermodynamic limit requires working with lattice sizes for which exact diagonalization is no longer feasible, and approximate methods like lattice variational Monte Carlo (VMC) will be employed to sample low-energy states as well as calculate energies and RDM elements.
-In the weak and strong coupling regimes typical choices for varitional wave functions are non-interacting and Gutzwiller trial functions respectively, but other post-Gutzwiller and generalized Hartree Fock (GHF) trial functions have been shown to give accurate results in the moderate coupling regimes.
-A rigorous prescription for finite size extrapolation specifically for the 1-band square lattice Hubbard model will be followed to extend our conclusions to the thermodynamic limit.
+While relatively simple, this molecule shares many similarities with and forms the basic unit of SLG.
 
-\begin{figure}[H]
-\centering
-\includegraphics[width=0.6\linewidth]{figs/eff_eig_prelim.pdf}
-\caption{Results of weighted single parameter model fits to the eigenstates of a Hubbard model for various U and temperature.
-Darker hues indicate a higher R$^2$ score.
-Vertical dashed line denotes the bulk Slater to Mott-Heisenberg transition U$^*$/t = 4.7 calculated in [ref].}
-\label{fig:Hubbard}
-\end{figure}
+As such, developing a model for benzene will help me become familiar with model fitting for 2-D carbon based systems.
 
-I will conclude by showing some very preliminary results on model fitting for the half-filled, 2x2 square Hubbard model for U: 0 - 12 which put the outlined approach to model fitting with exact diagonalization at T$\geq 0$ into practice.
-I started off by exactly diagonalizing each model, defining $\mathcal{LE} = \mathcal{H}$, and treating the full sample space as just the eigenstates of the model.
-I then fit three single parameter models to the eigenstates: A) 
-A "effective t" model $t_\text{eff} \sum_{\langle i, j \rangle, \sigma} c_{i\sigma}^\dagger c_{j\sigma} + h.c.$, B) An "effective U" model $U_\text{eff} \sum_{i} n_{i\uparrow} n_{i\downarrow}$, and C) An "effective J" model $J_\text{eff} \sum_{\langle i,j \rangle} \vec{S}_i \cdot \vec{S}_j$ using a weighted linear regression with weights $e^{-\beta E}$.
-The basis for all of these operators are just the bare Hubbard Hamiltonian basis.
-The weighted linear regression approximates the effect of generating samples using a Markov Chain Monte Carlo method with a Boltzmann distribution.
-At low-temperatures there is a crossover between the weak coupling regime, best described by an effective t model, and the strong coupling regime, best described by an effective J model, around U$^*/t \sim 5$ providing some evidence of a Slater to Mott-Heisenberg transition.
-In the crossover region T = 0 and U = 4 - 8 both models perform poorly and is a region of particular interest when developing our low-energy theories.
-At high temperatures there is a crossover between the effective t and effective U models which can also be investigated.
+The candidate descriptors I will use for DMD are motivated by a previous calculation on the benzene molecule.
+
+A new constrained variational Monte Carlo (CVMC) method will be used to sample the low-energy wave functions necessary for DMD.
+
+Note also that a model for benzene has already been developed using DMD.
+
+This makes benzene a useful system not only for my personal learning, but as a benchmarking system for new methods like CVMC which have been developed recently.
+
+The second system I will be interested in working on is SLG.
+
+The model I am interested in working on will be an extension of a previous model developed for SLG using DMD.
+
+I will extend the previous model by introducing long range density-density interactions into the effective theory.
+
+The candidate model I anticipate to work with is an extended Hubbard model of the form.
+
+CVMC will be used to sample the low-energy wave functions necessary to fit the model.
+
+The last systems I will develop models for are AA and AB stacked BLG.
+
+Being the simplest BLG configuration, these models will form a starting point for a more thorough development of models for TBLG.
+
+I anticipate that the candidate descriptors should include interlayer couplings as well as the terms seen in SLG.
+
+I will again use CVMC to generate the low-energy wave functions for the model regression.
+
+The final four models can be used to investigate the transferrability of model parameters between these different carbon-based systems.
+
+I will conclude by briefly discussing potential avenues of further study in using DMD to develop model Hamiltonians for TBLG.
+
+My proposed plan involves building a sequence of model Hamiltonians for increasingly complicated systems using DMD, working towards a model for TBLG.
+I will begin with the basic building block of SLG, the benzene molecule, as an exercise to become familiar with the structure and chemistry of graphene like materials.
+The benzene molecule consists of a hexagonal carbon ring capped by hydrogen atoms. 
+It shares many features with SLG such as delocalized $\pi$ electrons and long-range charge density-density interactions.
+Alongside the fact that previous DMD work has already developed an accurate model for benzene, this molecule provides a useful playground for building my familiarity with the chemistry and effective interactions in graphene like systems.
+
+The benzene model fitting will also serve as a benchmark study for new model fitting methods developed since the first DMD calculations on it.
+I will be particularly interested in testing a constrained variation Monte Carlo (CVMC) sampling method which allows us to target low-energy states with specified descriptor values.
+This sampling method will be the core sampling method used in all the proposed work, with the calculations on benzene serving as an familiarizing exercise for me.
+The results of my model fit using CVMC can be validated by comparing the effective theory I have built to the many-body model previous developed for benzene.
+
+The next step will be developing a model with long range interactions for SLG, expanding on previous work.
+An on-site effective Hubbard model has already been built using DMD for single-layer graphene with parameters $U/t \sim 2.00$ eV.
+Following the example of the benzene molecule, however, one would expect that longer range density-density interactions would be important in describing SLG more accurately.
+The goal of this step of the project would be to develop a model of the following form:
+\begin{equation}
+\sum_{ij} t_{ij} c_i^\dagger c_j + h.c. +  \sum_i U n_{i\uparrow} n_{i\downarrow} + \sum_{ij} V_{ij} n_i^\dagger n_j
+\end{equation}
+with hopping terms $t_{ij}$, an on-site Hubbard $U$, and long range density-density interaction $V_{ij}$.
+
+The last part of my plan involves developing a model for AA and AB stacked bilayer graphene.
+These stackings are the simplest forms of bilayer graphene as the supercell is much smaller than those for even moderate commensurate twist angles.
+The relative shift in distance among the atoms between the AA and AB stackings will serve as a first step to understand how the interlayer coupling between graphene layers may depend on the distance between atoms in the layers.
+The final model to fit for these cases would be identical to the one above, but with additional interlayer hopping terms.
+
+The end result of my proposed project is four model Hamiltonians for the  benzene molecule, SLG, AA stacked BLG, and AB stacked BLG.
+These four models will include, at least, nearest neighbor hoppings, on-site Hubbard interactions, and long range density-density interactions.
+For the AA and AB stacked BLG we will also consider interlayer couplings starting with just hoppings between the layers.
+Each model will come with fit parameter values, an appropriate single particle basis, and a measure of the goodness of fit for the model.
+These models can be solved and compared to experiment.
+
+With these models we can resolve an interesting question: are model parameters transferable between benzene, SLG and BLG?
+The question of transferability of model parameters between SLG and BLG is quite interesting because it may inform us about how much the behavior of each graphene sheet changes once we stack them.
+If the interlayer coupling does not affect the effective behavior on the single sheets, then one can easily treat the final model for bi-layer graphene in two independent pieces: an accurate model for the graphene sheets with a position dependent interlaying coupling.
+However, if the interlayer coupling plays a large role in the effective theory then it may be impossible to separate the BLG model into these two pieces.
+A similar study can be done for the single particle basis, giving some information about orbital relaxation going from SLG to BLG.
+
+My work will provide a useful baseline to begin studies on TBLG.
+The approach to TBLG is still very far away and will require all of the previous steps above. 
+Afterwards, work on TBLG will require building effective theories of BLG with multiple twist angles, or at least varying local distances between atoms.
+The goal here is to develop models for BLG where the parameters, like interlayer coupling, are defined continuously as a function of atomic distance.
+The resulting interacting lattice model can then be solved over very large supercell sizes and a second step of downfolding can be done.
+The goal of this second step would be to downfold the lattice model into a continuum model over the length scale of the Moire patterns seen in small twist angle BLG.
+Work on TBLG would be highly exploratory, and the steps detailed here are very speculative, outlining a possible avenue of work continuing after my thesis.
 

### 2019-07-26 outline.tex
worked on propsed outline. ready for writing, has TBLG stuff now

@@
 
-\section{Introduction and methods}
-\paragraph{Overall goal} We want to be able to systematically develop low-energy effective models for strongly correlated electron systems.
+\section{Introduction}
+\begin{itemize}
+\item Systematically developing model Hamiltonians which can accurately describe the energies and properties of the lowest lying eigenstates for strongly correlated electron systems is a pressing problem in modern condensed matter physics.
 
-\paragraph{Barrier to achieving goal} Non-perturbative interactions between various degrees of freedom - for example, lattice, spin and charge - makes writing down low-energy effective theories for strongly correlated materials very challenging.
+\item The necessity for non-perturbative, many-body models in describing the low-lying excited states of strongly correlated electron systems poses a difficult challenge when developing effective theories for real world systems.
 
-\paragraph{Current state of the art} \textcolor{red}{Talk about methods people use for model fitting now. Will fill out as I do literature review.}
+\item Common approaches to building suitable model Hamiltonians for strongly correlated materials involve using single particle theories like density functional theory (DFT).
 
-\paragraph{Advancement to the state of the art} We propose using a density matrix downfolding (DMD) procedure to develop low-energy effective theories of the 1-band nearest-neighbor Hubbard model on a square lattice as a stepping stone towards building effective theories for strongly correlated materials.
+\item We propose using a DMD procedure to develop accurate many-body effective theories for single (SLG) and bi-layer graphene (BLG) systems as a step towards building a many-body model for twisted bilayer graphene (TBLG).
+\end{itemize}
 
-\paragraph{Density matrix down folding} 
-The DMD procedure allows us to develop low-energy effective theories for physical systems in a systematic manner beginning with a higher energy Hamiltonian, like the \textit{ab-initio} Hamiltonian $H_{ab}$.
+\section{Methods}
+\begin{enumerate}
+\item Density matrix downfolding
+\begin{itemize}
+\item We begin by defining what we mean by a low-energy effective Hamiltonian and downfolding.
 
-\paragraph{Fixed node diffusion Monte Carlo}
-We will use fixed node diffusion Monte Carlo (FN-DMC) to generate wave functions in our chosen low-energy subspace and to accurately calculate \textit{ab-initio} energies and reduced density matrices.
+\item The key insight of DMD is that the Hamiltonian downfolding problem can be mapped onto an equivalent linear regression problem.
+
+\item This linear regression problem can be tackled in three steps, beginning with sampling states from $\mathcal{LE}$.
+
+\item Generally, one cannot sample the true low-energy subspace and an approximate subspace $\mathcal{LE}^\prime$ is sampled.
+
+\item Next, a set of candidate descriptors which form the effective Hamiltonian are selected.
+
+\item Typically, candidate operators are written in second quantized form, and an appropriate single particle basis to express these operators must be constructed.
+
+\item Finally, the sampled data are used to fit the coefficients $\{g_k\}$ by linear regression.
+
+\item Importantly, the fit effective Hamiltonian carries with it a quantitative measure of its validity.
+\end{itemize}
+
+\item Fixed-node diffusion Monte Carlo 
+\begin{itemize}
+\item Diffusion Monte Carlo (DMC) is a quantum Monte Carlo method which projects out the ground state of a real-space Hamiltonian given some initial trial wave function.
+
+\item This stochastic implementation of DMC suffers from a fermion sign problem which is alleviated via a fixed-node approximation.
+
+\item We take advantage of the variational nature of FN-DMC to sample the low-energy states necessary for DMD.
+\end{itemize}
+\end{enumerate}
 
@@
 \section{Proposed work}
-\paragraph{Hubbard: Effective models at half-filling} We want to develop low-energy effective theories for the half-filled Hubbard model over the space U: 0 - 12, 1/$k_B$T: 0 - 3 using the DMD method, particularly on the medium coupling region of U: 2 - 6 and 1/$k_B$T $\sim$ 1 where perturbative effective theories break down and a metal-insulator transition exists.
+\begin{itemize}
+\item My proposed plan involves building a sequence of model Hamiltonians for increasing complicated systems using DMD, working towards an accurate many-body model for TBLG.
 
-\paragraph{Hubbard: Effective models away from half-filling} A farther reaching goal would be to develop effective theories in the U-T-$\delta$ space using DMD, where more complex behavior similar to the cuprates have been found.
+\item The first system I will develop a model for is the benzene molecule.
 
-\paragraph{Hubbard: Finite size effects} Our initial calculations necessary for DMD can be done using exact diagonalization, but memory limitations will require us to use approximate methods like lattice quantum Monte Carlo to calculate energies and density matrix elements for states in our LES.
+\item While relatively simple, this molecule shares many similarities with and forms the basic unit of SLG.
 
+\item As such, developing a model for benzene will help me become familiar with model fitting for 2-D carbon based systems.
+
+\item The candidate descriptors I will use for DMD are motivated by a previous calculation on the benzene molecule.
+
+\item A new constrained variational Monte Carlo (CVMC) method will be used to sample the low-energy wave functions necessary for DMD.
+
+\item Note also that a model for benzene has already been developed using DMD.
+
+\item This makes benzene a useful system not only for my personal learning, but as a benchmarking system for new methods like CVMC which have been developed recently.
+
+\item The second system I will be interested in working on is SLG.
+
+\item The model I am interested in working on will be an extension of a previous model developed for SLG using DMD.
+
+\item I will extend the previous model by introducing long range density-density interactions into the effective theory.
+
+\item The candidate model I anticipate to work with is an extended Hubbard model of the form.
+
+\item CVMC will be used to sample the low-energy wave functions necessary to fit the model.
+
+\item The last systems I will develop models for are AA and AB stacked BLG.
+
+\item Being the simplest BLG configuration, these models will form a starting point for a more thorough development of models for TBLG.
+
+\item I anticipate that the candidate descriptors should include interlayer couplings as well as the terms seen in SLG.
+
+\item I will again use CVMC to generate the low-energy wave functions for the model regression.
+
+\item The final four models can be used to investigate the transferrability of model parameters between these different carbon-based systems.
+
+\item I will conclude by briefly discussing potential avenues of further study in using DMD to develop model Hamiltonians for TBLG.
+
+\end{itemize}
 \section{Proposed timeline}

### 2019-07-27 draft.tex
rewrote proposed work. need to redo intro, seems a bit off topic

@@
 Systematically developing model Hamiltonians which can accurately describe the energies and properties of the lowest lying eigenstates for strongly correlated electron systems is a pressing problem in modern condensed matter physics.
+At the forefront of modern research are materials such as the cuprate supercondutors, heavy fermion systems, and most recently, twisted bilayer graphene (TBLG).
+These materials exhibit various phases of matter which are far from understood, like unconventional superconductivity, psuedogap and bad metal phases, quantum critical points, and stripe ordering.
+While no definitive models for these phases have been introduced, it is expected that strong correlations between electronic degrees of freedom lead to these phases.
 
 The necessity for non-perturbative, many-body models in describing the low-lying excited states of strongly correlated electron systems poses a difficult challenge when developing effective theories for real world systems.
-
-Common approaches to building suitable model Hamiltonians for strongly correlated materials involve using single particle theories like density functional theory (DFT).
-
-We propose using a DMD procedure to develop accurate many-body effective theories for single (SLG) and bi-layer graphene (BLG) systems as a step towards building a many-body model for twisted bilayer graphene (TBLG).
-
-A core problem of condensed matter physics is understanding how to systematically develop low-energy effective models for strongly correlated electron systems. 
-In the forefront of modern research are materials like high temperature superconductors and heavy fermion systems for which simple effective theories have not yet been found. 
-For example, the cuprate family of high-Tc superconductors have an intricate phase diagram in the doping-temperature space with anti-ferromagnetic (AFM) insulating, superconducting, psuedo-gap, "bad metal", and Fermi-liquid like phases of which only the first and last are well understood.
-The iron-pnictides also have superconducting domes, non-Fermi liquid like metallic phases, a spin-density-wave antiferommagnetic phase, all under various dopings and temperatures. 
-Heavy-fermion materials have a purported quantum critical point between an AFM and Fermi-liquid phase which houses a superconducting dome, leading to another phase diagram for which building model Hamiltonians is very challenging.
-
-The fundamental barrier in writing down low-energy theories for strong correlated materials is the existence of non-perturbative interactions between various degrees of freedom - for example, lattice, spin and charge.
-Consider just the cuprate case for now, the two phases which are relatively well understood contain a singular important degree of freedom in the low-energy space. 
-Near zero doping the cuprates are charge-transfer insulators with localized spin-1/2 low-energy degrees of freedom and resemble a Heisenberg antiferromagnet away from the magnetic Brillouin zone. 
-In the heavily overdoped side, the lowest energy single-particle excitations look like well defined quasiparticles resembling a Fermi liquid apart from non-standard plasmonic modes. 
-In the intermediate doping region where the low-energy excitations cannot be easily sectioned into spin-like or electron-like only, we see the emergence of exotic phases and where most difficulties arise in writing down low-energy effective models.
+Consider for now just the case of TBLG, where for certain magic twist angles, bands near the Fermi surface flatten out to a bandwidth of O(10) meV.
+At these magic angles, a small doping can lead to superconductivity and many-body insulating phases in the material, indicative of electronic interaction effects which cannot be explained by a simple band theory.
+The effects of the interaction are also beyond just simple renormalization of the non-interacting parameters in a band theory.
+A non-perturbative, many-body effective theory is therefore necessary to describe these interesting phases in TBLG.
 
@@
 
-We propose using a density matrix downfolding (DMD) procedure to develop an accurate low-energy effective theory for twisted bilayer graphene (TBLG) beginning from the \textit{ab-initio} Hamiltonian.
-TBLG exhibits strong correlation effects like unconventional superconductivity and many-body insulating states for extremely small, gate-tunable, dopings near charge neutrality for certain magic twist angles.
-These strong correlation effects arise from the emergence of very flat bands near the Fermi level as one approaches the magic angles.
-While some DFT calculations have generated effective single particle theories, no accurate many-body model Hamiltonian for TBLG which can describe these strong electronic correlations exists.
-As discussed below, DMD provides us the unique ability to generate such accurate many-body models directly from the \textit{ab-initio} Hamiltonian.
+I propose using a quantum Monte Carlo (QMC) density matrix downfolding (DMD) procedure to develop accurate many-body effective theories for single (SLG) and bi-layer graphene (BLG) systems as a step towards building a many-body model for the strongly correlated TBLG system.
+DMD is a relatively new method which allows one to systematically downfold the \textit{ab-initio} Hamiltonian for a system to a many-body lattice model.
+While it can be applied to any system, the methods necessary to carry out the downfolding accurately and efficiently have not yet been developed and tested for strong correlated materials.
+Both BLG and SLG provide useful playgrounds for developing methods for DMD since they have moderate interaction effects, with non-interacting theories describing the lowest lying excitations relatively accurately.
+Fitting models for SLG and BLG from DMD will not only provide a useful stepping stone towards understanding TBLG, but will also help mature the DMD method for its application to other strongly correlated systems.
 
@@
 \section{Proposed work}
-My proposed plan involves building a sequence of model Hamiltonians for increasing complicated systems using DMD, working towards an accurate many-body model for TBLG.
+My proposed plan involves building a sequence of model Hamiltonians for increasing complicated systems using QMC DMD, working towards an accurate many-body model for TBLG.
+The systems I will be interested in are the benzene molecule, single layer graphene, AA stacked bilayer graphene and AB stacked bilayer graphene.
+Conceptually, the benzene molecule is a kind of building block for SLG, SLG is a building block for BLG, and BLG for TBLG, creating a sequence of systems from least to most complex.
+I do not believe I will be able to build a general model for TBLG in the time span of my thesis, but my work should lay the foundation for further studies using QMC DMD focused on bilayer models with twists.
 
 The first system I will develop a model for is the benzene molecule.
+The benzene molecule has chemical formula $C_6 H_6$ and contains a hexagonal ring of carbon atoms with hydrogen atoms sticking out from each carbon.
+These carbon atoms are linked by sp$^2$ hybridized $\sigma$ bonds and an alternating sequence of $\pi$ bonds from the C 2p$_z$ orbitals.
+Intuitively, the ground state of the system can be seen as a superposition of two configurations of alternating $\pi$ bonds as seen in Figure \textbf{ref} rotated 30$^o$ relative to each other.
 
 While relatively simple, this molecule shares many similarities with and forms the basic unit of SLG.
-
+SLG can be viewed as a tiling of benzene molecules where the external hydrogen legs are replaced by carbon atoms.
+The delocalized $\pi$ orbitals in benzene carry over directly into SLG where they are the primary contributors to the well known Dirac bands.
+Interaction effects seen in the benzene molecule are also reflected in SLG, such as on-site double occupation energies for the carbon atoms.
 As such, developing a model for benzene will help me become familiar with model fitting for 2-D carbon based systems.
@@
 The candidate descriptors I will use for DMD are motivated by a previous calculation on the benzene molecule.
+In that study, an extended Hubbard model built on the six $\pi$ orbitals was found to perform well in describing the low-lying excitations of the material.
+The final model took the form 
+\begin{equation}
+H_\text{benz} = -\sum_{\langle i,j \rangle} t_{ij}c_i^\dagger c_j + U \sum_i n_{i\uparrow}n_{i\downarrow}  + \sum_{\langle i,j \rangle}V_{ij} n_i n_j + \sum_{\langle \langle i,j \rangle\rangle}V_{ij} n_i n_j
+\label{Hbenz}
+\end{equation}
+where the brackets denote nearest neighbor and next-nearest neighbor summations.
+I will consider at least all the descriptors above and also longer range hopping terms.
+The single particle orbitals used to express these descriptors will be IAOs.
 
 A new constrained variational Monte Carlo (CVMC) method will be used to sample the low-energy wave functions necessary for DMD.
+The CVMC method works by minimizing the following cost function for a parameterized wave function $|\Psi(\vec{p})\rangle$:
+\begin{equation}
+\vec{p}^* = \text{argmin} \ \langle \Psi(\vec{p}) | \hat{H}_\text{ab} | \Psi(\vec{p}) \rangle - \vec{\lambda} \cdot \langle \Psi(\vec{p}) | (\vec{d} - \vec{d}_0)^2 | \Psi(\vec{p}) \rangle 
+\end{equation}
+with respect to the parameters $\vec{p}$.
+The quantity $\vec{d}$ is short hand for a list of candidate descriptors to sample variations in, $\vec{d}_0$ is the target values of those candidate descriptors, and $\vec{\lambda}$ controls the importance of finding a state with low-energy versus one with descriptor values near $\vec{d}_0$.
+Essentially this method allows us, for a given parameterization of $|\Psi(\vec{p})\rangle$, to generate low-energy wave functions which sit near a target point $\vec{d}_0$ in candidate descriptor space.
 
-Note also that a model for benzene has already been developed using DMD.
+I will use a multi-Slater-Jastrow (MSJ) parameterization in the CVMC sampling for the benzene molecule.
+The multi-Slater expansion will consist of determinants of single particle orbitals calculated using a self consistent field calculation like DFT.
+The determinants will include the DFT ground state as well as molecular orbital excitations among the $\pi$ and $\pi^*$ orbitals which form our active space, allowing us to probe variation in all the descriptors above.
+A three body Jastrow factor will be applied to ensure that we include correlations in our wave function and that our wave function is low in energy.
+The total set of parameters $\vec{p}$ will include both Jastrow and determinental coefficients.
 
-This makes benzene a useful system not only for my personal learning, but as a benchmarking system for new methods like CVMC which have been developed recently.
+Since a model for benzene has already been developed using DMD, my calculation will also serve as a benchmark for new methods like CVMC sampling.
+The CVMC method is relatively new and has not been tested extensively on realistic systems.
+The final model that I will produce should be exactly comparable to previous calculations, and as such can give insight into the workings of the CVMC method.
+A few questions of interest include whether the CVMC is more statistically efficient in terms of regression, how to select values like $\vec{\lambda}$ and $\vec{d}_0$, and whether the final model fit with CVMC is similar to the one fit previously using a different sampling technique.
 
 The second system I will be interested in working on is SLG.
+Single layer graphene is composed of a triangular Bravais lattice with two carbon atoms per unit cell.
+In terms of a non-interacting picture, the $\sigma$ orbitals composed of C $2s, 2p_x,$ and $2p_y$ lie below the Fermi level.
+The bands which cross the Fermi level, and compose the distinct Dirac cones seen in SLG, are primary $\pi$ bonded molecular orbitals composed of C $2p_z$ orbitals.
+The lack of significant screening in SLG leads to long range Coulomb interactions in the system.
 
 The model I am interested in working on will be an extension of a previous model developed for SLG using DMD.
+The fit model was a single band, on-site Hubbard model which contained only the $\pi$ orbital degrees of freedom in the system.
+The fit model parameters were $U/t \sim$ 2.00, which is smaller than the critical value of $(U/t)_c \sim$ 3.8 for the semimetal-insulator transition in the honeycomb lattice, as expected.
+It is interesting to note that if the $\sigma$ orbitals in graphene are treated as a constant negative charge background, then the effective $U$ on site nearly doubles, indicating the importance of including these degrees of freedom in the low-energy sampling.
 
 I will extend the previous model by introducing long range density-density interactions into the effective theory.
-
-The candidate model I anticipate to work with is an extended Hubbard model of the form.
+As mentioned previously, one expects to see long range Coulomb interactions in graphene due to the vanishing density of states near the Fermi level, in contrast to other metals where electron-hole pairs screen the long range interaction.
+I will work on developing a model for SLG of the following form:
+\begin{equation}
+H_\text{SLG} = -\sum_{i,j} t_{ij}c_i^\dagger c_j + U \sum_i n_{i\uparrow}n_{i\downarrow}  + \sum_{i,j} V_{ij} n_i n_j
+\label{Hslg}
+\end{equation}
+where the indices correspond to the $\pi$ orbitals on the copper atoms, represented by IAO basis elements.
+I will consider arbitrarily long range hopping and density-density interaction terms in my model, in contrast to the benzene calculation where the finite size of the system imposes a cutoff in the possible range of interaction.
 
 CVMC will be used to sample the low-energy wave functions necessary to fit the model.
+The trial wave function will also be of MSJ type as explained for the benzene molecule, with the determinants built from the single particle orbitals of DFT calculations.
+I will consider the DFT ground state determinant and all excitations built on in restricted to the $\pi, \pi^*$ molecular orbitals.
+A three-body Jastrow will be used to introduce correlations into the wave function and ensure our sampled states are low in energy.
 
 The last systems I will develop models for are AA and AB stacked BLG.
+AA stacked BLG is simply two sheets of graphene placed on top of one another where each atom is aligned with the atom beneath it.
+In AB stacked BLG, a carbon atom from the lower layer is present in the center of each carbon hexagon of the upper layer, and vice versa.
+As such, one can think of AB BLG as a shifted version of AA BLG, where the shift is half the size of a single carbonic hexagon.
 
 Being the simplest BLG configuration, these models will form a starting point for a more thorough development of models for TBLG.
+The ultimate goal of developing a model for TBLG would be a model which has the parameters as a function of the twist angle.
+The twist angle changes the local chemistry of a particular atom in a single layer of graphene by introducing neighbors on the other layer at various non-commensurate distances.
+As such, the first avenue of investigation would be to see how the model parameters change as one shifts the position of interlayer neighbor carbon atoms. 
+The AA and AB BLG configurations provide the simplest systems where this position dependence on model parameters can be probed.
 
 I anticipate that the candidate descriptors should include interlayer couplings as well as the terms seen in SLG.
+As such, a candidate model that we may consider initially for BLG would be
+\begin{equation}
+H_\text{BLG} = H_\text{SLG}^{(1)} + H_\text{SLG}^{(2)} - \sum_{i_1, i_2} t_{i_1, i_2}^\prime c_{i_1}^\dagger c_{i_2} + h.c.
+\label{Hblg}
+\end{equation}
+Here we have two models for the single layers from \eqref{Hslg} plus an interlayer hopping given by the term $t_{i_1, i_2}^\prime.$
+I suspect that the hopping values will be very different between the AA and AB models as the distance between interlayer neighbors differs greatly between the two stackings.
 
 I will again use CVMC to generate the low-energy wave functions for the model regression.
+As before, I think that an MSJ parameterization will serve the purpose well.
+Similarly, the determinants used will include a DFT ground state alongside all excitations within the $\pi, \pi^*$ orbitals of the system.
+Note that excitations between the $\pi, \pi^*$ orbitals from both sheets should be included in the multi-Slater expansion.
+A three-body Jastrow factor will be used for the reasons listed before.
 
-The final four models can be used to investigate the transferrability of model parameters between these different carbon-based systems.
+The final product of my thesis would be four models for these increasingly complex systems.
+The model for benzene should reproduce previous results and provide an intro calculation for me, as well as a benchmark test for the CVMC sampling method.
+The model for SLG will extend on previous studies, incorporating long range Coulomb interactions.
+The models for BLG will be a first step in understanding how interlaying coupling depends on interlayer neighbor distances in the bilayer systems, an important stepping stone towards a general model for TBLG.
 
-I will conclude by briefly discussing potential avenues of further study in using DMD to develop model Hamiltonians for TBLG.
+Apart from their individual value, the four models can be used to investigate the transferrability of model parameters between these different carbon-based systems.
+The first question would be whether the parameter values between a finite size system, like benzene, match those seen in the bulk SLG system.
+The second would be whether the intralayer parameter values change in each sheet of SLG when they are brought together in the AA and AB configuration.
+The latter question hints at a broader question, whether the SLG intralayer parameters are transferrable to TBLG, especially near the magic twist angles whre interlayer coupling becomes very large.
 
-My proposed plan involves building a sequence of model Hamiltonians for increasingly complicated systems using DMD, working towards a model for TBLG.
-I will begin with the basic building block of SLG, the benzene molecule, as an exercise to become familiar with the structure and chemistry of graphene like materials.
-The benzene molecule consists of a hexagonal carbon ring capped by hydrogen atoms. 
-It shares many features with SLG such as delocalized $\pi$ electrons and long-range charge density-density interactions.
-Alongside the fact that previous DMD work has already developed an accurate model for benzene, this molecule provides a useful playground for building my familiarity with the chemistry and effective interactions in graphene like systems.
-
-The benzene model fitting will also serve as a benchmark study for new model fitting methods developed since the first DMD calculations on it.
-I will be particularly interested in testing a constrained variation Monte Carlo (CVMC) sampling method which allows us to target low-energy states with specified descriptor values.
-This sampling method will be the core sampling method used in all the proposed work, with the calculations on benzene serving as an familiarizing exercise for me.
-The results of my model fit using CVMC can be validated by comparing the effective theory I have built to the many-body model previous developed for benzene.
-
-The next step will be developing a model with long range interactions for SLG, expanding on previous work.
-An on-site effective Hubbard model has already been built using DMD for single-layer graphene with parameters $U/t \sim 2.00$ eV.
-Following the example of the benzene molecule, however, one would expect that longer range density-density interactions would be important in describing SLG more accurately.
-The goal of this step of the project would be to develop a model of the following form:
-\begin{equation}
-\sum_{ij} t_{ij} c_i^\dagger c_j + h.c. +  \sum_i U n_{i\uparrow} n_{i\downarrow} + \sum_{ij} V_{ij} n_i^\dagger n_j
-\end{equation}
-with hopping terms $t_{ij}$, an on-site Hubbard $U$, and long range density-density interaction $V_{ij}$.
-
-The last part of my plan involves developing a model for AA and AB stacked bilayer graphene.
-These stackings are the simplest forms of bilayer graphene as the supercell is much smaller than those for even moderate commensurate twist angles.
-The relative shift in distance among the atoms between the AA and AB stackings will serve as a first step to understand how the interlayer coupling between graphene layers may depend on the distance between atoms in the layers.
-The final model to fit for these cases would be identical to the one above, but with additional interlayer hopping terms.
-
-The end result of my proposed project is four model Hamiltonians for the  benzene molecule, SLG, AA stacked BLG, and AB stacked BLG.
-These four models will include, at least, nearest neighbor hoppings, on-site Hubbard interactions, and long range density-density interactions.
-For the AA and AB stacked BLG we will also consider interlayer couplings starting with just hoppings between the layers.
-Each model will come with fit parameter values, an appropriate single particle basis, and a measure of the goodness of fit for the model.
-These models can be solved and compared to experiment.
-
-With these models we can resolve an interesting question: are model parameters transferable between benzene, SLG and BLG?
-The question of transferability of model parameters between SLG and BLG is quite interesting because it may inform us about how much the behavior of each graphene sheet changes once we stack them.
-If the interlayer coupling does not affect the effective behavior on the single sheets, then one can easily treat the final model for bi-layer graphene in two independent pieces: an accurate model for the graphene sheets with a position dependent interlaying coupling.
-However, if the interlayer coupling plays a large role in the effective theory then it may be impossible to separate the BLG model into these two pieces.
-A similar study can be done for the single particle basis, giving some information about orbital relaxation going from SLG to BLG.
-
-My work will provide a useful baseline to begin studies on TBLG.
-The approach to TBLG is still very far away and will require all of the previous steps above. 
-Afterwards, work on TBLG will require building effective theories of BLG with multiple twist angles, or at least varying local distances between atoms.
-The goal here is to develop models for BLG where the parameters, like interlayer coupling, are defined continuously as a function of atomic distance.
-The resulting interacting lattice model can then be solved over very large supercell sizes and a second step of downfolding can be done.
-The goal of this second step would be to downfold the lattice model into a continuum model over the length scale of the Moire patterns seen in small twist angle BLG.
-Work on TBLG would be highly exploratory, and the steps detailed here are very speculative, outlining a possible avenue of work continuing after my thesis.
+I conclude by briefly discussing potential avenues of further study in using DMD to develop model Hamiltonians for TBLG. 
+One would have to begin by building atomic length scale lattice models for TBLG using DMD, which itself is a monumental task.
+When dealing with small commensurate twists, the size of the Moire supercell is enormous and prohibitive to QMC calculations.
+As such, a computationally feasible method for probing small shifts, either by translation or angle, between the graphene sheets must be constructed.
+This short length scale Hamiltonian can then be downfolded again, using DMD, to a continuum Hamiltonian on the length scale of the Moire superlattice.
+This kind of downfolding will require using large-scale tight binding codes like \textbf{ref}.
+Iterating over various twist angles can then give insight into the behavior of the long-range effective theory of TBLG as a function of the twist angle.
 
 \subsection{Proposed timeline}
-I believe that developing effective theories for the half-filled Hubbard model should take 2 years. 
-16 months should be spent on T = 0 calculations.
-The first 8 months will be focused on exact diagonalization.
-In this period, the first two months would be spent developing an efficient sampling scheme for $\mathcal{LE}$.
-The following three months will be geared towards DMD for increasing system sizes from 2x2 to 4x4 lattices for U = 2, 12 in order to identify the existence of multiple zero temperature phases and will be primarily exploratory.
-The next three months will be focused on studying intermediate U and the transition between the two phases, if they exist.
-The following 8 months will be focused on finite size extrapolation.
-The first two months here will be dedicated to extending our ED sampling scheme to VMC.
-The following six months will involve extending our finite size results to the thermodynamic limit.
-At the end of the first 16 months I should have a phase diagram at T = 0 with effective theories at various U.
-The last 8 month chunk will be dedicated to the T $>$ 0 model fitting.
-The first four months will be dedicated to exact diagonalization calculations and should be a relatively simple extension of the T = 0 model fitting with samples drawn according to an additional importance criteria $e^{-\beta E}$.
-The last four months will be focused on extending our T = 0 VMC scheme to T $>$ 0 and conducting finite size extrapolation.
-At the end of the 2 years I should have a phase diagram in the U - T space with effective theories at various values of U, T and an understanding of the phases and transitions between them.
+I believe that developing effective theories for the benzene molecule, SLG, AA and AB stacked BLG should take 2 years. 
+The model fitting for benzene should take no more than four months.
+The goal is to reproduce prior DMD calculations for the same material while benchmarking and familiarizing myself with the new CVMC sampling technique that I will use extensively in my project.
+The SLG model fitting should take another eight months.
+This is an extension of a previous work, and the first three months will be dedicate to reproducing the results from that work, namely a single band on-site Hubbard model.
+Inclusion of long range Coulomb interactions into the effective theory should take another three months.
+The last two months will be geared towards finite size extrapolation in order to ensure our final model is applicable to bulk systems.
+The last year of the project will be geared towards the AA and AB stacked BLG, with each project taking about six months to complete.
+For each stacking, I expect four months will be necessary to build an effective theory, and the final two months will be dedicated to finite size extrapolation.
 \end{document}

### 2019-07-28 draft.tex
finished intro

@@
 
-I propose using a quantum Monte Carlo (QMC) density matrix downfolding (DMD) procedure to develop accurate many-body effective theories for single (SLG) and bi-layer graphene (BLG) systems as a step towards building a many-body model for the strongly correlated TBLG system.
-DMD is a relatively new method which allows one to systematically downfold the \textit{ab-initio} Hamiltonian for a system to a many-body lattice model.
-While it can be applied to any system, the methods necessary to carry out the downfolding accurately and efficiently have not yet been developed and tested for strong correlated materials.
-Both BLG and SLG provide useful playgrounds for developing methods for DMD since they have moderate interaction effects, with non-interacting theories describing the lowest lying excitations relatively accurately.
-Fitting models for SLG and BLG from DMD will not only provide a useful stepping stone towards understanding TBLG, but will also help mature the DMD method for its application to other strongly correlated systems.
+I propose using a quantum Monte Carlo (QMC) density matrix downfolding (DMD) procedure to systematically develop accurate many-body effective theories for single (SLG) and bi-layer graphene (BLG) systems.
+Recent observations of a fractional quantum Hall effect and plasmarons in SLG as well as unconventional superconductivity in TBLG have spurred an interest in understanding many-body interactions in graphene systems.
+Various attempts at characterizing and quantifying these interactions have been made, but the methods used to construct models typically depend on fitting to experimental data or making strong assumptions about electronic interactions, and usually do not come with a measure of their accuracy.
+The DMD method provides a structure for constructing a model Hamiltonian using only the \textit{ab-initio} Hamiltonian of the system, with the resultant model accompanied with a statistical measure of its ability to accurately describe the low-energy excitations of the system.
+While not considered strongly correlated systems themselves, building models for SLG and BLG using DMD will provide a useful stepping stone towards understanding TBLG and also help mature the DMD method for its application to other strongly correlated systems.
 

### 2019-07-29 draft.tex
cleaned preliminary work a lot based on CuO paper edits I just made. Need to cut 18 -> 13 pages. Planning 3, 6, 4 pages for intro, preliminary work, and proposed work. or 4, 5, 4, dunno

@@
 DMC provides a stochastic implementation of this projector method and involves moving samples from the trial function $\Psi_T(R)$ using the Green function $G(R, R^\prime, \tau) = \langle R | e^{-\tau(\hat{H} - E_T)} | R^\prime \rangle$.
-Since $\hat{H} = \hat{T} + \hat{V}$, kinetic and potential energy terms, the Green function is approximated by a Trotter expansion $G(R, R^\prime, \tau) = \langle R | e^{-\tau(\hat{H} - E_T)} | R^\prime \rangle \sim \Big[e^{-d\tau(\frac{V(R) + V(R^\prime)}{2} - E_T)} \langle R| e^{-d\tau\hat{T}}|R^\prime \rangle + O(d\tau^2) \Big]^N $ where $d\tau = \tau/N$.
+Since $\hat{H} = \hat{T} + \hat{V}$, kinetic and potential energy terms, the Green function is approximated by a Trotter expansion 
+$$G(R, R^\prime, \tau) = \langle R | e^{-\tau(\hat{H} - E_T)} | R^\prime \rangle \sim \Big[e^{-d\tau(\frac{V(R) + V(R^\prime)}{2} - E_T)} \langle R| e^{-d\tau\hat{T}}|R^\prime \rangle + O(d\tau^2) \Big]^N $$
+where $d\tau = \tau/N$.
 This expansion can be interpreted as an interative procedure where samples are moved N times with a small timestep $d\tau$ until convergence.
@@
 In order to become acquainted with our code QWalk \cite{WAGNER20093390} and QMC algorithms in general I worked on implementing and testing multi-Slater-Jastrow trial functions with optimized non-orthogonal determinants (MSJ+NO) in FN-DMC \cite{Pathak2018}.
+The MSJ+NO trial wave functions take the form
+\begin{equation}
+\Psi=e^{J(\vec{\alpha})} \sum_I e^{\hat{\Theta}(\vec{\theta_I})} C_I |D_I (\{ \phi\})\rangle
+\end{equation}
+where the determinants $|D_I\rangle$ are composed of single particle orbitals from a self-consistent field calculation, $e^J$ is a spin-independent Jastrow factor, $e^{\hat{\Theta}}$ is a spin-dependent orbital rotation matrix, and the wave function has variable parameters $\vec{\alpha}, \vec{\theta}, \vec{C}$ which can be optimized.
 We assessed the efficiency and compactness of this new trial function by calculating the ground state energy and single particle densities of a C$_2$ molecule using FN-DMC and comparing to the results when using multi-Slater-Jastrow trial functions with optimized orthogonal determinant trial functions (MSJ+O).
 The workflow involved constructing the un-optimized trial wave functions, optimizing the parameters using an energy optimization method \cite{Toulouse2007}, and finally using the optimized trial functions in both variational Monte Carlo (VMC) and FN-DMC calculations.
+
 We found the average improvement in variational energy when using optimized orthogonal determinants is $\langle E_\text{MSJ} - E_\text{MSJ+O} \rangle = $ 0.32 eV with a standard deviation of 0.08 eV, with an additional improvement from the non-orthogonal determinants of 0.031 eV with a standard deviation of 0.021 eV.
@@
 \subsection{Effective theory for CuO molecule using DMD}
-As a first step towards developing accurate models for extended systems like solids using DMD, we constructed a many-body effective model for the CuO molecule which accurately describes the eigenstates and spectra seen in experiment up to 2eV above the ground state.
+As a first step towards developing accurate models for extended systems like solids using DMD, I constructed a many-body effective model for the CuO molecule which accurately describes the lowest-lying eigenstates seen in experiment.
 The neutral CuO molecule has been a subject of intense theoretical and experimental study due to the complex structure of its low-energy space.
-While the low-energy excitations in the doublet sector of the CuO molecule are well understood, singlet selection rules for spectroscopic measurements and the doublet ground state leave the quartet sector of excitations unexplored.
-We construct our $\mathcal{LE}$ based on recent anion photoelectron spectroscopy (APES) measurements of the low-lying excited states of the nuetral CuO molecule. 
-The lowest lying excitations involve electrons in the Cu 3d, Cu 4s and O 2p orbitals.
-Given that the lowest energy excitations contain a fully-filled Cu 3d shell or a single hole in this shell we define our low-energy space as:
+Unlike lighter transition metal oxide molecules like TiO or FeO, the energy scales related to the active Cu 3d, 4s and O 2p orbitals are very similar.
+This leads to many low-energy eigenstates below $\sim $ 3eV which have radically different properties.
+As such, no accurate low-energy effective model for the molecule has been constructed successfully.
+
+In the following four sections I will detail the steps I took towards developing a model for the CuO molecule using DMD.
+First, I discuss how low-energy states were sampled for the model regression.
+This is followed by details of descriptor selection and single particle basis construction.
+With the sampled data and candidate descriptors in hand, the feature selection and model fitting procedures are described.
+I will conclude by comparing the solutions of the model Hamiltonian to experiment and previous theoretical calculations.
+
+\subsubsection{Sampling low-energy states}
+For the CuO molecule, low-energy states were generated by using FN-DMC projected multi-Slater-Jastrow trial wave functions.
+The trial wave functions took the form
 \begin{equation}
-\mathcal{LE} = \text{Span(}\{ |\Psi \rangle | \langle \Psi | \hat{n}_{3d} | \Psi \rangle \ge 9,\ \hat{H}_\text{ab}|\Psi\rangle = E |\Psi\rangle \}\text{)}.
-\label{eq:LE}
-\end{equation}
-
-Our sample space was generated by using multi-Slater-Jastrow trial wavefunctions in FN-DMC.
-The parameterization for our low-energy trial wave functions was
-\begin{equation}
-e^{J}\sum_{i} c_i|\text{D}_i\rangle = \sum_{i} c_i (e^J |\text{D}_i \rangle) \xrightarrow{FN-DMC} \ \sim \sum_i c_i |\Phi_i \rangle \in \mathcal{LE}
+|\Psi_T(\vec{p}) \rangle =  e^{J}\sum_{j} p_j|\text{D}_j\rangle,
 \label{eq:sampling}
 \end{equation}
-where $e^J$ is a three-body Jastrow factor and $|\text{D}_i\rangle$ is a Slater determinant whose nodes accurately approximate those of an eigenstate $|\Phi_i \rangle \in \mathcal{LE}$.
-Under the FN-DMC projection these trial wave functions map to states which are very close to linear combinations of eigenstates in $\mathcal{LE}$.
-We used symmetry-targeted unrestricted Kohn-Sham (UKS) to generate our determinants $|\text{D}_i\rangle$ using a B3LYP functional, a Trail-Needs pseudopotential, and the VTZ Trail-Needs basis using the package PySCF.
-This method allows us to access almost every excitation in $\mathcal{LE}$, except the few cases when two states have identical S$_z$ and number of electrons per irrep.
-An example would be the ground state and the state with an excitation from Cu 3d$_{xz} \rightarrow $  O 2p$_x$, where the latter would be inaccessible.
-The FN-DMC calculations were conducted with T-moves to make the calculation variational while using pseudopotentials, and a timestep of $\tau = 0.01$.
-The coefficients $\{c_j\}$ for our sample states were chosen via a shell-sampling method. 
-We begin by fixing the coefficient $c_0 = \sqrt{w}$ where $w \in \{1.0, 0.8,..., 0.2\}$. 
-For each choice of $w$ we sample n = 5 states by randomly selecting the unassigned coefficients from a uniform distribution such that $\sum_i c_i^2 = 1$. 
-We can then loop over all $i$ to generate shells of decreasing distance near each eigenstate $|\Phi_i\rangle$.
+where $e^J$ is a Jastrow factor, $|\text{D}_j\rangle$ are selected determinants which we discuss below, and $p_j$ are variable coefficients.
+A single three-body Jastrow factor was used for all sampled states optimized on the lowest energy determinant $|\text{D}_0 \rangle$.
+The choice to use a single Jastrow factor significantly reduced the computational cost of the DMD method and is justified by noting that large variations in the Jastrow factor would move the state outside the low-energy space.
 
-Parameterizing our effective model requires selecting a set of descriptors from which we would like to build our model and a basis on which to build our descriptors.
-Occupation energies of and hybridization between Cu 3d, Cu 4s and O 2p are necessary to describe the low-energy excitations in CuO and are best represented by the occupation energies in a molecular orbital basis (MO) which would neatly package both effects.
-The lack of doubly occupied Cu 4s states in the low-energy spectrum of CuO and the existence of a large Hund's coupling between the Cu 4s and 3d orbitals in the Cu atom indicate the necessity for a Hubbard U$_s$ and Hund's J$_{sd}$ respectively.
-These two interaction terms are best represented in a localized intrinsic atomic orbital (IAO).
-To account for orbital relaxation between excited states in the CuO molecule we will consider including additional hybridization between MOs.
-In total our space of canditate models is then 
+The determinants $|\text{D}_j \rangle$ were constructed from the orbitals of independent symmetry-targeted unrestricted Kohn Sham (UKS) calculations.
+This choice was made to ensure that the three major low-energy variations seen in the molecule can be captured in our samples: occupancy, relaxation and spin configuration of the Cu 3d, 4s and O 2p orbitals.
+The symmetry-targeting allows us to select the number of electrons in both the spin-up and spin-down channel for a given irreducible representation of the symmetry group for CuO, letting us generate determinants with variations in both total spin and orbital occupation.
+Orbital relaxation effects are captured by having each determinant, corresponding to a unique orbital and spin configuration, be the result of an independent self-consistent field calculation.
+The UKS calculations were done using PySCF, a B3LYP functional, Trail-Needs psuedopotential and VTZ Trail-Needs basis.
+
+Of these UKS determinants, only those which contribute to accurately describing the low-lying excitations of the molecule are selected.
+All determinants with at most a single hole in the Cu 3d shell and no doubly occupied Cu 4s orbital were included in \eqref{eq:sampling} to reflect the low-energy excitations observed in experiment.
+Only the lowest energy state with a Cu 4s$^2$ electronic configuration was included, corresponding to a double excitation 2 O ${2p_\pi} \rightarrow$ 2 Cu 4s with an energy $>7$ eV above the UKS ground state.
+Inclusion of this high energy determinant is necessary to ensure our model places similar Cu 4s$^2$ states high in energy.
+No other Cu 4s$^2$ determinants were included since we were not interested in describing the variation among these high energy states.
+
+A shell sampling scheme was used to sample states from the selected approximate low-energy space.
+We began by fixing the coefficient $p_0 = \sqrt{w}$ where $w \in \{1.0, 0.8, ..., 0.2\}$. 
+For each choice of $w$ we sampled n = 5 states by randomly selecting the unassigned coefficients from independent, standard normal distributions such that $\sum_j p_j^2 = 1$. 
+This procedure generated a uniform set of samples from spherical shells at decreasing radii in determinant space centered at $|\text{D}_0\rangle$.
+We repeated this process centered at every determinant in \eqref{eq:sampling} resulting in a sample set of overlapping shells which fill the space $\mathcal{LE}^\prime$.
+
+The final sampled states were projected using FN-DMC to ensure they are as low in energy as possible.
+The FN-DMC calculations were conducted with T-moves to ensure the method is variational while using pseudopotentials.
+A timestep of $\tau = 0.01$ was used.
+
+\subsubsection{Selecting candidate descriptors}
+We select the following set of 1- and 2-body operators to construct our candidate descriptors based on our understanding of the low-energy excitations of the CuO molecule and our sampled states:
 \begin{equation}
 \begin{split}
-\{\bar{n}_{4s}, \bar{n}_{p_\pi}, \bar{n}_{p_z}, \hat{n}_{4s\uparrow} \hat{n}_{4s\downarrow},\sum_{i \in \{xy, xz,...\}}\vec{S}_{4s}\cdot \vec{S}_{d_i}\} \ + \\
-\text{P}(\bar{c}_{d_\pi}^\dagger \bar{c}_{p_\pi} + h.c.,\ \bar{c}_{d_z^2}^\dagger \bar{c}_{p_z} + h.c.,\ \bar{c}_{4s}^\dagger \bar{c}_{p_z} + h.c.,\ \bar{c}_{d_z^2}^\dagger \bar{c}_{4s} + h.c.)
+\sum_{\sigma = \uparrow, \downarrow} \Bigg(\epsilon_{3d_\pi}\hat{n}_{3d_\pi,\sigma} + \epsilon_{3d_{z^2}}\hat{n}_{3d_{z^2},\sigma} +  \epsilon_{2p_z} \hat{n}_{2p_z,\sigma} + \epsilon_{2p_\pi}\hat{n}_{2p_\pi,\sigma} + \epsilon_{4s}\hat{n}_{4s,\sigma} +\\
+t_\pi \hat{c}_{3d_\pi,\sigma}^\dagger \hat{c}_{2p_\pi,\sigma} + t_{dz} \hat{c}_{3d_{z^2},\sigma}^\dagger \hat{c}_{2p_z,\sigma}  + t_{sz}\hat{c}_{4s,\sigma}^\dagger \hat{c}_{2p_z,\sigma} + t_{sd}\hat{c}_{4s,\sigma}^\dagger \hat{c}_{3d_{z^2},\sigma} + \text{h.c.} + \Bigg)  \\
+J_{sd}\sum_{i\in {\{\text{xy, xz, ...}}\}} \vec{S}_{4s} \cdot \vec{S}_{3d_i} + U_s \hat{n}_{4s,\uparrow}\hat{n}_{4s,\downarrow} + E_0.
 \end{split}
@@
 \end{equation}
-where P denotes the power set and operators are defined on the IAO basis unless denoted by a bar, in which case the operator is built on the MO basis. 
-The coefficients for each term above will be denoted $\bar{\epsilon}_{4s},\ \bar{\epsilon}_{p_\pi},\ \bar{\epsilon}_{p_z},\ U_s,\ J_{sd},\ \bar{t}_\pi,\ \bar{t}_{dz},\ \bar{t}_{sz},\ \bar{t}_{ds}$. 
-All energies will be relative to $\bar{\epsilon_{3d}} = 0$. 
-The symbol $\pi$ denotes a contraction over $x, y$ for p orbitals and $xz,\ yz$ for d orbitals. The symbol $\delta$, introduced later, denotes a contraction over $xy,\ x^2-y^2$ for d orbitals.
+The operator subscripts correspond to the intrinsic atomic orbital (IAO) basis elements constructed using the molecular orbitals in \eqref{eq:sampling}, and can be seen in \ref{fig:cuo_regr}a.
+Operators are contracted over symmetry-related pairs for the $3d_\pi, 2p_\pi$ orbitals, for example $\hat{n}_{3d_\pi} = \hat{n}_{3d_{xz}} + \hat{n}_{3d_{yz}}$.
 
-After fitting each of the potential models in \eqref{eq:models} using ordinary least squares (OLS) linear regression we find three models which describe the energy functional on our sample set accurately with low variance, but whose eigenstates and spectra differ as a consequence of intruder states below an energy of 2eV as illustrated in Figure \ref{fig:InitED}.
-All three models have a first excited state around 1eV, but only in the minimal model does this state fall near the sample set used for the fitting.
-The large distance of the eigenstates for the other two models from our sample set indicates that their predicted eigenvalues may suffer from large extrapolation errors, and as such we will call them \textit{intruder} states.
-In Figure \ref{fig:Intruder} we present all eigenstates among the three potential models which we classify as intruders using a k-nearest neighbor approach with k = 5.
+The parantheses of \eqref{eq:models} contain the full set of symmetry allowed 1-body operators within the Cu 3d, 4s and O 2p space.
+These operators are required to describe the most basic single particle features seen in molecules like orbital occupations and hybridizations.
+Two 2-body interaction terms are also considered: a Hund's coupling between Cu 4s and 3d orbitals and a Cu 4s Coulomb repulsion.
+The prior term is motivated by the existence of a $\sim $ 0.5 eV Hund's coupling on the neutral copper atom.
+Motivation for the latter comes from the FN-DMC energies of our sampled data.
+We find that states with a Cu $3d^{10} 4s^{2}$ configuration lie at least 7 eV above our lowest energy sampled state, while the highest energy states with a $3d^{10} 4s^{1}$ configuration lie near 2 eV.
 
-\begin{figure}[H]
+The operator $\hat{n}_{3d_\delta}$ is not included as it is linearly dependent on the rest of the occupation operators.
+The linear dependency arises from the fixed number of electrons in the model.
+As such, the operator $\hat{n}_{3d_\delta} = N_\text{elec} - \sum_{i \neq 3d_\delta} \hat{n}_i$.
+The first term can be absorbed into the constant $E_0$, and the latter terms shift the occupation energies $\epsilon_i$.
+
+\subsubsection{Fitting the model}
+A direct attempt at fitting all eleven terms in \eqref{eq:models} leads to a high goodness of fit but unexpectedly large coefficients for the fit model.
+The fit model had an R$^2= 0.99$, but the fit coefficients were very inconsistent with experimental measurements and our sampled data.
+For example, the model predicted an enormous occupation energy gap of $\epsilon_{2p_\pi} - \epsilon_{3d_\delta} = $ 5.87 eV, while in experiment and our sampled data this scale was between 2 $-$ 3 eV.
+This inflation of model parameters was found to be associated with multicollinearity among the candidate descriptors, quantified by a large condition number (CN) of 1470.
+
+A principal component analysis (PCR) analysis was used to alleviate this multicollinearity and fit the effective theory.
+PCR can be broken into three parts: conducting a principal component analysis (PCA) and selecting candidate principal components to work with (PCs), fitting a linear model to these PCs, and then inverting the coefficients back to the original descriptors.
+If the subset of PCs are chosen appropriately, the final regressed model will have a similar goodness of fit to one fit on the original data matrix but without the consequence of serious multicollinearity.
+
+From the PCA, the nine highest ranked PCs were selected to use in the PCR.
+In Figure \eqref{fig:cuo_regr}b we plot both the cumulative explained variance ratio (CEVR) and CN of a matrix composed of the $N_\text{comp}$ highest ranked PCs varying from one to eleven.
+The CEVR indicates how much of the variation in the descriptor values a subset of PCs capture, and the CN quantifies the multicollinearity among these PCs.
+The objective in selecting PCs is to the maximize the prior while simultaneously minimizing the latter, ensuring a high goodness of fit with little to no multicollinearity.
+Reducing the number of PCs from eleven rapidly decreases the CN to $\sim $ 200 while keeping the CEVR relatively fixed.
+Further reducing the number of PCs past nine does not affect the CN but steadily decreases the CEVR, indicating that nine principal components best satisfies the objective above with a CEVR 0.9995 and CN of 257.
+
+The model fit to these nine PCs maintains an excellent goodness of fit without serious multicollinearity.
+When trained on all the sampled data this model yields an $R^2$ = 0.98 with a CN of 257. 
+While the goodness of fit does decrease slightly compared to a fit with the full data matrix, the nearly six-fold reduced condition number readily compensates for this loss.
+An inverse PCA transformation was used to map the fit coefficients corresponding to the PC descriptors back into the original eleven coefficients of \eqref{eq:models}.
+These coefficients can be seen in Figure \ref{fig:cuo_regr}c with single standard deviation bootstrap error estimates.
+
+\begin{figure*}
 \centering
-	\subfigure[Results of exact diagonalization showing the energies and properties of the first twenty eigenstates for three candidate models fit using ordinary linear regression.
-The ground state and first excited state are colored and have enlarged markers for clarity.
-Error bars are 95\% CI calculated by a bootstrap estimate.]{
-	 	\label{fig:InitED}
-		\includegraphics[width=0.45\textwidth]{figs/init_ed.pdf}
-	}
-	\qquad
-	\subfigure[Collection of unique intruder states from our three potential models selected using a k-nearest neighbor approach with k = 5.]{	
-		\label{fig:Intruder}
-		\includegraphics[width=0.45\textwidth]{figs/intruder.pdf}
-	}
-	\caption{Results of OLS fitting on sampled data and selected intruder states.}
-\end{figure}
+\includegraphics[width=1.0\linewidth]{./figs/cuo_regr.pdf}
+\caption{Figures related to model regression for the CuO molecule: a) Intrinsic atomic orbitals used as a single particle basis for the model Hamiltonian, b) Results of PCA showing the CEVR and CN of a data set containing N$_{comp}$  highest ranked PCs, c) The final regressed parameters from the PCR.}
+\label{fig:cuo_regr}
+\end{figure*}
 
-Given that our intruder states correspond to states within $\mathcal{LE}$ but outside our sample set, they must live within the span of the eigenstates which were excluded in our sampling scheme as described before.
-This is evidenced by the fact that two of the intruder states correspond explicitly to excitations we could not access: Cu 3d$_{z^2} \rightarrow $ O 2p$_\pi$ and  Cu 3d$_\pi \rightarrow $ O 2p$_\pi$.
-We can approximate these states by single rigid MO excitations above the UKS ground state, which results in states with a FN-DMC energy above 4eV.
-Including a generous 2eV reduction in FN-DMC energy due to orbital relaxation, we estimate that these two states should lie at least 2eV in energy. 
-Since all the other inaccessible eigenstates are double or higher order excitations, we assert a prior: intruder states should lie above 2eV in energy in our effective theories, enforced via an augmented cost function
-\begin{equation}
-\text{Cost} = \sum_{i} (E_\text{eff}[\Psi_i] - E_\text{ab}[\Psi_i])^2 + \lambda \sum_{p}\text{QHL}(2 - E_\text{eff}[\Psi_p]),\ \text{QHL}(x) = \Theta(x)x^2
-\label{eq:cost}
-\end{equation}
-where $\lambda>0$ is a parameter which can be varied, QHL is a quadratic hinge loss, $\Theta$ is the Heaviside step function, the index $i$ is over our sampled states and $p$ over the selected intruder states.
+\subsubsection{Solutions of the model}
+The model is solved by exact diagonalization, and the solutions agree well with experimental spectra and state assignments.
+Shown in Figure \ref{fig:comp} is a comparison between the excitation energies calculated from our model and those seen in the latest anion photoelectron spectroscopy measurements.
+The model energies for these eigenstates are denoted by orange circles and agree with experiment up to a maximum deviation of 0.2 eV.
+The term symbols assigned in experiment also match the term symbols of the model eigenstates, which are constructed by looking at the multiplicity and electronic configuration of each model eigenstate.
 
-\begin{figure}[H]
+\begin{figure*}
 \centering
-	\subfigure[On the left, scores of the three candidate models at various $\lambda$ when fitting our effective theories using the cost function \eqref{eq:cost}.
-On the right we show predicted versus calculated energies for our low-energy sampled states for the minimal model fit at $\lambda = 20$.]{
-	 	\label{fig:Prior}
-		\includegraphics[width=0.45\textwidth]{figs/prior_and_regr.pdf}
-	}
-	\qquad
-	\subfigure[Results of exact diagonalization showing the energies and properties of the first thirty eigenstates for three candidate models using linear regression with cost function \ref{eq:cost} and $\lambda = 20$.
-The ground state and first excited state are colored and have enlarged markers for clarity.
-Error bars are 95\% CI calculated by a bootstrap estimate.]{
-		\label{fig:FinalED}
-		\includegraphics[width=0.45\textwidth]{figs/final_ed.pdf}
-	}
-	\caption{Results of linear regression using augmented cost function \eqref{eq:cost}.}
-\end{figure}
+\includegraphics{./figs/comp.pdf}
+\caption{Comparison of the four lowest lying eigenstates of the CuO moelcule between experiment and three computational methods.
+The model and UKS calculations are from this work.
+Text near the data points corresponds to the molecular term symbols of the eigenstates.}
+\label{fig:comp}
+\end{figure*}
 
-Using our new cost function we find a set of models which accurately describe our sample data, have very similar spectra and eigenvectors to eachother and experiment, and do not contain intruder states below the 2eV prior.
-Shown in Figure \ref{fig:Prior} are the R$^2$ scores and QHL for the three candidate models fit at $\lambda \in [0,20]$. 
-We believe that all three models at $\lambda = 20$ should be considered good models as they satisfy our prior and also explain our sample set accurately. 
-Also shown are $E_\text{eff}, E_\text{ab}$ for our samples using the minimal model at $\lambda = 20$ to illustrate the high fit quality even with a large $\lambda$.
-Figure \ref{fig:FinalED} presents the properties and eigenvalues of the lowest 30 eigenstates for the three fit models at $\lambda = 20$. 
-The additional constraint from asserting a prior has seemingly caused a single effective theory to emerge from our set of candidate models.
-This conclusion is verified by looking at the values of the fit parameters for our three models, presented in Table 1.
-All occupation energies are relative to $\epsilon_{d_\delta}$.
-In the doublet sector the ground and first excited state of any of the three models match experiment, as well as the block of states at 2eV corresponding to excitations out of the Cu 3d shell.
-The lowest energy quartet states are at 1, 2eV respectively and are in line with the observation of a quartet state at 2eV in APES measurements.
-
-\begin{table}[H]
-\begin{center}
-\begin{tabular}{l|lllll}
-Model &$\epsilon_{d_{z^2}}$ & $\epsilon_{d_\pi}$ & $\epsilon_{4s}$ & $\epsilon_{p_\pi}$ & $\epsilon_{p_z}$ \\ \hline 
-Min & 0.32(2)& 0.19(1)& 2.2(4)& 1.7(3)& 0.96(4)\\
-Min $ +\ \bar{c}_{d_\pi}^\dagger \bar{c}_{p_\pi}$& 0.32(1)& 0.10(2)& 2.23(3)& 1.8(2)& 0.97(3)\\
-Min $ +\ \bar{c}_{d_{z^2}}^\dagger \bar{c}_{p_z}$& 0.29(2)& 0.19(1)& 2.19(5)& 1.7(3)& 0.99(5)\\
-\end{tabular} \\
-
-\begin{tabular}{l|llllll}
-Model &$t_\pi$ & $t_{dz}$ & $t_{sz}$ & $t_{ds}$ & $J_{sd}$ & $U_s$ \\ \hline 
-Min &  -0.57(1)& 0.55(3)& 0.87(2)& 0.44(1)& -0.63(9)& 3.8(2)\\
-Min $ +\ \bar{c}_{d_\pi}^\dagger \bar{c}_{p_\pi}$& -0.45(2)& 0.56(2)& 0.89(2)& 0.45(1)& -0.78(7)& 3.7(2)\\
-Min $ +\ \bar{c}_{d_{z^2}}^\dagger \bar{c}_{p_z}$& -0.57(1)& 0.54(2)& 0.85(2)& 0.45(2)& -0.58(3)& 3.9(2)\\
-\end{tabular} \\
-
-\end{center}
-Table 1: Parameters in eV for our three potential models when fit using \eqref{eq:cost} at $\lambda = 20$, energies relative to $\epsilon_{d_\delta} = 0$.
-Error bars are 95\% CI calculated using a bootstrap estimate.
-\end{table}
+The model states also agree with recent theoretical calculations like UKS and CCSD(T).
+The energies for the UKS and CCSD(T) states are shown in Figure \ref{fig:comp} as green and red circles, respectively.
+The UKS energies were calculated in this work in en route to constructing the determinants $|\text{D}_j\rangle$ used in our sampling scheme \eqref{eq:sampling}.
+Our model energies agree with both theoretical techniques within errorbars for the first three states shown.
+The energy of the highest energy $^2\Delta$ is severely over estimated using both UKS and CCSD(T), by nearly 1 eV.
+This over estimation may arise from an inadequate relaxation of the $\sigma$ type orbitals which form a three orbital hybridization complex in this system.
 
@@
 These carbon atoms are linked by sp$^2$ hybridized $\sigma$ bonds and an alternating sequence of $\pi$ bonds from the C 2p$_z$ orbitals.
-Intuitively, the ground state of the system can be seen as a superposition of two configurations of alternating $\pi$ bonds as seen in Figure \textbf{ref} rotated 30$^o$ relative to each other.
+Intuitively, the ground state of the system can be seen as a superposition of two configurations of alternating $\pi$ bonds as seen in Figure \ref{fig:proposed}a rotated 30$^o$ relative to each other.
+
+\begin{figure*}
+\centering
+\includegraphics[width=1.0\linewidth]{./figs/proposed.pdf}
+\caption{Schematic diagrams for the a) benzene molecule, b) AB and AA stacked BLG \textbf{ref}.}
+%Voltage Tunable Plasmon Propagation in Dual Gated Bilayer Graphene
+\label{fig:proposed}
+\end{figure*}
 
@@
 As such, one can think of AB BLG as a shifted version of AA BLG, where the shift is half the size of a single carbonic hexagon.
+Schematic drawings of these two stackings can be seen in Figure \ref{fig:proposed}b.
 

### 2019-07-29 outline.tex
cleaned preliminary work a lot based on CuO paper edits I just made. Need to cut 18 -> 13 pages. Planning 3, 6, 4 pages for intro, preliminary work, and proposed work. or 4, 5, 4, dunno

@@
 
-\paragraph{CuO: Introduction} As a first step towards developing accurate models for extended systems like solids using DMD, we constructed a many-body effective model for the CuO molecule which accurately describes the eigenstates and spectra seen in experiment up to 2eV above the ground state.
+\begin{enumerate}
+\item Intro (no subsection)
+\begin{itemize}
+\item Our objective is to construct a model Hamiltonian for the low-energy electronic excitations of the CuO molecule.
 
-\paragraph{CuO: Methods} In order to fit a model using DMD we had to construct a low energy space (LES), choose a sampling scheme for states within the LES resulting in our sample space (SS), develop a basis which can accurately describe variations within the LES, identify and select descriptors built out of these basis elements, and appropriately fit the data from our SS onto our selected descriptors.
+\item We will construct such a low-energy Hamiltonian by downfolding the non-relativistic \textit{ab-initio} Hamiltonian subject to the Born-Oppenheimer approximation.
 
-\paragraph{CuO: Results and Discussion} The resultant model from our analysis accurately describes the various eigenstates as measured in experiment up to 2eV, and is quantitatively and qualitatively more accurate than competing methods like density functional theory.
+\item It is important to note that the approximations present in (eqref) will prevent our model Hamiltonian from describing effects like bond-length relaxation and spin-orbit splitting.
 
-\paragraph{Hubbard: Preliminary results} Our preliminary calculations on the 2x2 half-filled 1-band Hubbard model map out the limiting behavior in the U-T model space by investigating the effectiveness of the single descriptor models.
-\textcolor{red}{Will be changed based on how far along we get with preliminary results.}
+\item The following four sections are dedicated towards detailing the DMD procedure.
+\end{itemize}
+
+\item Sampling low-energy states
+\begin{itemize}
+\item Sampling low-energy states involves selecting an approximate low-energy space to work with and choosing a scheme for drawing wavefunctions from that space.
+
+\item We suggest two criteria for selecting $\mathcal{LE}^\prime$ so that a model fit on it approximates the target model fit on $\mathcal{LE}$ accurately.
+
+\item First, the approximate space should be able to capture most of the variation among states expected in $\mathcal{LE}$.
+
+\item Second, $\mathcal{LE}^\prime$ should be as low in energy as possible.
+
+\item The sampling scheme can introduce a sampling bias into the final regressed model, and must be carefully chosen.
+
+\item In this work, we consider an approximate low-energy space of FN-DMC projected multi-Slater-Jastrow trial wave functions.
+
+\item The first criteria above can be satisfied by selecting determinants which capture the low-energy variations seen in the CuO molecule.
+
+\item We can capture these variations by using determinants resulting from independent symmetry-targeted unrestricted Kohn Sham (UKS) calculations.
+
+\item Of these UKS determinants, only those which contribute to accurately describing the low-lying excitations of the molecule are selected.
+
+\item The second criteria is met by the FN-DMC projection.
+
+\item A shell sampling scheme was used to sample states from the selected approximate low-energy space.
+
+\item Extra samples were drawn for regions of descriptor space with sparse sample density to ensure a uniform distribution of samples across the space.
+\end{itemize}
+
+\item Selecting candidate descriptors 
+\begin{itemize}
+\item We start by constructing a single particle basis which will be used to express second quantized candidate operators based on two criteria.
+First, the basis elements should be localized, atomic-like, and mutually orthogonal.
+
+\item Second, the trace of the FN-DMC 1-RDM evaluated on the single particle basis should approximately equal the total number of electrons in the system.
+
+\item To satisfy the first criteria above, we construct a set of intrinsic atomic orbitals (IAOs) to use for our single particle basis.
+
+\item We validate the quality of the IAO basis by evaluating the trace of the FN-DMC 1-RDM on the basis over our sampled data.
+
+\item We select the following set of 1- and 2-body operators to construct our candidate descriptors based on our understanding of the low-energy excitations of the CuO molecule and our sampled states
+
+\item The parantheses of \textbf{eqref} contain the full set of symmetry allowed 1-body operators within the Cu 3d, 4s and O 2p space
+
+\item The operator $\hat{n}_{3d_\delta}$ is not included as it is linearly dependent on the rest of the occupation operators
+
+\item Two 2-body interaction terms are also considered: a Hund's coupling between Cu 4s and 3d orbitals and a Cu 4s Coulomb repulsion
+\end{itemize}
+
+\item Fitting the effective Hamiltonian 
+\begin{itemize}
+\item {A direct attempt at fitting all eleven terms in \textbf{eqref} leads to a high goodness of fit but unexpectedly large coefficients for the fit model.}
+
+\item {By studying the condition number (CN) associated with the fit we conclude the large coefficients are symptomatic of multicollinearity between our candidate descriptors.}
+
+\item {Alleviation of this multicollinearity is necessary as we will be interested in inferences related to the coefficients of our final model.}
+
+\item {We approach the fitting using a principal components regression (PCR) to reduce the impact of the multicollinearity among our candidate descriptors.}
+
+\item {Before moving forwards with the PCR we discuss two quantities used in our principal component selection: the cumulative explained variance ratio (CEVR) and CN.}
+
+\item {We conduct a PCA and select the nine highest ranked PCs to use in the PCR.}
+
+\item {A model fit to these nine PCs maintains an excellent goodness of fit without serious multicollinearity.}
+
+\item {An inverse PCA transformation is used to map the fit coefficients corresponding to the PC descriptors back into the original eleven coefficients of (eqref), resulting in our final model.}
+
+\item {We conclude by validating our model using a k-folds cross validation scheme.}
+
+\item {We conduct 100 independent 5-fold cross validation tests for our nine PC model and find no indication of overfitting.}
+\end{itemize}
+
+\item Solutions of effective Hamiltonian
+\begin{itemize}
+\item Solutions of our effective Hamiltonian are obtained by exact diagonalization.
+
+\item Before making comparisons to experiment, it was necessary to assign molecular term symbols to the model eigenstates.
+
+\item We constructed the molecular orbitals (MOs) used for state assignments by diagonalizing the 1-body part of the model Hamiltonian.
+
+\item We briefly discuss the constructed MOs, beginning with the $\delta$ and $\pi$ symmetry orbitals.
+
+\item The three $\sigma$ symmetry MOs do not follow a simple bonding, non-bonding, anti-bonding structure.
+
+\item In Figure \ref{fig:ed} we present the energies and occupations on this MO basis of the lowest-lying model eigenstates.
+
+\item The model eigenstates are assigned term symbols using these MO occupations and spin multiplicities.
+
+\item Equipped with term symbols, we compare our model solutions to experiment and find excellent agreement in state ordering and energies.
+
+\item Note that our model has a larger error in describing high-energy eigenstates of the CuO molecule than low-energy ones.
+
+\item The model states also agree with recent theoretical calculations like UKS and CCSD(T).
+
+\item We believe the overestimation of the $^2\Delta$ excitation energy in UKS arises from an inadequate relaxation of the $\sigma$ orbitals.
+
+\item We end by discussing the existence of a low-energy $^2\Sigma^-$ state for the CuO molecule which has not been observed in photoelectron spectroscopy.
+\end{itemize}
+\end{enumerate}
 

### 2019-07-30 draft.tex
heavy intro rewrite. looks a lot better now

@@
 \section{Introduction}
-Systematically developing model Hamiltonians which can accurately describe the energies and properties of the lowest lying eigenstates for strongly correlated electron systems is a pressing problem in modern condensed matter physics.
-At the forefront of modern research are materials such as the cuprate supercondutors, heavy fermion systems, and most recently, twisted bilayer graphene (TBLG).
-These materials exhibit various phases of matter which are far from understood, like unconventional superconductivity, psuedogap and bad metal phases, quantum critical points, and stripe ordering.
-While no definitive models for these phases have been introduced, it is expected that strong correlations between electronic degrees of freedom lead to these phases.
+Developing model Hamiltonians for realistic materials is a fundamental problem in modern condensed matter physics.
+A model Hamiltonian is a simplified description of a real material which can accurately reproduce the low-lying eigenstates and energies of the material.
+Accurate models give clear and direct indications of effective interactions within a material and underpin modern theoretical understanding of ideas like emergence and symmetry breaking.
+On a practical level, calculations related to these low-lying eigenstates can be carried out on the much simpler model system and still yield accurate results reflective of reality.
 
-The necessity for non-perturbative, many-body models in describing the low-lying excited states of strongly correlated electron systems poses a difficult challenge when developing effective theories for real world systems.
-Consider for now just the case of TBLG, where for certain magic twist angles, bands near the Fermi surface flatten out to a bandwidth of O(10) meV.
-At these magic angles, a small doping can lead to superconductivity and many-body insulating phases in the material, indicative of electronic interaction effects which cannot be explained by a simple band theory.
-The effects of the interaction are also beyond just simple renormalization of the non-interacting parameters in a band theory.
-A non-perturbative, many-body effective theory is therefore necessary to describe these interesting phases in TBLG.
+Yet unresolved is a paradigm for building many-body model Hamiltonians which can describe realistic systems.
+Band and Fermi liquid theory have been successful in accurately modeling systems where effective interactions are negligible at low energy scales.
+These theories, however, fail in describing the low-energy excitations of strongly correlated systems like high Tc superconductors and twisted bilayer graphene (TBLG).
+Many-body models like the Hubbard, Kanamori, t-J and Heisenberg models have 
+been proposed to describe such strongly correlated systems, but their effectiveness in describing realistic materials is unclear.
 
-Common approaches to building suitable model Hamiltonians for strongly correlated materials involve using single particle theories like density functional theory (DFT).
-One-body terms are typically obtained by projecting the solutions from the DFT band structure onto a localized single-particle basis, giving an estimate of hoppings and occupations on a lattice model.
-Two-body terms in the model are developed by assuming a screened Coulomb interaction based on constrained DFT or RPA.
-Aside from this method, L\''{o}wdin methods coupled to a stochastic method and canonical transformations have also been used to develop effective theories.
-These methods have the drawback that the constructed effective theory typically does not come with any estimate of the quality of the model, therefore making model validation a phenomenological task which can lead to unclear biases in the model construction.
+Common approaches to building many-body theories for real materials involve density functional theory (DFT).
+One-body terms are obtained by projecting the solutions from the DFT band structure onto a single-particle basis.
+Two-body terms are developed by assuming a screened Coulomb interaction based on constrained DFT, RPA, or other methods.
+L\''{o}wdin methods and canonical transformations have also been used to develop effective theories.
+While useful techniques, it is typically impossible to know whether models built from them accurately describe reality, as usually no clear estimate of model quality is provided.
 
-I propose using a quantum Monte Carlo (QMC) density matrix downfolding (DMD) procedure to systematically develop accurate many-body effective theories for single (SLG) and bi-layer graphene (BLG) systems.
-Recent observations of a fractional quantum Hall effect and plasmarons in SLG as well as unconventional superconductivity in TBLG have spurred an interest in understanding many-body interactions in graphene systems.
-Various attempts at characterizing and quantifying these interactions have been made, but the methods used to construct models typically depend on fitting to experimental data or making strong assumptions about electronic interactions, and usually do not come with a measure of their accuracy.
-The DMD method provides a structure for constructing a model Hamiltonian using only the \textit{ab-initio} Hamiltonian of the system, with the resultant model accompanied with a statistical measure of its ability to accurately describe the low-energy excitations of the system.
-While not considered strongly correlated systems themselves, building models for SLG and BLG using DMD will provide a useful stepping stone towards understanding TBLG and also help mature the DMD method for its application to other strongly correlated systems.
+I propose using a quantum Monte Carlo (QMC) density matrix downfolding (DMD) procedure to develop accurate many-body model Hamiltonians for single (SLG) and bi-layer (BLG) graphene systems.
+Recent observations of a fractional quantum Hall effect and plasmarons in SLG as well as unconventional superconductivity in TBLG have spurred an interest in developing many-body models for graphene.
+The DMD procedure provides a framework for developing such a many-body model using QMC calculations and starting from the \textit{ab-initio} Hamiltonian 
+\begin{equation}
+\hat{H}_\text{ab} = -\frac{1}{2} \sum_{i} \nabla_i^2 -\frac{1}{2\mu}\sum_{I} \nabla_I^2 + \frac{1}{2}\sum_{ij} \frac{1}{|r_i - r_j|} + \frac{1}{2}\sum_{IJ} \frac{Z_I Z_J}{|R_I - R_J|} - \sum_{iJ}\frac{Z_J}{|r_i - R_J|},\ \mu = (M/m)
+\end{equation}
+where $i$ index electrons, $I$ ions, and energy in Hartree (Ha).
+The procedure makes no assumptions about the form of microscopic electronic interactions in the system and reports a clear, statistical measure of model quality.
+By developing many-body models for SLG and BLG, I hope to further understanding of interactions in graphene like systems and the maturity of DMD for its application to other strongly correlated systems.
 

### 2019-07-30 outline.tex
heavy intro rewrite. looks a lot better now

@@
 \begin{itemize}
-\item Systematically developing model Hamiltonians which can accurately describe the energies and properties of the lowest lying eigenstates for strongly correlated electron systems is a pressing problem in modern condensed matter physics.
+\item Developing model Hamiltonians for realistic materials is a fundamental problem in modern condensed matter physics.
 
-\item The necessity for non-perturbative, many-body models in describing the low-lying excited states of strongly correlated electron systems poses a difficult challenge when developing effective theories for real world systems.
+\item Yet unresolved is a paradigm for building many-body model Hamiltonians which can describe realistic systems.
 
-\item Common approaches to building suitable model Hamiltonians for strongly correlated materials involve using single particle theories like density functional theory (DFT).
+\item Common approaches to building many-body theories for real materials involve density functional theory (DFT).
 
-\item We propose using a DMD procedure to develop accurate many-body effective theories for single (SLG) and bi-layer graphene (BLG) systems as a step towards building a many-body model for twisted bilayer graphene (TBLG).
+\item I propose using a DMD procedure to develop accurate many-body model Hamiltonians for single (SLG) and bi-layer (BLG) systems.
 \end{itemize}

### 2019-07-30 draft.tex
edited + shortened draft. have 1.5 pages for references now.

@@
 \author{Shivesh Pathak}
-\title{Developing effective theories for the square lattice Hubbard model using density matrix downfolding}
+\title{Developing effective theories for single and bilayer graphene using density matrix downfolding}
 
@@
 \begin{abstract}
-A core problem of condensed matter physics is understanding how to systematically develop low-energy effective models for strongly correlated electron systems. 
-We believe that a density matrix downfolding (DMD) method can be used to systematically build effective theories for such systems beginning with a high energy theory like the \textit{ab-initio} Hamiltonian H$_\text{ab}$.
-We present as a preliminary result an effective theory developed using DMD which can accurately reproduce the spectra and properties of the lowest lying excitations of the CuO molecule.
-As a step towards bulk strongly correlated materials, we propose developing effective theories using DMD for the simplest toy system which exhibits strongly correlated behavior: the 1-band square lattice Hubbard model at half-filling for various couplings U and temperatures T.
+A core problem of condensed matter physics is understanding how to develop low-energy effective models for realistic systems.
+I believe that a density matrix downfolding (DMD) method can be used to systematically build effective theories for such systems beginning with a high energy theory like the \textit{ab-initio} Hamiltonian.
+I present as a preliminary result an effective Hamiltonian developed using DMD which can accurately reproduce the spectra and properties of the lowest lying excitations of the CuO molecule.
+I propose using DMD to develop accurate model Hamiltonians for single and bilayer graphene.
+My work will serve to mature the DMD method and take steps towards a model for the strongly correlated twisted bilayer graphene system.
 \end{abstract}
-Committee: Lucas Wagner, ...
+Committee: Lucas Wagner, Smitha Vishveshwara, David Ceperley, Karin Dahmen
+
+Pool: \textit{David Ceperley},\textit{ Brian DeMarco}, Smitha Vishveshwara, Andre Schleife, \textit{Karin Dahmen} 
 \pagebreak
@@
 Recent observations of a fractional quantum Hall effect and plasmarons in SLG as well as unconventional superconductivity in TBLG have spurred an interest in developing many-body models for graphene.
-The DMD procedure provides a framework for developing such a many-body model using QMC calculations and starting from the \textit{ab-initio} Hamiltonian 
+The DMD procedure provides a framework for developing such a many-body model using QMC calculations starting from the \textit{ab-initio} Hamiltonian 
 \begin{equation}
 \hat{H}_\text{ab} = -\frac{1}{2} \sum_{i} \nabla_i^2 -\frac{1}{2\mu}\sum_{I} \nabla_I^2 + \frac{1}{2}\sum_{ij} \frac{1}{|r_i - r_j|} + \frac{1}{2}\sum_{IJ} \frac{Z_I Z_J}{|R_I - R_J|} - \sum_{iJ}\frac{Z_J}{|r_i - R_J|},\ \mu = (M/m)
+\label{Hab}
 \end{equation}
@@
 The procedure makes no assumptions about the form of microscopic electronic interactions in the system and reports a clear, statistical measure of model quality.
-By developing many-body models for SLG and BLG, I hope to further understanding of interactions in graphene like systems and the maturity of DMD for its application to other strongly correlated systems.
+By developing many-body models for SLG and BLG, I hope to further the understanding of interactions in graphene like systems and the maturity of DMD for its application to other strongly correlated systems like TBLG.
 
-\section{Methods}
 \subsection{Density Matrix Downfolding (DMD)}
-We begin by defining what we mean by a low-energy effective Hamiltonian and downfolding.
+I begin by defining what a low-energy effective Hamiltonian and downfolding are.
 Consider a Hamiltonian $\hat{H}$ defined on a Hilbert space $\mathcal{H}$.
@@
 \begin{split}
-\forall |\Psi\rangle \in \mathcal{LE},\ (E[\Psi] \equiv \langle \Psi|\hat{H} | \Psi \rangle)  = \\ (E_\text{eff}[\Psi] \equiv \langle \Psi | \hat{H}_\text{eff} | \Psi \rangle) + \epsilon[\Psi]
+\forall |\Psi\rangle \in \mathcal{LE},\ (E[\Psi] \equiv \langle \Psi|\hat{H} | \Psi \rangle)  = (E_\text{eff}[\Psi] \equiv \langle \Psi | \hat{H}_\text{eff} | \Psi \rangle) + \epsilon[\Psi]
 \end{split}
@@
 In practice, the space is too large, and instead a representative sample set is drawn and used for fitting.
+Generally, one cannot sample the true low-energy subspace, and low-energy wave functions are sampled from an approximate subspace $\mathcal{LE}^\prime$.
 For each sampled state $|\Psi_s\rangle$, the quantities $E[\Psi_s], \{d_k[\Psi_s]\}$ are calculated to be used in the linear regression.
-Optimal sampling schemes vary between systems and can significantly reduce the number of samples required to achieve robust statistics.
-
-Generally, however, one cannot sample the true low-energy subspace and low-energy wave functions are sampled from an approximate subspace $\mathcal{LE}^\prime$.
-A model fit to an arbitrary subspace of $\mathcal{H}$ will generally fail in describing the low-energy excitations of $\hat{H}$.
-Rather, the choice of approximate subspace must be made carefully to ensure the fit model is transferable to the true low-energy space.
-We describe in detail later how we selected $\mathcal{LE}^\prime$ for the CuO molecule and the scheme used to draw samples from it.
+The sampling and calculation are typically done in QMC, but less accurate methods like DFT and Hartree-Fock can be used as well, resulting is less accurate models.
 
@@
 In principle there is no restriction on the form of the candidate operators other than Hermiticity, but second quantized operators are commonly used.
-
-Typically, candidate operators are written in second quantized form, and an appropriate single particle basis to express these operators must be constructed.
-An appropriate choice of single particle basis can simplify the representation of the effective theory and typically depends on the system of interest.
-Common choices include Bloch waves and molecular orbitals for delocalized basis elements or Wannier and intrinsic atomic orbitals for localized ones.
-The selected single particle basis can be validated by ensuring it accurately describes single particle properties among the sampled states like the total number of electrons in the system.
+If so, an appropriate single particle basis to express the second quantized operators must be constructed.
 
@@
 
-We take advantage of the variational nature of FN-DMC to sample the low-energy states necessary for DMD.
+One can take advantage of the variational nature of FN-DMC to sample the low-energy states necessary for DMD.
 Consider a set of trial wave functions with varying nodal structures.
@@
 If the initial trial wave functions are appropriately chosen, the final projected states will be low-energy states within $\Psi_\text{DMC}$.
-We calculate the expectation values $E[\Psi_\text{DMC}], \{d_k[\Psi_\text{DMC}]\}$ necessary for the model regression using a mixed estimator.
+The expectation values $E[\Psi_\text{DMC}], \{d_k[\Psi_\text{DMC}]\}$ necessary for the model regression are calculated using a mixed estimator.
 Details for calculating reduced density matrix elements in FN-DMC can be found in Wagner.
@@
 where the determinants $|D_I\rangle$ are composed of single particle orbitals from a self-consistent field calculation, $e^J$ is a spin-independent Jastrow factor, $e^{\hat{\Theta}}$ is a spin-dependent orbital rotation matrix, and the wave function has variable parameters $\vec{\alpha}, \vec{\theta}, \vec{C}$ which can be optimized.
+
 We assessed the efficiency and compactness of this new trial function by calculating the ground state energy and single particle densities of a C$_2$ molecule using FN-DMC and comparing to the results when using multi-Slater-Jastrow trial functions with optimized orthogonal determinant trial functions (MSJ+O).
 The workflow involved constructing the un-optimized trial wave functions, optimizing the parameters using an energy optimization method \cite{Toulouse2007}, and finally using the optimized trial functions in both variational Monte Carlo (VMC) and FN-DMC calculations.
-
 We found the average improvement in variational energy when using optimized orthogonal determinants is $\langle E_\text{MSJ} - E_\text{MSJ+O} \rangle = $ 0.32 eV with a standard deviation of 0.08 eV, with an additional improvement from the non-orthogonal determinants of 0.031 eV with a standard deviation of 0.021 eV.
 The average improvement in the FN-DMC energy with the MSJ+O trial wave functions was 0.14 eV (standard deviation 0.03 eV) with an additional reduction of 0.032 eV (standard deviation 0.019 eV) when using the MSJ+NO trial wave function.
+
 While the benefit using the MSJ+NO trial wave function seems small, we saw that the FN-DMC energy calculated using an MSJ+NO trial function with only 24 determinants was lower than the FN-DMC energy using an MSJ+O trial function with 55 determinants.
@@
 \subsection{Effective theory for CuO molecule using DMD}
-As a first step towards developing accurate models for extended systems like solids using DMD, I constructed a many-body effective model for the CuO molecule which accurately describes the lowest-lying eigenstates seen in experiment.
-The neutral CuO molecule has been a subject of intense theoretical and experimental study due to the complex structure of its low-energy space.
-Unlike lighter transition metal oxide molecules like TiO or FeO, the energy scales related to the active Cu 3d, 4s and O 2p orbitals are very similar.
-This leads to many low-energy eigenstates below $\sim $ 3eV which have radically different properties.
-As such, no accurate low-energy effective model for the molecule has been constructed successfully.
-
+As a first step towards developing accurate models for extended systems like solids using DMD, I constructed a many-body effective model for the CuO molecule by downfolding \eqref{Hab} under a Born-Oppenheimer approximation with fixed bond length $r_e$ = 1.725\r{A}.
 In the following four sections I will detail the steps I took towards developing a model for the CuO molecule using DMD.
@@
 
-\subsubsection{Sampling low-energy states}
-For the CuO molecule, low-energy states were generated by using FN-DMC projected multi-Slater-Jastrow trial wave functions.
+\paragraph{Sampling low-energy states}
+For the CuO molecule, low-energy states were generated by using FN-DMC projected multi-Slater-Jastrow (MSJ) trial wave functions.
 The trial wave functions took the form
@@
 A single three-body Jastrow factor was used for all sampled states optimized on the lowest energy determinant $|\text{D}_0 \rangle$.
-The choice to use a single Jastrow factor significantly reduced the computational cost of the DMD method and is justified by noting that large variations in the Jastrow factor would move the state outside the low-energy space.
+This significantly reduced the computational cost of the DMD method and is justified by noting that large variations in the Jastrow factor would move the state outside the low-energy space.
 
 The determinants $|\text{D}_j \rangle$ were constructed from the orbitals of independent symmetry-targeted unrestricted Kohn Sham (UKS) calculations.
-This choice was made to ensure that the three major low-energy variations seen in the molecule can be captured in our samples: occupancy, relaxation and spin configuration of the Cu 3d, 4s and O 2p orbitals.
-The symmetry-targeting allows us to select the number of electrons in both the spin-up and spin-down channel for a given irreducible representation of the symmetry group for CuO, letting us generate determinants with variations in both total spin and orbital occupation.
-Orbital relaxation effects are captured by having each determinant, corresponding to a unique orbital and spin configuration, be the result of an independent self-consistent field calculation.
+This choice was made to ensure that the three major low-energy variations seen in the molecule can be captured in our samples: occupancy, spin configuration and relaxation of the Cu 3d, 4s and O 2p orbitals.
+All determinants with at most a single hole in the Cu 3d shell and no doubly occupied Cu 4s orbital were included in \eqref{eq:sampling} to reflect the low-energy excitations seen in experiment.
+Only the lowest energy state with a Cu 4s$^2$ electronic configuration was included, corresponding to a double excitation 2 O ${2p_\pi} \rightarrow$ 2 Cu 4s with an energy $>7$ eV above the UKS ground state.
+Inclusion of this high energy determinant is necessary so our model places similar Cu 4s$^2$ states high in energy.
 The UKS calculations were done using PySCF, a B3LYP functional, Trail-Needs psuedopotential and VTZ Trail-Needs basis.
-
-Of these UKS determinants, only those which contribute to accurately describing the low-lying excitations of the molecule are selected.
-All determinants with at most a single hole in the Cu 3d shell and no doubly occupied Cu 4s orbital were included in \eqref{eq:sampling} to reflect the low-energy excitations observed in experiment.
-Only the lowest energy state with a Cu 4s$^2$ electronic configuration was included, corresponding to a double excitation 2 O ${2p_\pi} \rightarrow$ 2 Cu 4s with an energy $>7$ eV above the UKS ground state.
-Inclusion of this high energy determinant is necessary to ensure our model places similar Cu 4s$^2$ states high in energy.
-No other Cu 4s$^2$ determinants were included since we were not interested in describing the variation among these high energy states.
 
@@
 This procedure generated a uniform set of samples from spherical shells at decreasing radii in determinant space centered at $|\text{D}_0\rangle$.
-We repeated this process centered at every determinant in \eqref{eq:sampling} resulting in a sample set of overlapping shells which fill the space $\mathcal{LE}^\prime$.
+We repeated this process centered at every determinant in \eqref{eq:sampling} resulting in a sample set of overlapping shells which fill $\mathcal{LE}^\prime$.
 
-The final sampled states were projected using FN-DMC to ensure they are as low in energy as possible.
-The FN-DMC calculations were conducted with T-moves to ensure the method is variational while using pseudopotentials.
+The final sampled MSJ states were projected using FN-DMC to generate our low-energy states.
+The FN-DMC calculations were conducted with T-moves to ensure the variational principle holds while using pseudopotentials.
 A timestep of $\tau = 0.01$ was used.
 
-\subsubsection{Selecting candidate descriptors}
-We select the following set of 1- and 2-body operators to construct our candidate descriptors based on our understanding of the low-energy excitations of the CuO molecule and our sampled states:
+\paragraph{Selecting candidate descriptors}
+We selected the following set of 1- and 2-body operators to construct our candidate descriptors based on our understanding of the low-energy excitations of the CuO molecule and our sampled states:
 \begin{equation}
@@
 The operator subscripts correspond to the intrinsic atomic orbital (IAO) basis elements constructed using the molecular orbitals in \eqref{eq:sampling}, and can be seen in \ref{fig:cuo_regr}a.
-Operators are contracted over symmetry-related pairs for the $3d_\pi, 2p_\pi$ orbitals, for example $\hat{n}_{3d_\pi} = \hat{n}_{3d_{xz}} + \hat{n}_{3d_{yz}}$.
 
@@
 
-The operator $\hat{n}_{3d_\delta}$ is not included as it is linearly dependent on the rest of the occupation operators.
-The linear dependency arises from the fixed number of electrons in the model.
-As such, the operator $\hat{n}_{3d_\delta} = N_\text{elec} - \sum_{i \neq 3d_\delta} \hat{n}_i$.
-The first term can be absorbed into the constant $E_0$, and the latter terms shift the occupation energies $\epsilon_i$.
-
-\subsubsection{Fitting the model}
-A direct attempt at fitting all eleven terms in \eqref{eq:models} leads to a high goodness of fit but unexpectedly large coefficients for the fit model.
-The fit model had an R$^2= 0.99$, but the fit coefficients were very inconsistent with experimental measurements and our sampled data.
-For example, the model predicted an enormous occupation energy gap of $\epsilon_{2p_\pi} - \epsilon_{3d_\delta} = $ 5.87 eV, while in experiment and our sampled data this scale was between 2 $-$ 3 eV.
+\paragraph{Fitting the model}
+A direct attempt at fitting all eleven terms in \eqref{eq:models} leads to a high goodness of fit, R$^2$ = 0.99, but unexpectedly large coefficients for the fit model, like an enormous occupation energy gap of $\epsilon_{2p_\pi} - \epsilon_{3d_\delta} = $ 5.87 eV, which is 2 $-$ 3 eV in experiment and our sampled data.
 This inflation of model parameters was found to be associated with multicollinearity among the candidate descriptors, quantified by a large condition number (CN) of 1470.
 
-A principal component analysis (PCR) analysis was used to alleviate this multicollinearity and fit the effective theory.
+A principal component regression (PCR) was used to alleviate this multicollinearity and fit the effective theory.
 PCR can be broken into three parts: conducting a principal component analysis (PCA) and selecting candidate principal components to work with (PCs), fitting a linear model to these PCs, and then inverting the coefficients back to the original descriptors.
@@
 The objective in selecting PCs is to the maximize the prior while simultaneously minimizing the latter, ensuring a high goodness of fit with little to no multicollinearity.
-Reducing the number of PCs from eleven rapidly decreases the CN to $\sim $ 200 while keeping the CEVR relatively fixed.
-Further reducing the number of PCs past nine does not affect the CN but steadily decreases the CEVR, indicating that nine principal components best satisfies the objective above with a CEVR 0.9995 and CN of 257.
+Nine principal components best satisfies the objective above with a CEVR 0.9995 and CN of 257.
 
@@
 When trained on all the sampled data this model yields an $R^2$ = 0.98 with a CN of 257. 
-While the goodness of fit does decrease slightly compared to a fit with the full data matrix, the nearly six-fold reduced condition number readily compensates for this loss.
 An inverse PCA transformation was used to map the fit coefficients corresponding to the PC descriptors back into the original eleven coefficients of \eqref{eq:models}.
@@
 \centering
-\includegraphics[width=1.0\linewidth]{./figs/cuo_regr.pdf}
-\caption{Figures related to model regression for the CuO molecule: a) Intrinsic atomic orbitals used as a single particle basis for the model Hamiltonian, b) Results of PCA showing the CEVR and CN of a data set containing N$_{comp}$  highest ranked PCs, c) The final regressed parameters from the PCR.}
+\hspace*{-0.25in}
+\includegraphics[width=1.1\linewidth]{./figs/cuo_regr_2.pdf}
+\caption{Figures related to model regression for the CuO molecule: a) Intrinsic atomic orbitals used as a single particle basis for the model Hamiltonian, b) Results of PCA showing the CEVR and CN of a data set containing N$_{comp}$  highest ranked PCs, c) The final regressed parameters from the PCR, d) Comparison of four lowest lying eigenstates of CuO molecule between experiment and three computation methods.}
 \label{fig:cuo_regr}
@@
 
-\subsubsection{Solutions of the model}
+\paragraph{Solutions of the model}
 The model is solved by exact diagonalization, and the solutions agree well with experimental spectra and state assignments.
-Shown in Figure \ref{fig:comp} is a comparison between the excitation energies calculated from our model and those seen in the latest anion photoelectron spectroscopy measurements.
+Shown in Figure \ref{fig:cuo_regr}d is a comparison between the excitation energies calculated from our model and those seen in the latest anion photoelectron spectroscopy measurements.
 The model energies for these eigenstates are denoted by orange circles and agree with experiment up to a maximum deviation of 0.2 eV.
@@
 
-\begin{figure*}
-\centering
-\includegraphics{./figs/comp.pdf}
-\caption{Comparison of the four lowest lying eigenstates of the CuO moelcule between experiment and three computational methods.
-The model and UKS calculations are from this work.
-Text near the data points corresponds to the molecular term symbols of the eigenstates.}
-\label{fig:comp}
-\end{figure*}
-
 The model states also agree with recent theoretical calculations like UKS and CCSD(T).
-The energies for the UKS and CCSD(T) states are shown in Figure \ref{fig:comp} as green and red circles, respectively.
+The energies for the UKS and CCSD(T) states are shown in Figure \ref{fig:cuo_regr}d as green and red circles, respectively.
 The UKS energies were calculated in this work in en route to constructing the determinants $|\text{D}_j\rangle$ used in our sampling scheme \eqref{eq:sampling}.
@@
 \section{Proposed work}
-My proposed plan involves building a sequence of model Hamiltonians for increasing complicated systems using QMC DMD, working towards an accurate many-body model for TBLG.
+My proposed plan involves building a sequence of model Hamiltonians for increasingly complicated graphene-like systems using QMC DMD, ending with BLG.
 The systems I will be interested in are the benzene molecule, single layer graphene, AA stacked bilayer graphene and AB stacked bilayer graphene.
-Conceptually, the benzene molecule is a kind of building block for SLG, SLG is a building block for BLG, and BLG for TBLG, creating a sequence of systems from least to most complex.
-I do not believe I will be able to build a general model for TBLG in the time span of my thesis, but my work should lay the foundation for further studies using QMC DMD focused on bilayer models with twists.
+Conceptually, the benzene molecule is a kind of building block for SLG, and SLG is a building block for BLG, creating a sequence of systems from least to most complex.
 
-The first system I will develop a model for is the benzene molecule.
-The benzene molecule has chemical formula $C_6 H_6$ and contains a hexagonal ring of carbon atoms with hydrogen atoms sticking out from each carbon.
-These carbon atoms are linked by sp$^2$ hybridized $\sigma$ bonds and an alternating sequence of $\pi$ bonds from the C 2p$_z$ orbitals.
-Intuitively, the ground state of the system can be seen as a superposition of two configurations of alternating $\pi$ bonds as seen in Figure \ref{fig:proposed}a rotated 30$^o$ relative to each other.
+The first system I will develop a model for is the benzene molecule, shown in Figure \ref{fig:proposed}a.
+While relatively simple, this molecule shares many similarities with and forms the basic unit of SLG.
+The delocalized $\pi$ orbitals in benzene carry over directly into SLG where they are the primary contributors to the well known Dirac bands.
+Interaction effects seen in the benzene molecule are also reflected in SLG, such as on-site double occupation energies for the carbon atoms.
+As such, developing a model for benzene will help me become familiar with model fitting for 2-D carbon based systems.
 
@@
 \end{figure*}
-
-While relatively simple, this molecule shares many similarities with and forms the basic unit of SLG.
-SLG can be viewed as a tiling of benzene molecules where the external hydrogen legs are replaced by carbon atoms.
-The delocalized $\pi$ orbitals in benzene carry over directly into SLG where they are the primary contributors to the well known Dirac bands.
-Interaction effects seen in the benzene molecule are also reflected in SLG, such as on-site double occupation energies for the carbon atoms.
-As such, developing a model for benzene will help me become familiar with model fitting for 2-D carbon based systems.
 
@@
 I will consider at least all the descriptors above and also longer range hopping terms.
-The single particle orbitals used to express these descriptors will be IAOs.
+
+The second system I will be interested in working on is SLG.
+Single layer graphene is composed of a triangular Bravais lattice with two carbon atoms per unit cell.
+In terms of a non-interacting picture, the $\sigma$ orbitals composed of C $2s, 2p_x,$ and $2p_y$ lie below the Fermi level.
+The bands which cross the Fermi level, and compose the distinct Dirac cones seen in SLG, are primary $\pi$ bonded molecular orbitals composed of C $2p_z$ orbitals.
+The lack of significant screening in SLG leads to long range Coulomb interactions in the system.
+
+In particular, I will work to extend a previous on-site Hubbard model developed for SLG by including long range density-density interactions.
+The candidate model will be of the form:
+\begin{equation}
+H_\text{SLG} = -\sum_{i,j} t_{ij}c_i^\dagger c_j + U \sum_i n_{i\uparrow}n_{i\downarrow}  + \sum_{i,j} V_{ij} n_i n_j
+\label{Hslg}
+\end{equation}
+where the indices correspond to the $\pi$ orbitals on the copper atoms.
+I will consider both long range hopping and density-density interaction terms extending across the computational unit cell with the final goal of developing position dependent parameter values $t(r), V(r)$.
+
+The last systems I will develop models for are AA and AB stacked BLG, shown in Figure \ref{fig:proposed}b.
+AA stacked BLG is simply two sheets of graphene placed on top of one another where each atom is aligned with the atom beneath it.
+In AB stacked BLG, a carbon atom from the lower layer is present in the center of each carbon hexagon of the upper layer, and vice versa.
+As such, one can think of AB BLG as a shifted version of AA BLG, where the shift is half the size of a single carbonic hexagon.
+
+I anticipate that the candidate descriptors should include interlayer couplings as well as the terms seen in SLG.
+As such, the candidate model for BLG would be
+\begin{equation}
+H_\text{BLG} = H_\text{SLG}^{(1)} + H_\text{SLG}^{(2)} - \sum_{i_1, i_2} t_{i_1, i_2}^\prime c_{i_1}^\dagger c_{i_2} + h.c.
+\label{Hblg}
+\end{equation}
+There are two terms for the single layers from \eqref{Hslg} plus an interlayer hopping given by the term $t_{i_1, i_2}^\prime.$
+I suspect that the hopping values will be very different between the AA and AB models as the distance between interlayer neighbors differs greatly between the two stackings.
+This difference in value will provide the first steps into developing model parameters for BLG which depend on the relative orientation of the sheets.
 
@@
 
-I will use a multi-Slater-Jastrow (MSJ) parameterization in the CVMC sampling for the benzene molecule.
+I will use a multi-Slater-Jastrow (MSJ) parameterization in the CVMC sampling for each system.
 The multi-Slater expansion will consist of determinants of single particle orbitals calculated using a self consistent field calculation like DFT.
-The determinants will include the DFT ground state as well as molecular orbital excitations among the $\pi$ and $\pi^*$ orbitals which form our active space, allowing us to probe variation in all the descriptors above.
-A three body Jastrow factor will be applied to ensure that we include correlations in our wave function and that our wave function is low in energy.
+The determinants will include the DFT ground state as well as molecular orbital excitations among the $\pi$ and $\pi^*$ orbitals which form the active space.
+A three body Jastrow factor will be applied to ensure that correlations are included in the wave function and that the wave function is low in energy.
 The total set of parameters $\vec{p}$ will include both Jastrow and determinental coefficients.
-
-Since a model for benzene has already been developed using DMD, my calculation will also serve as a benchmark for new methods like CVMC sampling.
-The CVMC method is relatively new and has not been tested extensively on realistic systems.
-The final model that I will produce should be exactly comparable to previous calculations, and as such can give insight into the workings of the CVMC method.
-A few questions of interest include whether the CVMC is more statistically efficient in terms of regression, how to select values like $\vec{\lambda}$ and $\vec{d}_0$, and whether the final model fit with CVMC is similar to the one fit previously using a different sampling technique.
-
-The second system I will be interested in working on is SLG.
-Single layer graphene is composed of a triangular Bravais lattice with two carbon atoms per unit cell.
-In terms of a non-interacting picture, the $\sigma$ orbitals composed of C $2s, 2p_x,$ and $2p_y$ lie below the Fermi level.
-The bands which cross the Fermi level, and compose the distinct Dirac cones seen in SLG, are primary $\pi$ bonded molecular orbitals composed of C $2p_z$ orbitals.
-The lack of significant screening in SLG leads to long range Coulomb interactions in the system.
-
-The model I am interested in working on will be an extension of a previous model developed for SLG using DMD.
-The fit model was a single band, on-site Hubbard model which contained only the $\pi$ orbital degrees of freedom in the system.
-The fit model parameters were $U/t \sim$ 2.00, which is smaller than the critical value of $(U/t)_c \sim$ 3.8 for the semimetal-insulator transition in the honeycomb lattice, as expected.
-It is interesting to note that if the $\sigma$ orbitals in graphene are treated as a constant negative charge background, then the effective $U$ on site nearly doubles, indicating the importance of including these degrees of freedom in the low-energy sampling.
-
-I will extend the previous model by introducing long range density-density interactions into the effective theory.
-As mentioned previously, one expects to see long range Coulomb interactions in graphene due to the vanishing density of states near the Fermi level, in contrast to other metals where electron-hole pairs screen the long range interaction.
-I will work on developing a model for SLG of the following form:
-\begin{equation}
-H_\text{SLG} = -\sum_{i,j} t_{ij}c_i^\dagger c_j + U \sum_i n_{i\uparrow}n_{i\downarrow}  + \sum_{i,j} V_{ij} n_i n_j
-\label{Hslg}
-\end{equation}
-where the indices correspond to the $\pi$ orbitals on the copper atoms, represented by IAO basis elements.
-I will consider arbitrarily long range hopping and density-density interaction terms in my model, in contrast to the benzene calculation where the finite size of the system imposes a cutoff in the possible range of interaction.
-
-CVMC will be used to sample the low-energy wave functions necessary to fit the model.
-The trial wave function will also be of MSJ type as explained for the benzene molecule, with the determinants built from the single particle orbitals of DFT calculations.
-I will consider the DFT ground state determinant and all excitations built on in restricted to the $\pi, \pi^*$ molecular orbitals.
-A three-body Jastrow will be used to introduce correlations into the wave function and ensure our sampled states are low in energy.
-
-The last systems I will develop models for are AA and AB stacked BLG.
-AA stacked BLG is simply two sheets of graphene placed on top of one another where each atom is aligned with the atom beneath it.
-In AB stacked BLG, a carbon atom from the lower layer is present in the center of each carbon hexagon of the upper layer, and vice versa.
-As such, one can think of AB BLG as a shifted version of AA BLG, where the shift is half the size of a single carbonic hexagon.
-Schematic drawings of these two stackings can be seen in Figure \ref{fig:proposed}b.
-
-Being the simplest BLG configuration, these models will form a starting point for a more thorough development of models for TBLG.
-The ultimate goal of developing a model for TBLG would be a model which has the parameters as a function of the twist angle.
-The twist angle changes the local chemistry of a particular atom in a single layer of graphene by introducing neighbors on the other layer at various non-commensurate distances.
-As such, the first avenue of investigation would be to see how the model parameters change as one shifts the position of interlayer neighbor carbon atoms. 
-The AA and AB BLG configurations provide the simplest systems where this position dependence on model parameters can be probed.
-
-I anticipate that the candidate descriptors should include interlayer couplings as well as the terms seen in SLG.
-As such, a candidate model that we may consider initially for BLG would be
-\begin{equation}
-H_\text{BLG} = H_\text{SLG}^{(1)} + H_\text{SLG}^{(2)} - \sum_{i_1, i_2} t_{i_1, i_2}^\prime c_{i_1}^\dagger c_{i_2} + h.c.
-\label{Hblg}
-\end{equation}
-Here we have two models for the single layers from \eqref{Hslg} plus an interlayer hopping given by the term $t_{i_1, i_2}^\prime.$
-I suspect that the hopping values will be very different between the AA and AB models as the distance between interlayer neighbors differs greatly between the two stackings.
-
-I will again use CVMC to generate the low-energy wave functions for the model regression.
-As before, I think that an MSJ parameterization will serve the purpose well.
-Similarly, the determinants used will include a DFT ground state alongside all excitations within the $\pi, \pi^*$ orbitals of the system.
-Note that excitations between the $\pi, \pi^*$ orbitals from both sheets should be included in the multi-Slater expansion.
-A three-body Jastrow factor will be used for the reasons listed before.
-
-The final product of my thesis would be four models for these increasingly complex systems.
-The model for benzene should reproduce previous results and provide an intro calculation for me, as well as a benchmark test for the CVMC sampling method.
-The model for SLG will extend on previous studies, incorporating long range Coulomb interactions.
-The models for BLG will be a first step in understanding how interlaying coupling depends on interlayer neighbor distances in the bilayer systems, an important stepping stone towards a general model for TBLG.
 
@@
 The second would be whether the intralayer parameter values change in each sheet of SLG when they are brought together in the AA and AB configuration.
-The latter question hints at a broader question, whether the SLG intralayer parameters are transferrable to TBLG, especially near the magic twist angles whre interlayer coupling becomes very large.
+The latter question hints at a broader question, whether the SLG intralayer parameters are transferrable to TBLG, especially near the magic twist angles where interlayer coupling becomes very large.
 
-I conclude by briefly discussing potential avenues of further study in using DMD to develop model Hamiltonians for TBLG. 
-One would have to begin by building atomic length scale lattice models for TBLG using DMD, which itself is a monumental task.
-When dealing with small commensurate twists, the size of the Moire supercell is enormous and prohibitive to QMC calculations.
-As such, a computationally feasible method for probing small shifts, either by translation or angle, between the graphene sheets must be constructed.
-This short length scale Hamiltonian can then be downfolded again, using DMD, to a continuum Hamiltonian on the length scale of the Moire superlattice.
-This kind of downfolding will require using large-scale tight binding codes like \textbf{ref}.
-Iterating over various twist angles can then give insight into the behavior of the long-range effective theory of TBLG as a function of the twist angle.
+Further, my calculations will help mature methods like CVMC for further usage on strongly correlated systems.
+A few questions of interest include how one should select values like $\vec{\lambda}$ and $\vec{d}_0$, and whether final models fit with CVMC are consistent with those fit previously using a different sampling technique.
+The prior question will be naturally investigated while using CVMC in building models for the systems above.
+The latter can be directly probed by comparing my fit models for benzene and on-site SLG to those which were developed previously using DMD.
 
-\subsection{Proposed timeline}
-I believe that developing effective theories for the benzene molecule, SLG, AA and AB stacked BLG should take 2 years. 
+Lastly, my calculations will lay the groundwork for model development for the strongly correlated TBLG system using DMD.
+The final goal for TBLG would be an interacting continuum model on the Moire superlattice length scale.
+From a DMD perspective, this can accomplished by first downfolding the \textit{ab-initio} Hamiltonian with a given twist to an atomic length scale model, and then downfolding a second time to the Moire length scale.
+Developing the atomic length scale models requires a strong understanding of intra- and inter-layer interactions in this material.
+The prior will be thoroughly investigated by my model fitting on SLG, and I will begin to answer the latter with the BLG calculations.
+
+\paragraph{Proposed timeline} I believe that developing effective theories for the benzene molecule, SLG, AA and AB stacked BLG should take 2 years. 
 The model fitting for benzene should take no more than four months.
@@
 Inclusion of long range Coulomb interactions into the effective theory should take another three months.
-The last two months will be geared towards finite size extrapolation in order to ensure our final model is applicable to bulk systems.
+The last two months will be geared towards finite size extrapolation in order to ensure the final model is applicable to bulk systems.
 The last year of the project will be geared towards the AA and AB stacked BLG, with each project taking about six months to complete.

### 2019-07-30 outline.tex
edited + shortened draft. have 1.5 pages for references now.

--- 2019-07-26 outline.tex
@@
 \begin{itemize}
-\item Systematically developing model Hamiltonians which can accurately describe the energies and properties of the lowest lying eigenstates for strongly correlated electron systems is a pressing problem in modern condensed matter physics.
+\item Developing model Hamiltonians for realistic materials is a fundamental problem in modern condensed matter physics.
 
-\item The necessity for non-perturbative, many-body models in describing the low-lying excited states of strongly correlated electron systems poses a difficult challenge when developing effective theories for real world systems.
+\item Yet unresolved is a paradigm for building many-body model Hamiltonians which can describe realistic systems.
 
-\item Common approaches to building suitable model Hamiltonians for strongly correlated materials involve using single particle theories like density functional theory (DFT).
+\item Common approaches to building many-body theories for real materials involve density functional theory (DFT).
 
-\item We propose using a DMD procedure to develop accurate many-body effective theories for single (SLG) and bi-layer graphene (BLG) systems as a step towards building a many-body model for twisted bilayer graphene (TBLG).
+\item I propose using a quantum Monte Carlo (QMC) density matrix downfolding (DMD) procedure to develop accurate many-body model Hamiltonians for single (SLG) and bi-layer (BLG) graphene systems.
 \end{itemize}
@@
 \begin{itemize}
-\item We begin by defining what we mean by a low-energy effective Hamiltonian and downfolding.
+\item I begin by defining what a low-energy effective Hamiltonian and downfolding are.
 
-\item The key insight of DMD is that the Hamiltonian downfolding problem can be mapped onto an equivalent linear regression problem.
+\item The key insight of DMD is that the Hamiltonian downfolding problem can be mapped onto a linear regression problem.
 
@@
 
-\item Generally, one cannot sample the true low-energy subspace and an approximate subspace $\mathcal{LE}^\prime$ is sampled.
-
 \item Next, a set of candidate descriptors which form the effective Hamiltonian are selected.
-
-\item Typically, candidate operators are written in second quantized form, and an appropriate single particle basis to express these operators must be constructed.
 
@@
 
-\item We take advantage of the variational nature of FN-DMC to sample the low-energy states necessary for DMD.
+\item One can take advantage of the variational nature of FN-DMC to sample the low-energy states necessary for DMD.
 \end{itemize}
@@
 
-\paragraph{CuO: Introduction} As a first step towards developing accurate models for extended systems like solids using DMD, we constructed a many-body effective model for the CuO molecule which accurately describes the eigenstates and spectra seen in experiment up to 2eV above the ground state.
+\begin{enumerate}
+\item Intro (no subsection)
+\begin{itemize}
+\item As a first step towards developing accurate models for extended systems like solids using DMD, I constructed a many-body effective model for the CuO molecule by downfolding (1) under a Born-Oppenheimer approximation with fixed bond length $r_e$ = 1.725\r{A}.
+\end{itemize}
 
-\paragraph{CuO: Methods} In order to fit a model using DMD we had to construct a low energy space (LES), choose a sampling scheme for states within the LES resulting in our sample space (SS), develop a basis which can accurately describe variations within the LES, identify and select descriptors built out of these basis elements, and appropriately fit the data from our SS onto our selected descriptors.
+\item Sampling low-energy states
+\begin{itemize}
 
-\paragraph{CuO: Results and Discussion} The resultant model from our analysis accurately describes the various eigenstates as measured in experiment up to 2eV, and is quantitatively and qualitatively more accurate than competing methods like density functional theory.
+\item For the CuO molecule, low-energy states were generated by
+using FN-DMC projected multi-Slater-Jastrow (MSJ) trial wave functions.
 
-\paragraph{Hubbard: Preliminary results} Our preliminary calculations on the 2x2 half-filled 1-band Hubbard model map out the limiting behavior in the U-T model space by investigating the effectiveness of the single descriptor models.
-\textcolor{red}{Will be changed based on how far along we get with preliminary results.}
+\item The determinants were constructed from the orbitals of independent symmetry-targeted unrestricted Kohn Sham (UKS) calculations.
+
+\item A shell sampling scheme was used to sample states from the selected approximate low-energy space.
+
+\item The final sampled MSJ states were projected using FN-DMC to generate our low-energy states.
+\end{itemize}
+
+\item Selecting candidate descriptors 
+\begin{itemize}
+\item We selected the following set of 1- and 2-body operators to construct our candidate descriptors based on our understanding of the low-energy excitations of the CuO molecule and our sampled states
+
+\item The parantheses of \textbf{eqref} contain the full set of symmetry allowed 1-body operators within the Cu 3d, 4s and O 2p space
+\end{itemize}
+
+\item Fitting the effective Hamiltonian 
+\begin{itemize}
+\item {A direct attempt at fitting all eleven terms in \textbf{eqref} leads to a high goodness of fit but unexpectedly large coefficients for the fit model.}
+
+\item A principal component regression (PCR) was used to alleviate this multicollinearity and fit the effective theory.
+
+\item From the PCA, the nine highest ranked PCs were selected to use in the PCR.
+
+\item {The model fit to these nine PCs maintains an excellent goodness of fit without serious multicollinearity.}
+\end{itemize}
+
+\item Solutions of effective Hamiltonian
+\begin{itemize}
+\item Solutions of our effective Hamiltonian are obtained by exact diagonalization.
+
+\item The model is solved by exact diagonalization, and the solutions
+agree well with experimental spectra and state assignments.
+
+\item The model states also agree with recent theoretical calculations like UKS and CCSD(T).
+\end{itemize}
+\end{enumerate}
 
@@
 \begin{itemize}
-\item My proposed plan involves building a sequence of model Hamiltonians for increasing complicated systems using DMD, working towards an accurate many-body model for TBLG.
+\item My proposed plan involves building a sequence of model Hamiltonians for increasingly complicated graphene-like systems using QMC DMD, ending with BLG.
 
-\item The first system I will develop a model for is the benzene molecule.
-
-\item While relatively simple, this molecule shares many similarities with and forms the basic unit of SLG.
-
-\item As such, developing a model for benzene will help me become familiar with model fitting for 2-D carbon based systems.
+\item The first system I will develop a model for is the benzene molecule, shown in Figure 2a.
 
@@
 
-\item A new constrained variational Monte Carlo (CVMC) method will be used to sample the low-energy wave functions necessary for DMD.
-
-\item Note also that a model for benzene has already been developed using DMD.
-
-\item This makes benzene a useful system not only for my personal learning, but as a benchmarking system for new methods like CVMC which have been developed recently.
-
 \item The second system I will be interested in working on is SLG.
 
-\item The model I am interested in working on will be an extension of a previous model developed for SLG using DMD.
+\item In particular, I will work to extend a previous on-site Hubbard model developed for SLG by including long range density-density interactions.
 
-\item I will extend the previous model by introducing long range density-density interactions into the effective theory.
-
-\item The candidate model I anticipate to work with is an extended Hubbard model of the form.
-
-\item CVMC will be used to sample the low-energy wave functions necessary to fit the model.
-
-\item The last systems I will develop models for are AA and AB stacked BLG.
-
-\item Being the simplest BLG configuration, these models will form a starting point for a more thorough development of models for TBLG.
+\item The last systems I will develop models for are AA and AB stacked BLG, shown in Figure 2b.
 
@@
 
-\item I will again use CVMC to generate the low-energy wave functions for the model regression.
+\item A new constrained variational Monte Carlo (CVMC) method will be used to sample the low-energy wave functions necessary for DMD.
 
-\item The final four models can be used to investigate the transferrability of model parameters between these different carbon-based systems.
+\item I will use a multi-Slater-Jastrow (MSJ) parameterization in the CVMC sampling for each system.
 
-\item I will conclude by briefly discussing potential avenues of further study in using DMD to develop model Hamiltonians for TBLG.
+\item Apart from their individual value, the four models can be used to investigate the transferrability of model parameters between these different carbon-based systems.
 
+\item Further, my calculations will help mature methods like CVMC for further usage on strongly correlated systems.
+
+\item Lastly, my calculations will lay the groundwork for model development for the strongly correlated TBLG system using DMD.
 \end{itemize}
 \section{Proposed timeline}
-\textit{Do this after proposed work nailed down}.
 

### 2019-07-31 draft.tex
added references, all fits in 15 + 1 pages now.

@@
 \usepackage{float}
-\usepackage{ulem}
 \usepackage{amsmath}
@@
 \usepackage{wrapfig}
+\usepackage{setspace}
+\usepackage{multicol}
 \renewcommand{\baselinestretch}{1.5}
 \usepackage[utf8]{inputenc}
+\usepackage{etoolbox}
+\patchcmd{\thebibliography}{\section*{\refname}}{}{}{}
 \author{Shivesh Pathak}
@@
 A core problem of condensed matter physics is understanding how to develop low-energy effective models for realistic systems.
-I believe that a density matrix downfolding (DMD) method can be used to systematically build effective theories for such systems beginning with a high energy theory like the \textit{ab-initio} Hamiltonian.
+I believe that a density matrix downfolding (DMD) method \cite{Zheng2017} can be used to systematically build effective theories for such systems beginning with a high energy theory like the \textit{ab-initio} Hamiltonian.
 I present as a preliminary result an effective Hamiltonian developed using DMD which can accurately reproduce the spectra and properties of the lowest lying excitations of the CuO molecule.
@@
 These theories, however, fail in describing the low-energy excitations of strongly correlated systems like high Tc superconductors and twisted bilayer graphene (TBLG).
-Many-body models like the Hubbard, Kanamori, t-J and Heisenberg models have 
-been proposed to describe such strongly correlated systems, but their effectiveness in describing realistic materials is unclear.
+Many-body models like the Hubbard \cite{Hubbard1963}, Kanamori \cite{10.1143/PTP.30.275}, t-J \cite{Chao_1977} and Heisenberg models have been proposed to describe such strongly correlated systems, but their effectiveness in describing realistic materials is unclear.
 
-Common approaches to building many-body theories for real materials involve density functional theory (DFT).
-One-body terms are obtained by projecting the solutions from the DFT band structure onto a single-particle basis.
-Two-body terms are developed by assuming a screened Coulomb interaction based on constrained DFT, RPA, or other methods.
-L\''{o}wdin methods and canonical transformations have also been used to develop effective theories.
+Common approaches to building many-body theories for real materials involve density functional theory (DFT) \cite{PhysRev.140.A1133}.
+One-body terms are obtained by projecting the solutions from the DFT band structure onto a single-particle basis \cite{PhysRevLett.87.047003}.
+Two-body terms are developed by assuming a screened Coulomb interaction based on constrained DFT, RPA, or other methods \cite{PhysRevB.62.R16219, PhysRevB.70.195104, PhysRevB.88.075106}.
+L\''{o}wdin methods and canonical transformations have also been used to develop effective theories \cite{doi:10.1063/1.4802766, PhysRevA.81.013402, doi:10.1063/1.2196410}.
 While useful techniques, it is typically impossible to know whether models built from them accurately describe reality, as usually no clear estimate of model quality is provided.
@@
 I propose using a quantum Monte Carlo (QMC) density matrix downfolding (DMD) procedure to develop accurate many-body model Hamiltonians for single (SLG) and bi-layer (BLG) graphene systems.
-Recent observations of a fractional quantum Hall effect and plasmarons in SLG as well as unconventional superconductivity in TBLG have spurred an interest in developing many-body models for graphene.
+Recent observations of a fractional quantum Hall effect \cite {Bolotin2009, Du2009} and plasmarons \cite{Bostwick999} in SLG as well as unconventional superconductivity \cite{Cao2018_2} and many-body insulating states \cite{Cao2018} in TBLG have spurred an interest in developing many-body models for graphene.
 The DMD procedure provides a framework for developing such a many-body model using QMC calculations starting from the \textit{ab-initio} Hamiltonian 
@@
 A typical fitting workflow involves feature selection, model regression and model validation.
-Feature selection methods for choosing appropriate descriptors can range from wrapper methods like orthogonal matching pursuit and principal component analysis to embedded methods like LASSO, ridge and elastic net regularizations.
-Fitting the effective Hamiltonian usually involves an ordinary least squares linear regression, but the cost function can be altered as in the cases of the embedded methods above or weighted least squares linear regression.
+Feature selection methods for choosing appropriate descriptors can range from wrapper methods like orthogonal matching pursuit \cite{Cai2011} and principal component analysis \cite{pearson_karl_1901_1430636} to embedded methods like LASSO \cite{10.2307/2346178}, ridge \cite{doi:10.1137/S0895479897326432} and elastic net regularizations \cite{Zou05regularizationand}.
+Fitting the effective Hamiltonian usually involves an ordinary least squares linear regression, but the cost function can be altered.
 Model validation can be carried out by calculating cross validated single parameter measures like R$^2$ scores which carry information about goodness of fit and overfitting.
@@
 Given the statistical nature of the fitting procedure, model quality assessment can be carried out and reported in exhaustive detail if one is interested.
-Typically, however, simple single parameter measures like cross validated R$^2$ scores and root mean squared errors are used to quantify the accuracy of the effective Hamiltonian.
+Typically, however, single parameter measures like cross validated R$^2$ scores are used to quantify the accuracy of the effective Hamiltonian.
 
@@
 The final projected $\Psi_\text{DMC}$ for these different trial functions will be the lowest energy states in $\mathcal{H}$ for the given nodal structures.
-If the initial trial wave functions are appropriately chosen, the final projected states will be low-energy states within $\Psi_\text{DMC}$.
+If the initial trial wave functions are appropriately chosen, the final projected states will be low-energy states within $\mathcal{H}$.
 The expectation values $E[\Psi_\text{DMC}], \{d_k[\Psi_\text{DMC}]\}$ necessary for the model regression are calculated using a mixed estimator.
-Details for calculating reduced density matrix elements in FN-DMC can be found in Wagner.
 
@@
 \subsection{Non-orthogonal determinants in FN-DMC trial wave functions}
-In order to become acquainted with our code QWalk \cite{WAGNER20093390} and QMC algorithms in general I worked on implementing and testing multi-Slater-Jastrow trial functions with optimized non-orthogonal determinants (MSJ+NO) in FN-DMC \cite{Pathak2018}.
+In order to become acquainted with our code QWalk \cite{Wagner2009} and QMC algorithms in general I worked on implementing and testing multi-Slater-Jastrow trial functions with optimized non-orthogonal determinants (MSJ+NO) in FN-DMC \cite{doi:10.1063/1.5052906}.
 The MSJ+NO trial wave functions take the form
@@
 Inclusion of this high energy determinant is necessary so our model places similar Cu 4s$^2$ states high in energy.
-The UKS calculations were done using PySCF, a B3LYP functional, Trail-Needs psuedopotential and VTZ Trail-Needs basis.
+The UKS calculations were done using PySCF \cite{Sun2018}, a B3LYP functional \cite{doi:10.1063/1.464304, PhysRevB.37.785}, Trail-Needs psuedopotential and VTZ Trail-Needs basis \cite{doi:10.1063/1.4811651}.
 
@@
 A timestep of $\tau = 0.01$ was used.
+These calculations were done in QWalk \cite{Wagner2009}.
 
@@
 
+\begin{figure*}
+\centering
+%\hspace*{-0.25in}
+\includegraphics[width=1.0\linewidth]{./figs/cuo_regr_2.pdf}
+\caption{Figures related to model regression for the CuO molecule: a) Intrinsic atomic orbitals used as a single particle basis for the model Hamiltonian, b) Results of PCA showing the CEVR and CN of a data set containing N$_{comp}$  highest ranked PCs, c) The final regressed parameters from the PCR, d) Comparison of four lowest lying eigenstates of CuO molecule between experiment and three computation methods.}
+\label{fig:cuo_regr}
+\end{figure*}
+
 \paragraph{Fitting the model}
@@
 
-A principal component regression (PCR) was used to alleviate this multicollinearity and fit the effective theory.
+A principal component regression (PCR) was used to alleviate this multicollinearity and fit the effective theory \cite{10.2307/2348005}.
 PCR can be broken into three parts: conducting a principal component analysis (PCA) and selecting candidate principal components to work with (PCs), fitting a linear model to these PCs, and then inverting the coefficients back to the original descriptors.
@@
 
-\begin{figure*}
-\centering
-\hspace*{-0.25in}
-\includegraphics[width=1.1\linewidth]{./figs/cuo_regr_2.pdf}
-\caption{Figures related to model regression for the CuO molecule: a) Intrinsic atomic orbitals used as a single particle basis for the model Hamiltonian, b) Results of PCA showing the CEVR and CN of a data set containing N$_{comp}$  highest ranked PCs, c) The final regressed parameters from the PCR, d) Comparison of four lowest lying eigenstates of CuO molecule between experiment and three computation methods.}
-\label{fig:cuo_regr}
-\end{figure*}
-
 \paragraph{Solutions of the model}
 The model is solved by exact diagonalization, and the solutions agree well with experimental spectra and state assignments.
-Shown in Figure \ref{fig:cuo_regr}d is a comparison between the excitation energies calculated from our model and those seen in the latest anion photoelectron spectroscopy measurements.
+Shown in Figure \ref{fig:cuo_regr}d is a comparison between the excitation energies calculated from our model and those seen in the latest anion photoelectron spectroscopy measurements \cite{Wu1997}.
 The model energies for these eigenstates are denoted by orange circles and agree with experiment up to a maximum deviation of 0.2 eV.
@@
 The model states also agree with recent theoretical calculations like UKS and CCSD(T).
-The energies for the UKS and CCSD(T) states are shown in Figure \ref{fig:cuo_regr}d as green and red circles, respectively.
+The energies for the UKS and CCSD(T) \cite{Xian2000} states are shown in Figure \ref{fig:cuo_regr}d as green and red circles, respectively.
 The UKS energies were calculated in this work in en route to constructing the determinants $|\text{D}_j\rangle$ used in our sampling scheme \eqref{eq:sampling}.
@@
 The delocalized $\pi$ orbitals in benzene carry over directly into SLG where they are the primary contributors to the well known Dirac bands.
-Interaction effects seen in the benzene molecule are also reflected in SLG, such as on-site double occupation energies for the carbon atoms.
+Interaction effects seen in the benzene molecule are also reflected in SLG, such as on-site double occupation energies for the carbon atoms \cite{Zheng2017, Wagner2015}.
 As such, developing a model for benzene will help me become familiar with model fitting for 2-D carbon based systems.
@@
 
-The candidate descriptors I will use for DMD are motivated by a previous calculation on the benzene molecule.
+The candidate descriptors I will use for DMD are motivated by a previous calculation on the benzene molecule \cite{Wagner2015}.
 In that study, an extended Hubbard model built on the six $\pi$ orbitals was found to perform well in describing the low-lying excitations of the material.
@@
 The bands which cross the Fermi level, and compose the distinct Dirac cones seen in SLG, are primary $\pi$ bonded molecular orbitals composed of C $2p_z$ orbitals.
-The lack of significant screening in SLG leads to long range Coulomb interactions in the system.
+Further, reduced screening in SLG leads to long range Coulomb interactions in the system \cite{Elias2012, Yu2013}.
 
-In particular, I will work to extend a previous on-site Hubbard model developed for SLG by including long range density-density interactions.
+In particular, I will work to extend a previous on-site Hubbard model developed for SLG \cite{Zheng2017, Wagner2015} by including long range density-density interactions .
 The candidate model will be of the form:
@@
 
-A new constrained variational Monte Carlo (CVMC) method will be used to sample the low-energy wave functions necessary for DMD.
-The CVMC method works by minimizing the following cost function for a parameterized wave function $|\Psi(\vec{p})\rangle$:
+A new constrained variational Monte Carlo (cVMC) method will be used to sample the low-energy wave functions necessary for DMD.
+The cVMC method works by minimizing the following cost function for a parameterized wave function $|\Psi(\vec{p})\rangle$:
 \begin{equation}
@@
 
-I will use a multi-Slater-Jastrow (MSJ) parameterization in the CVMC sampling for each system.
-The multi-Slater expansion will consist of determinants of single particle orbitals calculated using a self consistent field calculation like DFT.
+I will use a multi-Slater-Jastrow (MSJ) parameterization in the cVMC sampling for each system.
+The multi-Slater expansion will consist of determinants of single particle orbitals calculated using DFT.
 The determinants will include the DFT ground state as well as molecular orbital excitations among the $\pi$ and $\pi^*$ orbitals which form the active space.
@@
 The second would be whether the intralayer parameter values change in each sheet of SLG when they are brought together in the AA and AB configuration.
-The latter question hints at a broader question, whether the SLG intralayer parameters are transferrable to TBLG, especially near the magic twist angles where interlayer coupling becomes very large.
+The latter question hints at a broader question, whether the SLG intralayer parameters are transferrable to TBLG, especially near the magic twist angles where interlayer coupling becomes very large \cite{Bistritzer2011}.
 
-Further, my calculations will help mature methods like CVMC for further usage on strongly correlated systems.
-A few questions of interest include how one should select values like $\vec{\lambda}$ and $\vec{d}_0$, and whether final models fit with CVMC are consistent with those fit previously using a different sampling technique.
-The prior question will be naturally investigated while using CVMC in building models for the systems above.
+Further, my calculations will help mature methods like cVMC for further usage on strongly correlated systems.
+A few questions of interest include how one should select values like $\vec{\lambda}$ and $\vec{d}_0$, and whether final models fit with cVMC are consistent with those fit previously using a different sampling technique.
+The prior question will be naturally investigated while using cVMC in building models for the systems above.
 The latter can be directly probed by comparing my fit models for benzene and on-site SLG to those which were developed previously using DMD.
@@
 The model fitting for benzene should take no more than four months.
-The goal is to reproduce prior DMD calculations for the same material while benchmarking and familiarizing myself with the new CVMC sampling technique that I will use extensively in my project.
+The goal is to reproduce prior DMD calculations for the same material while benchmarking and familiarizing myself with the new cVMC sampling technique that I will use extensively in my project.
 The SLG model fitting should take another eight months.
@@
 For each stacking, I expect four months will be necessary to build an effective theory, and the final two months will be dedicated to finite size extrapolation.
+
+\begin{multicols}{2}
+\begin{spacing}{0.5}
+{\fontsize{10}{5}\selectfont
+\bibliography{biblio}
+\bibliographystyle{unsrt}
+}
+\end{spacing}
+\end{multicols}
 \end{document}

### 2019-08-10 draft.tex
updated commitee
rearranged last paragraph of intro, based on joaos comments

@@
 \end{abstract}
-Committee: Lucas Wagner, Smitha Vishveshwara, David Ceperley, Karin Dahmen
-
-Pool: \textit{David Ceperley},\textit{ Brian DeMarco}, Smitha Vishveshwara, Andre Schleife, \textit{Karin Dahmen} 
+Committee: Lucas Wagner, Brian DeMarco, Andre Schleife, Karin Dahmen
 \pagebreak
@@
 
-I propose using a quantum Monte Carlo (QMC) density matrix downfolding (DMD) procedure to develop accurate many-body model Hamiltonians for single (SLG) and bi-layer (BLG) graphene systems.
-Recent observations of a fractional quantum Hall effect \cite {Bolotin2009, Du2009} and plasmarons \cite{Bostwick999} in SLG as well as unconventional superconductivity \cite{Cao2018_2} and many-body insulating states \cite{Cao2018} in TBLG have spurred an interest in developing many-body models for graphene.
-The DMD procedure provides a framework for developing such a many-body model using QMC calculations starting from the \textit{ab-initio} Hamiltonian 
+A relatively new technique is density matrix downfolding (DMD), which provides a framework for developing many-body models directly from calculations on the \textit{ab-initio} Hamiltonian
 \begin{equation}
@@
 The procedure makes no assumptions about the form of microscopic electronic interactions in the system and reports a clear, statistical measure of model quality.
+I propose using a quantum Monte Carlo (QMC) DMD procedure to develop accurate many-body model Hamiltonians for single (SLG) and bi-layer (BLG) graphene systems.
+Recent observations of a fractional quantum Hall effect \cite {Bolotin2009, Du2009} and plasmarons \cite{Bostwick999} in SLG as well as unconventional superconductivity \cite{Cao2018_2} and many-body insulating states \cite{Cao2018} in TBLG have spurred an interest in developing many-body models for graphene.
 By developing many-body models for SLG and BLG, I hope to further the understanding of interactions in graphene like systems and the maturity of DMD for its application to other strongly correlated systems like TBLG.
@@
 \begin{equation}
-\vec{p}^* = \text{argmin} \ \langle \Psi(\vec{p}) | \hat{H}_\text{ab} | \Psi(\vec{p}) \rangle - \vec{\lambda} \cdot \langle \Psi(\vec{p}) | (\vec{d} - \vec{d}_0)^2 | \Psi(\vec{p}) \rangle 
+\vec{p}^* = \text{argmin} \ \langle \Psi(\vec{p}) | \hat{H}_\text{ab} | \Psi(\vec{p}) \rangle - \sum_k [\lambda_k \langle \Psi(\vec{p}) | d_k - d_{k}^0 | \Psi(\vec{p}) \rangle]^2
 \end{equation}
 with respect to the parameters $\vec{p}$.
-The quantity $\vec{d}$ is short hand for a list of candidate descriptors to sample variations in, $\vec{d}_0$ is the target values of those candidate descriptors, and $\vec{\lambda}$ controls the importance of finding a state with low-energy versus one with descriptor values near $\vec{d}_0$.
-Essentially this method allows us, for a given parameterization of $|\Psi(\vec{p})\rangle$, to generate low-energy wave functions which sit near a target point $\vec{d}_0$ in candidate descriptor space.
+The quantity $\vec{d}$ is short hand for a list of candidate descriptors to sample variations in, $\vec{d}^0$ is the target values of those candidate descriptors, and $\vec{\lambda}$ controls the importance of finding a state with low-energy versus one with descriptor values near $\vec{d}_0$.
+Essentially this method allows us, for a given parameterization of $|\Psi(\vec{p})\rangle$, to generate low-energy wave functions which sit near a target point $\vec{d}^0$ in candidate descriptor space.
 
@@
 The second would be whether the intralayer parameter values change in each sheet of SLG when they are brought together in the AA and AB configuration.
-The latter question hints at a broader question, whether the SLG intralayer parameters are transferrable to TBLG, especially near the magic twist angles where interlayer coupling becomes very large \cite{Bistritzer2011}.
+The latter question hints at a broader question, whether the SLG intralayer parameters are transferrable to TBLG, especially near the magic twist angles where interlayer coupling becomes very large \cite{Bistritzer2	011}.
 

### 2019-08-19 draft.tex
Lowdin

@@
 Two-body terms are developed by assuming a screened Coulomb interaction based on constrained DFT, RPA, or other methods \cite{PhysRevB.62.R16219, PhysRevB.70.195104, PhysRevB.88.075106}.
-L\''{o}wdin methods and canonical transformations have also been used to develop effective theories \cite{doi:10.1063/1.4802766, PhysRevA.81.013402, doi:10.1063/1.2196410}.
+L\"owdin methods and canonical transformations have also been used to develop effective theories \cite{doi:10.1063/1.4802766, PhysRevA.81.013402, doi:10.1063/1.2196410}.
 While useful techniques, it is typically impossible to know whether models built from them accurately describe reality, as usually no clear estimate of model quality is provided.

### 2019-09-07 reoutline.tex
outline being redone, introduction complete!

\documentclass{article}
\begin{document}
\section{Introduction}
\begin{enumerate}
\item Developing accurate model Hamiltonians for strongly correlated materials remains an outstanding problem in modern condensed matter physics.

\item To bridge this gap, a framework for model development using solutions from \textit{ab-initio} simulations has emerged.

\item A recently developed density matrix downfolding (DMD) method addresses the need for a more systematic link between \textit{ab-initio} and model Hamiltonians.
%No QMC here

\item The surge in model development for twisted bilayer graphene (TBLG) exemplifies the necessity for such a systematic framework.

\item I propose using DMD to systematically develop accurate, many-body model Hamiltonians for single- (SLG) and simple bi-layer (BLG) graphene systems.
%Start this paragraph with the grant proposal, have my thesis statement in the middle makes more sense!
%Name drop QMC
\end{enumerate}

\section{Density Matrix Downfolding (DMD)}
\begin{enumerate}
\item I will begin by defining carefully the terms model Hamiltonian and Hamiltonian downfolding.

\item The key insight of DMD is that the Hamiltonian downfolding problem is equivalent to a linear regression problem.

\item Being cast to a linear regression, the downfolding problem can be handled highly systematically.

\item Practically, the linear regression is tackled in three steps.

\item The through line to \textit{ab-initio} is maintained via the training data.
%QMC because accurate + efficient, but details later
\end{enumerate}

\section{Fixed-node diffusion Monte Carlo (FN-DMC)}
\begin{enumerate}
\item Diffusion Monte Carlo (DMC) is a quantum Monte Carlo method which projects out the ground state of a real-space Hamiltonian given some initial trial wave function.
%Talk about efficiency here

\item DMC, however, suffers from a fermion sign problem which is alleviated via a fixed-node approximation.
%Talk about accuracy here, compare to other methods

\item While seemingly disadvantageous, the variational nature of FN-DMC plays a key role in generating the low-energy states needed for DMD.

\item The training data required for downfolding are extracted from the projected low-energy states via a mixed estimator.
%Start paragraph with something else, probably the stochastic representation of the wave function
\end{enumerate}

\end{document}

### 2019-09-08 draft.tex
intro done, need to do abstract + references

@@
 
-Yet unresolved is a paradigm for building many-body model Hamiltonians which can describe realistic systems.
+Yet unresolved is a paradigm for building many-body model Hamiltonians which can describe strongly correlated systems.
 Band and Fermi liquid theory have been successful in accurately modeling systems where effective interactions are negligible at low energy scales.
@@
 
-Common approaches to building many-body theories for real materials involve density functional theory (DFT) \cite{PhysRev.140.A1133}.
-One-body terms are obtained by projecting the solutions from the DFT band structure onto a single-particle basis \cite{PhysRevLett.87.047003}.
-Two-body terms are developed by assuming a screened Coulomb interaction based on constrained DFT, RPA, or other methods \cite{PhysRevB.62.R16219, PhysRevB.70.195104, PhysRevB.88.075106}.
+In an attempt to build accurate many-body Hamiltonians for realistic systems, a framework for model development using solutions from \textit{ab-initio} simulations has emerged.
+Common approaches involve density functional theory (DFT) \cite{PhysRev.140.A1133} with one-body terms obtained from the DFT band structure \cite{PhysRevLett.87.047003} and two-body terms developed using constrained DFT, RPA or other methods \cite{PhysRevB.62.R16219, PhysRevB.70.195104, PhysRevB.88.075106}.
 L\"owdin methods and canonical transformations have also been used to develop effective theories \cite{doi:10.1063/1.4802766, PhysRevA.81.013402, doi:10.1063/1.2196410}.
-While useful techniques, it is typically impossible to know whether models built from them accurately describe reality, as usually no clear estimate of model quality is provided.
+While useful, it is typically impossible to know whether models built from these techniques accurately describe reality, as measures of model quality are rarely reported.
+Instead, out-of-method comparisons of model predictions to experiment must be conducted for model validation.
 
-A relatively new technique is density matrix downfolding (DMD), which provides a framework for developing many-body models directly from calculations on the \textit{ab-initio} Hamiltonian
-\begin{equation}
-\hat{H}_\text{ab} = -\frac{1}{2} \sum_{i} \nabla_i^2 -\frac{1}{2\mu}\sum_{I} \nabla_I^2 + \frac{1}{2}\sum_{ij} \frac{1}{|r_i - r_j|} + \frac{1}{2}\sum_{IJ} \frac{Z_I Z_J}{|R_I - R_J|} - \sum_{iJ}\frac{Z_J}{|r_i - R_J|},\ \mu = (M/m)
-\label{Hab}
-\end{equation}
-where $i$ index electrons, $I$ ions, and energy in Hartree (Ha).
-The procedure makes no assumptions about the form of microscopic electronic interactions in the system and reports a clear, statistical measure of model quality.
-I propose using a quantum Monte Carlo (QMC) DMD procedure to develop accurate many-body model Hamiltonians for single (SLG) and bi-layer (BLG) graphene systems.
-Recent observations of a fractional quantum Hall effect \cite {Bolotin2009, Du2009} and plasmarons \cite{Bostwick999} in SLG as well as unconventional superconductivity \cite{Cao2018_2} and many-body insulating states \cite{Cao2018} in TBLG have spurred an interest in developing many-body models for graphene.
-By developing many-body models for SLG and BLG, I hope to further the understanding of interactions in graphene like systems and the maturity of DMD for its application to other strongly correlated systems like TBLG.
+A recently developed density matrix downfolding (DMD) method provides a more direct link between \textit{ab-initio} and many-body effective Hamiltonians.
+Similar to the methods above, DMD uses information from \textit{ab-initio} simulations on the system to fit the best parameters for a given many-body model \textit{ansatz}.
+Unlike the methods above, the resultant fit model is accompanied by a clear measure of model quality, quantifying the ability of the model to reproduce the low-energy spectrum of the system.
+Since model quality is reported, DMD allows for direct comparison between different model \textit{ansatzes} as well, an ability which can be used to learn the most important interaction terms for a given system directly from \textit{ab-initio} calculations.
+For example, DMD calculations on benzene and single layer graphene determined that long-range coulomb interactions are necessary in modeling both systems.
+
+The utility of DMD is made apparent when studying the recent surge in model development for twisted bilayer graphene (TBLG).
+\textit{Ab-initio} model development has so far been restricted to tight-binding models which accurately describe single particle effects like band flattening, but without a report of model quality cannot determine the role of interactions in the system.
+Minimal interacting models have been proposed using traditional theoretical methods, but without a link to \textit{ab-initio} it is unclear whether these simplified models can capture the complexity of the real system.
+Starting with the tight-binding \textit{ansatz}, DMD can be used to learn the most important interactions in the system, resulting in a final many-body model which can accurately describe TBLG.
+
+As such, one portion of a project between research groups at UIUC, UCSB and Rice University is dedicated towards the development of an accurate interacting model for TBLG.
+The model development consists of two layers of downfolding: first, developing an atomic scale lattice model for TBLG using \textit{ab-initio} simulations, and second, downfolding the resultant lattice model to a continuum length scale.
+My thesis will be focused on the first layer of downfolding, which I will conduct within the DMD framework using accurate \textit{ab-initio} fixed-node diffusion Monte Carlo calculations.
+I propose a two year plan which will result in atomic scale, many-body lattice Hamiltonians for single- (SLG) and simple bi-layer (BLG) graphene systems.
+My work will lay the necessary ground work for the development of a fully twist angle dependent lattice model for TBLG, and the eventual continuum model described above.
 

### 2019-09-08 reoutline.tex
intro done, need to do abstract + references

@@
 \begin{enumerate}
-\item Developing accurate model Hamiltonians for strongly correlated materials remains an outstanding problem in modern condensed matter physics.
+\item Developing model Hamiltonians for realistic materials is a fundamental problem in modern condensed matter physics.
 
-\item To bridge this gap, a framework for model development using solutions from \textit{ab-initio} simulations has emerged.
+\item Yet unresolved is a paradigm for building many-body model Hamiltonians which can describe strongly correlated systems.
 
-\item A recently developed density matrix downfolding (DMD) method addresses the need for a more systematic link between \textit{ab-initio} and model Hamiltonians.
-%No QMC here
+\item In an attempt to build accurate many-body Hamiltonians for realistic systems, a framework for model development using solutions from \textit{ab-initio} simulations has emerged.
 
-\item The surge in model development for twisted bilayer graphene (TBLG) exemplifies the necessity for such a systematic framework.
+\item A recently developed density matrix downfolding (DMD) method provides a more direct link between \textit{ab-initio} and many-body effective Hamiltonians.
 
-\item I propose using DMD to systematically develop accurate, many-body model Hamiltonians for single- (SLG) and simple bi-layer (BLG) graphene systems.
-%Start this paragraph with the grant proposal, have my thesis statement in the middle makes more sense!
-%Name drop QMC
+\item The utility of DMD is made apparent when studying the recent surge in model development for twisted bilayer graphene (TBLG).
+
+\item As such, one portion of a project between research groups at UIUC, UCSB and Rice University is dedicated towards the development of an accurate interacting model for TBLG.
 \end{enumerate}

### 2019-09-08 draft.tex
added extra refs to intro + rewrote abstract

@@
 \begin{abstract}
-A core problem of condensed matter physics is understanding how to develop low-energy effective models for realistic systems.
-I believe that a density matrix downfolding (DMD) method \cite{Zheng2017} can be used to systematically build effective theories for such systems beginning with a high energy theory like the \textit{ab-initio} Hamiltonian.
-I present as a preliminary result an effective Hamiltonian developed using DMD which can accurately reproduce the spectra and properties of the lowest lying excitations of the CuO molecule.
-I propose using DMD to develop accurate model Hamiltonians for single and bilayer graphene.
-My work will serve to mature the DMD method and take steps towards a model for the strongly correlated twisted bilayer graphene system.
+Building accurate model Hamiltonians for realistic materials is a fundamental problem in modern condensed matter physics.
+A recently developed density matrix downfolding (DMD) \cite{Zheng2017} method provides a very direct framework for downfolding the \textit{ab-initio} Hamiltonian to a many-body effective Hamiltonian using only \textit{ab-initio} calculations.
+Here I propose a two year thesis plan for developing atomic scale many-body model Hamiltonians for single- and simple bi-layer graphene systems within the DMD framework using high accuracy \textit{ab-initio} quantum Monte Carlo calculations.
+The results of my thesis will lay the groundwork for a continuing four year project to build a twist dependent many-body continuum model for the twisted bilayer graphene (TBLG) system.
+To demonstrate my preparedness I present a model Hamiltonian developed using DMD which accurately reproduces the spectra and properties of the lowest lying excitations of the CuO molecule.
 \end{abstract}
@@
 Band and Fermi liquid theory have been successful in accurately modeling systems where effective interactions are negligible at low energy scales.
-These theories, however, fail in describing the low-energy excitations of strongly correlated systems like high Tc superconductors and twisted bilayer graphene (TBLG).
-Many-body models like the Hubbard \cite{Hubbard1963}, Kanamori \cite{10.1143/PTP.30.275}, t-J \cite{Chao_1977} and Heisenberg models have been proposed to describe such strongly correlated systems, but their effectiveness in describing realistic materials is unclear.
+These theories, however, fail in describing the low-energy excitations of strongly correlated systems like high Tc superconductors and TBLG.
+Minimal many-body models like the Hubbard \cite{Hubbard1963}, Kanamori \cite{10.1143/PTP.30.275}, t-J \cite{Chao_1977} and Heisenberg models have been proposed to describe such strongly correlated systems, their effectiveness in describing realistic materials is unclear.
 
 In an attempt to build accurate many-body Hamiltonians for realistic systems, a framework for model development using solutions from \textit{ab-initio} simulations has emerged.
-Common approaches involve density functional theory (DFT) \cite{PhysRev.140.A1133} with one-body terms obtained from the DFT band structure \cite{PhysRevLett.87.047003} and two-body terms developed using constrained DFT, RPA or other methods \cite{PhysRevB.62.R16219, PhysRevB.70.195104, PhysRevB.88.075106}.
+Common approaches involve density functional theory (DFT) \cite{PhysRev.140.A1133} with one-body terms obtained from the DFT band structure \cite{PhysRevLett.87.047003} and two-body terms developed using constrained DFT, random phase approximation (RPA) or other methods \cite{PhysRevB.62.R16219, PhysRevB.70.195104, PhysRevB.88.075106}.
 L\"owdin methods and canonical transformations have also been used to develop effective theories \cite{doi:10.1063/1.4802766, PhysRevA.81.013402, doi:10.1063/1.2196410}.
 While useful, it is typically impossible to know whether models built from these techniques accurately describe reality, as measures of model quality are rarely reported.
-Instead, out-of-method comparisons of model predictions to experiment must be conducted for model validation.
+Instead, out-of-method comparisons of model predictions to experiment are conducted for model validation.
 
-A recently developed density matrix downfolding (DMD) method provides a more direct link between \textit{ab-initio} and many-body effective Hamiltonians.
+The recently developed DMD \cite{Zheng2017} method provides a more direct link between \textit{ab-initio} and many-body effective Hamiltonians.
 Similar to the methods above, DMD uses information from \textit{ab-initio} simulations on the system to fit the best parameters for a given many-body model \textit{ansatz}.
@@
 Since model quality is reported, DMD allows for direct comparison between different model \textit{ansatzes} as well, an ability which can be used to learn the most important interaction terms for a given system directly from \textit{ab-initio} calculations.
-For example, DMD calculations on benzene and single layer graphene determined that long-range coulomb interactions are necessary in modeling both systems.
+DMD has been used to develop accurate many-body Hamiltonians for benzene, a 1-d hydrogen chain, and a preliminary on-site Hubbard model for graphene \cite{Zheng2017, Wagner2015}.
 
-The utility of DMD is made apparent when studying the recent surge in model development for twisted bilayer graphene (TBLG).
-\textit{Ab-initio} model development has so far been restricted to tight-binding models which accurately describe single particle effects like band flattening, but without a report of model quality cannot determine the role of interactions in the system.
-Minimal interacting models have been proposed using traditional theoretical methods, but without a link to \textit{ab-initio} it is unclear whether these simplified models can capture the complexity of the real system.
+The utility of DMD is made apparent when studying the recent surge in model development for TBLG.
+\textit{Ab-initio} model development has so far been restricted to tight-binding models which accurately describe single particle effects like band flattening, but without a report of model quality cannot determine the role of interactions in the system \cite{Fang, Fang2016, PhysRevX.8.031088, PhysRevX.8.031087}.
+Minimal interacting models have been proposed using traditional theoretical methods, but without a link to \textit{ab-initio} it is unclear whether these simplified models can capture the complexity of the real system \cite{PhysRevB.98.045103, PhysRevB.100.125116, PhysRevX.8.031089}.
 Starting with the tight-binding \textit{ansatz}, DMD can be used to learn the most important interactions in the system, resulting in a final many-body model which can accurately describe TBLG.
 
-As such, one portion of a project between research groups at UIUC, UCSB and Rice University is dedicated towards the development of an accurate interacting model for TBLG.
+As such, one portion of a project between research groups at UIUC, UCSB and Rice University is dedicated towards the development of an accurate continuum-scale interacting model for TBLG.
 The model development consists of two layers of downfolding: first, developing an atomic scale lattice model for TBLG using \textit{ab-initio} simulations, and second, downfolding the resultant lattice model to a continuum length scale.
@@
 I propose a two year plan which will result in atomic scale, many-body lattice Hamiltonians for single- (SLG) and simple bi-layer (BLG) graphene systems.
-My work will lay the necessary ground work for the development of a fully twist angle dependent lattice model for TBLG, and the eventual continuum model described above.
+My thesis will lay the necessary ground work for the development of a fully twist angle dependent lattice model for TBLG, and the eventual continuum model described above.
 

### 2019-09-08 draft.tex
almost done with intro + methods

@@
 I begin by defining what a low-energy effective Hamiltonian and downfolding are.
-Consider a Hamiltonian $\hat{H}$ defined on a Hilbert space $\mathcal{H}$.
+Consider a high-energy Hamiltonian $\hat{H}$ and Hilbert space $\mathcal{H}$.
 A low-energy effective Hamiltonian is an operator $\hat{H}_\text{eff}$ which can accurately reproduce the spectra, up to a constant, and eigenstates of $\hat{H}$ on a subpsace $\mathcal{LE}$ of the Hilbert space $\mathcal{H}$.
@@
 The objective of Hamiltonian downfolding is to construct such an $\hat{H}_\text{eff}$ given $\hat{H}, \mathcal{H}$.
+Throughout this proposal the Hamiltonian $\hat{H}$ will be taken to be the \textit{ab-initio} Hamiltonian under the Born-Oppenheimer approximation
+\begin{equation}
+\hat{H}_\text{ab} = \sum_i -\frac{1}{2} \nabla_i^2 + \sum_{iI}\frac{Z_I}{|r_i - R_I|} + \frac{1}{2} \sum_{ij} \frac{1}{|r_i - r_j|} + E_\text{ion}(\{R_I\})  
+\label{eq:Hab}
+\end{equation} 
+where index $i,j$ correspond to electrons, $I$ to ions, and $E_\text{ion}$ is a constant energy equal to the ionic part of the Hamiltonian. 
+The unit of energy is Hartree (Ha).
 
 The key insight of DMD is that the Hamiltonian downfolding problem can be mapped onto a linear regression problem.
-It has been shown that the above definition of $\hat{H}_\text{eff}$ is equivalent to the following: a low-energy effective Hamiltonian is an operator on a subspace $\mathcal{LE}$ of $\mathcal{H}$ such that
+It has been shown \cite{Zheng2017} that the above definition of $\hat{H}_\text{eff}$ is equivalent to the following: a low-energy effective Hamiltonian is an operator on a subspace $\mathcal{LE}$ of $\mathcal{H}$ such that
 \begin{equation}
 \begin{split}
-\forall |\Psi\rangle \in \mathcal{LE},\ (E[\Psi] \equiv \langle \Psi|\hat{H} | \Psi \rangle)  = (E_\text{eff}[\Psi] \equiv \langle \Psi | \hat{H}_\text{eff} | \Psi \rangle) + \epsilon[\Psi]
+\forall |\Psi\rangle \in \mathcal{LE},\ E[\Psi] = E_\text{eff}[\Psi]  + \epsilon[\Psi]
 \end{split}
@@
 \end{equation}
-where $\epsilon[\Psi]$ is the error in the effective theory.
+where $\epsilon[\Psi]$ is the error in the effective theory and the energy functionals are defined as $E[\Psi] \equiv \langle \Psi | \hat{H} |\Psi \rangle,\  E_\text{eff}[\Psi] \equiv \langle \Psi | \hat{H}_\text{eff} |\Psi \rangle$.
 Given the general linear form of $\hat{H}_\text{eff}$ seen in \eqref{eq:Heff}, 
@@
 
-This linear regression problem can be tackled in three steps, beginning with sampling states from $\mathcal{LE}$.
-Ideally the linear regression would be conducted over the entire space $\mathcal{LE}$. 
-In practice, the space is too large, and instead a representative sample set is drawn and used for fitting.
-Generally, one cannot sample the true low-energy subspace, and low-energy wave functions are sampled from an approximate subspace $\mathcal{LE}^\prime$.
-For each sampled state $|\Psi_s\rangle$, the quantities $E[\Psi_s], \{d_k[\Psi_s]\}$ are calculated to be used in the linear regression.
-The sampling and calculation are typically done in QMC, but less accurate methods like DFT and Hartree-Fock can be used as well, resulting is less accurate models.
+This linear regression problem can be tackled in three steps.
+First, low-energy states are sampled from the \textit{ab-initio} Hilbert space.
+Next, a pool of candidate operators $\{\hat{d}_k\}$ which may be included in the final model are selected.
+Finally, a linear regression is carried out with descriptor and \textit{ab-initio} energies evaluated on the sampled states as the independent and target variables.
 
-Next, a set of candidate descriptors which form the effective Hamiltonian are selected.
-Descriptors will typically be selected if their values co-vary with the energy over the set of sampled states. 
-Knowledge of the low-energy degrees of the freedom from experiment and physical intuition can help restrict the set of candidate descriptors further.
-In principle there is no restriction on the form of the candidate operators other than Hermiticity, but second quantized operators are commonly used.
+Sampling low-energy states  %Last, hard paragraph!
+
+Candidate descriptors are typically selected using a mix of physical intuition and statistical methods.
+An essential first pass involves selecting a large set of descriptors using knowledge of the low-energy degrees of the freedom from experiment and physical principles like symmetry.
+Further filtering follows by examining the covariance between descriptors and the \textit{ab-initio} energy over the low-energy sampled states: descriptors will typically be discarded if they do not exhibit covariance with the energy.
+Note, while there is no restriction on the form of the candidate operators other than Hermiticity, 1- and 2-body second quantized operators are commonly used.
 If so, an appropriate single particle basis to express the second quantized operators must be constructed.
 
-Finally, the sampled data are used to fit the coefficients $\{g_k\}$ by linear regression.
 A typical fitting workflow involves feature selection, model regression and model validation.
-Feature selection methods for choosing appropriate descriptors can range from wrapper methods like orthogonal matching pursuit \cite{Cai2011} and principal component analysis \cite{pearson_karl_1901_1430636} to embedded methods like LASSO \cite{10.2307/2346178}, ridge \cite{doi:10.1137/S0895479897326432} and elastic net regularizations \cite{Zou05regularizationand}.
+Feature selection methods for choosing descriptors from a candidate pool can range from wrapper methods like orthogonal matching pursuit \cite{Cai2011} and principal component analysis \cite{pearson_karl_1901_1430636} to embedded methods like LASSO \cite{10.2307/2346178}, ridge \cite{doi:10.1137/S0895479897326432} and elastic net regularizations \cite{Zou05regularizationand}.
 Fitting the effective Hamiltonian usually involves an ordinary least squares linear regression, but the cost function can be altered.
-Model validation can be carried out by calculating cross validated single parameter measures like R$^2$ scores which carry information about goodness of fit and overfitting.
+Model validation can be carried out by calculating cross validated single parameter measures like $R^2$ scores which carry information about goodness of fit and overfitting.
 
-Importantly, the fit effective Hamiltonian carries with it a quantitative measure of its validity.
-The ability to quantify the accuracy of an effective Hamiltonian sets DMD apart from contemporary downfolding approaches which typically report only a final model.
-Given the statistical nature of the fitting procedure, model quality assessment can be carried out and reported in exhaustive detail if one is interested.
-Typically, however, single parameter measures like cross validated R$^2$ scores are used to quantify the accuracy of the effective Hamiltonian.
+The results of the linear regression are an estimate of the best parameter values $g_k$ for the candidate operators and an estimate of the model error $\epsilon[\Psi]$.
+The error estimate usually takes the form of a goodness of fit, such as a cross-validated $R^2$ score or root mean squared error (RMSE).
+These error estimates present a direct measurement of model quality, ensuring that DMD not only gives the best fit parameters but also an assessment of model validity.
+Further, by quantifying model quality through statistical error estimates, it is possible to compare the accuracy of various low-energy \textit{ansatzes} for a given system.
 
@@
 
-One can take advantage of the variational nature of FN-DMC to sample the low-energy states necessary for DMD.
+While seemingly disadvantageous, the variational nature of FN-DMC plays
+a key role in generating the low-energy states needed for DMD.
 Consider a set of trial wave functions with varying nodal structures.
 The final projected $\Psi_\text{DMC}$ for these different trial functions will be the lowest energy states in $\mathcal{H}$ for the given nodal structures.
-If the initial trial wave functions are appropriately chosen, the final projected states will be low-energy states within $\mathcal{H}$.
-The expectation values $E[\Psi_\text{DMC}], \{d_k[\Psi_\text{DMC}]\}$ necessary for the model regression are calculated using a mixed estimator.
+If the initial trial wave functions are appropriately chosen, the final projected states will be within the target low-energy subspace or $\mathcal{H}$.
 
+The training data required for downfolding are extracted from the projected low-energy states via a mixed estimator.
+Due to the fixed-node approximation, the final distribution returned by the projection is $f(R) = \Psi_T(R) \Psi_\text{DMC}(\tau \rightarrow \infty, R)$, and Monte Carlo estimates of operator expectation values correspond to the mixed estimator:
+\begin{equation}
+\frac{\langle \Psi_T | \hat{O} | \Psi_\text{DMC} (\tau \rightarrow \infty, R) \rangle}{\langle \Psi_T | \Psi_\text{DMC} (\tau \rightarrow \infty, R) \rangle}.
+\end{equation}
+which does not equal the expectation value we are after: $\langle \hat{O} \rangle_{\Psi_\text{DMC}(\tau \rightarrow \infty, R)}$.
+This results in a mixed estimator error which can be reduced using the results of an additional variational Monte Carlo calculation.
+Details regarding the mixed estimator error in DMD and evaluation of expectation values for 1- and 2-body second quantized operators can be found in the literature \cite{Wagner2015, Wagner2017}.
 \section{Preliminary Work}

### 2019-09-08 draft.tex
finished cleaneup for intro + methods

@@
 \patchcmd{\thebibliography}{\section*{\refname}}{}{}{}
+\usepackage{sectsty}
+\sectionfont{\fontsize{12}{15}\selectfont}
+\subsectionfont{\fontsize{12}{15}\selectfont}
+\usepackage{titlesec}
+\titlespacing*{\section}{0pt}{0.2\baselineskip}{0.2\baselineskip}
+\titlespacing*{\subsection}{0pt}{0.2\baselineskip}{0.2\baselineskip}
+\setlength{\abovedisplayskip}{0pt}
+\setlength{\belowdisplayskip}{0pt}
 \author{Shivesh Pathak}
@@
 A recently developed density matrix downfolding (DMD) \cite{Zheng2017} method provides a very direct framework for downfolding the \textit{ab-initio} Hamiltonian to a many-body effective Hamiltonian using only \textit{ab-initio} calculations.
-Here I propose a two year thesis plan for developing atomic scale many-body model Hamiltonians for single- and simple bi-layer graphene systems within the DMD framework using high accuracy \textit{ab-initio} quantum Monte Carlo calculations.
+Here I propose a two year thesis plan for developing atomic scale many-body model Hamiltonians for single- and simple bi-layer graphene systems within the DMD framework using high accuracy \textit{ab-initio} quantum Monte Carlo (QMC) calculations.
 The results of my thesis will lay the groundwork for a continuing four year project to build a twist dependent many-body continuum model for the twisted bilayer graphene (TBLG) system.
@@
 These theories, however, fail in describing the low-energy excitations of strongly correlated systems like high Tc superconductors and TBLG.
-Minimal many-body models like the Hubbard \cite{Hubbard1963}, Kanamori \cite{10.1143/PTP.30.275}, t-J \cite{Chao_1977} and Heisenberg models have been proposed to describe such strongly correlated systems, their effectiveness in describing realistic materials is unclear.
+Minimal many-body models like the Hubbard \cite{Hubbard1963}, Kanamori \cite{10.1143/PTP.30.275}, t-J \cite{Chao_1977} and Heisenberg models have been proposed to describe such systems, their effectiveness in describing realistic materials is unclear.
 
 In an attempt to build accurate many-body Hamiltonians for realistic systems, a framework for model development using solutions from \textit{ab-initio} simulations has emerged.
-Common approaches involve density functional theory (DFT) \cite{PhysRev.140.A1133} with one-body terms obtained from the DFT band structure \cite{PhysRevLett.87.047003} and two-body terms developed using constrained DFT, random phase approximation (RPA) or other methods \cite{PhysRevB.62.R16219, PhysRevB.70.195104, PhysRevB.88.075106}.
+Common approaches involve density functional theory (DFT) \cite{PhysRev.140.A1133, PhysRevLett.87.047003} alongside random phase approximation (RPA) or other methods \cite{PhysRevB.62.R16219, PhysRevB.70.195104, PhysRevB.88.075106}.
 L\"owdin methods and canonical transformations have also been used to develop effective theories \cite{doi:10.1063/1.4802766, PhysRevA.81.013402, doi:10.1063/1.2196410}.
@@
 Unlike the methods above, the resultant fit model is accompanied by a clear measure of model quality, quantifying the ability of the model to reproduce the low-energy spectrum of the system.
-Since model quality is reported, DMD allows for direct comparison between different model \textit{ansatzes} as well, an ability which can be used to learn the most important interaction terms for a given system directly from \textit{ab-initio} calculations.
-DMD has been used to develop accurate many-body Hamiltonians for benzene, a 1-d hydrogen chain, and a preliminary on-site Hubbard model for graphene \cite{Zheng2017, Wagner2015}.
+Since model quality is reported, DMD allows for direct comparison between different model \textit{ansatzes} as well, an ability which can be used to deduce important model terms using \textit{ab-initio} calculations.
+DMD has been used to develop accurate many-body models for benzene, a 1-d hydrogen chain, and an on-site Hubbard model for graphene \cite{Zheng2017, Wagner2015}.
 
 The utility of DMD is made apparent when studying the recent surge in model development for TBLG.
-\textit{Ab-initio} model development has so far been restricted to tight-binding models which accurately describe single particle effects like band flattening, but without a report of model quality cannot determine the role of interactions in the system \cite{Fang, Fang2016, PhysRevX.8.031088, PhysRevX.8.031087}.
-Minimal interacting models have been proposed using traditional theoretical methods, but without a link to \textit{ab-initio} it is unclear whether these simplified models can capture the complexity of the real system \cite{PhysRevB.98.045103, PhysRevB.100.125116, PhysRevX.8.031089}.
-Starting with the tight-binding \textit{ansatz}, DMD can be used to learn the most important interactions in the system, resulting in a final many-body model which can accurately describe TBLG.
+\textit{Ab-initio} model development has so far been restricted to tight-binding models which accurately describe single particle effects like band flattening, but do not address interactions \cite{Fang, Fang2016, PhysRevX.8.031088, PhysRevX.8.031087}.
+Minimal interacting models have been proposed using traditional theoretical methods, but without a link to \textit{ab-initio} it is unclear whether they can describe the real system \cite{PhysRevB.98.045103, PhysRevB.100.125116, PhysRevX.8.031089}.
+DMD can bridge the gap between these methods by providing a framework to build an accurate many-body model directly from \textit{ab-initio} simulations on TBLG.
 
 As such, one portion of a project between research groups at UIUC, UCSB and Rice University is dedicated towards the development of an accurate continuum-scale interacting model for TBLG.
-The model development consists of two layers of downfolding: first, developing an atomic scale lattice model for TBLG using \textit{ab-initio} simulations, and second, downfolding the resultant lattice model to a continuum length scale.
-My thesis will be focused on the first layer of downfolding, which I will conduct within the DMD framework using accurate \textit{ab-initio} fixed-node diffusion Monte Carlo calculations.
-I propose a two year plan which will result in atomic scale, many-body lattice Hamiltonians for single- (SLG) and simple bi-layer (BLG) graphene systems.
-My thesis will lay the necessary ground work for the development of a fully twist angle dependent lattice model for TBLG, and the eventual continuum model described above.
+The model development consists of two layers of downfolding: first, developing an atomic scale lattice model using \textit{ab-initio} calculations, and second, downfolding the resultant lattice model to a continuum length scale.
+My thesis will be focused on the first layer of downfolding.
+I propose a two year plan for developing many-body lattice models for single- (SLG) and simple bi-layer (BLG) graphene systems using the DMD method and \textit{ab-initio} fixed-node diffusion Monte Carlo (FN-DMC) calculations.
+My thesis will lay the necessary ground work for the development of a fully twist angle dependent lattice model for TBLG, and an eventual continuum model.
 
@@
 Consider a high-energy Hamiltonian $\hat{H}$ and Hilbert space $\mathcal{H}$.
-A low-energy effective Hamiltonian is an operator $\hat{H}_\text{eff}$ which can accurately reproduce the spectra, up to a constant, and eigenstates of $\hat{H}$ on a subpsace $\mathcal{LE}$ of the Hilbert space $\mathcal{H}$.
-Here $\mathcal{LE}$ is taken to be the span of the N lowest energy eigenstates of $\hat{H}$, and $\hat{H}_\text{eff}$ is a linear combination of Hermitian operators $\hat{d}_k$ with a constant energy shift $E_0$
+A low-energy effective Hamiltonian is an operator $\hat{H}_\text{eff}$ which can accurately reproduce the spectra and eigenstates of $\hat{H}$ on a subpsace $\mathcal{LE}$ of the Hilbert space $\mathcal{H}$.
+Here $\mathcal{LE}$ is the span of the N lowest energy eigenstates of $\hat{H}$, and $\hat{H}_\text{eff}$ is a linear combination of Hermitian operators $\hat{d}_k$ with a constant shift $E_0$
 \begin{equation}
@@
 The objective of Hamiltonian downfolding is to construct such an $\hat{H}_\text{eff}$ given $\hat{H}, \mathcal{H}$.
-Throughout this proposal the Hamiltonian $\hat{H}$ will be taken to be the \textit{ab-initio} Hamiltonian under the Born-Oppenheimer approximation
+Throughout this proposal $\hat{H}$ is the \textit{ab-initio} Hamiltonian under the Born-Oppenheimer approximation
 \begin{equation}
@@
 The key insight of DMD is that the Hamiltonian downfolding problem can be mapped onto a linear regression problem.
-It has been shown \cite{Zheng2017} that the above definition of $\hat{H}_\text{eff}$ is equivalent to the following: a low-energy effective Hamiltonian is an operator on a subspace $\mathcal{LE}$ of $\mathcal{H}$ such that
+It has been shown \cite{Zheng2017} that the above definition of $\hat{H}_\text{eff}$ is equivalent to the following: a low-energy effective Hamiltonian is an operator such that 
 \begin{equation}
@@
 
-Sampling low-energy states  %Last, hard paragraph!
+Low-energy states can be accessed using any wave function based \textit{ab-initio} simulation technique, but two broad principles are used for selecting a method.
+First, the method should be able to generate accurate low-energy wave functions in the \textit{ab-initio} Hilbert space, as the accuracy of the low-energy states dictate the accuracy of the final model.
+Second, the method should be relatively inexpensive, as large amounts of training data is preferred for a more robust regression.
+QMC methods have typically been selected due to their polynomial scaling computational cost and high accuracy, but DFT can also be used as an initial step trading accuracy for computational ease.
 
-Candidate descriptors are typically selected using a mix of physical intuition and statistical methods.
-An essential first pass involves selecting a large set of descriptors using knowledge of the low-energy degrees of the freedom from experiment and physical principles like symmetry.
-Further filtering follows by examining the covariance between descriptors and the \textit{ab-initio} energy over the low-energy sampled states: descriptors will typically be discarded if they do not exhibit covariance with the energy.
+Candidate descriptors are typically selected using both physical intuition and statistical methods.
+An essential first pass involves selecting a large set of descriptors using knowledge of the low-energy degrees of the freedom of the system and physical principles like symmetry.
+Further filtering follows by examining the relationship between descriptors and the \textit{ab-initio} energy over the low-energy sampled states: descriptors will typically be discarded if they do not exhibit covariance with the energy.
 Note, while there is no restriction on the form of the candidate operators other than Hermiticity, 1- and 2-body second quantized operators are commonly used.
-If so, an appropriate single particle basis to express the second quantized operators must be constructed.
 
-A typical fitting workflow involves feature selection, model regression and model validation.
-Feature selection methods for choosing descriptors from a candidate pool can range from wrapper methods like orthogonal matching pursuit \cite{Cai2011} and principal component analysis \cite{pearson_karl_1901_1430636} to embedded methods like LASSO \cite{10.2307/2346178}, ridge \cite{doi:10.1137/S0895479897326432} and elastic net regularizations \cite{Zou05regularizationand}.
+A standard fitting workflow involves feature selection, model regression and model validation.
+Feature selection methods for choosing descriptors from the candidate pool can range from wrapper methods like principal component analysis \cite{pearson_karl_1901_1430636} to regularized embedded methods LASSO \cite{10.2307/2346178, doi:10.1137/S0895479897326432, Zou05regularizationand}.
 Fitting the effective Hamiltonian usually involves an ordinary least squares linear regression, but the cost function can be altered.
@@
 
-The results of the linear regression are an estimate of the best parameter values $g_k$ for the candidate operators and an estimate of the model error $\epsilon[\Psi]$.
+The results of the linear regression are an estimate of the parameter values $g_k$ and model error $\epsilon[\Psi]$.
 The error estimate usually takes the form of a goodness of fit, such as a cross-validated $R^2$ score or root mean squared error (RMSE).
 These error estimates present a direct measurement of model quality, ensuring that DMD not only gives the best fit parameters but also an assessment of model validity.
-Further, by quantifying model quality through statistical error estimates, it is possible to compare the accuracy of various low-energy \textit{ansatzes} for a given system.
+Further, by quantifying model quality through statistical error estimates, it is possible to compare the accuracy of various low-energy \textit{ansatzes}.
 
 \subsection{Fixed-node diffusion Monte Carlo (FN-DMC)}
-Diffusion Monte Carlo (DMC) is a quantum Monte Carlo method which projects out the ground state of a real-space Hamiltonian given some initial trial wave function.
+Diffusion Monte Carlo (DMC) is a QMC method which projects out the ground state of a real-space Hamiltonian given some initial trial wave function.
 Consider a trial wave function $|\Psi_T\rangle$ and the Hamiltonian $\hat{H}$ with ground state $|\Phi_0\rangle$. 
@@
 projects out the ground state as long as the trial wave function is not orthogonal to the ground state. 
-DMC provides a stochastic implementation of this projector method and involves moving samples from the trial function $\Psi_T(R)$ using the Green function $G(R, R^\prime, \tau) = \langle R | e^{-\tau(\hat{H} - E_T)} | R^\prime \rangle$.
-Since $\hat{H} = \hat{T} + \hat{V}$, kinetic and potential energy terms, the Green function is approximated by a Trotter expansion 
-$$G(R, R^\prime, \tau) = \langle R | e^{-\tau(\hat{H} - E_T)} | R^\prime \rangle \sim \Big[e^{-d\tau(\frac{V(R) + V(R^\prime)}{2} - E_T)} \langle R| e^{-d\tau\hat{T}}|R^\prime \rangle + O(d\tau^2) \Big]^N $$
+DMC provides a stochastic implementation of the projection by moving samples from $\Psi_T(R)$ using the Green function $G(R, R^\prime, \tau) = \langle R | e^{-\tau(\hat{H} - E_T)} | R^\prime \rangle$, which is approximated using a Trotter expansion $G(R, R^\prime, \tau) \sim \Big[e^{-d\tau(\frac{V(R) + V(R^\prime)}{2} - E_T)} \langle R| e^{-d\tau\hat{T}}|R^\prime \rangle + O(d\tau^2) \Big]^N $
 where $d\tau = \tau/N$.
@@
 The final projected $\Psi_\text{DMC}$ for these different trial functions will be the lowest energy states in $\mathcal{H}$ for the given nodal structures.
-If the initial trial wave functions are appropriately chosen, the final projected states will be within the target low-energy subspace or $\mathcal{H}$.
+If the initial trial wave functions are appropriately chosen, the final projected states will be within the target low-energy subspace of $\mathcal{H}$.
 
 The training data required for downfolding are extracted from the projected low-energy states via a mixed estimator.
-Due to the fixed-node approximation, the final distribution returned by the projection is $f(R) = \Psi_T(R) \Psi_\text{DMC}(\tau \rightarrow \infty, R)$, and Monte Carlo estimates of operator expectation values correspond to the mixed estimator:
-\begin{equation}
-\frac{\langle \Psi_T | \hat{O} | \Psi_\text{DMC} (\tau \rightarrow \infty, R) \rangle}{\langle \Psi_T | \Psi_\text{DMC} (\tau \rightarrow \infty, R) \rangle}.
-\end{equation}
-which does not equal the expectation value we are after: $\langle \hat{O} \rangle_{\Psi_\text{DMC}(\tau \rightarrow \infty, R)}$.
-This results in a mixed estimator error which can be reduced using the results of an additional variational Monte Carlo calculation.
-Details regarding the mixed estimator error in DMD and evaluation of expectation values for 1- and 2-body second quantized operators can be found in the literature \cite{Wagner2015, Wagner2017}.
+Due to the fixed-node approximation, the final distribution returned by the projection is $f(R) = \Psi_T(R) \Psi_\text{DMC}$, and Monte Carlo estimates of operator expectation values correspond to the mixed estimation $\frac{\langle \Psi_T | \hat{O} | \Psi_\text{DMC} \rangle}{\langle \Psi_T | \Psi_\text{DMC}\rangle}$.
+This does not equal the target expectation value $\frac{\langle \Psi_\text{DMC} | \hat{O} | \Psi_\text{DMC} \rangle}{\langle \Psi_\text{DMC} | \Psi_\text{DMC}\rangle}$, resulting in a mixed estimator error.
+Details on alleviating this error in DMD and evaluation of expectation values for 1- and 2-body second quantized operators can be found in the literature \cite{Wagner2015, Wagner2017}.
 \section{Preliminary Work}
@@
 
+\vspace{-10pt}
 \paragraph{Sampling low-energy states}
@@
 
+\vspace{-10pt}
 \paragraph{Selecting candidate descriptors}
@@
 
+\vspace{-10pt}
 \paragraph{Fitting the model}
@@
 
+\vspace{-10pt}
 \paragraph{Solutions of the model}
@@
 
+\vspace{-10pt}
 \paragraph{Proposed timeline} I believe that developing effective theories for the benzene molecule, SLG, AA and AB stacked BLG should take 2 years. 

### 2019-09-08 draft.tex
small edits via yang

@@
 The objective of Hamiltonian downfolding is to construct such an $\hat{H}_\text{eff}$ given $\hat{H}, \mathcal{H}$.
-Throughout this proposal $\hat{H}$ is the \textit{ab-initio} Hamiltonian under the Born-Oppenheimer approximation
-\begin{equation}
-\hat{H}_\text{ab} = \sum_i -\frac{1}{2} \nabla_i^2 + \sum_{iI}\frac{Z_I}{|r_i - R_I|} + \frac{1}{2} \sum_{ij} \frac{1}{|r_i - r_j|} + E_\text{ion}(\{R_I\})  
-\label{eq:Hab}
-\end{equation} 
-where index $i,j$ correspond to electrons, $I$ to ions, and $E_\text{ion}$ is a constant energy equal to the ionic part of the Hamiltonian. 
-The unit of energy is Hartree (Ha).
 
@@
 This linear regression problem can be tackled in three steps.
-First, low-energy states are sampled from the \textit{ab-initio} Hilbert space.
+First, low-energy states are sampled from $\mathcal{LE}$.
 Next, a pool of candidate operators $\{\hat{d}_k\}$ which may be included in the final model are selected.
-Finally, a linear regression is carried out with descriptor and \textit{ab-initio} energies evaluated on the sampled states as the independent and target variables.
+Finally, a linear regression is carried out with descriptor and energies evaluated on the sampled states as the independent and target variables.
+
+When developing models for realistic system, $\hat{H}$ is taken to be the \textit{ab-initio} Hamiltonian
+\begin{equation}
+\hat{H}_\text{ab} = \sum_i -\frac{1}{2} \nabla_i^2 + \sum_{iI}\frac{Z_I}{|r_i - R_I|} + \frac{1}{2} \sum_{ij} \frac{1}{|r_i - r_j|} + E_\text{ion}(\{R_I\})  
+\label{eq:Hab}
+\end{equation} 
+where index $i,j$ correspond to electrons, $I$ to ions, and $E_\text{ion}$ corresponds to the ionic kinetic and potential energies. 
+The unit of energy is Hartree (Ha).
+To reduce the computational expense of \textit{ab-initio} calculations one works under the Born-Oppenheimer approximation where $E_\text{ion}$ becomes a constant.
+Ionic effects related to phonon modes can be recovered by altering the ionic geometry and observing the effects on the electronic state.
 
@@
 First, the method should be able to generate accurate low-energy wave functions in the \textit{ab-initio} Hilbert space, as the accuracy of the low-energy states dictate the accuracy of the final model.
-Second, the method should be relatively inexpensive, as large amounts of training data is preferred for a more robust regression.
+Second, the method should be relatively inexpensive, as large amounts of training data are preferred for a more robust regression.
 QMC methods have typically been selected due to their polynomial scaling computational cost and high accuracy, but DFT can also be used as an initial step trading accuracy for computational ease.
@@
 A standard fitting workflow involves feature selection, model regression and model validation.
-Feature selection methods for choosing descriptors from the candidate pool can range from wrapper methods like principal component analysis \cite{pearson_karl_1901_1430636} to regularized embedded methods LASSO \cite{10.2307/2346178, doi:10.1137/S0895479897326432, Zou05regularizationand}.
+Feature selection methods for choosing descriptors from the candidate pool can range from wrapper methods like principal component analysis \cite{pearson_karl_1901_1430636} to regularized embedded methods like LASSO \cite{10.2307/2346178, doi:10.1137/S0895479897326432, Zou05regularizationand}.
 Fitting the effective Hamiltonian usually involves an ordinary least squares linear regression, but the cost function can be altered.

### 2019-09-09 draft.tex
updated preliminary work: MSJNO, CuO intro, fitting and results

@@
 \subsection{Non-orthogonal determinants in FN-DMC trial wave functions}
-In order to become acquainted with our code QWalk \cite{Wagner2009} and QMC algorithms in general I worked on implementing and testing multi-Slater-Jastrow trial functions with optimized non-orthogonal determinants (MSJ+NO) in FN-DMC \cite{doi:10.1063/1.5052906}.
+In order to become acquainted with QMC algorithms I worked on implementing and testing multi-Slater-Jastrow trial functions with optimized non-orthogonal determinants (MSJ+NO) in FN-DMC \cite{doi:10.1063/1.5052906}.
 The MSJ+NO trial wave functions take the form
@@
 \end{equation}
-where the determinants $|D_I\rangle$ are composed of single particle orbitals from a self-consistent field calculation, $e^J$ is a spin-independent Jastrow factor, $e^{\hat{\Theta}}$ is a spin-dependent orbital rotation matrix, and the wave function has variable parameters $\vec{\alpha}, \vec{\theta}, \vec{C}$ which can be optimized.
+where the determinants $|D_I\rangle$ are composed of single particle orbitals from a self-consistent field calculation, $e^J$ is a spin-independent Jastrow factor, and $e^{\hat{\Theta}}$ is a spin-dependent orbital rotation matrix.
+The wave function has variable parameters $\vec{\alpha}, \vec{\theta}, \vec{C}$ which can be optimized.
 
-We assessed the efficiency and compactness of this new trial function by calculating the ground state energy and single particle densities of a C$_2$ molecule using FN-DMC and comparing to the results when using multi-Slater-Jastrow trial functions with optimized orthogonal determinant trial functions (MSJ+O).
-The workflow involved constructing the un-optimized trial wave functions, optimizing the parameters using an energy optimization method \cite{Toulouse2007}, and finally using the optimized trial functions in both variational Monte Carlo (VMC) and FN-DMC calculations.
-We found the average improvement in variational energy when using optimized orthogonal determinants is $\langle E_\text{MSJ} - E_\text{MSJ+O} \rangle = $ 0.32 eV with a standard deviation of 0.08 eV, with an additional improvement from the non-orthogonal determinants of 0.031 eV with a standard deviation of 0.021 eV.
-The average improvement in the FN-DMC energy with the MSJ+O trial wave functions was 0.14 eV (standard deviation 0.03 eV) with an additional reduction of 0.032 eV (standard deviation 0.019 eV) when using the MSJ+NO trial wave function.
+I assessed the efficiency and compactness of the new function by calculating the ground state energy of a C$_2$ molecule using FN-DMC and comparing to the results when using multi-Slater-Jastrow trial functions with orthogonal determinants (MSJ+O).
+The workflow involved constructing the un-optimized trial wave functions, optimizing the parameters using an energy optimization method \cite{Toulouse2007}, and finally using the optimized trial functions in FN-DMC calculations.
+The average improvement in the FN-DMC energy with the MSJ+O trial wave functions was $\langle E_{MSJ} - E_{MSJ,O}\rangle =$ 0.14eV with an additional reduction of 0.032 eV when using the MSJ+NO trial wave function.
+While the benefit using the MSJ+NO trial wave function seems small, the energy calculated using an MSJ+NO trial function with only 24 determinants was lower than the energy using an MSJ+O trial 
+function with 55 determinants.
 
-While the benefit using the MSJ+NO trial wave function seems small, we saw that the FN-DMC energy calculated using an MSJ+NO trial function with only 24 determinants was lower than the FN-DMC energy using an MSJ+O trial function with 55 determinants.
 Further, the FN-DMC charge density calculated using MSJ+NO trial functions had stronger bonding character than when using MSJ+O trial functions, a reasonable result as introducing correlations into trial functions allows for electrons to avoid each other while still occupying the same bonding region.
-Our results indicated that using non-orthogonal determinants may lead to more compact multi-Slater-Jastrow trial wave functions for small molecules.
+The results indicated that using non-orthogonal determinants may lead to more compact trial wave functions for small molecules.
 
 \subsection{Effective theory for CuO molecule using DMD}
-As a first step towards developing accurate models for extended systems like solids using DMD, I constructed a many-body effective model for the CuO molecule by downfolding \eqref{Hab} under a Born-Oppenheimer approximation with fixed bond length $r_e$ = 1.725\r{A}.
-In the following four sections I will detail the steps I took towards developing a model for the CuO molecule using DMD.
-First, I discuss how low-energy states were sampled for the model regression.
-This is followed by details of descriptor selection and single particle basis construction.
-With the sampled data and candidate descriptors in hand, the feature selection and model fitting procedures are described.
-I will conclude by comparing the solutions of the model Hamiltonian to experiment and previous theoretical calculations.
+To become acquainted with DMD, I constructed a many-body effective model for the CuO molecule with fixed bond length $r_e$ = 1.725\r{A}.
+In the following three sections I will detail the steps I took towards developing a model for the CuO molecule using DMD, followed by a comparison of model spectra to experiment.
 
@@
 \paragraph{Fitting the model}
-A direct attempt at fitting all eleven terms in \eqref{eq:models} leads to a high goodness of fit, R$^2$ = 0.99, but unexpectedly large coefficients for the fit model, like an enormous occupation energy gap of $\epsilon_{2p_\pi} - \epsilon_{3d_\delta} = $ 5.87 eV, which is 2 $-$ 3 eV in experiment and our sampled data.
+A direct attempt at fitting all eleven terms in \eqref{eq:models} leads to a high goodness of fit, R$^2$ = 0.99, but unexpectedly large coefficients for the fit model.
+This is seen in an enormous occupation energy gap of $\epsilon_{2p_\pi} - \epsilon_{3d_\delta} = $ 5.87 eV, which is 2 $-$ 3 eV in experiment and our sampled data.
 This inflation of model parameters was found to be associated with multicollinearity among the candidate descriptors, quantified by a large condition number (CN) of 1470.
 
-A principal component regression (PCR) was used to alleviate this multicollinearity and fit the effective theory \cite{10.2307/2348005}.
-PCR can be broken into three parts: conducting a principal component analysis (PCA) and selecting candidate principal components to work with (PCs), fitting a linear model to these PCs, and then inverting the coefficients back to the original descriptors.
-If the subset of PCs are chosen appropriately, the final regressed model will have a similar goodness of fit to one fit on the original data matrix but without the consequence of serious multicollinearity.
+A principal components analysis (PCA) on the descriptors illustrates the multicollinearity in the data set.
+In Figure \eqref{fig:cuo_regr}b I show the cumulative explained variance ratio (CEVR) and CN of a matrix composed of the $N_\text{comp}$ highest ranked PCs varying from one to eleven.
+The CEVR indicates how much of the variation in the descriptor values a subset of PCs capture, and the CN quantifies the multicollinearity among these PCs.
+The sharp decrease in the condition number and sustained high CEVR up to nine principal components indicates the presence of only nine linearly independent variations within the descriptor set.
 
-From the PCA, the nine highest ranked PCs were selected to use in the PCR.
-In Figure \eqref{fig:cuo_regr}b we plot both the cumulative explained variance ratio (CEVR) and CN of a matrix composed of the $N_\text{comp}$ highest ranked PCs varying from one to eleven.
-The CEVR indicates how much of the variation in the descriptor values a subset of PCs capture, and the CN quantifies the multicollinearity among these PCs.
-The objective in selecting PCs is to the maximize the prior while simultaneously minimizing the latter, ensuring a high goodness of fit with little to no multicollinearity.
-Nine principal components best satisfies the objective above with a CEVR 0.9995 and CN of 257.
-
-The model fit to these nine PCs maintains an excellent goodness of fit without serious multicollinearity.
+To reduce multicollinearity while maintaining a high goodness of fit, I fit a model to these nine PCs.
 When trained on all the sampled data this model yields an $R^2$ = 0.98 with a CN of 257. 
@@
 These coefficients can be seen in Figure \ref{fig:cuo_regr}c with single standard deviation bootstrap error estimates.
+The procedure carried out here is well established and called principal components regression (PCR) \cite{10.2307/2348005}.
 

### 2019-09-09 reoutline.tex
updated preliminary work: MSJNO, CuO intro, fitting and results

@@
 
+\section{Non-orthogonal determinants in FN-DMC trial wave functions}
+\begin{enumerate}
+\item In order to become acquainted with our code QWalk and QMC algorithms in general I worked on implementing and testing multi-Slater-Jastrow trial functions with optimized nonorthogonal determinants (MSJ+NO) in FN-DMC.
+
+\item We assessed the efficiency and compactness of this new trial function by calculating the ground state energy and single particle densities of a C$_2$ molecule using FN-DMC and comparing to the results when using multi-Slater-Jastrow trial functions with optimized orthogonal determinant trial functions (MSJ+O).
+
+\item Our results indicated that using non-orthogonal determinants may
+lead to more compact multi-Slater-Jastrow trial wave functions for small molecules.
+\end{enumerate}
+
+\section{Effective theory for CuO molecule using DMD}
+\begin{enumerate}
+\item To become acquainted with DMD, I constructed a many-body effective model for the CuO molecule with fixed bond length r$_e$ = 1.725A.
+
+%sampling low energy states
+\item Explain why MSJ was chosen (Probably the hardest one)
+
+\item How determinants selected
+
+\item How coefficients selected
+
+\item How Jastrow optimized and FN-DMC done
+
+%candidate descriptors 
+\item Physical intuition and symmetry
+
+\item Statistics
+
+\item Final set and basis elements
+
+%Fitting the model 
+\item A direct attempt at fitting all eleven terms in \textbf{eqref} leads to a high goodness of fit, R$^2$ = 0.99, but unexpectedly large coefficients for the fit model.
+
+\item A principal components analysis (PCA) on the descriptors illustrates the multicollinearity in the data set.
+
+\item To reduce multicollinearity while maintaining a high goodness of fit, I fit a model to these nine PCs.
+
+%Solutions of the model 
+\item The model is solved by exact diagonalization, and the solutions agree well with experimental spectra and state assignments.
+
+\item The model states also agree with recent theoretical calculations like UKS and CCSD(T).
+
+\end{enumerate}
+
 \end{document}

### 2019-09-10 draft.tex
candidate descriptors cleaned, need to just do wf sampling

@@
 \paragraph{Selecting candidate descriptors}
-We selected the following set of 1- and 2-body operators to construct our candidate descriptors based on our understanding of the low-energy excitations of the CuO molecule and our sampled states:
+The following set of 1- and 2-body operators were selected as candidates for the model based on understanding of the low-energy excitations of the CuO molecule and properties of the sampled low-energy states:
 \begin{equation}
@@
 
-The parantheses of \eqref{eq:models} contain the full set of symmetry allowed 1-body operators within the Cu 3d, 4s and O 2p space.
-These operators are required to describe the most basic single particle features seen in molecules like orbital occupations and hybridizations.
-Two 2-body interaction terms are also considered: a Hund's coupling between Cu 4s and 3d orbitals and a Cu 4s Coulomb repulsion.
-The prior term is motivated by the existence of a $\sim $ 0.5 eV Hund's coupling on the neutral copper atom.
-Motivation for the latter comes from the FN-DMC energies of our sampled data.
-We find that states with a Cu $3d^{10} 4s^{2}$ configuration lie at least 7 eV above our lowest energy sampled state, while the highest energy states with a $3d^{10} 4s^{1}$ configuration lie near 2 eV.
+The first chunk of descriptors were selected using physical intuition and symmetry principles.
+As described above, the low-lying excitations in CuO observed in experiment involve the Cu 3d, 4s and O 2p atomic orbitals.
+Given that orbitals also hybridize, the occupation energies of and hopping terms between these orbitals were selected as candidate operators.
+Symmetry considerations help narrow down the set as hopping terms are only non-zero between same symmetry orbitals, and occupation energies can be traced over symmetry related orbitals, for example $n_{2p_\pi} = n_{2p_x} + n_{2p_y}$.
+The term $\sum_{i\in {\{\text{xy, xz, ...}}\}} \vec{S}_{4s} \cdot \vec{S}_{3d_i}$ was also considered, motivated by the existence of a $\sim $0.5 eV Hund's coupling observed on the copper atom \textbf{ref}.
+
+The inclusion of a double occupancy term $n_{4s,\uparrow} n_{4s,\downarrow}$ was motivated directly from the sampled low-energy states.
+Within the sampled data, states with a Cu $4s^{2}$ configuration lie at least 7 eV above the lowest energy sampled state, while the lowest energy state with a $4s^{1}$ configuration lies at 1.6 eV.
+If the double occupancy term was not included, the effective theory would predict the prior state to have an energy of $\sim $ 3eV, an underestimation of 4 eV from the \textit{ab-initio} result.
+The effective on-site Coulomb repulsion $U_s$ can ensure our model places all $4s^{2}$ states high in energy.
 

### 2019-09-10 reoutline.tex
candidate descriptors cleaned, need to just do wf sampling

@@
 %candidate descriptors 
-\item Physical intuition and symmetry
+\item The following set of 1- and 2-body operators were selected as candidates for the model based on understanding of the low-energy excitations of the CuO molecule and properties of the sampled low-energy states.
 
-\item Statistics
+\item The first chunk of descriptors were selected using physical intuition and symmetry principles.
 
-\item Final set and basis elements
+\item The inclusion of a double occupancy term $n_{4s,\uparrow} n_{4s,\downarrow}$ was motivated directly from the sampled low-energy states.
 

### 2019-09-10 draft.tex
updated biblio

@@
 Symmetry considerations help narrow down the set as hopping terms are only non-zero between same symmetry orbitals, and occupation energies can be traced over symmetry related orbitals, for example $n_{2p_\pi} = n_{2p_x} + n_{2p_y}$.
-The term $\sum_{i\in {\{\text{xy, xz, ...}}\}} \vec{S}_{4s} \cdot \vec{S}_{3d_i}$ was also considered, motivated by the existence of a $\sim $0.5 eV Hund's coupling observed on the copper atom \textbf{ref}.
+The term $\sum_{i\in {\{\text{xy, xz, ...}}\}} \vec{S}_{4s} \cdot \vec{S}_{3d_i}$ was also considered, motivated by the existence of a $\sim $0.5 eV Hund's coupling observed on the copper atom \cite{Data2009}.
 
@@
 If the double occupancy term was not included, the effective theory would predict the prior state to have an energy of $\sim $ 3eV, an underestimation of 4 eV from the \textit{ab-initio} result.
-The effective on-site Coulomb repulsion $U_s$ can ensure our model places all $4s^{2}$ states high in energy.
+The effective on-site Coulomb repulsion $U_s$ ensures the model places all $4s^{2}$ states high in energy.
 
@@
 
-To reduce multicollinearity while maintaining a high goodness of fit, I fit a model to these nine PCs.
+To reduce multicollinearity while maintaining a high goodness of fit, model was fit to these nine PCs.
 When trained on all the sampled data this model yields an $R^2$ = 0.98 with a CN of 257. 

### 2019-09-10 draft.tex
finished preliminary work. need to make bib shorter...

@@
 Building accurate model Hamiltonians for realistic materials is a fundamental problem in modern condensed matter physics.
-A recently developed density matrix downfolding (DMD) \cite{Zheng2017} method provides a very direct framework for downfolding the \textit{ab-initio} Hamiltonian to a many-body effective Hamiltonian using only \textit{ab-initio} calculations.
-Here I propose a two year thesis plan for developing atomic scale many-body model Hamiltonians for single- and simple bi-layer graphene systems within the DMD framework using high accuracy \textit{ab-initio} quantum Monte Carlo (QMC) calculations.
+A recently developed density matrix downfolding (DMD) \cite{Zheng2017} method provides a very direct framework for downfolding the \textit{ab initio} Hamiltonian to a many-body effective Hamiltonian using only \textit{ab initio} calculations.
+Here I propose a two year thesis plan for developing atomic scale many-body model Hamiltonians for single- and simple bi-layer graphene systems within the DMD framework using high accuracy \textit{ab initio} quantum Monte Carlo (QMC) calculations.
 The results of my thesis will lay the groundwork for a continuing four year project to build a twist dependent many-body continuum model for the twisted bilayer graphene (TBLG) system.
@@
 Accurate models give clear and direct indications of effective interactions within a material and underpin modern theoretical understanding of ideas like emergence and symmetry breaking.
-On a practical level, calculations related to these low-lying eigenstates can be carried out on the much simpler model system and still yield accurate results reflective of reality.
+On a practical level, calculations related to these low-lying eigenstates can be carried out on the much simpler model and still yield accurate results reflective of reality.
 
@@
 
-In an attempt to build accurate many-body Hamiltonians for realistic systems, a framework for model development using solutions from \textit{ab-initio} simulations has emerged.
+In an attempt to build accurate many-body Hamiltonians for realistic systems, a framework for model development using solutions from \textit{ab initio} simulations has emerged.
 Common approaches involve density functional theory (DFT) \cite{PhysRev.140.A1133, PhysRevLett.87.047003} alongside random phase approximation (RPA) or other methods \cite{PhysRevB.62.R16219, PhysRevB.70.195104, PhysRevB.88.075106}.
@@
 While useful, it is typically impossible to know whether models built from these techniques accurately describe reality, as measures of model quality are rarely reported.
-Instead, out-of-method comparisons of model predictions to experiment are conducted for model validation.
+Instead, comparisons of model predictions to experiment are conducted for model validation.
 
-The recently developed DMD \cite{Zheng2017} method provides a more direct link between \textit{ab-initio} and many-body effective Hamiltonians.
-Similar to the methods above, DMD uses information from \textit{ab-initio} simulations on the system to fit the best parameters for a given many-body model \textit{ansatz}.
+The recently developed DMD \cite{Zheng2017} method provides a more direct link between \textit{ab initio} and many-body effective Hamiltonians.
+Similar to the methods above, DMD uses information from \textit{ab initio} simulations on the system to fit the best parameters for a given many-body model \textit{ansatz}.
 Unlike the methods above, the resultant fit model is accompanied by a clear measure of model quality, quantifying the ability of the model to reproduce the low-energy spectrum of the system.
-Since model quality is reported, DMD allows for direct comparison between different model \textit{ansatzes} as well, an ability which can be used to deduce important model terms using \textit{ab-initio} calculations.
+Since model quality is reported, DMD allows for direct comparison between different model \textit{ansatzes} as well, an ability which can be used to deduce important model terms using \textit{ab initio} calculations.
 DMD has been used to develop accurate many-body models for benzene, a 1-d hydrogen chain, and an on-site Hubbard model for graphene \cite{Zheng2017, Wagner2015}.
@@
 The utility of DMD is made apparent when studying the recent surge in model development for TBLG.
-\textit{Ab-initio} model development has so far been restricted to tight-binding models which accurately describe single particle effects like band flattening, but do not address interactions \cite{Fang, Fang2016, PhysRevX.8.031088, PhysRevX.8.031087}.
-Minimal interacting models have been proposed using traditional theoretical methods, but without a link to \textit{ab-initio} it is unclear whether they can describe the real system \cite{PhysRevB.98.045103, PhysRevB.100.125116, PhysRevX.8.031089}.
-DMD can bridge the gap between these methods by providing a framework to build an accurate many-body model directly from \textit{ab-initio} simulations on TBLG.
+\textit{ab initio} model development has so far been restricted to tight-binding models which accurately describe single particle effects like band flattening, but do not address interactions \cite{Fang, Fang2016, PhysRevX.8.031088, PhysRevX.8.031087}.
+Minimal interacting models have been proposed using traditional theoretical methods, but without a link to \textit{ab initio} it is unclear whether they can describe the real system \cite{PhysRevB.98.045103, PhysRevB.100.125116, PhysRevX.8.031089}.
+DMD can bridge the gap between these methods by providing a framework to build an accurate many-body model directly from \textit{ab initio} simulations on TBLG.
 
 As such, one portion of a project between research groups at UIUC, UCSB and Rice University is dedicated towards the development of an accurate continuum-scale interacting model for TBLG.
-The model development consists of two layers of downfolding: first, developing an atomic scale lattice model using \textit{ab-initio} calculations, and second, downfolding the resultant lattice model to a continuum length scale.
+The model development consists of two layers of downfolding: first, developing an atomic scale lattice model using \textit{ab initio} calculations, and second, downfolding the resultant lattice model to a continuum length scale.
 My thesis will be focused on the first layer of downfolding.
-I propose a two year plan for developing many-body lattice models for single- (SLG) and simple bi-layer (BLG) graphene systems using the DMD method and \textit{ab-initio} fixed-node diffusion Monte Carlo (FN-DMC) calculations.
+I propose a two year plan for developing many-body lattice models for single- (SLG) and simple bi-layer (BLG) graphene systems using the DMD method and \textit{ab initio} fixed-node diffusion Monte Carlo (FN-DMC) calculations.
 My thesis will lay the necessary ground work for the development of a fully twist angle dependent lattice model for TBLG, and an eventual continuum model.
@@
 
-When developing models for realistic system, $\hat{H}$ is taken to be the \textit{ab-initio} Hamiltonian
+When developing models for realistic system, $\hat{H}$ is taken to be the \textit{ab initio} Hamiltonian
 \begin{equation}
@@
 The unit of energy is Hartree (Ha).
-To reduce the computational expense of \textit{ab-initio} calculations one works under the Born-Oppenheimer approximation where $E_\text{ion}$ becomes a constant.
+To reduce the computational expense of \textit{ab initio} calculations one works under the Born-Oppenheimer approximation where $E_\text{ion}$ becomes a constant.
 Ionic effects related to phonon modes can be recovered by altering the ionic geometry and observing the effects on the electronic state.
 
-Low-energy states can be accessed using any wave function based \textit{ab-initio} simulation technique, but two broad principles are used for selecting a method.
-First, the method should be able to generate accurate low-energy wave functions in the \textit{ab-initio} Hilbert space, as the accuracy of the low-energy states dictate the accuracy of the final model.
+Low-energy states can be accessed using any wave function based \textit{ab initio} simulation technique, but two broad principles are used for selecting a method.
+First, the method should be able to generate accurate low-energy wave functions in the \textit{ab initio} Hilbert space, as the accuracy of the low-energy states dictate the accuracy of the final model.
 Second, the method should be relatively inexpensive, as large amounts of training data are preferred for a more robust regression.
@@
 An essential first pass involves selecting a large set of descriptors using knowledge of the low-energy degrees of the freedom of the system and physical principles like symmetry.
-Further filtering follows by examining the relationship between descriptors and the \textit{ab-initio} energy over the low-energy sampled states: descriptors will typically be discarded if they do not exhibit covariance with the energy.
+Further filtering follows by examining the relationship between descriptors and the \textit{ab initio} energy over the low-energy sampled states: descriptors will typically be discarded if they do not exhibit covariance with the energy.
 Note, while there is no restriction on the form of the candidate operators other than Hermiticity, 1- and 2-body second quantized operators are commonly used.
@@
 \end{equation}
-where $e^J$ is a Jastrow factor, $|\text{D}_j\rangle$ are selected determinants which we discuss below, and $p_j$ are variable coefficients.
-A single three-body Jastrow factor was used for all sampled states optimized on the lowest energy determinant $|\text{D}_0 \rangle$.
-This significantly reduced the computational cost of the DMD method and is justified by noting that large variations in the Jastrow factor would move the state outside the low-energy space.
+where $e^J$ is a Jastrow factor, $|\text{D}_j\rangle$ are selected determinants, and $p_j$ are variable coefficients.
+By varying $\{p_j\}$ one can move around the low-energy space if the set $\{|\text{D}_j \rangle\}$ is representative of the low-energy excitations in the system.
+The Jastrow factor explicitly introduces electronic correlation and reduces the localization error within FN-DMC \cite{Foulkes2001}.
 
 The determinants $|\text{D}_j \rangle$ were constructed from the orbitals of independent symmetry-targeted unrestricted Kohn Sham (UKS) calculations.
-This choice was made to ensure that the three major low-energy variations seen in the molecule can be captured in our samples: occupancy, spin configuration and relaxation of the Cu 3d, 4s and O 2p orbitals.
-All determinants with at most a single hole in the Cu 3d shell and no doubly occupied Cu 4s orbital were included in \eqref{eq:sampling} to reflect the low-energy excitations seen in experiment.
-Only the lowest energy state with a Cu 4s$^2$ electronic configuration was included, corresponding to a double excitation 2 O ${2p_\pi} \rightarrow$ 2 Cu 4s with an energy $>7$ eV above the UKS ground state.
-Inclusion of this high energy determinant is necessary so our model places similar Cu 4s$^2$ states high in energy.
+This choice was made to ensure that the three major low-energy variations seen in the molecule were captured: occupancy, spin configuration and relaxation of the Cu 3d, 4s and O 2p orbitals.
+Only determinants with electrons in the Cu 3d, 4s, and O 2p orbitals and at most a single hole in the Cu 3d shell were included in \eqref{eq:sampling} to reflect the low-energy excitations seen in experiment.
 The UKS calculations were done using PySCF \cite{Sun2018}, a B3LYP functional \cite{doi:10.1063/1.464304, PhysRevB.37.785}, Trail-Needs psuedopotential and VTZ Trail-Needs basis \cite{doi:10.1063/1.4811651}.
 
-A shell sampling scheme was used to sample states from the selected approximate low-energy space.
+A shell sampling scheme was used to select coefficients $p_j$ for the trial wave functions.
 We began by fixing the coefficient $p_0 = \sqrt{w}$ where $w \in \{1.0, 0.8, ..., 0.2\}$. 
-For each choice of $w$ we sampled n = 5 states by randomly selecting the unassigned coefficients from independent, standard normal distributions such that $\sum_j p_j^2 = 1$. 
+For each choice of $w$, n = 5 states were selected by sampling the unassigned coefficients from independent, standard normal distributions such that $\sum_j p_j^2 = 1$. 
 This procedure generated a uniform set of samples from spherical shells at decreasing radii in determinant space centered at $|\text{D}_0\rangle$.
-We repeated this process centered at every determinant in \eqref{eq:sampling} resulting in a sample set of overlapping shells which fill $\mathcal{LE}^\prime$.
+The process was repeated by centering at every determinant in \eqref{eq:sampling}, resulting in a sample set of overlapping shells which fill the low-energy space.
 
-The final sampled MSJ states were projected using FN-DMC to generate our low-energy states.
-The FN-DMC calculations were conducted with T-moves to ensure the variational principle holds while using pseudopotentials.
-A timestep of $\tau = 0.01$ was used.
-These calculations were done in QWalk \cite{Wagner2009}.
+A single three-body Jastrow factor was used for all trial wave functions, and was optimized on the lowest energy determinant $|\text{D}_0 \rangle$ using an energy optimization scheme \cite{Toulouse2007}.
+This choice significantly reduced the computational cost of sampling low-energy states as linear optimization was not necessary for every sampled state.
+The resulting MSJ trial function remained low in energy as variations of the Jastrow factor among low-energy states is expected to be small.
+
+The sampled MSJ trial wave functions were projected using FN-DMC to generate the final set of low-energy sample states.
+FN-DMC calculations were conducted with T-moves \cite{PhysRevB.74.161102} to ensure the variational principle holds while using pseudopotentials, and a timestep of $\tau = 0.01$.
+All QMC calculations were done in QWalk \cite{Wagner2009}.
 
@@
 \paragraph{Selecting candidate descriptors}
-The following set of 1- and 2-body operators were selected as candidates for the model based on understanding of the low-energy excitations of the CuO molecule and properties of the sampled low-energy states:
+The following set of 1- and 2-body operators were selected as candidates for the model based on an understanding of the low-energy excitations of the CuO molecule and properties of the sampled low-energy states:
 \begin{equation}
 \begin{split}
-\sum_{\sigma = \uparrow, \downarrow} \Bigg(\epsilon_{3d_\pi}\hat{n}_{3d_\pi,\sigma} + \epsilon_{3d_{z^2}}\hat{n}_{3d_{z^2},\sigma} +  \epsilon_{2p_z} \hat{n}_{2p_z,\sigma} + \epsilon_{2p_\pi}\hat{n}_{2p_\pi,\sigma} + \epsilon_{4s}\hat{n}_{4s,\sigma} +\\
-t_\pi \hat{c}_{3d_\pi,\sigma}^\dagger \hat{c}_{2p_\pi,\sigma} + t_{dz} \hat{c}_{3d_{z^2},\sigma}^\dagger \hat{c}_{2p_z,\sigma}  + t_{sz}\hat{c}_{4s,\sigma}^\dagger \hat{c}_{2p_z,\sigma} + t_{sd}\hat{c}_{4s,\sigma}^\dagger \hat{c}_{3d_{z^2},\sigma} + \text{h.c.} + \Bigg)  \\
+\sum_{\sigma = \uparrow, \downarrow} \Bigg(\epsilon_{3d_\pi}\hat{n}_{3d_\pi,\sigma} + \epsilon_{3d_{z^2}}\hat{n}_{3d_{z^2},\sigma} +  \epsilon_{2p_z} \hat{n}_{2p_z,\sigma} + \epsilon_{2p_\pi}\hat{n}_{2p_\pi,\sigma} + \epsilon_{4s}\hat{n}_{4s,\sigma} +
+t_\pi \hat{c}_{3d_\pi,\sigma}^\dagger \hat{c}_{2p_\pi,\sigma} + t_{dz} \hat{c}_{3d_{z^2},\sigma}^\dagger \hat{c}_{2p_z,\sigma}\\
+  + t_{sz}\hat{c}_{4s,\sigma}^\dagger \hat{c}_{2p_z,\sigma} + t_{sd}\hat{c}_{4s,\sigma}^\dagger \hat{c}_{3d_{z^2},\sigma} + \text{h.c.} \Bigg)  +
 J_{sd}\sum_{i\in {\{\text{xy, xz, ...}}\}} \vec{S}_{4s} \cdot \vec{S}_{3d_i} + U_s \hat{n}_{4s,\uparrow}\hat{n}_{4s,\downarrow} + E_0.
@@
 Within the sampled data, states with a Cu $4s^{2}$ configuration lie at least 7 eV above the lowest energy sampled state, while the lowest energy state with a $4s^{1}$ configuration lies at 1.6 eV.
-If the double occupancy term was not included, the effective theory would predict the prior state to have an energy of $\sim $ 3eV, an underestimation of 4 eV from the \textit{ab-initio} result.
-The effective on-site Coulomb repulsion $U_s$ ensures the model places all $4s^{2}$ states high in energy.
+If the double occupancy term was not included, the effective theory would predict the prior state to have an energy of $\sim $ 3eV, an underestimation of 4 eV from the \textit{ab initio} result.
+The effective on-site Coulomb repulsion $U_s$ ensures the model places all such $4s^{2}$ states high in energy.
 
@@
 
-To reduce multicollinearity while maintaining a high goodness of fit, model was fit to these nine PCs.
+To reduce multicollinearity while maintaining a high goodness of fit, a model was fit to these nine PCs.
 When trained on all the sampled data this model yields an $R^2$ = 0.98 with a CN of 257. 
@@
 \paragraph{Solutions of the model}
-The model is solved by exact diagonalization, and the solutions agree well with experimental spectra and state assignments.
-Shown in Figure \ref{fig:cuo_regr}d is a comparison between the excitation energies calculated from our model and those seen in the latest anion photoelectron spectroscopy measurements \cite{Wu1997}.
+The final model in \ref{fig:cuo_regr}c was solved by exact diagonalization, and the solutions agree well with experimental spectra and state assignments.
+Shown in Figure \ref{fig:cuo_regr}d is a comparison between the excitation energies calculated from the model and those seen in the latest anion photoelectron spectroscopy measurements \cite{Wu1997}.
 The model energies for these eigenstates are denoted by orange circles and agree with experiment up to a maximum deviation of 0.2 eV.
-The term symbols assigned in experiment also match the term symbols of the model eigenstates, which are constructed by looking at the multiplicity and electronic configuration of each model eigenstate.
+Further, the term symbols assigned in experiment match the term symbols of the model eigenstates, which are constructed by looking at the multiplicity and electronic configuration of each model eigenstate.
 
-The model states also agree with recent theoretical calculations like UKS and CCSD(T).
-The energies for the UKS and CCSD(T) \cite{Xian2000} states are shown in Figure \ref{fig:cuo_regr}d as green and red circles, respectively.
+The energies for the UKS and CCSD(T) \cite{Xian2000} states are also shown in Figure \ref{fig:cuo_regr}d as green and red circles, respectively.
 The UKS energies were calculated in this work in en route to constructing the determinants $|\text{D}_j\rangle$ used in our sampling scheme \eqref{eq:sampling}.
-Our model energies agree with both theoretical techniques within errorbars for the first three states shown.
-The energy of the highest energy $^2\Delta$ is severely over estimated using both UKS and CCSD(T), by nearly 1 eV.
+The model energies agree with both theoretical techniques within errorbars for the first three states shown.
+The energy of the highest energy $^2\Delta$, however, is severely over estimated using both UKS and CCSD(T), by nearly 1 eV.
 This over estimation may arise from an inadequate relaxation of the $\sigma$ type orbitals which form a three orbital hybridization complex in this system.
@@
 The final goal for TBLG would be an interacting continuum model on the Moire superlattice length scale.
-From a DMD perspective, this can accomplished by first downfolding the \textit{ab-initio} Hamiltonian with a given twist to an atomic length scale model, and then downfolding a second time to the Moire length scale.
+From a DMD perspective, this can accomplished by first downfolding the \textit{ab initio} Hamiltonian with a given twist to an atomic length scale model, and then downfolding a second time to the Moire length scale.
 Developing the atomic length scale models requires a strong understanding of intra- and inter-layer interactions in this material.

### 2019-09-11 reoutline.tex
finished preliminary work. need to make bib shorter...
updated reoutline

@@
 \item Diffusion Monte Carlo (DMC) is a quantum Monte Carlo method which projects out the ground state of a real-space Hamiltonian given some initial trial wave function.
-%Talk about efficiency here
 
 \item DMC, however, suffers from a fermion sign problem which is alleviated via a fixed-node approximation.
-%Talk about accuracy here, compare to other methods
 
@@
 \item The training data required for downfolding are extracted from the projected low-energy states via a mixed estimator.
-%Start paragraph with something else, probably the stochastic representation of the wave function
 \end{enumerate}
@@
 
-%sampling low energy states
-\item Explain why MSJ was chosen (Probably the hardest one)
+%Sampling LE states
+\item For the CuO molecule, low-energy states were generated by using FN-DMC projected multi-Slater-Jastrow (MSJ) trial wave functions.
 
-\item How determinants selected
+\item The determinants $|D_j \rangle$ were constructed from the orbitals of independent symmetry-targeted unrestricted Kohn Sham (UKS) calculations.
 
-\item How coefficients selected
-
-\item How Jastrow optimized and FN-DMC done
+\item A shell sampling scheme was used to sample states from the selected approximate low-energy space.
 
@@
 \item The inclusion of a double occupancy term $n_{4s,\uparrow} n_{4s,\downarrow}$ was motivated directly from the sampled low-energy states.
+
+\item A single three-body Jastrow factor was used for all trial wave functions, and was optimized on the lowest energy determinant $|D_0 \rangle$ using an energy optimization scheme \cite{Toulouse2007}.
 
@@
 %Solutions of the model 
-\item The model is solved by exact diagonalization, and the solutions agree well with experimental spectra and state assignments.
+\item The model was solved by exact diagonalization, and the solutions agree well with experimental spectra and state assignments.
 
 \item The model states also agree with recent theoretical calculations like UKS and CCSD(T).
-
 \end{enumerate}

### 2019-09-12 draft.tex
updated draft. preliminary work all done + custom bibliography is just 1 page1
fully revised prelim, going to go through it once tomorrow before sending to lucas

@@
 \setlength{\belowdisplayskip}{0pt}
+\usepackage{etoolbox}
+\patchcmd{\thebibliography}
+  {\settowidth}
+  {\setlength{\itemsep}{0pt plus 0.01pt}\settowidth}
+  {}{}
+\apptocmd{\thebibliography}
+  {\small}
+  {}{}
 \author{Shivesh Pathak}
@@
 
-As such, one portion of a project between research groups at UIUC, UCSB and Rice University is dedicated towards the development of an accurate continuum-scale interacting model for TBLG.
+As such, a project named QMC High Accuracy Multiscale Modeling (QMC HAMM) between research groups at UIUC, UCSB and Rice University is dedicated towards the development of an accurate continuum-scale interacting model for TBLG.
 The model development consists of two layers of downfolding: first, developing an atomic scale lattice model using \textit{ab initio} calculations, and second, downfolding the resultant lattice model to a continuum length scale.
@@
 A standard fitting workflow involves feature selection, model regression and model validation.
-Feature selection methods for choosing descriptors from the candidate pool can range from wrapper methods like principal component analysis \cite{pearson_karl_1901_1430636} to regularized embedded methods like LASSO \cite{10.2307/2346178, doi:10.1137/S0895479897326432, Zou05regularizationand}.
+Feature selection methods for choosing descriptors from the candidate pool can range from wrapper methods like principal component analysis \cite{pearson_karl_1901_1430636} to regularized embedded methods like LASSO \cite{10.2307/2346178}.
 Fitting the effective Hamiltonian usually involves an ordinary least squares linear regression, but the cost function can be altered.
@@
 \section{Proposed work}
-My proposed plan involves building a sequence of model Hamiltonians for increasingly complicated graphene-like systems using QMC DMD, ending with BLG.
-The systems I will be interested in are the benzene molecule, single layer graphene, AA stacked bilayer graphene and AB stacked bilayer graphene.
-Conceptually, the benzene molecule is a kind of building block for SLG, and SLG is a building block for BLG, creating a sequence of systems from least to most complex.
+I propose a two year thesis plan for developing atomic scale many-body
+model Hamiltonians for SLG and simple BLG systems within the DMD framework using high accuracy \textit{ab initio} QMC calculations.
+To do so, I will be building a sequence of four many-body model Hamiltonians for increasing complex systems: the benzene molecule, SLG, AA stacked BLG, and AB stacked BLG.
+The following sections detail the sampling scheme and candidate models that will be used for downfolding each system.
+I conclude with a brief statement about the impact of my work within the QMC HAMM project and modeling TBLG and a timeline.
 
+\subsection{Constrained variational Monte Carlo (cVMC)}
+A new constrained cVMC method will be used to sample the low-energy wave functions necessary for DMD.
+The cVMC method works by minimizing the following cost function for a parameterized wave function $|\Psi(\vec{p})\rangle$:
+\begin{equation}
+\vec{p}^* = \text{argmin} \ \langle \Psi(\vec{p}) | \hat{H}_\text{ab} | \Psi(\vec{p}) \rangle - \sum_k [\lambda_k \langle \Psi(\vec{p}) | d_k - d_{k}^0 | \Psi(\vec{p}) \rangle]^2
+\end{equation}
+with respect to the parameters $\vec{p}$.
+The quantity $\vec{d}$ is short hand for a list of candidate descriptors to sample variations in, $\vec{d}^*$ is the target values of those descriptors, and $\vec{\lambda}$ controls the importance of finding a state with low-energy versus one with descriptor values near $\vec{d}^*$.
+This method allows one to target low-energy wave functions which sit near a target point $\vec{d}^*$ in descriptor space, allowing for a more systematic sampling of the low-energy space.
+
+I will use a multi-Slater-Jastrow (MSJ) parameterization in the cVMC sampling.
+The multi-Slater expansion will consist of determinants of single particle orbitals calculated using DFT.
+Given the nature of graphene-like systems, as described below, the determinants will include the DFT ground state as well as molecular orbital excitations among the $\pi$ and $\pi^*$ orbitals which form the active space.
+A three body Jastrow factor will be applied to ensure that correlations are included in the wave function and that the wave function is low in energy.
+The total set of parameters $\vec{p}$ will be Jastrow and determinental coefficients.
+
+\subsection{Benzene}
 The first system I will develop a model for is the benzene molecule, shown in Figure \ref{fig:proposed}a.
@@
 Interaction effects seen in the benzene molecule are also reflected in SLG, such as on-site double occupation energies for the carbon atoms \cite{Zheng2017, Wagner2015}.
-As such, developing a model for benzene will help me become familiar with model fitting for 2-D carbon based systems.
 
@@
 
-The candidate descriptors I will use for DMD are motivated by a previous calculation on the benzene molecule \cite{Wagner2015}.
-In that study, an extended Hubbard model built on the six $\pi$ orbitals was found to perform well in describing the low-lying excitations of the material.
+My objective will be to reproduce a previously developed many-body model for the benzene molecule using QMC DMD \cite{Wagner2015}.
+In this study, an extended Hubbard model built on the six $\pi$ orbitals was found to perform well in describing the low-lying excitations of the material.
 The final model took the form 
@@
 where the brackets denote nearest neighbor and next-nearest neighbor summations.
-I will consider at least all the descriptors above and also longer range hopping terms.
+As such, the model \textit{ansatz} that I will try to fit will be \eqref{Hbenz}.
 
-The second system I will be interested in working on is SLG.
+This first stage of model development provides a playground for benchmarking new model fitting procedures.
+Downfolding techniques have advanced significantly since the initial model for benzene was developed in 2015.
+Methods like constrained VMC for low-energy sampling have been posited but not thoroughly tested.
+By comparing models built using new methods directly to the 2015 model, it will be possible to quantify the efficiency and effectiveness of these methods.
+
+\subsection{SLG}
+The second system I will work on is SLG.
 Single layer graphene is composed of a triangular Bravais lattice with two carbon atoms per unit cell.
@@
 
-In particular, I will work to extend a previous on-site Hubbard model developed for SLG \cite{Zheng2017, Wagner2015} by including long range density-density interactions .
+I will work on extending an on-site Hubbard model for SLG \cite{Zheng2017, Wagner2015} to one with long-range Coulomb interactions.
 The candidate model will be of the form:
@@
 where the indices correspond to the $\pi$ orbitals on the copper atoms.
-I will consider both long range hopping and density-density interaction terms extending across the computational unit cell with the final goal of developing position dependent parameter values $t(r), V(r)$.
+I will consider both long range hopping and density-density interaction terms extending across the computational unit cell, with the final goal of developing position dependent parameter values $t(r), V(r)$.
 
+The resultant model will be able to non-parametrically determine the form
+of long-range interactions in graphene.
+This is an important contribution to the study of interaction effects in graphene as many potential properties of the system depend on details of these long-range interactions.
+Such effects include edge magentism \cite{PhysRevB.95.195420} and non-local electrodynamics \cite{PhysRevB.90.045137}.
+
+\subsection{BLG}
 The last systems I will develop models for are AA and AB stacked BLG, shown in Figure \ref{fig:proposed}b.
@@
 There are two terms for the single layers from \eqref{Hslg} plus an interlayer hopping given by the term $t_{i_1, i_2}^\prime.$  
-I suspect that the hopping values will be very different between the AA and AB models as the distance between interlayer neighbors differs greatly between the two stackings.
-This difference in value will provide the first steps into developing model parameters for BLG which depend on the relative orientation of the sheets.
+The index $i_1$ refers to layer 1 and $i_2$ to layer two.
+While this is an initial \textit{ansatz}, additional inter- and intra-layer terms can be included if necessary.
 
-A new constrained variational Monte Carlo (cVMC) method will be used to sample the low-energy wave functions necessary for DMD.
-The cVMC method works by minimizing the following cost function for a parameterized wave function $|\Psi(\vec{p})\rangle$:
-\begin{equation}
-\vec{p}^* = \text{argmin} \ \langle \Psi(\vec{p}) | \hat{H}_\text{ab} | \Psi(\vec{p}) \rangle - \sum_k [\lambda_k \langle \Psi(\vec{p}) | d_k - d_{k}^0 | \Psi(\vec{p}) \rangle]^2
-\end{equation}
-with respect to the parameters $\vec{p}$.
-The quantity $\vec{d}$ is short hand for a list of candidate descriptors to sample variations in, $\vec{d}^0$ is the target values of those candidate descriptors, and $\vec{\lambda}$ controls the importance of finding a state with low-energy versus one with descriptor values near $\vec{d}_0$.
-Essentially this method allows us, for a given parameterization of $|\Psi(\vec{p})\rangle$, to generate low-energy wave functions which sit near a target point $\vec{d}^0$ in candidate descriptor space.
+With these two models, one can begin to answer some prominent questions about the inter-layer coupling in TBLG.
+First, one can determine, at least for the simplest orientations, whether the inter-layer coupling takes the form of a hopping or interaction term.
+While current studies tend to assume the prior \cite{Fang, Fang2016, PhysRevX.8.031088, PhysRevX.8.031087}, there is no good reason to discount the latter.
+Further, the quantitative details of spatial dependence, which are intimately related to the strongly correlated behavior in TBLG, can begin to be addressed.
 
-I will use a multi-Slater-Jastrow (MSJ) parameterization in the cVMC sampling for each system.
-The multi-Slater expansion will consist of determinants of single particle orbitals calculated using DFT.
-The determinants will include the DFT ground state as well as molecular orbital excitations among the $\pi$ and $\pi^*$ orbitals which form the active space.
-A three body Jastrow factor will be applied to ensure that correlations are included in the wave function and that the wave function is low in energy.
-The total set of parameters $\vec{p}$ will include both Jastrow and determinental coefficients.
-
-Apart from their individual value, the four models can be used to investigate the transferrability of model parameters between these different carbon-based systems.
-The first question would be whether the parameter values between a finite size system, like benzene, match those seen in the bulk SLG system.
-The second would be whether the intralayer parameter values change in each sheet of SLG when they are brought together in the AA and AB configuration.
-The latter question hints at a broader question, whether the SLG intralayer parameters are transferrable to TBLG, especially near the magic twist angles where interlayer coupling becomes very large \cite{Bistritzer2	011}.
-
-Further, my calculations will help mature methods like cVMC for further usage on strongly correlated systems.
-A few questions of interest include how one should select values like $\vec{\lambda}$ and $\vec{d}_0$, and whether final models fit with cVMC are consistent with those fit previously using a different sampling technique.
-The prior question will be naturally investigated while using cVMC in building models for the systems above.
-The latter can be directly probed by comparing my fit models for benzene and on-site SLG to those which were developed previously using DMD.
-
-Lastly, my calculations will lay the groundwork for model development for the strongly correlated TBLG system using DMD.
-The final goal for TBLG would be an interacting continuum model on the Moire superlattice length scale.
-From a DMD perspective, this can accomplished by first downfolding the \textit{ab initio} Hamiltonian with a given twist to an atomic length scale model, and then downfolding a second time to the Moire length scale.
-Developing the atomic length scale models requires a strong understanding of intra- and inter-layer interactions in this material.
-The prior will be thoroughly investigated by my model fitting on SLG, and I will begin to answer the latter with the BLG calculations.
+\vspace{-10pt}
+\paragraph{QMC HAMM}
+A major goal of the QMC HAMM project is to develop a twist dependent, Moire length scale, interacting model for TBLG using QMC \textit{ab initio} calculations.
+The first step in doing so is developing an atomic scale model for TBLG, and the necessary prerequisites for this step are a strong understanding of effective theories of the basic constituents of the system.
+The models I propose to develop for SLG and AA/AB BLG provide exactly this kind of understanding.
+As such, my work will lay the foundation for the four year QMC HAMM project to build an effective theory for TBLG.
 
@@
 \paragraph{Proposed timeline} I believe that developing effective theories for the benzene molecule, SLG, AA and AB stacked BLG should take 2 years. 
-The model fitting for benzene should take no more than four months.
-The goal is to reproduce prior DMD calculations for the same material while benchmarking and familiarizing myself with the new cVMC sampling technique that I will use extensively in my project.
+The model fitting for benzene should take no more than four months as it is a reproduction of previous work.
 The SLG model fitting should take another eight months.
-This is an extension of a previous work, and the first three months will be dedicate to reproducing the results from that work, namely a single band on-site Hubbard model.
+As an extension of a previous work, the first three months will be dedicate to reproducing the on-site Hubbard model from that work.
 Inclusion of long range Coulomb interactions into the effective theory should take another three months.
 The last two months will be geared towards finite size extrapolation in order to ensure the final model is applicable to bulk systems.
-The last year of the project will be geared towards the AA and AB stacked BLG, with each project taking about six months to complete.
-For each stacking, I expect four months will be necessary to build an effective theory, and the final two months will be dedicated to finite size extrapolation.
+The last year of the project will be focused on the AA and AB BLG, with each project taking six months to complete.
+For each stacking, need four months to build the model and two months for finite size extrapolation.
 
+\vspace{-10pt}
 \begin{multicols}{2}
 \begin{spacing}{0.5}
-{\fontsize{10}{5}\selectfont
+{\footnotesize
 \bibliography{biblio}
-\bibliographystyle{unsrt}
+\bibliographystyle{prelim}
 }

### 2019-09-12 reoutline.tex
reoutline
fully revised prelim, going to go through it once tomorrow before sending to lucas

@@
 \end{enumerate}
+\section{Proposed Work}
+\begin{enumerate}
+\item I propose a two year thesis plan for developing atomic scale many-body
+model Hamiltonians for SLG and simple BLG systems within the DMD framework using high accuracy ab initio quantum Monte Carlo (QMC)
+calculations.
+%Start with some filler
 
+%Benzene molecule
+\item The first system I will develop a model for is the benzene molecule, shown in Figure 2a.
+
+\item My objective will be to reproduce the many-body model for the benzene molecule developed in \textbf{ref}.
+
+\item This first stage of model development provides a playground for benchmarking new model fitting procedures.
+
+%SLG
+\item The second system I will be interested in working on is SLG.
+
+\item I will work on the extension of an on-site Hubbard model for SLG \textbf{ref} to one with long-range Coulomb interactions.
+
+\item The resultant model will be able to non-parametrically determine the form of long-range interactions in graphene.
+%Could say something more precise here about necessity of LRI
+
+%BLG
+\item The last systems I will develop models for are AA and AB stacked BLG, shown in Figure 2b.
+
+\item I anticipate that the candidate descriptors should include interlayer couplings as well as the terms seen in SLG.
+
+\item The BLG models provide a starting point in understanding the position dependence of model parameters in TBLG.
+%Could say something more careful
+
+\item The final four models can also be used to investigate the transferrability
+of model parameters between these different carbon-based systems.
+
+%CVMC
+\item A new constrained variational Monte Carlo (cVMC) method will be used to sample the low-energy wave functions necessary for DMD.
+
+\item The multi-Slater-Jastrow (MSJ) parameterization of \textbf{eqref} will be used within cVMC for each system.
+\end{enumerate}
+
+\section{Proposed Timeline}
 \end{document}

### 2019-09-13 draft.tex
removed some unecessary acronyms

@@
 In an attempt to build accurate many-body Hamiltonians for realistic systems, a framework for model development using solutions from \textit{ab initio} simulations has emerged.
-Common approaches involve density functional theory (DFT) \cite{PhysRev.140.A1133, PhysRevLett.87.047003} alongside random phase approximation (RPA) or other methods \cite{PhysRevB.62.R16219, PhysRevB.70.195104, PhysRevB.88.075106}.
+Common approaches involve density functional theory (DFT) \cite{PhysRev.140.A1133, PhysRevLett.87.047003} alongside random phase approximation or other methods \cite{PhysRevB.62.R16219, PhysRevB.70.195104, PhysRevB.88.075106}.
 L\"owdin methods and canonical transformations have also been used to develop effective theories \cite{doi:10.1063/1.4802766, PhysRevA.81.013402, doi:10.1063/1.2196410}.
@@
 
-As such, a project named QMC High Accuracy Multiscale Modeling (QMC HAMM) between research groups at UIUC, UCSB and Rice University is dedicated towards the development of an accurate continuum-scale interacting model for TBLG.
+As such, a project named QMC High-Accuracy Multiscale Modeling (QMC HAMM) between research groups at UIUC, UCSB and Rice University is dedicated towards the development of an accurate continuum-scale interacting model for TBLG.
 The model development consists of two layers of downfolding: first, developing an atomic scale lattice model using \textit{ab initio} calculations, and second, downfolding the resultant lattice model to a continuum length scale.
@@
 The results of the linear regression are an estimate of the parameter values $g_k$ and model error $\epsilon[\Psi]$.
-The error estimate usually takes the form of a goodness of fit, such as a cross-validated $R^2$ score or root mean squared error (RMSE).
+The error estimate usually takes the form of a goodness of fit, such as a cross-validated $R^2$ score or root mean squared error.
 These error estimates present a direct measurement of model quality, ensuring that DMD not only gives the best fit parameters but also an assessment of model validity.
@@
 \includegraphics[width=1.0\linewidth]{./figs/cuo_regr_2.pdf}
-\caption{Figures related to model regression for the CuO molecule: a) Intrinsic atomic orbitals used as a single particle basis for the model Hamiltonian, b) Results of PCA showing the CEVR and CN of a data set containing N$_{comp}$  highest ranked PCs, c) The final regressed parameters from the PCR, d) Comparison of four lowest lying eigenstates of CuO molecule between experiment and three computation methods.}
+\caption{Figures related to model regression for the CuO molecule: a) Intrinsic atomic orbitals used as a single particle basis for the model Hamiltonian, b) Results of PCA showing the CEVR and CN of a data set containing N$_{comp}$  highest ranked PCs, c) The final regressed model parameters, d) Comparison of four lowest lying eigenstates of CuO molecule between experiment and three computation methods.}
 \label{fig:cuo_regr}
@@
 The quantity $\vec{d}$ is short hand for a list of candidate descriptors to sample variations in, $\vec{d}^*$ is the target values of those descriptors, and $\vec{\lambda}$ controls the importance of finding a state with low-energy versus one with descriptor values near $\vec{d}^*$.
-This method allows one to target low-energy wave functions which sit near a target point $\vec{d}^*$ in descriptor space, allowing for a more systematic sampling of the low-energy space.
+This method allows one to target low-energy wave functions which sit near a target point $\vec{d}^*$ in descriptor space, providing a more systematic method for sampling the low-energy space.
 
@@
 \paragraph{QMC HAMM}
-A major goal of the QMC HAMM project is to develop a twist dependent, Moire length scale, interacting model for TBLG using QMC \textit{ab initio} calculations.
+A major goal of the QMC HAMM project is to develop a twist dependent, Moire length scale, interacting model for TBLG using \textit{ab initio} QMC calculations.
 The first step in doing so is developing an atomic scale model for TBLG, and the necessary prerequisites for this step are a strong understanding of effective theories of the basic constituents of the system.

### 2019-09-17 draft.tex
added a missing reference
smallest changes based on lucas response

@@
 A recently developed density matrix downfolding (DMD) \cite{Zheng2017} method provides a very direct framework for downfolding the \textit{ab initio} Hamiltonian to a many-body effective Hamiltonian using only \textit{ab initio} calculations.
-Here I propose a two year thesis plan for developing atomic scale many-body model Hamiltonians for single- and simple bi-layer graphene systems within the DMD framework using high accuracy \textit{ab initio} quantum Monte Carlo (QMC) calculations.
+Here I propose a thesis plan for developing atomic scale many-body model Hamiltonians for single- and simple bi-layer graphene systems within the DMD framework using high accuracy \textit{ab initio} quantum Monte Carlo (QMC) calculations.
 The results of my thesis will lay the groundwork for a continuing four year project to build a twist dependent many-body continuum model for the twisted bilayer graphene (TBLG) system.
@@
 \textit{ab initio} model development has so far been restricted to tight-binding models which accurately describe single particle effects like band flattening, but do not address interactions \cite{Fang, Fang2016, PhysRevX.8.031088, PhysRevX.8.031087}.
-Minimal interacting models have been proposed using traditional theoretical methods, but without a link to \textit{ab initio} it is unclear whether they can describe the real system \cite{PhysRevB.98.045103, PhysRevB.100.125116, PhysRevX.8.031089}.
+Minimal interacting models have been proposed using traditional theoretical methods, but without a link to the \textit{ab initio} Hamiltonian it is unclear whether they can describe the real system \cite{PhysRevB.98.045103, PhysRevB.100.125116, PhysRevX.8.031089}.
 DMD can bridge the gap between these methods by providing a framework to build an accurate many-body model directly from \textit{ab initio} simulations on TBLG.
 
-As such, a project named QMC High-Accuracy Multiscale Modeling (QMC HAMM) between research groups at UIUC, UCSB and Rice University is dedicated towards the development of an accurate continuum-scale interacting model for TBLG.
+As such, a project named QMC High-Accuracy Multiscale Modeling (QMC HAMM) between research groups at UIUC and Rice University is dedicated towards the development of an accurate continuum-scale interacting model for TBLG.
 The model development consists of two layers of downfolding: first, developing an atomic scale lattice model using \textit{ab initio} calculations, and second, downfolding the resultant lattice model to a continuum length scale.
-My thesis will be focused on the first layer of downfolding.
+My thesis will be focused on the first stage of downfolding.
 I propose a two year plan for developing many-body lattice models for single- (SLG) and simple bi-layer (BLG) graphene systems using the DMD method and \textit{ab initio} fixed-node diffusion Monte Carlo (FN-DMC) calculations.
@@
 
-This linear regression problem can be tackled in three steps.
+This linear regression problem is tackled in three steps.
 First, low-energy states are sampled from $\mathcal{LE}$.
@@
 Note, while there is no restriction on the form of the candidate operators other than Hermiticity, 1- and 2-body second quantized operators are commonly used.
-
-A standard fitting workflow involves feature selection, model regression and model validation.
-Feature selection methods for choosing descriptors from the candidate pool can range from wrapper methods like principal component analysis \cite{pearson_karl_1901_1430636} to regularized embedded methods like LASSO \cite{10.2307/2346178}.
-Fitting the effective Hamiltonian usually involves an ordinary least squares linear regression, but the cost function can be altered.
-Model validation can be carried out by calculating cross validated single parameter measures like $R^2$ scores which carry information about goodness of fit and overfitting.
 
@@
 The workflow involved constructing the un-optimized trial wave functions, optimizing the parameters using an energy optimization method \cite{Toulouse2007}, and finally using the optimized trial functions in FN-DMC calculations.
-The average improvement in the FN-DMC energy with the MSJ+O trial wave functions was $\langle E_{MSJ} - E_{MSJ,O}\rangle =$ 0.14eV with an additional reduction of 0.032 eV when using the MSJ+NO trial wave function.
+The average improvement in the FN-DMC energy with the MSJ+O trial wave functions was $\langle E_{MSJ} - E_{MSJ,O}\rangle =$ 0.14 eV with an additional reduction of 0.032 eV when using the MSJ+NO trial wave function.
 While the benefit using the MSJ+NO trial wave function seems small, the energy calculated using an MSJ+NO trial function with only 24 determinants was lower than the energy using an MSJ+O trial 
@@
 
-Further, the FN-DMC charge density calculated using MSJ+NO trial functions had stronger bonding character than when using MSJ+O trial functions, a reasonable result as introducing correlations into trial functions allows for electrons to avoid each other while still occupying the same bonding region.
+Further, the FN-DMC charge density calculated using MSJ+NO trial functions had stronger bonding character than when using MSJ+O trial functions, a reasonable result since introducing correlations into trial functions allows for electrons to avoid each other while still occupying the same bonding region.
 The results indicated that using non-orthogonal determinants may lead to more compact trial wave functions for small molecules.
@@
 \subsection{Effective theory for CuO molecule using DMD}
-To become acquainted with DMD, I constructed a many-body effective model for the CuO molecule with fixed bond length $r_e$ = 1.725\r{A}.
+To become acquainted with DMD, I constructed a many-body effective model for the CuO molecule with fixed bond length $r_e$ = 1.725 \r{A}.
 In the following three sections I will detail the steps I took towards developing a model for the CuO molecule using DMD, followed by a comparison of model spectra to experiment.
@@
 The following sections detail the sampling scheme and candidate models that will be used for downfolding each system.
-I conclude with a brief statement about the impact of my work within the QMC HAMM project and modeling TBLG and a timeline.
 
@@
 \includegraphics[width=1.0\linewidth]{./figs/proposed.pdf}
-\caption{Schematic diagrams for the a) benzene molecule, b) AB and AA stacked BLG \textbf{ref}.}
-%Voltage Tunable Plasmon Propagation in Dual Gated Bilayer Graphene
+\caption{Schematic diagrams for the a) benzene molecule, b) AB and AA stacked BLG \cite{doi:10.1063/1.5007713}.}
 \label{fig:proposed}
@@
 
-\subsection{SLG}
-The second system I will work on is SLG.
+\subsection{Single layer graphene (SLG)}
 Single layer graphene is composed of a triangular Bravais lattice with two carbon atoms per unit cell.
@@
 
-\subsection{BLG}
+\subsection{Bilayer graphene (BLG)}
 The last systems I will develop models for are AA and AB stacked BLG, shown in Figure \ref{fig:proposed}b.
@@
 
-With these two models, one can begin to answer some prominent questions about the inter-layer coupling in TBLG.
+With these two models, one can begin to answer some questions about the inter-layer coupling in TBLG.
 First, one can determine, at least for the simplest orientations, whether the inter-layer coupling takes the form of a hopping or interaction term.

### 2019-09-17 draft.tex
finished fndmc edits

@@
 If the initial trial wave functions are appropriately chosen, the final projected states will be within the target low-energy subspace of $\mathcal{H}$.
+The training data for DMD are then extracted by evaluating the expectation values of the Hamiltonian and candidate operators using the final distributions of walkers from the FN-DMC calculations \cite{Wagner2015}.
 
-The training data required for downfolding are extracted from the projected low-energy states via a mixed estimator.
-Due to the fixed-node approximation, the final distribution returned by the projection is $f(R) = \Psi_T(R) \Psi_\text{DMC}$, and Monte Carlo estimates of operator expectation values correspond to the mixed estimation $\frac{\langle \Psi_T | \hat{O} | \Psi_\text{DMC} \rangle}{\langle \Psi_T | \Psi_\text{DMC}\rangle}$.
-This does not equal the target expectation value $\frac{\langle \Psi_\text{DMC} | \hat{O} | \Psi_\text{DMC} \rangle}{\langle \Psi_\text{DMC} | \Psi_\text{DMC}\rangle}$, resulting in a mixed estimator error.
-Details on alleviating this error in DMD and evaluation of expectation values for 1- and 2-body second quantized operators can be found in the literature \cite{Wagner2015, Wagner2017}.
 \section{Preliminary Work}

### 2019-09-18 draft.tex
sharpening dmd methods

@@
 \begin{abstract}
-Building accurate model Hamiltonians for realistic materials is a fundamental problem in modern condensed matter physics.
+Building accurate model Hamiltonians for materials is a fundamental problem in modern condensed matter physics.
 A recently developed density matrix downfolding (DMD) \cite{Zheng2017} method provides a very direct framework for downfolding the \textit{ab initio} Hamiltonian to a many-body effective Hamiltonian using only \textit{ab initio} calculations.
@@
 \section{Introduction}
-Developing model Hamiltonians for realistic materials is a fundamental problem in modern condensed matter physics.
+Developing model Hamiltonians for materials is a fundamental problem in modern condensed matter physics.
 A model Hamiltonian is a simplified description of a real material which can accurately reproduce the low-lying eigenstates and energies of the material.
@@
 \subsection{Density Matrix Downfolding (DMD)}
-I begin by defining what a low-energy effective Hamiltonian and downfolding are.
-Consider a high-energy Hamiltonian $\hat{H}$ and Hilbert space $\mathcal{H}$.
-A low-energy effective Hamiltonian is an operator $\hat{H}_\text{eff}$ which can accurately reproduce the spectra and eigenstates of $\hat{H}$ on a subpsace $\mathcal{LE}$ of the Hilbert space $\mathcal{H}$.
-Here $\mathcal{LE}$ is the span of the N lowest energy eigenstates of $\hat{H}$, and $\hat{H}_\text{eff}$ is a linear combination of Hermitian operators $\hat{d}_k$ with a constant shift $E_0$
+The starting point of model development for materials is the \textit{ab initio} Hamiltonian defined on the Hilbert space $\mathcal{H}$
+\begin{equation}
+\hat{H}_\text{ab} = \sum_i -\frac{1}{2} \nabla_i^2 + \sum_{iI}\frac{Z_I}{|r_i - R_I|} + \frac{1}{2} \sum_{ij} \frac{1}{|r_i - r_j|} + E_\text{ion}(\{R_I\})  
+\label{eq:Hab}
+\end{equation}
+where index $i,j$ correspond to electrons, $I$ to ions, $E_\text{ion}$ corresponds to the total ionic energy, and the unit of energy is Hartree (Ha).
+The goal is to develop an effective Hamiltonian $\hat{H}_\text{eff}$ defined on a low-energy subspace $\mathcal{LE} \subset \mathcal{H}$ which can accurately reproduce the lowest lying eigenstates and spectra of $\hat{H}_\text{ab}$.
+$\hat{H}_\text{eff}$ is taken to be a linear operator on $\mathcal{LE}$ and can be expressed as a linear combination of Hermitian operators $\hat{d}_k$, with coupling constants $g_k$ and energy offset $E_0$
 \begin{equation}
@@
 \end{equation}
-The objective of Hamiltonian downfolding is to construct such an $\hat{H}_\text{eff}$ given $\hat{H}, \mathcal{H}$.
+The process of taking $(\hat{H}_\text{ab}, \mathcal{H}) \rightarrow (\hat{H}_\text{eff}, \mathcal{LE})$ is called downfolding of $\hat{H}_\text{ab}$.
 
-The key insight of DMD is that the Hamiltonian downfolding problem can be mapped onto a linear regression problem.
-It has been shown \cite{Zheng2017} that the above definition of $\hat{H}_\text{eff}$ is equivalent to the following: a low-energy effective Hamiltonian is an operator such that 
+DMD provides a framework for this downfolding by considering energy functionals.
+In their seminal paper on DMD, Zheng \textit{et al.} \cite{Zheng2017} show that $\hat{H}_\text{eff}$ will exactly reproduce the low-lying eigenstates of $\hat{H}_\text{ab}$ if
 \begin{equation}
 \begin{split}
-\forall |\Psi\rangle \in \mathcal{LE},\ E[\Psi] = E_\text{eff}[\Psi]  + \epsilon[\Psi]
+\forall |\Psi\rangle \in \mathcal{LE},\ E_\text{ab}[\Psi] = E_\text{eff}[\Psi].
 \end{split}
@@
 \end{equation}
-where $\epsilon[\Psi]$ is the error in the effective theory and the energy functionals are defined as $E[\Psi] \equiv \langle \Psi | \hat{H} |\Psi \rangle,\  E_\text{eff}[\Psi] \equiv \langle \Psi | \hat{H}_\text{eff} |\Psi \rangle$.
+Here $E_\text{ab}[\Psi] \equiv \langle \Psi | \hat{H}_\text{ab} | \Psi \rangle$ is the known \textit{ab initio} energy functional and $E_\text{eff}[\Psi] \equiv \langle \Psi | \hat{H}_\text{eff} | \Psi \rangle$ is the to be determined effective energy functional.
+Deviation from equality manifests as an error $\epsilon[\Psi] \equiv E_\text{ab}[\Psi] - E_\text{eff}[\Psi]$ which quantifies the model quality.
+Therefore, the key insight of DMD is that downfolding $\hat{H}_\text{ab}$ can be cast to a functional appoximation of $E_\text{ab}[\Psi]$ on the low-energy space $\mathcal{LE}$.
+
 Given the general linear form of $\hat{H}_\text{eff}$ seen in \eqref{eq:Heff}, 
-the task of building an effective Hamiltonian then reduces to fitting a linear model that minimizes the error $\epsilon[\Psi]$ over $\mathcal{LE}$.
-The independent variables (descriptors) in the fitting are expectation values of the operators $d_k[\Psi] \equiv \langle \Psi |\hat{d}_k|\Psi \rangle$ and the target variable is $E[\Psi]$.
-
-This linear regression problem is tackled in three steps.
+the functional approximation reduces to fitting a linear model that minimizes the error $\epsilon[\Psi]$ over $\mathcal{LE}$.
+The fitting is tackled in three steps.
 First, low-energy states are sampled from $\mathcal{LE}$.
 Next, a pool of candidate operators $\{\hat{d}_k\}$ which may be included in the final model are selected.
-Finally, a linear regression is carried out with descriptor and energies evaluated on the sampled states as the independent and target variables.
-
-When developing models for realistic system, $\hat{H}$ is taken to be the \textit{ab initio} Hamiltonian
-\begin{equation}
-\hat{H}_\text{ab} = \sum_i -\frac{1}{2} \nabla_i^2 + \sum_{iI}\frac{Z_I}{|r_i - R_I|} + \frac{1}{2} \sum_{ij} \frac{1}{|r_i - r_j|} + E_\text{ion}(\{R_I\})  
-\label{eq:Hab}
-\end{equation} 
-where index $i,j$ correspond to electrons, $I$ to ions, and $E_\text{ion}$ corresponds to the ionic kinetic and potential energies. 
-The unit of energy is Hartree (Ha).
-To reduce the computational expense of \textit{ab initio} calculations one works under the Born-Oppenheimer approximation where $E_\text{ion}$ becomes a constant.
-Ionic effects related to phonon modes can be recovered by altering the ionic geometry and observing the effects on the electronic state.
+Finally, a linear regression is carried out with the expectation values $\{\langle \Psi | \hat{d}_k|\Psi \rangle, E_\text{ab}[\Psi] \}$  evaluated on the sampled states as the independent and target variables.
 

### 2019-09-18 draft.tex
methods done

@@
 
-DMD provides a framework for this downfolding by considering energy functionals.
+DMD provides a framework for this downfolding using energy functionals.
 In their seminal paper on DMD, Zheng \textit{et al.} \cite{Zheng2017} show that $\hat{H}_\text{eff}$ will exactly reproduce the low-lying eigenstates of $\hat{H}_\text{ab}$ if
@@
 \end{equation}
-Here $E_\text{ab}[\Psi] \equiv \langle \Psi | \hat{H}_\text{ab} | \Psi \rangle$ is the known \textit{ab initio} energy functional and $E_\text{eff}[\Psi] \equiv \langle \Psi | \hat{H}_\text{eff} | \Psi \rangle$ is the to be determined effective energy functional.
+Here $E_\text{ab}[\Psi] \equiv \langle \Psi | \hat{H}_\text{ab} | \Psi \rangle$ is the known \textit{ab initio} energy functional and $E_\text{eff}[\Psi] \equiv \langle \Psi | \hat{H}_\text{eff} | \Psi \rangle$ is the to-be-determined effective energy functional.
 Deviation from equality manifests as an error $\epsilon[\Psi] \equiv E_\text{ab}[\Psi] - E_\text{eff}[\Psi]$ which quantifies the model quality.
-Therefore, the key insight of DMD is that downfolding $\hat{H}_\text{ab}$ can be cast to a functional appoximation of $E_\text{ab}[\Psi]$ on the low-energy space $\mathcal{LE}$.
+In short, the key insight of DMD is that downfolding $\hat{H}_\text{ab}$ can be cast to a functional appoximation of $E_\text{ab}[\Psi]$ on the low-energy space $\mathcal{LE}$.
 
@@
 
-Low-energy states can be accessed using any wave function based \textit{ab initio} simulation technique, but two broad principles are used for selecting a method.
+Low-energy states are accessed using any wave function based \textit{ab initio} simulation technique, but two broad principles are used for selecting a method.
 First, the method should be able to generate accurate low-energy wave functions in the \textit{ab initio} Hilbert space, as the accuracy of the low-energy states dictate the accuracy of the final model.
 Second, the method should be relatively inexpensive, as large amounts of training data are preferred for a more robust regression.
-QMC methods have typically been selected due to their polynomial scaling computational cost and high accuracy, but DFT can also be used as an initial step trading accuracy for computational ease.
+QMC methods have been popular choices in previous DMD calculations \cite{Zheng2017, Wagner2015} due to their polynomial scaling computational cost, high accuracy, and applicability to both molecular and lattice systems.
 
-Candidate descriptors are typically selected using both physical intuition and statistical methods.
+Candidate descriptors are selected using both physical intuition and statistical
+methods, seen clearly in a DMD study on the stretched FeSe molecule \cite{Zheng2017}.
 An essential first pass involves selecting a large set of descriptors using knowledge of the low-energy degrees of the freedom of the system and physical principles like symmetry.
-Further filtering follows by examining the relationship between descriptors and the \textit{ab initio} energy over the low-energy sampled states: descriptors will typically be discarded if they do not exhibit covariance with the energy.
-Note, while there is no restriction on the form of the candidate operators other than Hermiticity, 1- and 2-body second quantized operators are commonly used.
+This set is then further reduced by discarding descriptors which do not exhibit covariance with the energy over the low-energy sampled states.
+Note, while there is no restriction on the candidate operators other than Hermiticity, 1- and 2-body second quantized operators are commonly used.
 

### 2019-09-18 draft.tex
finished intro revisions. one sentence left need ot udpate

@@
 While useful, it is typically impossible to know whether models built from these techniques accurately describe reality, as measures of model quality are rarely reported.
-Instead, comparisons of model predictions to experiment are conducted for model validation.
+Instead, model validation rely on comparisons to experiment.
 
@@
 The utility of DMD is made apparent when studying the recent surge in model development for TBLG.
-\textit{ab initio} model development has so far been restricted to tight-binding models which accurately describe single particle effects like band flattening, but do not address interactions \cite{Fang, Fang2016, PhysRevX.8.031088, PhysRevX.8.031087}.
+\textit{Ab initio} model development has so far been restricted to tight-binding models which accurately describe single particle effects like band flattening, but do not address interactions \cite{Fang, Fang2016, PhysRevX.8.031088, PhysRevX.8.031087}.
 Minimal interacting models have been proposed using traditional theoretical methods, but without a link to the \textit{ab initio} Hamiltonian it is unclear whether they can describe the real system \cite{PhysRevB.98.045103, PhysRevB.100.125116, PhysRevX.8.031089}.
@@
 
-As such, a project named QMC High-Accuracy Multiscale Modeling (QMC HAMM) between research groups at UIUC and Rice University is dedicated towards the development of an accurate continuum-scale interacting model for TBLG.
-The model development consists of two layers of downfolding: first, developing an atomic scale lattice model using \textit{ab initio} calculations, and second, downfolding the resultant lattice model to a continuum length scale.
-My thesis will be focused on the first stage of downfolding.
-I propose a two year plan for developing many-body lattice models for single- (SLG) and simple bi-layer (BLG) graphene systems using the DMD method and \textit{ab initio} fixed-node diffusion Monte Carlo (FN-DMC) calculations.
-My thesis will lay the necessary ground work for the development of a fully twist angle dependent lattice model for TBLG, and an eventual continuum model.
+%Rework this paragraph, could be much better with the new idea i have!
+Instead of jumping head on into TBLG, however, I propose a two year plan for modeling the AA and AB stacked BLG systems using DMD with \textit{ab initio} fixed-node diffusion Monte Carlo (FN-DMC) calculations.
+I will first develop a model for single-layer graphene (SLG) with long-range Coulomb and electron-phonon interactions, and then extend this to the BLG systems by introducing inter-layer couplings.
+The SLG model will serve as an important tool in understanding the interplay of electronic and geometric degrees of freedom in graphene, as seen in straining experiments \cite{Naumis_2017}.
+The AA/AB BLG models will address the form and magnitude of inter-layer coupling in the bi-layer, and the effect of a single layer on the intra-layer interactions of its neighbor.
+My work will also provide a baseline for follow-ups on the effects of commensurate twists on inter- and intra-layer coupling in TBLG.
 
@@
 If the initial trial wave functions are appropriately chosen, the final projected states will be within the target low-energy subspace of $\mathcal{H}$.
-The training data for DMD are then extracted by evaluating the expectation values of the Hamiltonian and candidate operators using the final distributions of walkers from the FN-DMC calculations \cite{Wagner2015}.
+The training data for DMD are then extracted by evaluating the expectation values of the Hamiltonian and candidate operators using the final distributions of walkers from the FN-DMC calculations \cite{Wagner2017}.
 
@@
 \subsection{Benzene}
-The first system I will develop a model for is the benzene molecule, shown in Figure \ref{fig:proposed}a.
-While relatively simple, this molecule shares many similarities with and forms the basic unit of SLG.
+While relatively simple, the benzene molecule shares many similarities with and forms the basic unit of SLG.
+A schematic drawing of the molecule is shown in Figure \ref{fig:proposed}a.
 The delocalized $\pi$ orbitals in benzene carry over directly into SLG where they are the primary contributors to the well known Dirac bands.
@@
 
-\subsection{Single layer graphene (SLG)}
-Single layer graphene is composed of a triangular Bravais lattice with two carbon atoms per unit cell.
+\subsection{Single layer graphene}
+SLG is composed of a triangular Bravais lattice with two carbon atoms per unit cell.
 In terms of a non-interacting picture, the $\sigma$ orbitals composed of C $2s, 2p_x,$ and $2p_y$ lie below the Fermi level.
@@
 
-\subsection{Bilayer graphene (BLG)}
+\subsection{Bilayer graphene}
 The last systems I will develop models for are AA and AB stacked BLG, shown in Figure \ref{fig:proposed}b.

### 2019-09-18 draft.tex
changed a sentence

@@
 While useful, it is typically impossible to know whether models built from these techniques accurately describe reality, as measures of model quality are rarely reported.
-Instead, model validation rely on comparisons to experiment.
+Instead, model validation relies on comparisons to experiment.
 
@@
 
-%Rework this paragraph, could be much better with the new idea i have!
-Instead of jumping head on into TBLG, however, I propose a two year plan for modeling the AA and AB stacked BLG systems using DMD with \textit{ab initio} fixed-node diffusion Monte Carlo (FN-DMC) calculations.
+Instead of jumping head first into TBLG, however, I propose a two year plan for modeling the AA and AB stacked BLG systems using DMD with \textit{ab initio} fixed-node diffusion Monte Carlo (FN-DMC) calculations.
 I will first develop a model for single-layer graphene (SLG) with long-range Coulomb and electron-phonon interactions, and then extend this to the BLG systems by introducing inter-layer couplings.
-The SLG model will serve as an important tool in understanding the interplay of electronic and geometric degrees of freedom in graphene, as seen in straining experiments \cite{Naumis_2017}.
+The SLG model will demonstrate the importance of long range interactions and geometric relaxation in describing the low-energy excitations of graphene.
 The AA/AB BLG models will address the form and magnitude of inter-layer coupling in the bi-layer, and the effect of a single layer on the intra-layer interactions of its neighbor.
-My work will also provide a baseline for follow-ups on the effects of commensurate twists on inter- and intra-layer coupling in TBLG.
+My work will also provide a baseline for follow-up studies on the effects of commensurate twists on inter- and intra-layer coupling in TBLG.
 

### 2019-09-19 draft.tex
updated preliminary + plan

@@
 Here I propose a thesis plan for developing atomic scale many-body model Hamiltonians for single- and simple bi-layer graphene systems within the DMD framework using high accuracy \textit{ab initio} quantum Monte Carlo (QMC) calculations.
-The results of my thesis will lay the groundwork for a continuing four year project to build a twist dependent many-body continuum model for the twisted bilayer graphene (TBLG) system.
-To demonstrate my preparedness I present a model Hamiltonian developed using DMD which accurately reproduces the spectra and properties of the lowest lying excitations of the CuO molecule.
+The results of my thesis will contribute significantly to the understanding of low-energy effective interactions in single and bilayer graphene, and will provide a baseline for continuing study of the twisted bi-layer graphene (TBLG) system.
+To demonstrate my preparedness I present progress on a many-body model for the CuO molecule and my work in QMC methods development.
 \end{abstract}
@@
 
+\begin{figure*}
+\centering
+%\hspace*{-0.25in}
+\includegraphics[width=1.0\linewidth]{./figs/cuo_regr_2.pdf}
+\caption{Figures related to model regression for the CuO molecule: a) Intrinsic atomic orbitals used as a single particle basis for the model Hamiltonian, b) Results of PCA showing the CEVR and CN of a data set containing N$_{comp}$  highest ranked PCs, c) The final regressed model parameters, d) Comparison of four lowest lying eigenstates of CuO molecule between experiment and three computation methods.}
+\label{fig:cuo_regr}
+\end{figure*}
+
 The first chunk of descriptors were selected using physical intuition and symmetry principles.
@@
 The effective on-site Coulomb repulsion $U_s$ ensures the model places all such $4s^{2}$ states high in energy.
-
-\begin{figure*}
-\centering
-%\hspace*{-0.25in}
-\includegraphics[width=1.0\linewidth]{./figs/cuo_regr_2.pdf}
-\caption{Figures related to model regression for the CuO molecule: a) Intrinsic atomic orbitals used as a single particle basis for the model Hamiltonian, b) Results of PCA showing the CEVR and CN of a data set containing N$_{comp}$  highest ranked PCs, c) The final regressed model parameters, d) Comparison of four lowest lying eigenstates of CuO molecule between experiment and three computation methods.}
-\label{fig:cuo_regr}
-\end{figure*}
 
@@
 This over estimation may arise from an inadequate relaxation of the $\sigma$ type orbitals which form a three orbital hybridization complex in this system.
+
+\vspace{-10pt}
+\paragraph{Next steps}
+I have shown that the my model has comparable accuracy to modern computational techniques.
+This poses the question, however: why use DMD if one can reproduce eigenstates with simpler methods? 
+The next steps in the project will address this question by demonstrating the unique interpretability of a model.
+I plan on accomplishing this by refitting my model after dropping specific interactions, like $n_{4s,\uparrow} n_{4s,\downarrow}$, and quantitatively describe the effect on the model spectra and eigenstates, as well as renormalization of non-interacting parameters.
+In doing so, I will be able to categorize the precise effects of certain interactions on the low-energy spectra of CuO, a level of detailed interpretation unavailable to methods like CCSD(T).
+
+\subsection{Methods development and coding}
+In addition to directly applying DMD, I have worked on the methods and code development needed to efficiently carry out downfolding.
+My first contribution was developing the theory and implementation of an efficient MC estimator for wave function parameter gradients.
+The parameter gradient of a wave function, $\langle \Psi(\vec{p}) | \Psi_p (\vec{p}) \rangle$ where $| \Psi_p (\vec{p}) \rangle \equiv \partial_{p} |\Psi (\vec{p}) \rangle$ , is a quantity that pops up frequently in QMC, from optimization algorithms \cite{Toulouse2007} to density matrix downfolding \cite{2019APS..MARF18008W}.
+Unfortunately, a naive Monte Carlo estimate of the gradient has an infinite fourth moment due to divergences near the nodal surface of $|\Psi \rangle$, so MC error estimates for the gradient are very unstable.
+I proved that the fourth moment can be made finite when using a stable estimator 
+\begin{equation} 
+\hat{\theta}=\begin{cases}
+\frac{\Psi_p(\textbf{r})}{\Psi(\textbf{r})} & \text{away from node} \\
+ \frac{\Psi_p(\textbf{r})-\Psi_p(\textbf{N})}{\Psi(\textbf{r})} & \text{near node}.      
+\end{cases}
+\end{equation}
+where $\vec{N}$ is a nodal point.
+I implemented, tested and integrated this estimator into QWalk.
+A publication with performance details of the estimator is in progress.
+
+I also made significant contributions to a new QMC package called PyQMC. 
+This package is oriented towards methods development, and will be the platform for most of my future DMD calculations.
+I implemented multi-Slater wave functions, ran consistency checks against QWalk, and worked on execution speedups for costly methods like VMC and wave function optimization.
+I plan on implementing 2-body RDM evaluation and Ewald summations too, thus providing the baseline tools for DMD on molecular and periodic systems.
 
@@
 
-I will work on extending an on-site Hubbard model for SLG \cite{Zheng2017, Wagner2015} to one with long-range Coulomb interactions.
+I will work on extending an on-site Hubbard model for SLG \cite{Zheng2017, Wagner2015} to one with long-range Coulomb interactions and electron-phonon coupling.
 The candidate model will be of the form:
 \begin{equation}
-H_\text{SLG} = -\sum_{i,j} t_{ij}c_i^\dagger c_j + U \sum_i n_{i\uparrow}n_{i\downarrow}  + \sum_{i,j} V_{ij} n_i n_j
+H_\text{SLG} = -\sum_{i,j} t_{ij}c_i^\dagger c_j + U \sum_i n_{i\uparrow}n_{i\downarrow}  + \sum_{i,j} V_{ij} n_i n_j + E_0
 \label{Hslg}
@@
 I will consider both long range hopping and density-density interaction terms extending across the computational unit cell, with the final goal of developing position dependent parameter values $t(r), V(r)$.
+Phononic contributions will be incorporated by fitting a sequence of models to altered geometries, leading to displacement dependent parameters similar to a previous spin-phonon calculation \cite{Wagner2014}.
 
-The resultant model will be able to non-parametrically determine the form
-of long-range interactions in graphene.
-This is an important contribution to the study of interaction effects in graphene as many potential properties of the system depend on details of these long-range interactions.
-Such effects include edge magentism \cite{PhysRevB.95.195420} and non-local electrodynamics \cite{PhysRevB.90.045137}.
+An important result from this study would be an accurate evaluation of $V(r)$ directly from \textit{ab initio} simulations.
+This result would be able to validate commonly used models for the spatial dependence, like the Ohno form \cite{Ohno1964}.
+Further, it will help develop understanding of interaction effects in graphene which depend on the details of $V(r)$, like edge magentism \cite{PhysRevB.95.195420} and non-local electrodynamics \cite{PhysRevB.90.045137}.
+Additionally, by introducing displacement dependent parameters, one can determine whether phonons in graphene serve to simply renormalize interactions, or drastically alter them.
+This determination is integral to understanding effects of strain on electronic properties as seen in experiment \cite{Naumis_2017}.
 
@@
 
-With these two models, one can begin to answer some questions about the inter-layer coupling in TBLG.
+With these two models, one can begin to answer some questions about the inter-layer coupling in BLG.
 First, one can determine, at least for the simplest orientations, whether the inter-layer coupling takes the form of a hopping or interaction term.
-While current studies tend to assume the prior \cite{Fang, Fang2016, PhysRevX.8.031088, PhysRevX.8.031087}, there is no good reason to discount the latter.
 Further, the quantitative details of spatial dependence, which are intimately related to the strongly correlated behavior in TBLG, can begin to be addressed.
-
-\vspace{-10pt}
-\paragraph{QMC HAMM}
-A major goal of the QMC HAMM project is to develop a twist dependent, Moire length scale, interacting model for TBLG using \textit{ab initio} QMC calculations.
-The first step in doing so is developing an atomic scale model for TBLG, and the necessary prerequisites for this step are a strong understanding of effective theories of the basic constituents of the system.
-The models I propose to develop for SLG and AA/AB BLG provide exactly this kind of understanding.
-As such, my work will lay the foundation for the four year QMC HAMM project to build an effective theory for TBLG.
+Lastly, when combined with the SLG model, it will be possible to determine how the intra-layer model is affected by the presence of a neighboring graphene sheet.
+The simple BLG models will also enable follow-up studies on the twist dependence of model parameters for TBLG.
 
@@
 For each stacking, need four months to build the model and two months for finite size extrapolation.
+In total, my work will contribute significantly to the understanding of low-energy effective interactions in single and bilayer graphene.
 

### 2019-09-19 draft.tex
updated some small lines! almost done

@@
 Instead of jumping head first into TBLG, however, I propose a two year plan for modeling the AA and AB stacked BLG systems using DMD with \textit{ab initio} fixed-node diffusion Monte Carlo (FN-DMC) calculations.
-I will first develop a model for single-layer graphene (SLG) with long-range Coulomb and electron-phonon interactions, and then extend this to the BLG systems by introducing inter-layer couplings.
-The SLG model will demonstrate the importance of long range interactions and geometric relaxation in describing the low-energy excitations of graphene.
+I will first develop a model for single-layer graphene (SLG) and then extend this to the BLG systems by introducing inter-layer couplings.
+The SLG model will provide an accurate evaluation of the effective Coulomb interaction spatial dependence without reliance on analytic forms like Ohno \cite{Ohno1964}, and will describe the role of phonons in modifying these interactions.
 The AA/AB BLG models will address the form and magnitude of inter-layer coupling in the bi-layer, and the effect of a single layer on the intra-layer interactions of its neighbor.
@@
 \paragraph{Sampling low-energy states}
-For the CuO molecule, low-energy states were generated by using FN-DMC projected multi-Slater-Jastrow (MSJ) trial wave functions.
+Low-energy states were generated by using FN-DMC projected multi-Slater-Jastrow (MSJ) trial wave functions.
 The trial wave functions took the form
@@
 
-The determinants $|\text{D}_j \rangle$ were constructed from the orbitals of independent symmetry-targeted unrestricted Kohn Sham (UKS) calculations.
-This choice was made to ensure that the three major low-energy variations seen in the molecule were captured: occupancy, spin configuration and relaxation of the Cu 3d, 4s and O 2p orbitals.
-Only determinants with electrons in the Cu 3d, 4s, and O 2p orbitals and at most a single hole in the Cu 3d shell were included in \eqref{eq:sampling} to reflect the low-energy excitations seen in experiment.
-The UKS calculations were done using PySCF \cite{Sun2018}, a B3LYP functional \cite{doi:10.1063/1.464304, PhysRevB.37.785}, Trail-Needs psuedopotential and VTZ Trail-Needs basis \cite{doi:10.1063/1.4811651}.
+The determinants were constructed from the orbitals of symmetry-targeted unrestricted Kohn Sham (UKS) calculations.
+Only determinants composed of Cu 3d, 4s, and O 2p orbitals with at most a single hole in the Cu 3d shell were included in \eqref{eq:sampling} to match the low-energy excitations seen in experiment.
+The UKS calculations were done using PySCF \cite{Sun2018}, a B3LYP functional \cite{doi:10.1063/1.464304}, Trail-Needs psuedopotential and VTZ Trail-Needs basis \cite{doi:10.1063/1.4811651}.
 
@@
 
-A single three-body Jastrow factor was used for all trial wave functions, and was optimized on the lowest energy determinant $|\text{D}_0 \rangle$ using an energy optimization scheme \cite{Toulouse2007}.
-This choice significantly reduced the computational cost of sampling low-energy states as linear optimization was not necessary for every sampled state.
+A single three-body Jastrow factor was used for all trial wave functions, and was optimized on the lowest energy determinant using an energy optimization scheme \cite{Toulouse2007}.
+This choice greatly reduced the computational cost of sampling low-energy states as linear optimization was not necessary for every sampled state.
 The resulting MSJ trial function remained low in energy as variations of the Jastrow factor among low-energy states is expected to be small.
@@
 \paragraph{Selecting candidate descriptors}
-The following set of 1- and 2-body operators were selected as candidates for the model based on an understanding of the low-energy excitations of the CuO molecule and properties of the sampled low-energy states:
+The following set of 1- and 2-body operators were selected as candidates for the model :
 \begin{equation}
@@
 \end{equation}
-The operator subscripts correspond to the intrinsic atomic orbital (IAO) basis elements constructed using the molecular orbitals in \eqref{eq:sampling}, and can be seen in \ref{fig:cuo_regr}a.
+The operator subscripts correspond to the intrinsic atomic orbital (IAO) basis elements constructed using the molecular orbitals in \eqref{eq:sampling}, and can be seen in Figure \ref{fig:cuo_regr}a.
+
+The first chunk of descriptors were selected using physical intuition.
+As described above, the low-lying excitations in CuO observed in experiment involve the Cu 3d, 4s and O 2p atomic orbitals.
+Given that orbitals also hybridize, the occupation energies of and hopping terms between these orbitals were selected as candidate operators.
+Symmetry considerations help narrow down the set as hopping terms are only non-zero between same symmetry orbitals, and occupation energies can be traced over symmetry related orbitals, for example $n_{2p_\pi} = n_{2p_x} + n_{2p_y}$.
+The term $\sum_{i\in {\{\text{xy, xz, ...}}\}} \vec{S}_{4s} \cdot \vec{S}_{3d_i}$ was also considered, motivated by the existence of a $\sim $0.5 eV Hund's coupling observed on the copper atom \cite{Data2009}.
+
+The inclusion of a double occupancy term $n_{4s,\uparrow} n_{4s,\downarrow}$ was motivated directly from the sampled states.
+Within the sampled data, states with a Cu $4s^{2}$ configuration lie at least 7 eV above the lowest energy sampled state, while the lowest energy state with a $4s^{1}$ configuration lies at 1.6 eV.
+If the double occupancy term was not included, the effective theory would predict the prior state to have an energy of $\sim $ 3 eV, an underestimation of 4 eV.
+The effective on-site Coulomb repulsion $U_s$ ensures the model places all such $4s^{2}$ states high in energy.
 
@@
 
-The first chunk of descriptors were selected using physical intuition and symmetry principles.
-As described above, the low-lying excitations in CuO observed in experiment involve the Cu 3d, 4s and O 2p atomic orbitals.
-Given that orbitals also hybridize, the occupation energies of and hopping terms between these orbitals were selected as candidate operators.
-Symmetry considerations help narrow down the set as hopping terms are only non-zero between same symmetry orbitals, and occupation energies can be traced over symmetry related orbitals, for example $n_{2p_\pi} = n_{2p_x} + n_{2p_y}$.
-The term $\sum_{i\in {\{\text{xy, xz, ...}}\}} \vec{S}_{4s} \cdot \vec{S}_{3d_i}$ was also considered, motivated by the existence of a $\sim $0.5 eV Hund's coupling observed on the copper atom \cite{Data2009}.
-
-The inclusion of a double occupancy term $n_{4s,\uparrow} n_{4s,\downarrow}$ was motivated directly from the sampled low-energy states.
-Within the sampled data, states with a Cu $4s^{2}$ configuration lie at least 7 eV above the lowest energy sampled state, while the lowest energy state with a $4s^{1}$ configuration lies at 1.6 eV.
-If the double occupancy term was not included, the effective theory would predict the prior state to have an energy of $\sim $ 3eV, an underestimation of 4 eV from the \textit{ab initio} result.
-The effective on-site Coulomb repulsion $U_s$ ensures the model places all such $4s^{2}$ states high in energy.
-
 \vspace{-10pt}
 \paragraph{Fitting the model}
-A direct attempt at fitting all eleven terms in \eqref{eq:models} leads to a high goodness of fit, R$^2$ = 0.99, but unexpectedly large coefficients for the fit model.
-This is seen in an enormous occupation energy gap of $\epsilon_{2p_\pi} - \epsilon_{3d_\delta} = $ 5.87 eV, which is 2 $-$ 3 eV in experiment and our sampled data.
+A direct attempt at fitting all eleven terms in \eqref{eq:models} leads to a high goodness of fit, R$^2$ = 0.99, but unreasonably large coefficients for model.
+This is seen in a huge energy gap of $\epsilon_{2p_\pi} - \epsilon_{3d_\delta} = $ 5.87 eV, which is 2 $-$ 3 eV in experiment and our sampled data.
 This inflation of model parameters was found to be associated with multicollinearity among the candidate descriptors, quantified by a large condition number (CN) of 1470.
@@
 \paragraph{Solutions of the model}
-The final model in \ref{fig:cuo_regr}c was solved by exact diagonalization, and the solutions agree well with experimental spectra and state assignments.
+The final model in \ref{fig:cuo_regr}c was solved by exact diagonalization.
 Shown in Figure \ref{fig:cuo_regr}d is a comparison between the excitation energies calculated from the model and those seen in the latest anion photoelectron spectroscopy measurements \cite{Wu1997}.
@@
 \paragraph{Next steps}
-I have shown that the my model has comparable accuracy to modern computational techniques.
-This poses the question, however: why use DMD if one can reproduce eigenstates with simpler methods? 
+I have shown that my model has comparable accuracy to modern computational techniques.
+This poses the question, however: why use DMD if simpler methods are just as accurate?
 The next steps in the project will address this question by demonstrating the unique interpretability of a model.
-I plan on accomplishing this by refitting my model after dropping specific interactions, like $n_{4s,\uparrow} n_{4s,\downarrow}$, and quantitatively describe the effect on the model spectra and eigenstates, as well as renormalization of non-interacting parameters.
-In doing so, I will be able to categorize the precise effects of certain interactions on the low-energy spectra of CuO, a level of detailed interpretation unavailable to methods like CCSD(T).
+I plan on accomplishing this by refitting my model after dropping specific interactions, like $n_{4s,\uparrow} n_{4s,\downarrow}$, and quantitatively evaluate the effect on the model spectra, eigenstates, and renormalization of non-interacting parameters.
+In doing so, I will be able to describe the precise effects of certain interactions on the low-energy spectra of CuO, a level of detailed interpretation unavailable to methods like CCSD(T).
 
@@
 I will consider both long range hopping and density-density interaction terms extending across the computational unit cell, with the final goal of developing position dependent parameter values $t(r), V(r)$.
-Phononic contributions will be incorporated by fitting a sequence of models to altered geometries, leading to displacement dependent parameters similar to a previous spin-phonon calculation \cite{Wagner2014}.
+Phononic contributions will be incorporated by fitting a sequence of models to altered geometries, leading to displacement dependent parameters similar to a spin-phonon model for the cuprates \cite{Wagner2014}.
 
@@
 The SLG model fitting should take another eight months.
-As an extension of a previous work, the first three months will be dedicate to reproducing the on-site Hubbard model from that work.
+The first three months will be dedicate to reproducing the on-site Hubbard model from that work.
 Inclusion of long range Coulomb interactions into the effective theory should take another three months.
-The last two months will be geared towards finite size extrapolation in order to ensure the final model is applicable to bulk systems.
+The last two months will be geared towards finite size extrapolation to ensure the final model is applicable to bulk systems.
 The last year of the project will be focused on the AA and AB BLG, with each project taking six months to complete.
 For each stacking, need four months to build the model and two months for finite size extrapolation.
-In total, my work will contribute significantly to the understanding of low-energy effective interactions in single and bilayer graphene.
+In summary, my work will contribute significantly to the understanding of low-energy effective interactions in single and bilayer graphene.
 

### 2019-09-19 draft.tex
grammar
number start after title page

@@
 \begin{abstract}
+\thispagestyle{empty}
 Building accurate model Hamiltonians for materials is a fundamental problem in modern condensed matter physics.
@@
 Committee: Lucas Wagner, Brian DeMarco, Andre Schleife, Karin Dahmen
+\clearpage
 \pagebreak
-
+\setcounter{page}{1}
 \section{Introduction}
@@
 The last year of the project will be focused on the AA and AB BLG, with each project taking six months to complete.
-For each stacking, need four months to build the model and two months for finite size extrapolation.
+For each stacking, I will need four months to build the model and two months for finite size extrapolation.
 In summary, my work will contribute significantly to the understanding of low-energy effective interactions in single and bilayer graphene.

### 2019-09-20 draft.tex
updated main body text

@@
 Building accurate model Hamiltonians for materials is a fundamental problem in modern condensed matter physics.
-A recently developed density matrix downfolding (DMD) \cite{Zheng2017} method provides a very direct framework for downfolding the \textit{ab initio} Hamiltonian to a many-body effective Hamiltonian using only \textit{ab initio} calculations.
+A recently developed density matrix downfolding (DMD) \cite{Zheng2017} method provides a direct framework for downfolding the \textit{ab initio} Hamiltonian to a many-body effective Hamiltonian using only \textit{ab initio} calculations.
 Here I propose a thesis plan for developing atomic scale many-body model Hamiltonians for single- and simple bi-layer graphene systems within the DMD framework using high accuracy \textit{ab initio} quantum Monte Carlo (QMC) calculations.
-The results of my thesis will contribute significantly to the understanding of low-energy effective interactions in single and bilayer graphene, and will provide a baseline for continuing study of the twisted bi-layer graphene (TBLG) system.
-To demonstrate my preparedness I present progress on a many-body model for the CuO molecule and my work in QMC methods development.
+The results of my thesis will contribute significantly to the understanding of low-energy effective interactions in single and bilayer graphene, and will provide a baseline for a continuing study of the twisted bi-layer graphene (TBLG) system.
+To demonstrate my preparedness, I present progress on a many-body model for the CuO molecule and my work in QMC methods development.
 \end{abstract}
@@
 Developing model Hamiltonians for materials is a fundamental problem in modern condensed matter physics.
-A model Hamiltonian is a simplified description of a real material which can accurately reproduce the low-lying eigenstates and energies of the material.
+A model Hamiltonian is a simplified description of a material which can accurately reproduce the low-lying eigenstates and energies of that material.
 Accurate models give clear and direct indications of effective interactions within a material and underpin modern theoretical understanding of ideas like emergence and symmetry breaking.
@@
 where index $i,j$ correspond to electrons, $I$ to ions, $E_\text{ion}$ corresponds to the total ionic energy, and the unit of energy is Hartree (Ha).
-The goal is to develop an effective Hamiltonian $\hat{H}_\text{eff}$ defined on a low-energy subspace $\mathcal{LE} \subset \mathcal{H}$ which can accurately reproduce the lowest lying eigenstates and spectra of $\hat{H}_\text{ab}$.
+The goal is to develop an effective Hamiltonian $\hat{H}_\text{eff}$ defined on a low-energy subspace $\mathcal{LE} \subset \mathcal{H}$ which can accurately reproduce the lowest lying eigenstates and energies of $\hat{H}_\text{ab}$.
 $\hat{H}_\text{eff}$ is taken to be a linear operator on $\mathcal{LE}$ and can be expressed as a linear combination of Hermitian operators $\hat{d}_k$, with coupling constants $g_k$ and energy offset $E_0$
@@
 Deviation from equality manifests as an error $\epsilon[\Psi] \equiv E_\text{ab}[\Psi] - E_\text{eff}[\Psi]$ which quantifies the model quality.
-In short, the key insight of DMD is that downfolding $\hat{H}_\text{ab}$ can be cast to a functional appoximation of $E_\text{ab}[\Psi]$ on the low-energy space $\mathcal{LE}$.
+In short, the key insight of DMD is that the downfolding $\hat{H}_\text{ab}$ of can be cast to a functional appoximation of $E_\text{ab}[\Psi]$ on $\mathcal{LE}$.
 
 Given the general linear form of $\hat{H}_\text{eff}$ seen in \eqref{eq:Heff}, 
-the functional approximation reduces to fitting a linear model that minimizes the error $\epsilon[\Psi]$ over $\mathcal{LE}$.
+the functional approximation reduces to fitting a linear model that minimizes $\epsilon[\Psi]$ over $\mathcal{LE}$.
 The fitting is tackled in three steps.
@@
 The final projected $\Psi_\text{DMC}$ for these different trial functions will be the lowest energy states in $\mathcal{H}$ for the given nodal structures.
-If the initial trial wave functions are appropriately chosen, the final projected states will be within the target low-energy subspace of $\mathcal{H}$.
+If the initial trial wave functions are appropriately chosen, the final projected states will be within the target $\mathcal{LE}$.
 The training data for DMD are then extracted by evaluating the expectation values of the Hamiltonian and candidate operators using the final distributions of walkers from the FN-DMC calculations \cite{Wagner2017}.
@@
 A shell sampling scheme was used to select coefficients $p_j$ for the trial wave functions.
-We began by fixing the coefficient $p_0 = \sqrt{w}$ where $w \in \{1.0, 0.8, ..., 0.2\}$. 
+I began by fixing the coefficient $p_0 = \sqrt{w}$ where $w \in \{1.0, 0.8, ..., 0.2\}$. 
 For each choice of $w$, n = 5 states were selected by sampling the unassigned coefficients from independent, standard normal distributions such that $\sum_j p_j^2 = 1$. 
@@
 A single three-body Jastrow factor was used for all trial wave functions, and was optimized on the lowest energy determinant using an energy optimization scheme \cite{Toulouse2007}.
-This choice greatly reduced the computational cost of sampling low-energy states as linear optimization was not necessary for every sampled state.
+This choice greatly reduced the computational cost of sampling low-energy states as optimization was not necessary for every sampled state.
 The resulting MSJ trial function remained low in energy as variations of the Jastrow factor among low-energy states is expected to be small.
@@
 A direct attempt at fitting all eleven terms in \eqref{eq:models} leads to a high goodness of fit, R$^2$ = 0.99, but unreasonably large coefficients for model.
-This is seen in a huge energy gap of $\epsilon_{2p_\pi} - \epsilon_{3d_\delta} = $ 5.87 eV, which is 2 $-$ 3 eV in experiment and our sampled data.
+This is seen in a huge energy gap of $\epsilon_{2p_\pi} - \epsilon_{3d_\delta} = $ 5.87 eV, which is 2 $-$ 3 eV in experiment and the sampled data.
 This inflation of model parameters was found to be associated with multicollinearity among the candidate descriptors, quantified by a large condition number (CN) of 1470.
@@
 The CEVR indicates how much of the variation in the descriptor values a subset of PCs capture, and the CN quantifies the multicollinearity among these PCs.
-The sharp decrease in the condition number and sustained high CEVR up to nine principal components indicates the presence of only nine linearly independent variations within the descriptor set.
+The sharp decrease in the condition number and sustained high CEVR up to nine PCs indicates the presence of only nine linearly independent variations within the descriptor set.
 
 To reduce multicollinearity while maintaining a high goodness of fit, a model was fit to these nine PCs.
-When trained on all the sampled data this model yields an $R^2$ = 0.98 with a CN of 257. 
+This model yields an $R^2$ = 0.98 with a CN of 257.
 An inverse PCA transformation was used to map the fit coefficients corresponding to the PC descriptors back into the original eleven coefficients of \eqref{eq:models}.
@@
 The energies for the UKS and CCSD(T) \cite{Xian2000} states are also shown in Figure \ref{fig:cuo_regr}d as green and red circles, respectively.
-The UKS energies were calculated in this work in en route to constructing the determinants $|\text{D}_j\rangle$ used in our sampling scheme \eqref{eq:sampling}.
+The UKS energies were calculated in this work en route to constructing the determinants $|\text{D}_j\rangle$ used in \eqref{eq:sampling}.
 The model energies agree with both theoretical techniques within errorbars for the first three states shown.
@@
 This poses the question, however: why use DMD if simpler methods are just as accurate?
-The next steps in the project will address this question by demonstrating the unique interpretability of a model.
-I plan on accomplishing this by refitting my model after dropping specific interactions, like $n_{4s,\uparrow} n_{4s,\downarrow}$, and quantitatively evaluate the effect on the model spectra, eigenstates, and renormalization of non-interacting parameters.
+In the next steps of the project I will address this question by demonstrating the unique interpretability of a model.
+I plan on accomplishing this by refitting my model after dropping specific interactions, like $n_{4s,\uparrow} n_{4s,\downarrow}$, and quantitatively evaluating the effect on the model spectra, eigenstates, and renormalization of non-interacting parameters.
 In doing so, I will be able to describe the precise effects of certain interactions on the low-energy spectra of CuO, a level of detailed interpretation unavailable to methods like CCSD(T).
@@
 A new constrained cVMC method will be used to sample the low-energy wave functions necessary for DMD.
-The cVMC method works by minimizing the following cost function for a parameterized wave function $|\Psi(\vec{p})\rangle$:
+The cVMC method works by minimizing the following cost function for a parameterized wave function $|\Psi(\vec{p})\rangle$ with respect to the parameters $\vec{p}$:
 \begin{equation}
-\vec{p}^* = \text{argmin} \ \langle \Psi(\vec{p}) | \hat{H}_\text{ab} | \Psi(\vec{p}) \rangle - \sum_k [\lambda_k \langle \Psi(\vec{p}) | d_k - d_{k}^0 | \Psi(\vec{p}) \rangle]^2
+\vec{p}^* = \text{argmin} \ \langle \Psi(\vec{p}) | \hat{H}_\text{ab} | \Psi(\vec{p}) \rangle - \sum_k [\lambda_k \langle \Psi(\vec{p}) | d_k - d_{k}^0 | \Psi(\vec{p}) \rangle]^2.
 \end{equation}
-with respect to the parameters $\vec{p}$.
 The quantity $\vec{d}$ is short hand for a list of candidate descriptors to sample variations in, $\vec{d}^*$ is the target values of those descriptors, and $\vec{\lambda}$ controls the importance of finding a state with low-energy versus one with descriptor values near $\vec{d}^*$.
-This method allows one to target low-energy wave functions which sit near a target point $\vec{d}^*$ in descriptor space, providing a more systematic method for sampling the low-energy space.
+This method allows one to target low-energy wave functions which sit near $\vec{d}^*$ in descriptor space, providing a more systematic method for sampling the low-energy space.
 
-I will use a multi-Slater-Jastrow (MSJ) parameterization in the cVMC sampling.
+I will use an MSJ parameterization in the cVMC sampling.
 The multi-Slater expansion will consist of determinants of single particle orbitals calculated using DFT.
@@
 My objective will be to reproduce a previously developed many-body model for the benzene molecule using QMC DMD \cite{Wagner2015}.
-In this study, an extended Hubbard model built on the six $\pi$ orbitals was found to perform well in describing the low-lying excitations of the material.
+In that study, an extended Hubbard model built on the six $\pi$ orbitals was found to perform well in describing the low-lying excitations of the material.
 The final model took the form 
 \begin{equation}
-H_\text{benz} = -\sum_{\langle i,j \rangle} t_{ij}c_i^\dagger c_j + U \sum_i n_{i\uparrow}n_{i\downarrow}  + \sum_{\langle i,j \rangle}V_{ij} n_i n_j + \sum_{\langle \langle i,j \rangle\rangle}V_{ij} n_i n_j
+H_\text{benz} = -\sum_{\langle i,j \rangle} t_{ij}c_i^\dagger c_j + U \sum_i n_{i\uparrow}n_{i\downarrow}  + \sum_{\langle i,j \rangle}V_{ij} n_i n_j + \sum_{\langle \langle i,j \rangle\rangle}V_{ij} n_i n_j + E_0
 \label{Hbenz}
@@
 Downfolding techniques have advanced significantly since the initial model for benzene was developed in 2015.
-Methods like constrained VMC for low-energy sampling have been posited but not thoroughly tested.
+For example, methods like constrained VMC for low-energy sampling have been posited, but not thoroughly tested.
 By comparing models built using new methods directly to the 2015 model, it will be possible to quantify the efficiency and effectiveness of these methods.
@@
 This result would be able to validate commonly used models for the spatial dependence, like the Ohno form \cite{Ohno1964}.
-Further, it will help develop understanding of interaction effects in graphene which depend on the details of $V(r)$, like edge magentism \cite{PhysRevB.95.195420} and non-local electrodynamics \cite{PhysRevB.90.045137}.
+It will also help develop understanding of interaction effects in graphene which depend on the details of $V(r)$, like edge magentism \cite{PhysRevB.95.195420} and non-local electrodynamics \cite{PhysRevB.90.045137}.
 Additionally, by introducing displacement dependent parameters, one can determine whether phonons in graphene serve to simply renormalize interactions, or drastically alter them.
@@
 \begin{equation}
-H_\text{BLG} = H_\text{SLG}^{(1)} + H_\text{SLG}^{(2)} - \sum_{i_1, i_2} t_{i_1, i_2}^\prime c_{i_1}^\dagger c_{i_2} + h.c.
+H_\text{BLG} = H_\text{SLG}^{(1)} + H_\text{SLG}^{(2)} - \sum_{i_1, i_2} t_{i_1, i_2}^\prime c_{i_1}^\dagger c_{i_2} + h.c. + E_0
 \label{Hblg}
@@
 With these two models, one can begin to answer some questions about the inter-layer coupling in BLG.
-First, one can determine, at least for the simplest orientations, whether the inter-layer coupling takes the form of a hopping or interaction term.
-Further, the quantitative details of spatial dependence, which are intimately related to the strongly correlated behavior in TBLG, can begin to be addressed.
-Lastly, when combined with the SLG model, it will be possible to determine how the intra-layer model is affected by the presence of a neighboring graphene sheet.
-The simple BLG models will also enable follow-up studies on the twist dependence of model parameters for TBLG.
+First, one can determine, at least for the simplest orientations, whether the inter-layer coupling takes the form of a hopping or an interaction term.
+Quantitative details of the interlayer coupling spatial dependence, which are intimately related to the strongly correlated behavior in TBLG, can also begin to be addressed.
+Further, when combined with the SLG model, it will be possible to determine how the intra-layer model is affected by the presence of a neighboring graphene sheet.
+Lastly, the simple BLG models will enable follow-up studies on the twist dependence of model parameters for TBLG.
 
@@
 The SLG model fitting should take another eight months.
-The first three months will be dedicate to reproducing the on-site Hubbard model from that work.
-Inclusion of long range Coulomb interactions into the effective theory should take another three months.
+The first two months will be dedicate to reproducing the on-site Hubbard model from that work.
+Including long range Coulomb and electron-phonon interactions should take another four months.
 The last two months will be geared towards finite size extrapolation to ensure the final model is applicable to bulk systems.

### 2019-09-21 draft.tex
small change

@@
 \author{Shivesh Pathak}
-\title{Developing effective theories for single and bilayer graphene using density matrix downfolding}
+\title{Developing model Hamiltonians for single and bilayer graphene using density matrix downfolding}
 
@@
 Yet unresolved is a paradigm for building many-body model Hamiltonians which can describe strongly correlated systems.
-Band and Fermi liquid theory have been successful in accurately modeling systems where effective interactions are negligible at low energy scales.
+Band and Fermi liquid theory have been successful in modeling systems where effective interactions are negligible at low energy scales.
 These theories, however, fail in describing the low-energy excitations of strongly correlated systems like high Tc superconductors and TBLG.
-Minimal many-body models like the Hubbard \cite{Hubbard1963}, Kanamori \cite{10.1143/PTP.30.275}, t-J \cite{Chao_1977} and Heisenberg models have been proposed to describe such systems, their effectiveness in describing realistic materials is unclear.
+Minimal many-body models like the Hubbard \cite{Hubbard1963}, Kanamori \cite{10.1143/PTP.30.275},  and t-J \cite{Chao_1977} models have been proposed to describe such systems.
+However, their effectiveness in describing realistic materials is unclear as they have no direct link to the \textit{ab initio} Hamiltonian.
 
 In an attempt to build accurate many-body Hamiltonians for realistic systems, a framework for model development using solutions from \textit{ab initio} simulations has emerged.
-Common approaches involve density functional theory (DFT) \cite{PhysRev.140.A1133, PhysRevLett.87.047003} alongside random phase approximation or other methods \cite{PhysRevB.62.R16219, PhysRevB.70.195104, PhysRevB.88.075106}.
+Common approaches use density functional theory (DFT) \cite{PhysRev.140.A1133, PhysRevLett.87.047003} alongside the random phase approximation \cite{PhysRevB.62.R16219, PhysRevB.70.195104, PhysRevB.88.075106}.
 L\"owdin methods and canonical transformations have also been used to develop effective theories \cite{doi:10.1063/1.4802766, PhysRevA.81.013402, doi:10.1063/1.2196410}.
@@
 \begin{equation}
-\hat{H}_\text{ab} = \sum_i -\frac{1}{2} \nabla_i^2 + \sum_{iI}\frac{Z_I}{|r_i - R_I|} + \frac{1}{2} \sum_{ij} \frac{1}{|r_i - r_j|} + E_\text{ion}(\{R_I\})  
+\hat{H}_\text{ab} = \sum_i -\frac{1}{2} \nabla_i^2 + \sum_{iI}\frac{Z_I}{|r_i - R_I|} + \frac{1}{2} \sum_{ij} \frac{1}{|r_i - r_j|} + hat{H}_\text{ion}(\{R_I\})  
 \label{eq:Hab}
 \end{equation}
-where index $i,j$ correspond to electrons, $I$ to ions, $E_\text{ion}$ corresponds to the total ionic energy, and the unit of energy is Hartree (Ha).
+where index $i,j$ correspond to electrons, $I$ to ions, $\hat{H}_\text{ion}$ corresponds to the total ionic energy, and the unit of energy is Hartree (Ha).
 The goal is to develop an effective Hamiltonian $\hat{H}_\text{eff}$ defined on a low-energy subspace $\mathcal{LE} \subset \mathcal{H}$ which can accurately reproduce the lowest lying eigenstates and energies of $\hat{H}_\text{ab}$.
@@
 Deviation from equality manifests as an error $\epsilon[\Psi] \equiv E_\text{ab}[\Psi] - E_\text{eff}[\Psi]$ which quantifies the model quality.
-In short, the key insight of DMD is that the downfolding $\hat{H}_\text{ab}$ of can be cast to a functional appoximation of $E_\text{ab}[\Psi]$ on $\mathcal{LE}$.
+In short, the key insight of DMD is that downfolding $\hat{H}_\text{ab}$ can be cast to a functional appoximation of $E_\text{ab}[\Psi]$ on $\mathcal{LE}$.
 
@@
 Consider a set of trial wave functions with varying nodal structures.
-The final projected $\Psi_\text{DMC}$ for these different trial functions will be the lowest energy states in $\mathcal{H}$ for the given nodal structures.
+The final projected $\Psi_\text{DMC}$ for these trial functions will be the lowest energy states in $\mathcal{H}$ for the given nodes.
 If the initial trial wave functions are appropriately chosen, the final projected states will be within the target $\mathcal{LE}$.

### 2019-09-22 draft.tex
updated a line;

@@
 It will also help develop understanding of interaction effects in graphene which depend on the details of $V(r)$, like edge magentism \cite{PhysRevB.95.195420} and non-local electrodynamics \cite{PhysRevB.90.045137}.
-Additionally, by introducing displacement dependent parameters, one can determine whether phonons in graphene serve to simply renormalize interactions, or drastically alter them.
+Additionally, by introducing displacement dependent parameters, one can determine how phonons affect the effective electronic couplings in graphene.
 This determination is integral to understanding effects of strain on electronic properties as seen in experiment \cite{Naumis_2017}.

### 2019-09-24 draft.tex
updated
need to change next steps for cuo

@@
 Here I propose a thesis plan for developing atomic scale many-body model Hamiltonians for single- and simple bi-layer graphene systems within the DMD framework using high accuracy \textit{ab initio} quantum Monte Carlo (QMC) calculations.
-The results of my thesis will contribute significantly to the understanding of low-energy effective interactions in single and bilayer graphene, and will provide a baseline for a continuing study of the twisted bi-layer graphene (TBLG) system.
-To demonstrate my preparedness, I present progress on a many-body model for the CuO molecule and my work in QMC methods development.
+I present progress on a many-body model for the CuO molecule, development and testing of QMC wave functions and parameter derivatives, and work on a new QMC package PyQMC.
+The results of my thesis will contribute to the understanding of low-energy effective interactions in single and bilayer graphene, and will provide a baseline for a continuing study of the twisted bi-layer graphene (TBLG) system.
+
 \end{abstract}
@@
 
-Instead of jumping head first into TBLG, however, I propose a two year plan for modeling the AA and AB stacked BLG systems using DMD with \textit{ab initio} fixed-node diffusion Monte Carlo (FN-DMC) calculations.
+To approach this complex problem systematically, I propose a two year plan for modeling the AA and AB stacked BLG systems using DMD with \textit{ab initio} fixed-node diffusion Monte Carlo (FN-DMC) calculations.
 I will first develop a model for single-layer graphene (SLG) and then extend this to the BLG systems by introducing inter-layer couplings.
@@
 \subsection{Density Matrix Downfolding (DMD)}
-The starting point of model development for materials is the \textit{ab initio} Hamiltonian defined on the Hilbert space $\mathcal{H}$
+The starting point of model development for materials is the \textit{ab initio} Hamiltonian defined on the \textit{ab initio} Hilbert space $\mathcal{H}$
 \begin{equation}
@@
 
-While seemingly disadvantageous, the variational nature of FN-DMC plays
+While seemingly disadvantageous, the fixed-node approximation plays
 a key role in generating the low-energy states needed for DMD.
@@
 
-A shell sampling scheme was used to select coefficients $p_j$ for the trial wave functions.
+A random sampling scheme was used to select coefficients $p_j$ for the trial wave functions.
 I began by fixing the coefficient $p_0 = \sqrt{w}$ where $w \in \{1.0, 0.8, ..., 0.2\}$. 
@@
 The parameter gradient of a wave function, $\langle \Psi(\vec{p}) | \Psi_p (\vec{p}) \rangle$ where $| \Psi_p (\vec{p}) \rangle \equiv \partial_{p} |\Psi (\vec{p}) \rangle$ , is a quantity that pops up frequently in QMC, from optimization algorithms \cite{Toulouse2007} to density matrix downfolding \cite{2019APS..MARF18008W}.
-Unfortunately, a naive Monte Carlo estimate of the gradient has an infinite fourth moment due to divergences near the nodal surface of $|\Psi \rangle$, so MC error estimates for the gradient are very unstable.
+Unfortunately, a na\"{i}ve Monte Carlo estimate of the gradient has an infinite fourth moment due to divergences near the nodal surface of $|\Psi \rangle$, so MC error estimates for the gradient are very unstable.
 I proved that the fourth moment can be made finite when using a stable estimator 
@@
 This package is oriented towards methods development, and will be the platform for most of my future DMD calculations.
-I implemented multi-Slater wave functions, ran consistency checks against QWalk, and worked on execution speedups for costly methods like VMC and wave function optimization.
+I implemented multi-Slater wave functions, ran consistency checks against QWalk, and worked on 2 $-$ 3 fold execution speedups for costly methods like VMC and wave function optimization.
 I plan on implementing 2-body RDM evaluation and Ewald summations too, thus providing the baseline tools for DMD on molecular and periodic systems.
@@
 \begin{equation}
-\vec{p}^* = \text{argmin} \ \langle \Psi(\vec{p}) | \hat{H}_\text{ab} | \Psi(\vec{p}) \rangle - \sum_k [\lambda_k \langle \Psi(\vec{p}) | d_k - d_{k}^0 | \Psi(\vec{p}) \rangle]^2.
+\vec{p}^* = \text{argmin} \ \langle \Psi(\vec{p}) | \hat{H}_\text{ab} | \Psi(\vec{p}) \rangle - \sum_k [\lambda_k \langle \Psi(\vec{p}) | d_k - d_{k}^* | \Psi(\vec{p}) \rangle]^2.
 \end{equation}
@@
 First, one can determine, at least for the simplest orientations, whether the inter-layer coupling takes the form of a hopping or an interaction term.
-Quantitative details of the interlayer coupling spatial dependence, which are intimately related to the strongly correlated behavior in TBLG, can also begin to be addressed.
+Quantitative details of the interlayer coupling spatial dependence, which are intimately related to the strongly correlated behavior in TBLG, can also be addressed.
 Further, when combined with the SLG model, it will be possible to determine how the intra-layer model is affected by the presence of a neighboring graphene sheet.

### 2019-09-24 draft.tex
addressed lucas comments. going to review + send tomorrow

@@
 These theories, however, fail in describing the low-energy excitations of strongly correlated systems like high Tc superconductors and TBLG.
-Minimal many-body models like the Hubbard \cite{Hubbard1963}, Kanamori \cite{10.1143/PTP.30.275},  and t-J \cite{Chao_1977} models have been proposed to describe such systems.
-However, their effectiveness in describing realistic materials is unclear as they have no direct link to the \textit{ab initio} Hamiltonian.
+Minimal models like the Hubbard \cite{Hubbard1963}, Kanamori \cite{10.1143/PTP.30.275},  and t-J \cite{Chao_1977} models have been proposed to describe such systems.
+However, their effectiveness in describing materials is unclear as they have no direct link to the \textit{ab initio} Hamiltonian.
 
@@
 Since model quality is reported, DMD allows for direct comparison between different model \textit{ansatzes} as well, an ability which can be used to deduce important model terms using \textit{ab initio} calculations.
-DMD has been used to develop accurate many-body models for benzene, a 1-d hydrogen chain, and an on-site Hubbard model for graphene \cite{Zheng2017, Wagner2015}.
+DMD has been used to develop accurate many-body models for benzene, a 1-d hydrogen chain, and a Hubbard model for graphene \cite{Zheng2017, Wagner2015}.
 
-The utility of DMD is made apparent when studying the recent surge in model development for TBLG.
-\textit{Ab initio} model development has so far been restricted to tight-binding models which accurately describe single particle effects like band flattening, but do not address interactions \cite{Fang, Fang2016, PhysRevX.8.031088, PhysRevX.8.031087}.
-Minimal interacting models have been proposed using traditional theoretical methods, but without a link to the \textit{ab initio} Hamiltonian it is unclear whether they can describe the real system \cite{PhysRevB.98.045103, PhysRevB.100.125116, PhysRevX.8.031089}.
+The utility of DMD is made clear when studying the recent surge in model development for TBLG.
+\textit{Ab initio} models have been restricted to tight-binding models which accurately describe single particle effects like band flattening, but do not address interactions \cite{Fang, PhysRevX.8.031087}.
+Minimal interacting models have been proposed using traditional theoretical methods, but without a link to the \textit{ab initio} Hamiltonian it is unclear whether they can describe the real system \cite{PhysRevB.98.045103, PhysRevB.100.125116}.
 DMD can bridge the gap between these methods by providing a framework to build an accurate many-body model directly from \textit{ab initio} simulations on TBLG.
@@
 DMC, however, suffers from a fermion sign problem which is alleviated via a fixed-node approximation.
-Under the fixed-node approximation the nodal surface of $\Psi_\text{DMC}(\tau, R)$ is forced to match that of the initial trial wave function for all $\tau$.
+Under this approximation the nodal surface of $\Psi_\text{DMC}(\tau, R)$ is forced to match that of $\Psi_T(R)$ for all $\tau$.
 This approximation makes FN-DMC variational, and will only return the exact ground state of $\hat{H}$ if the nodal surfaces of $|\Psi_T\rangle$ and $|\Phi_0\rangle$ are identical.
@@
 \Psi=e^{J(\vec{\alpha})} \sum_I e^{\hat{\Theta}(\vec{\theta_I})} C_I |D_I (\{ \phi\})\rangle
+\label{eq:MSJNO}
 \end{equation}
@@
 I assessed the efficiency and compactness of the new function by calculating the ground state energy of a C$_2$ molecule using FN-DMC and comparing to the results when using multi-Slater-Jastrow trial functions with orthogonal determinants (MSJ+O).
-The workflow involved constructing the un-optimized trial wave functions, optimizing the parameters using an energy optimization method \cite{Toulouse2007}, and finally using the optimized trial functions in FN-DMC calculations.
+The workflow involved constructing the state \eqref{eq:MSJNO}, optimizing the parameters using an energy optimization method \cite{Toulouse2007}, and finally using the optimized trial functions in FN-DMC calculations.
 The average improvement in the FN-DMC energy with the MSJ+O trial wave functions was $\langle E_{MSJ} - E_{MSJ,O}\rangle =$ 0.14 eV with an additional reduction of 0.032 eV when using the MSJ+NO trial wave function.
@@
 \paragraph{Next steps}
-I have shown that my model has comparable accuracy to modern computational techniques.
-This poses the question, however: why use DMD if simpler methods are just as accurate?
-In the next steps of the project I will address this question by demonstrating the unique interpretability of a model.
-I plan on accomplishing this by refitting my model after dropping specific interactions, like $n_{4s,\uparrow} n_{4s,\downarrow}$, and quantitatively evaluating the effect on the model spectra, eigenstates, and renormalization of non-interacting parameters.
-In doing so, I will be able to describe the precise effects of certain interactions on the low-energy spectra of CuO, a level of detailed interpretation unavailable to methods like CCSD(T).
+While PCR was demonstrated here, I also tried feature selection using LASSO and Ridge regression. 
+I found a high quality of fit on the sampled data using all three methods, $R^2$ = 0.96 - 0.98, but the model parameters, eigenstates and energies differed significantly between them.
+I realized this was because portions of the low-energy space were not accessible with the sampling scheme described above.
+%Rephrase
+In particular, the DFT calculations were unable to generate two single particle excitations on the ground state, $\text{Cu } 3d_\pi \rightarrow \text{O } 2p_\pi$ and $\text{Cu } 3d_{z^2} \rightarrow \text{O } 2p_\pi$.
+Therefore, eigenstates corresponding to these excitations were not constrained by the sampled data, leading to large variations in the eigenvalues between different feature selection schemes.
+
+A new variational Monte Carlo algorithm, described later, was developed by our group to address the need for a method which can target such excitations.
+The next step in the project involves using this algorithm to fill in the missing parts of the sample space.
+By doing so, I hope to demonstrate that DMD is robust to the choice of feature selection scheme if the low-energy space is fully sampled.
 
@@
 
-I also made significant contributions to a new QMC package called PyQMC. 
-This package is oriented towards methods development, and will be the platform for most of my future DMD calculations.
+I also made significant contributions to the new QMC package PyQMC. 
+This package is oriented towards methods development, and will be the platform for my DMD calculations.
 I implemented multi-Slater wave functions, ran consistency checks against QWalk, and worked on 2 $-$ 3 fold execution speedups for costly methods like VMC and wave function optimization.
@@
 where the indices correspond to the $\pi$ orbitals on the copper atoms.
-I will consider both long range hopping and density-density interaction terms extending across the computational unit cell, with the final goal of developing position dependent parameter values $t(r), V(r)$.
-Phononic contributions will be incorporated by fitting a sequence of models to altered geometries, leading to displacement dependent parameters similar to a spin-phonon model for the cuprates \cite{Wagner2014}.
+I will consider long range hopping and density-density interaction terms extending across the computational unit cell, with the final goal of developing position dependent parameter values $t(r), V(r)$.
+Phononic contributions will be included by fitting a sequence of models to altered geometries, leading to displacement dependent parameters similar to a spin-phonon model for cuprates \cite{Wagner2014}.
 
 An important result from this study would be an accurate evaluation of $V(r)$ directly from \textit{ab initio} simulations.
-This result would be able to validate commonly used models for the spatial dependence, like the Ohno form \cite{Ohno1964}.
+This result would be able to validate common models like the Ohno form \cite{Ohno1964}.
 It will also help develop understanding of interaction effects in graphene which depend on the details of $V(r)$, like edge magentism \cite{PhysRevB.95.195420} and non-local electrodynamics \cite{PhysRevB.90.045137}.
@@
 \begin{multicols}{2}
-\begin{spacing}{0.5}
-{\footnotesize
+\begin{spacing}{0.2}
+{\fontsize{5}{12}\selectfont
 \bibliography{biblio}

### 2019-09-25 draft.tex
finished!

@@
   {}{}
-\author{Shivesh Pathak}
+\usepackage{datetime}
+\newdate{date}{09}{10}{2019}
+\date{\displaydate{date}}
 \title{Developing model Hamiltonians for single and bilayer graphene using density matrix downfolding}
@@
 \begin{equation}
-\hat{H}_\text{ab} = \sum_i -\frac{1}{2} \nabla_i^2 + \sum_{iI}\frac{Z_I}{|r_i - R_I|} + \frac{1}{2} \sum_{ij} \frac{1}{|r_i - r_j|} + hat{H}_\text{ion}(\{R_I\})  
+\hat{H}_\text{ab} = \sum_i -\frac{1}{2} \nabla_i^2 + \sum_{iI}\frac{Z_I}{|r_i - R_I|} + \frac{1}{2} \sum_{ij} \frac{1}{|r_i - r_j|} + \hat{H}_\text{ion}(\{R_I\})  
 \label{eq:Hab}
@@
 \paragraph{Solutions of the model}
-The final model in \ref{fig:cuo_regr}c was solved by exact diagonalization.
+The final model in Figure \ref{fig:cuo_regr}c was solved by exact diagonalization.
 Shown in Figure \ref{fig:cuo_regr}d is a comparison between the excitation energies calculated from the model and those seen in the latest anion photoelectron spectroscopy measurements \cite{Wu1997}.
@@
 I realized this was because portions of the low-energy space were not accessible with the sampling scheme described above.
-%Rephrase
-In particular, the DFT calculations were unable to generate two single particle excitations on the ground state, $\text{Cu } 3d_\pi \rightarrow \text{O } 2p_\pi$ and $\text{Cu } 3d_{z^2} \rightarrow \text{O } 2p_\pi$.
-Therefore, eigenstates corresponding to these excitations were not constrained by the sampled data, leading to large variations in the eigenvalues between different feature selection schemes.
+In particular, the DFT method used was unable to generate determinants corresponding to two single particle excitations on the ground state, $\text{Cu } 3d_\pi \rightarrow \text{O } 2p_\pi$ and $\text{Cu } 3d_{z^2} \rightarrow \text{O } 2p_\pi$.
+Therefore, eigenstates corresponding to these excitations were not constrained by sampled data, leading to large variations in the eigenvalues between different feature selection schemes.
 
